### 2019-10-11 risk-percption.Rmd
recommit rmd in new repo
also add bayesian learning

---
title: "risk-perception"
author: "Nick Gauthier"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(zoo)
```

The basic risk assessment agorithm. Let's look at the risk of a drought year being 0.2. In a wet year you get 3 wheat and 2 barley, and in a dry year you get 0 wheat and 1 barley. What does that look like?

```{r}
wheat_dry <- 0
wheat_wet <- 1.6
barley_dry <- 0.93
barley_wet <- 1.18

calc_yield <- function(wheat_prop, risk) {
  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
  barley + wheat
}

util_wet <- wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)
util_dry <- wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)

util_wet * (1 - risk)

calc_yield(.92, .5)

calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)

calc_yield(1, .2)

plot(calc_yield((0:10)/10, .2))

expand_grid(wheat_prop = (0:10)/10,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

expand_grid(wheat_prop = (0:10)/1,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(risk, util, color = type)) + geom_line() + facet_wrap(~wheat_prop)

calc_prop <- function(risk) {
  pb <- barley_wet * (1 - risk) - barley_dry * risk
  pw <- wheat_wet * (1 - risk) - wheat_dry * risk

  pb / (pb + pw) 
}

plot(calc_prop((0:10)/10))
```

```{r}
calcNE <- function(risk){
  (118 - 25 * risk) / (278 - 185 * risk)
}

(barley_wet - .2 * (barley_dry + wheat_dry + wheat_wet + barley_wet)) / (barley_wet + wheat_wet)

calcNE(1)

plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)
```

so we're imagining a yield of 1 for barley and 2 for wheat, with equal prportions of each. then in the next time step we'd revise the proportion to take away .25
```{r}
reallocate <- function(wheat_prop, risk) {
  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
}

accumulate(1:10, ~reallocate(.x), .init = .4) %>% plot

# so this works as long as you're between 0 and 1
```

What about memory of risk?
```{r}
test <- runif(7) < .5

remember <- function(x, y) {
  x +  * .1 - .01 * x
}

1

(1-1 )/1

remember
accumulate(test, ~remember(.x, .y)) %>% plot

```

```{r}

prior <- function(m,n){
a = n * m
b = n * (1 - m)
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}

prior(.5, 2) %>% plot
```

```{r}
likelihood <- function(N,Y){
a <- Y + 1
b <- N - Y + 1
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}

likelihood(1, 1)%>% plot
```

```{r}
posterior <- function(m,n,N,Y){
a <- Y + (n*m) -1
b <- N - Y + (n*(1-m)) - 1
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}

-1
1 - 0 - 1

dbeta(seq(0,1,0.005), -1, 0)
prior(.5, 2) %>% plot
posterior(.5, 3, 1, 1) %>% plot
```

```{r}
test <- tibble(agent = 1:5,
       alpha = 1.5, #runif(5),
       beta = 1.5) %>%
  mutate(mu = alpha / (alpha + beta))

test2 <- runif(50) < .2

a <- Y + (n*m) -1
b <- N - Y + (n*(1-m)) - 1
a = n * m
b = n * (1 - m)

.5 * 3

learn <- function(x, event) {
  x %>%
    mutate(alpha = event + alpha,
           beta = beta + 1 - event,
           mu = alpha / (alpha + beta),
           estimate = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
}
learn(test, TRUE)

accumulate(test2, ~learn(.x, .y), .init = test) %>% bind_rows %>% pull(estimate) %>% plot
```

so now we get the same yields, but the proportions are different
```{r}
(1 *  0.625 - 2 * .375) * .375 * .625
```

```{r}
1- .375
```

```{r}
-1 * 0.0625 * (1 - 0.0625)
```

```{r}
yield_max_barley <- 2500
yield_max_wheat <- 3500
kcal_barley <- 3000
kcal_wheat <- 3540

calc_wheat_yield <- function(rainfall) {
  yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
}

calc_barley_yield <- function(rainfall) {
  yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51)  # annual rainfall impact on yields
}

tibble(rainfall = (0:200) / 100,
       barley = calc_barley_yield(rainfall),
       wheat = calc_wheat_yield(rainfall)) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + geom_line()

```

```{r}
# generate some forgetting curves
forgetting_curves <- tibble(x = c(.01, .1, 1, 0)) %>%
  mutate(memory = map(x, ~exp(-1 * . *1:50))) %>%
  unnest(cols= c(memory)) %>%
  group_by(x) %>%
  mutate(time = 1:n())

ggplot(forgetting_curves, aes(time, memory)) +
  geom_line(aes(color = as.factor(x)), size = 1.2) +
  labs(title = 'Forgetting curves', x = 'Years', y = 'Relative memory strength\n') +
  scale_color_viridis_d(name = 'Decay rate') +
  theme_classic()
```

```{r}
farm <- function(households){
  households %>%
    mutate(yield = climatic_yield, #* n_inhabitants ^ labor_elasticity,
           yield_memory = yield, #map2(yield_memory, yield, remember),
           harvest = land * yield * .5 - land * sowing_rate) #%>%  # *.5 is for fallow
  #select(-yield)
}

remember <- function(yield_memory, yield){
  # rnorm(1, yield, yield * 0.0333) %>%  #memory is fuzzy
  append(yield_memory[-length(yield_memory)], yield, after = 0) # remove the last entry in the vector and add new yield to the begining
}

#Agents use the peak-end rule when accessing memory.
peak_end <- function(x){
  map_dbl(x, ~mean(c(.x[1], min(.x))))
}

```

```{r}
obs.na <- read.csv('salihli_p.csv')[10:620, 5]  / 10  # raw data in mm*10. subset time series as water years (i.e. first observation is october, last is september)

get.winter <- function(x){seq(0, length(x) - 12, 12) %>%
    sapply(function(y) y + 1:6) %>%
    as.numeric %>% extract(x, .)}# function to get Oct-Mar values from
obs<- obs.na %>% 
  rev %>% 
  na.locf %>% 
  rev %>% 
  `+`(na.locf(obs.na)) %>% 
  `/`(2) %>%  # fill NA with avg of adjacent values
  get.winter

# Now import TraCE-21k GCM data. Extract a time series for the grid cell covering the Salihli weather station. Add the model-output convective and large-scale precipitation variables to get total precipitation. Convert from m/s to mm/month. Subset time series to the calibration period (10 BP - AD 1940) and prediction period (10000 BP - 11 BP).

salihli.pt <- c(28.14, 38.48) %>% matrix(nrow = 1) %>% SpatialPoints  # create a point in Salihli to extract data from GCM

pc <- list.files('GCM Data/PRECC', full.names = T) %>%  # get a list of all GCM output files
  lapply(brick) %>%                                 # import GCM files as raster bricks
  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
  unlist                                            # combine into a single vector

pl <- list.files('GCM Data/PRECL', full.names = T) %>%  # get a list of all GCM output files
  lapply(brick) %>%                                 # import GCM files as raster bricks
  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
  unlist                                            # combine into a single vector

pt <- pc %>% add(pl) %>% multiply_by(60 * 60 * 24 * 30 * 1000)  # add precip components and convert from m/s to mm/month

cal <- pt %>% extract(122278:122877) %>% get.winter

pred <- pt %>% extract(2398:122397) %>% get.winter

pred.byfifty <- pred %>% seq_along %>% divide_by(300) %>% ceiling %>% split(pred,.)

calc.cdf <- function(x){
  pred <- x
  ct <- CDFt(obs, cal, pred)
  ct$DS
}
monthly.recon <- lapply(pred.byfifty, calc.cdf) %>% unlist
annual.recon <- monthly.recon %>% seq_along %>% divide_by(6) %>% ceiling %>% split(monthly.recon,.) %>%
  lapply(sum) %>% unlist %>% data.frame(date = seq(-10000,-1,1), p = ., color = cut(., c(0,200,300, 800)))

```

### 2019-10-11 risk-percption.Rmd
crop yields

@@
 library(zoo)
+```
+
+```{r}
+test <- tibble(agent = 1:5,
+       alpha = 1.5, #runif(5),
+       beta = 1.5) %>%
+  mutate(mu = alpha / (alpha + beta))
+
+test2 <- c((runif(25) < .2),(runif(25) < .5), runif(25) < .1)
+tibble(time = 1:75,
+       drought = test2) %>%
+  mutate(period = round(time, -1)) %>%
+  add_count(period) %>%
+  filter(period >= 10) %>%
+  group_by(period) %>%
+  summarise(drought = sum(drought)) %>%
+  ggplot(aes(period, drought)) +
+  geom_col()
+```
+
+```{r}
+learn <- function(agents, event) {
+  agents %>%
+    mutate(alpha = event + alpha,
+           beta = beta + 1 - event,
+           mu = alpha / (alpha + beta),
+           estimate = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
+}
+
+accumulate(test2, ~learn(.x, .y), .init = test) %>% bind_rows(.id = 'time') %>% qplot(as.numeric(time), estimate, data = ., geom = 'point')
+```
+
+```{r}
+learn(test, TRUE)
 ```
@@
 
-What about memory of risk?
-```{r}
-test <- runif(7) < .5
-
-remember <- function(x, y) {
-  x +  * .1 - .01 * x
-}
-
-1
-
-(1-1 )/1
-
-remember
-accumulate(test, ~remember(.x, .y)) %>% plot
-
-```
-
 ```{r}
@@
 
-```{r}
-test <- tibble(agent = 1:5,
-       alpha = 1.5, #runif(5),
-       beta = 1.5) %>%
-  mutate(mu = alpha / (alpha + beta))
-
-test2 <- runif(50) < .2
-
-a <- Y + (n*m) -1
-b <- N - Y + (n*(1-m)) - 1
-a = n * m
-b = n * (1 - m)
-
-.5 * 3
-
-learn <- function(x, event) {
-  x %>%
-    mutate(alpha = event + alpha,
-           beta = beta + 1 - event,
-           mu = alpha / (alpha + beta),
-           estimate = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
-}
-learn(test, TRUE)
-
-accumulate(test2, ~learn(.x, .y), .init = test) %>% bind_rows %>% pull(estimate) %>% plot
-```
-
 so now we get the same yields, but the proportions are different
@@
 
-calc_wheat_yield <- function(rainfall) {
-  yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
+calc_yield <- function(rainfall, crop = 'wheat') {
+  if (crop == 'wheat') {
+    yield <- yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
   }
 
-calc_barley_yield <- function(rainfall) {
-  yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51)  # annual rainfall impact on yields
+  if(crop == 'barley') {
+    yield <- yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51) 
+  }
+
+  return(yield)
 }
@@
 tibble(rainfall = (0:200) / 100,
-       barley = calc_barley_yield(rainfall),
-       wheat = calc_wheat_yield(rainfall)) %>%
+       barley = calc_yield(rainfall, 'barley'),
+       wheat = calc_yield(rainfall, 'wheat')) %>%
   gather(crop, yield, barley:wheat) %>%
-  ggplot(aes(rainfall, yield, color = crop)) + geom_line()
-
+  ggplot(aes(rainfall, yield, color = crop)) + 
+  geom_line() +
+  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
+  geom_hline(yintercept = yield_max_barley * kcal_barley)
 ```

### 2019-10-15 risk-percption.Rmd
plots

@@
 
-calc_yield <- function(wheat_prop, risk) {
+calc_payoff <- function(wheat_prop, risk) {
   barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
@@
 
-util_wet <- wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)
-util_dry <- wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)
+tibble(wheat_prop)
+```
 
-util_wet * (1 - risk)
-
-calc_yield(.92, .5)
+```{r}
 
@@
 calc_ratio(.2)
+```
 
-calc_yield(1, .2)
+So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
+```{r}
+test_dat <- expand_grid(wheat_prop = (0:10)/10,
+       risk = seq(0, 1, 0.2)) %>%
+  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
+         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
+         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
+         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk)+ barley_dry * risk),
+         payoff = calc_payoff(wheat_prop, risk))
 
-plot(calc_yield((0:10)/10, .2))
+test_dat %>%
+  group_by(risk) %>%
+  filter(payoff == max(payoff)) %>%
+  ggplot(aes(risk, wheat_prop)) + geom_point()
 
-expand_grid(wheat_prop = (0:10)/10,
-       risk = (0:10)/10) %>%
-  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
-         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
-  gather(type, util, util_wet:util_dry) %>%
+test_dat %>%
+  select(-(util_dry:util_wet)) %>%
+  gather(type, util, util_wheat:util_barley) %>%
+    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
+test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  gather(type, util, util_dry:util_wet) %>%
     ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
 
-expand_grid(wheat_prop = (0:10)/1,
-       risk = (0:10)/10) %>%
-  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
-         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
-  gather(type, util, util_wet:util_dry) %>%
-  ggplot(aes(risk, util, color = type)) + geom_line() + facet_wrap(~wheat_prop)
+test_dat %>%
+  ggplot(aes(wheat_prop, util_wet + util_dry)) + geom_point() + facet_wrap(~risk)
 
@@
 plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)
+
+util_wet <- wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)
+util_dry <- wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)
+
+tibble()
+
 ```
@@
   geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
-  geom_hline(yintercept = yield_max_barley * kcal_barley)
+  geom_hline(yintercept = yield_max_barley * kcal_barley) +
+  geom_hline(yintercept = 1016525* 5, color ='red') +
+  geom_vline(xintercept = 0.3)
+```
+
+How do we define the "drought year" cutoff? Below 300mm there isn't enough wheat growing in a hectare to feed a family of 5, which seems appropriate.
+```{r}
+test <- tibble(time = 1:100,
+       rain = rnorm(100, mean = 1, sd = 1) %>% pmax(0),
+       yield = calc_yield(rain))
+
+ggplot(test, aes(time, yield)) + geom_point() + geom_hline(yintercept = 1016525)
+
+test %>%
+  mutate(drought = if_else(yield < 5e+6, TRUE, FALSE))%>%
+  filter(drought == TRUE)
 ```

### 2019-10-23 risk-percption.Rmd
lpsolve code

@@
 barley_wet <- 1.18
-
-calc_payoff <- function(wheat_prop, risk) {
-  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
-  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
-  barley + wheat
-}
-
-tibble(wheat_prop)
 ```
@@
 ```{r}
+library(lpSolve) #Loading lpSolve library
+risk <- 0.6
+wheat_exp <- wheat_wet * (1 - risk) + wheat_dry * risk
+barley_exp <- barley_wet * (1 - risk) + barley_dry * risk
+obj.fun <- c(wheat_exp, barley_exp)
 
+constr=matrix(c(1,1,1,0,0,1,wheat_exp,barley_exp),ncol = 2, byrow = TRUE) #Loading the constraints
+constr.dir=c("=",">=",">=",">=")
+constr.rhs=c(1,0,0,1)
+
+mod=lp("max",obj.fun,constr,constr.dir,constr.rhs,compute.sens = TRUE) #Using lp() to solve our problem
+
+mod$solution
+```
+
+```{r}
+calc_payoff <- function(x, drought) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
+
+  x %>%
+    mutate(yield = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop))
+}
+
+update_prop <- function(x, aspiration) {
+  x %>%
+    mutate(new_prop = round(runif(n()), 2),
+           wheat_prop = if_else(yield > aspiration, wheat_prop, new_prop)) %>%
+    select(-new_prop)
+}
+
+test3 <- tibble(agent = 1:1000) %>%
+  mutate(wheat_prop = round(runif(n()), 2)) %>%
+  calc_payoff(drought = TRUE)
+test3 %>% update_prop(aspiration = .6)
+
+accumulate((runif(25) < .3), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .5), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(as.numeric(time), yield, group = agent)) + geom_line(alpha = .2)
+
+```
+
+```{r}
+library(ggridges)
+accumulate((runif(25) < .2), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .6), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(wheat_prop, y = as.numeric(time), group = as.numeric(time))) +
+  geom_density_ridges() +
+  theme_minimal()
+```
+
+```{r}
 calc_ratio <- function(risk) {
@@
 So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
+
+```{r}
+tibble(wheat_prop = (0:10)/10,
+       risk = 0.5,
+       util_wet = wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop),
+      util_dry = wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop),
+      util_expected = util_wet * (1 - risk) + util_dry * risk,
+      min_yield = pmin(util_wet, util_dry)) %>%
+    ggplot(aes(util_expected, min_yield)) + geom_line() +
+  geom_hline(yintercept = .6, linetype = 2)
+```
+
+```{r}
+tibble(wheat_prop = (0:10)/10,
+       risk = 0.5,
+       util_wet = wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop),
+      util_dry = wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop),
+      util_expected = util_wet * (1 - risk) + util_dry * risk) %>%
+    gather(type, util, util_wet:util_expected) %>%
+    ggplot(aes(wheat_prop, util, color = type)) + geom_line()
+```
+
 ```{r}
@@
 test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  gather(type, util, c(util_dry, util_wet, payoff)) %>%
+    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
+
+test_dat %>%
   group_by(risk) %>%
@@
   gather(type, util, util_wheat:util_barley) %>%
-    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
-test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  gather(type, util, util_dry:util_wet) %>%
     ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
@@
 plot(calc_prop((0:10)/10))
+```
+
+```{r}
+tibble(x)
 ```
@@
 tibble()
+
+plot(seq(.1,1,.1) * (1 - seq(.1,1,.1)))
 

### 2019-10-23 risk-percption.Rmd
rearrange opener with more explanatory plots

--- 2019-10-15 risk-percption.Rmd
@@
 library(tidyverse)
-library(zoo)
+library(lpSolve)
+
+#library(zoo)
+```
+
+Start with a population of agents with the same subjective beliefs about drought risk.
+
+```{r}
+agents <- tibble(id = 1:5, 
+                 alpha = 1.5,
+                 beta = 1.5,
+                 risk = alpha / (alpha + beta))
 ```
@@
 ```{r}
-test <- tibble(agent = 1:5,
-       alpha = 1.5, #runif(5),
-       beta = 1.5) %>%
-  mutate(mu = alpha / (alpha + beta))
+agents
+```
+
+And payoffs for different crop types in different years.
+```{r}
+wheat_dry <- 0
+wheat_wet <- 1.6
+barley_dry <- 0.93
+barley_wet <- 1.18
+
+wheat_dry <- .2
+wheat_wet <- .4
+barley_dry <- 0.6
+barley_wet <- .3
+```
+
+The expected yield is thus calculated as.
+```{r}
+expected_utility <- function(risk) {
+  risk * (wheat_dry + barley_dry) + (1 - risk) * (wheat_wet + barley_wet) 
+}
+```
+
+```{r}
+expected_utility(0.2) # assuming a drought risk of 0.2
+```
+
+Given these payoffs, what's the point of indifference between the two crop types?
+
+```{r, echo = FALSE}
+plt_dat <- tibble(risk = seq(0, 1, .1),
+       barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
+      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
+  gather(type, utility, barley_exp:wheat_exp)
+```
+
+The pure maximin strategy is to plant barley:
+```{r}
+max(wheat_dry, barley_dry)
+```
+
+```{r, echo = FALSE}
+indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
+                 utility = risk * wheat_dry + (1 - risk) * wheat_wet)
+
+plt_dat %>%
+  ggplot(aes(risk, utility)) +
+  geom_hline(yintercept = indiff$utility, linetype = 2)+
+  geom_vline(xintercept = indiff$risk, linetype = 2) +
+  geom_line(aes(color = type), size = 1.2) +
+  geom_point(data = indiff, color = 'black', size = 3) +
+  geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
+  theme_classic()
+```
+
+The maximin strategy here is to plant barley all the time, but this entails a huge opportunity cost over planting wheat or a mix of the two
+
+```{r, echo = FALSE}
+plt_dat2 <- tibble(wheat_prop = seq(0, 1, .1),
+       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
+      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
+  gather(type, utility, dry_exp:wet_exp)
+```
+
+So the intersection point of the next plot is the mixed strategy equilbrium(indifference point) in a risk-free environment
+
+```{r, echo = FALSE}
+nash_eq <- tibble(wheat_prop = abs(barley_dry - barley_wet) / (abs(barley_dry - barley_wet) + abs(wheat_dry - wheat_wet)),
+                  utility = wheat_prop * (wheat_dry + wheat_wet))
+
+plt_dat2 %>%
+  ggplot(aes(wheat_prop, utility)) +
+  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
+  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
+  geom_line(aes(color = type), size = 1.2) +
+  geom_point(data = nash_eq, color = 'black', size = 3) +
+  geom_text(data = nash_eq, 
+            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
+            nudge_x = 0.1, nudge_y = 0.05) +
+  theme_classic()
+
+```
+
+when you sample from the expected value, use that to choose a strategy. but if you sample lots, use that to create a mixed streateg
+
+And the subjective expected yields for the agents are thus.
+```{r}
+agents %>%
+  mutate(risk = runif(n())) %>%
+  mutate(yield_exp = expected_utility(risk),
+         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
+         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
+         favorite = case_when(barley_exp == wheat_exp ~ 'either',
+                              barley_exp > wheat_exp ~ 'barley',
+                              barley_exp < wheat_exp ~ 'wheat')) 
+```
+
+```{r}
+calc_ratio <- function(risk) {
+  pmax(0, (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1)))
+}
+```
+
+Now create a population of agents with different subjective beliefs about the drought risk.
+```{r}
 
@@
 
-The basic risk assessment agorithm.
+The basic risk assessment algorithm.
 Let's look at the risk of a drought year being 0.2.
@@
 
+maximax
 ```{r}
-wheat_dry <- 0
-wheat_wet <- 1.6
-barley_dry <- 0.93
-barley_wet <- 1.18
+risk <- 0.9
+wheat_exp <- wheat_wet * (1 - risk) + wheat_dry * risk
+barley_exp <- barley_wet * (1 - risk) + barley_dry * risk
+obj.fun <- c(wheat_exp, barley_exp)
 
-calc_payoff <- function(wheat_prop, risk) {
-  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
-  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
-  barley + wheat
-}
+constr=matrix(c(1,1,1,0,0,1, wheat_exp, barley_exp),ncol = 2, byrow = TRUE) #Loading the constraints
+constr.dir=c("=",">=",">=", ">=")
+constr.rhs=c(1,0,0, .5)
 
-tibble(wheat_prop)
+mod=lp("min",obj.fun,constr,constr.dir,constr.rhs) #Using lp() to solve our problem
+mod$solution
 ```
@@
 ```{r}
+calc_payoff <- function(x, drought) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
 
-calc_ratio <- function(risk) {
-  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
+  x %>%
+    mutate(yield = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop))
 }
-calc_ratio(.2)
+
+update_prop <- function(x, aspiration) {
+  x %>%
+    mutate(new_prop = round(runif(n()), 2),
+           wheat_prop = if_else(yield > aspiration, wheat_prop, new_prop)) %>%
+    select(-new_prop)
+}
+
+test3 <- tibble(agent = 1:1000) %>%
+  mutate(wheat_prop = round(runif(n()), 2)) %>%
+  calc_payoff(drought = TRUE)
+test3 %>% update_prop(aspiration = .6)
+
+accumulate((runif(25) < .3), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .5), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(as.numeric(time), yield, group = agent)) + geom_line(alpha = .2)
+
+```
+
+```{r}
+library(ggridges)
+accumulate((runif(25) < .2), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .6), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(wheat_prop, y = as.numeric(time), group = as.numeric(time))) +
+  geom_density_ridges() +
+  theme_minimal()
 ```
@@
 So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
+
 ```{r}
 test_dat <- expand_grid(wheat_prop = (0:10)/10,
-       risk = seq(0, 1, 0.2)) %>%
+       risk = seq(0, 1, 0.1)) %>%
   mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
@@
          util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
-         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk)+ barley_dry * risk),
-         payoff = calc_payoff(wheat_prop, risk))
+         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
+         util_payoff = util_dry + util_wet)
+```
 
+```{r}
+test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  gather(type, util, c(util_dry, util_wet, util_payoff)) %>%
+    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk) + geom_hline(yintercept = .5, alpha = .3, linetype = 2) + geom_vline(xintercept = .14, alpha = 0.3,linetype = 2)
+```
+```{r}
+test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  mutate(util_min = pmin(util_wet, util_dry))%>%
+    ggplot(aes(util_payoff, util_min)) + geom_line() + facet_wrap(~risk) + geom_hline(yintercept = .3, alpha = .3, linetype = 2)
+```
+
+```{r}
 test_dat %>%
@@
   gather(type, util, util_wheat:util_barley) %>%
-    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
-test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  gather(type, util, util_dry:util_wet) %>%
-    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
+    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk) + geom_vline(xintercept = .14)
 

### 2019-10-24 risk-percption.Rmd
knitable document

@@
 ```{r}
-agents <- tibble(id = 1:5, 
+agents <- tibble(id = 1:100, 
                  alpha = 1.5,
@@
 barley_wet <- 1.18
-
-wheat_dry <- .2
-wheat_wet <- .4
-barley_dry <- 0.6
-barley_wet <- .3
 ```
@@
 
-The maximin strategy here is to plant barley all the time, but this entails a huge opportunity cost over planting wheat or a mix of the two
-
-```{r, echo = FALSE}
-plt_dat2 <- tibble(wheat_prop = seq(0, 1, .1),
-       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
-      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
-  gather(type, utility, dry_exp:wet_exp)
-```
-
+The maximin strategy here is to plant barley all the time, but this entails a huge opportunity cost over planting wheat or a mix of the two.
 So the intersection point of the next plot is the mixed strategy equilbrium(indifference point) in a risk-free environment
@@
 ```{r, echo = FALSE}
-nash_eq <- tibble(wheat_prop = abs(barley_dry - barley_wet) / (abs(barley_dry - barley_wet) + abs(wheat_dry - wheat_wet)),
-                  utility = wheat_prop * (wheat_dry + wheat_wet))
+nash_eq <- tibble(wheat_prop = round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2),
+                  utility = round(wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop), 2))
 
-plt_dat2 %>%
+tibble(wheat_prop = seq(0, 1, .1),
+       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
+      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
+  gather(type, utility, dry_exp:wet_exp) %>%
   ggplot(aes(wheat_prop, utility)) +
@@
   theme_classic()
+```
 
+```{r}
+# farmer's ne
+pmax(0, round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2))
+
+#nature's ne
+round((barley_wet - wheat_wet) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2)
 ```
+
+0.3111 is also the indifference point for the farmer with respect to crops.
+That is, if we imagine nature playing strategically, a drought would occur with probability 0.31111 because that would make nature indifferent to whether the farmer planted wheat or barley.
 
@@
 
-Now create a population of agents with different subjective beliefs about the drought risk.
+Simulate a time series with varying drought frequency.
 ```{r}
-
-test2 <- c((runif(25) < .2),(runif(25) < .5), runif(25) < .1)
-tibble(time = 1:75,
-       drought = test2) %>%
-  mutate(period = round(time, -1)) %>%
-  add_count(period) %>%
-  filter(period >= 10) %>%
-  group_by(period) %>%
-  summarise(drought = sum(drought)) %>%
-  ggplot(aes(period, drought)) +
-  geom_col()
+droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .1)
 ```
 
+Allow the agents to learn the drought risk over time.
 ```{r}
@@
            mu = alpha / (alpha + beta),
-           estimate = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
+           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
 }
+```
 
-accumulate(test2, ~learn(.x, .y), .init = test) %>% bind_rows(.id = 'time') %>% qplot(as.numeric(time), estimate, data = ., geom = 'point')
+Run an experiment.
+```{r}
+simulation <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
 ```
@@
 ```{r}
-learn(test, TRUE)
+ggplot(simulation, aes(time, risk)) +
+  geom_point(alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
 ```
 
-The basic risk assessment algorithm.
-Let's look at the risk of a drought year being 0.2.
-In a wet year you get 3 wheat and 2 barley, and in a dry year you get 0 wheat and 1 barley.
-What does that look like?
+Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
+
+```{r}
+knitr::knit_exit()
+```
 
@@
 test_dat <- expand_grid(wheat_prop = (0:10)/10,
-       risk = seq(0, 1, 0.1)) %>%
+       risk = seq(.1, .9, 0.1)) %>%
   mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
@@
          util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
-         util_payoff = util_dry + util_wet)
+         payoff_exp = util_dry + util_wet,
+         payoff_max = pmax(util_wheat, util_barley))
+```
+
+```{r}
+nash_eq2 <- tibble(risk = seq(0.1, 0.9, 0.1),
+                   wheat_prop = pmax(round(((1 - risk) * barley_wet - risk * barley_dry) / (risk * wheat_dry + (1 - risk) * barley_wet - ((1 - risk) * wheat_wet + risk * barley_dry)), 2), 0),
+                  utility = round(risk * wheat_dry * wheat_prop + risk * barley_dry * (1 - wheat_prop), 2))
+
+test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  gather(type, utility, c(util_dry, util_wet)) %>%
+    ggplot(aes(wheat_prop, utility)) + 
+  geom_line(aes(color = type)) + 
+  facet_wrap(~risk) + 
+  geom_hline(data = nash_eq2, aes(yintercept = utility), alpha = .3, linetype = 2) + 
+  geom_vline(data = nash_eq2, aes(xintercept = wheat_prop), alpha = 0.3,linetype = 2) +
+  geom_point(data = nash_eq2) +
+  theme_classic()
 ```
@@
 test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  gather(type, util, c(util_dry, util_wet, util_payoff)) %>%
-    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk) + geom_hline(yintercept = .5, alpha = .3, linetype = 2) + geom_vline(xintercept = .14, alpha = 0.3,linetype = 2)
+  select(-(util_dry:util_wet)) %>%
+  gather(type, util, util_wheat:util_barley) %>%
+  ggplot(aes(wheat_prop, util, color = type)) + 
+  geom_line() + 
+  facet_wrap(~risk) + 
+  geom_vline(xintercept = nash_eq$wheat_prop)
 ```
+
+Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.
 ```{r}
 test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  mutate(util_min = pmin(util_wet, util_dry))%>%
-    ggplot(aes(util_payoff, util_min)) + geom_line() + facet_wrap(~risk) + geom_hline(yintercept = .3, alpha = .3, linetype = 2)
+  gather(type, utility, payoff_exp:payoff_max) %>%
+  ggplot(aes(wheat_prop, utility)) + 
+  geom_line(aes(color = type)) +
+  facet_wrap(~risk)
 ```
 
-```{r}
-test_dat %>%
-  group_by(risk) %>%
-  filter(payoff == max(payoff)) %>%
-  ggplot(aes(risk, wheat_prop)) + geom_point()
+So both the Wald and Savage regret criteria tell us to choose barley all the time.
 
-test_dat %>%
-  select(-(util_dry:util_wet)) %>%
-  gather(type, util, util_wheat:util_barley) %>%
-    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk) + geom_vline(xintercept = .14)
+If we're using the empirical frequency distirubtion of nature's moves, then we are playing a fictitious game.
+The strategy of planting the crop with the highest expected yield, given the frequency distribution of nature is known, will only work when nature is stationary.
 
-test_dat %>%
-  ggplot(aes(wheat_prop, util_wet + util_dry)) + geom_point() + facet_wrap(~risk)
-
-calc_prop <- function(risk) {
-  pb <- barley_wet * (1 - risk) - barley_dry * risk
-  pw <- wheat_wet * (1 - risk) - wheat_dry * risk
-
-  pb / (pb + pw) 
-}
-
-plot(calc_prop((0:10)/10))
-```
+We could think of other algorithms that would minimize the cumulative or average regret of a player over many years in a situation with nonstationary rainfall, but we have to remain within cognitive limits.
 
 ```{r}
-tibble(x)
-```
-
-```{r}
+# this is for nature
 calcNE <- function(risk){
@@
 
-(barley_wet - .2 * (barley_dry + wheat_dry + wheat_wet + barley_wet)) / (barley_wet + wheat_wet)
-
-calcNE(1)
-
-plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)
-
-util_wet <- wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)
-util_dry <- wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)
-
-tibble()
-
-plot(seq(.1,1,.1) * (1 - seq(.1,1,.1)))
-
+plot(0:10, calcNE((0:10)/10))
 ```

### 2019-10-24 risk-percption.Rmd
simulation with mixed strategies

@@
 
-Given these payoffs, what's the point of indifference between the two crop types?
-
-```{r, echo = FALSE}
-plt_dat <- tibble(risk = seq(0, 1, .1),
-       barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
-      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
-  gather(type, utility, barley_exp:wheat_exp)
-```
-
 The pure maximin strategy is to plant barley:
@@
 
+Given these payoffs, what's the point of indifference between the two crop types?
 ```{r, echo = FALSE}
@@
 
-plt_dat %>%
+tibble(risk = seq(0, 1, .1),
+       barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
+      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
+  gather(type, utility, barley_exp:wheat_exp) %>%
   ggplot(aes(risk, utility)) +
@@
 pmax(0, round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2))
-
 #nature's ne
@@
 when you sample from the expected value, use that to choose a strategy. but if you sample lots, use that to create a mixed streateg
-
-And the subjective expected yields for the agents are thus.
-```{r}
-agents %>%
-  mutate(risk = runif(n())) %>%
-  mutate(yield_exp = expected_utility(risk),
-         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
-         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
-         favorite = case_when(barley_exp == wheat_exp ~ 'either',
-                              barley_exp > wheat_exp ~ 'barley',
-                              barley_exp < wheat_exp ~ 'wheat')) 
-```
 
@@
 
-Allow the agents to learn the drought risk over time.
+Allow the agents to learn the drought risk over time. Also need to calculate the actual payoffs for each agent.
+
 ```{r}
-learn <- function(agents, event) {
+learn <- function(agents, drought) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
+
   agents %>%
-    mutate(alpha = event + alpha,
-           beta = beta + 1 - event,
+    # farmer's turn
+      mutate(yield_exp = expected_utility(risk),
+         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
+         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
+         crop = case_when(barley_exp > wheat_exp ~ 'barley',
+                              barley_exp < wheat_exp ~ 'wheat',
+                              barley_exp == wheat_exp ~ 'either'),
+         prediction = map2(alpha, beta, ~rbeta(20, .x, .y)),
+         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
+      select(-prediction) %>%
+  # nature's turn
+    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
+           alpha = drought + alpha,
+           beta = beta + 1 - drought,
            mu = alpha / (alpha + beta),
@@
 
-```{r}
+```{r, echo = TRUE}
 ggplot(simulation, aes(time, risk)) +
@@
   theme_classic()
+```
+
+```{r, echo = TRUE}
+simulation %>%
+  group_by(time, crop) %>%
+  tally() %>%
+ggplot(aes(time, n, fill = crop)) +
+  geom_bar(width = 1, stat = 'identity') +
+  scale_fill_manual(values = c('yellow', 'orange')) +
+  theme_classic()
+```
+
+These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
+```{r}
+ggplot(simulation, aes(time, wheat_prop)) +
+  geom_point(alpha = .3) +
+  theme_classic()
+```
+
+```{r}
+simulation %>%
+  group_by(time) %>%
+  summarise(famine = sum(payoff == 0)) %>%
+ggplot(aes(time, famine)) +
+  geom_line()
 ```
@@
 ```{r}
-calc_payoff <- function(x, drought) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
-
-  x %>%
-    mutate(yield = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop))
-}
 

### 2019-10-30 risk-percption.Rmd
minor tweaks to setup

@@
 ```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .1)
+droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
 ```
@@
                               barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(20, .x, .y)),
+         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)),
          wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
@@
   geom_bar(width = 1, stat = 'identity') +
-  scale_fill_manual(values = c('yellow', 'orange')) +
+  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
   theme_classic()
@@
 ggplot(aes(time, famine)) +
-  geom_line()
+  geom_line() +
+  theme_classic()
 ```

### 2020-01-17 risk-percption.Rmd
incorporate manuscript text, rearrange, and add downscaling

@@
 ---
-title: "risk-perception"
+title: "Crop diversification in a game against nature"
 author: "Nick Gauthier"
-date: "9/18/2019"
+date: ""
+bibliography: refs.bib 
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 library(tidyverse)
-library(lpSolve)
-
-#library(zoo)
 ```
 
-Start with a population of agents with the same subjective beliefs about drought risk.
+# Introduction
 
-```{r}
-agents <- tibble(id = 1:100, 
-                 alpha = 1.5,
-                 beta = 1.5,
-                 risk = alpha / (alpha + beta))
-```
+Over the past 10,000 years, subsistence farmers in the Mediterranean basin have developed sophisticated strategies to maintain stable food supplies given uncertain rainfall.
+Their suite of strategies include practices like crop diversification, storage, and exchange [@Halstead1989]).
+These risk-management strategies all represent forms of \emph{regulatory feedback} used to manage complex socio-ecological systems.
+Regulatory feedbacks convert a system's high-variance inputs into low-variance outputs by dynamically monitoring the outputs and adjusting internal properties of the system [@Anderies2015].
+Risk-managing strategies require farmers to make specific resource allocation decisions in response to environmental risks, and as regulatory feedbacks they are thus sensitive to variability in both the spatiotemporal patterns of risk and the ability of farmers to perceive and act on those patterns.
 
-```{r}
-agents
-```
+Crop diversification is an excellent example of a widespread and effective regulatory feedback strategy (Figure 1).
+Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@helmers2001separating,Anderies2006].
+In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,Marston2011190].
+Wheat generally has higher yields but is sensitive to water deficits, while barley is less productive but drought tolerant.
+Planting a mix of wheat and barley, either in the same plot or in a combination of plots, is an effective means of diversifying grain supplies.
+By adjusting the ratio of wheat to barley, farmers can adapt to changing drought risks.
 
-And payoffs for different crop types in different years.
+![Conceptual model of crop diversification as a regulatory feedback system.
+Uncertain rainfall leads to uncertain returns from land-use systems.
+To keep food supplies stable, farmers must monitor year-to-year variations and decide what combination of crops works best to minimize that variance.](figures/regulatory_feedback.png)
+Are these crop-diversification strategies vulnerable to the same dynamics as other social-ecological systems with similar regulatory feedbacks mechanisms?
+One way to address this is to model the influence of imperfect monitoring and biased decision making in uncertain environments.
+In this study, I accomplish this by answering two main questions:
+
+1. What was the temporal pattern of climate and climate change during the Holocene in the eastern Mediterranean?
+2. How well could crop diversification strategies cope with this pattern, given farmers’ imperfect risk perception and decision-making?
+
+I couple a climate model with high temporal resolution to a simple game theoretic model of agricultural decision making under uncertainty, in order to examine the performance of optimal and suboptimal wheat-barley diversification practices.
+I model the year-to-year crop diversification strategies made by farmers as an iterated game of fictitious play against nature [@GOULD1963a].
+Farmers are boundedly rational, in that they seek to maximize their objective functions given the perceived probabilities of different actions by nature, but work with flawed mental models of climate-related risks due to imperfect recall of past events.
+Furthermore, by using a climate model to represent the actions of nature, I can more precisely capture the characteristic autocorrelated patterns of rainfall variability, rather than simply drawing rainfall values from a static distribution.
+This framework allows me to address the sensitivity of diversification strategies to changes and climatic variability and imperfect human perceptions of those changes.
+
+# Decision-Making in a Game Against Nature
+
+Assume two different crops, a low yield drought tolerant crop (barley) and a high yield drought susceptible crop (wheat).
+
+Estimates of yield volume (t/ha) for each crop type in wet and dry years were derived from isotopic studies of ancient wheat and barley samples from archaeological sites in the Mediterranean [@Slafer1999] (Table 1).
+
+||Drought Year|Normal Year|
+|-------|-----|-----|----|
+|Barley Yield|0.93|1.18|
+|Wheat Yield|0|1.60|
+
+Table: Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].
+
 ```{r}
@@
 
-The expected yield is thus calculated as.
-```{r}
-expected_utility <- function(risk) {
-  risk * (wheat_dry + barley_dry) + (1 - risk) * (wheat_wet + barley_wet) 
-}
-```
+Optimal wheat/barley proportions can be calculated using two different objective functions representing different decision making types of farmers.
 
-```{r}
-expected_utility(0.2) # assuming a drought risk of 0.2
-```
+## Strategies
+
+1. Wald Maximin -- pessimist, maximize the worst case scenario
+2. Savage Regret -- minimize regret
+3. Hurewicz - balance optimism and pessimism
+4. Laplace or "naive" - 50/50
+
+So both the Wald and Savage regret criteria tell us to choose barley all the time.
+
+If we're using the empirical frequency distirubtion of nature's moves, then we are playing a fictitious game.
+The strategy of planting the crop with the highest expected yield, given the frequency distribution of nature is known, will only work when nature is stationary.
+
+We could think of other algorithms that would minimize the cumulative or average regret of a player over many years in a situation with nonstationary rainfall, but we have to remain within cognitive limits.
+
+* Maximax
+2. Maximize *subjective* expected utility
+3. Satisficing learning -- win-stay lose-move
+4. Nash equilibrium -- minimax (risk free)
+5. Fictitious play learning rule -- use empirical frequency distribution.
+A. Good against environment when environment is stationary but not when its nonstationary.
+B. Bayesian belief based learning
 
@@
 ```
+
+Alternatively, yield maximizing farmers simply seek to maximize the function for their expected utility given all possible states of nature
+\begin{equation}
+    E\left[ U_{all} \right] = E \left[U_{dry} \right] \times P_{dry}^{*} + E \left[U_{wet} \right] \times \left(1-P_{dry}^{*}\right).
+\end{equation}
+
+Given this information on the potential payoffs for planting each crop type, and information on the (perceived) probability of a drought occurring, the crop-diversification decision practice can be thought of as a simple two-player game against nature [@GOULD1963a].
+The expected yields for a farmer given a fixed proportion of barley and wheat can be calculated as
+
+\begin{equation}
+    E \left[Y_{dry} \right] = W_{dry} \times  P + B_{dry} \times (1 - P)
+\end{equation}
+if a drought occurs and likewise
+\begin{equation}
+    E \left[Y_{wet} \right] = W_{wet} \times P + B_{wet} \times (1 - P)
+\end{equation}
+
+if a drought does not occur. This specification allows us to calculate the optimal proportion of wheat to barley one should plant in a diversified strategy as a mixed strategy profile.
+
+Risk minimizing farmers play the equivalent of a mixed strategy Nash equilibrium, setting the expected yields in drought and normal years equal to each other, weighted by risk, and solving for the wheat/barley proportion
+\begin{equation}
+    E \left[Y_{dry} \right] \times P_{dry}^{*} = E \left[U_{wet} \right] \times \left(1- P_{dry}^{*}\right), 
+\end{equation}
+where $P_{dry}^{*}$ is the \emph{perceived}, not realized, risk of a drought in a particular year.
+The solution to this function represents the ratio of wheat to barley that would make a farmer indifferent to whether a drought would occur in a given year.
 
@@
   geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
-  theme_classic()
+  theme_classic() +
+  ggtitle('Point of indifference between wheat and barley', 'Given varying drought risk')
+```
+
+The expected yield is thus calculated as.
+```{r}
+expected_utility <- function(risk) {
+  risk * (wheat_dry + barley_dry) + (1 - risk) * (wheat_wet + barley_wet) 
+}
+```
+
+```{r}
+expected_utility(0.2) # assuming a drought risk of 0.2
 ```
@@
 The maximin strategy here is to plant barley all the time, but this entails a huge opportunity cost over planting wheat or a mix of the two.
-So the intersection point of the next plot is the mixed strategy equilbrium(indifference point) in a risk-free environment
+So the intersection point of the next plot is the mixed strategy equilbrium (indifference point) in a risk-free environment
 
@@
 
+So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
+
+```{r}
+test_dat <- expand_grid(wheat_prop = (0:10)/10,
+       risk = seq(.1, .9, 0.1)) %>%
+  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
+         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
+         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
+         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
+         payoff_exp = util_dry + util_wet,
+         payoff_max = pmax(util_wheat, util_barley))
+```
+
+```{r}
+nash_eq2 <- tibble(risk = seq(0.1, 0.9, 0.1),
+                   wheat_prop = pmax(round(((1 - risk) * barley_wet - risk * barley_dry) / (risk * wheat_dry + (1 - risk) * barley_wet - ((1 - risk) * wheat_wet + risk * barley_dry)), 2), 0),
+                  utility = round(risk * wheat_dry * wheat_prop + risk * barley_dry * (1 - wheat_prop), 2))
+
+test_dat %>%
+  select(-(util_wheat:util_barley)) %>%
+  gather(type, utility, c(util_dry, util_wet)) %>%
+    ggplot(aes(wheat_prop, utility)) + 
+  geom_line(aes(color = type)) + 
+  facet_wrap(~risk) + 
+  geom_hline(data = nash_eq2, aes(yintercept = utility), alpha = .3, linetype = 2) + 
+  geom_vline(data = nash_eq2, aes(xintercept = wheat_prop), alpha = 0.3,linetype = 2) +
+  geom_point(data = nash_eq2) +
+  theme_minimal()
+# hmmm, on second thought this might not make any sense, what's actually informative here is tha maximum value, which is the same as saying whether it is above or below the indifference threshold.
+but beyond that the idea of the nash equilibrium at different risk levels doesn't make sense, becuase then the nash equilibrium is, under the hood, adding another risk assessment to nature's moves.
+```
+
+```{r}
+test_dat %>%
+  select(-(util_dry:util_wet)) %>%
+  gather(type, util, util_wheat:util_barley) %>%
+  ggplot(aes(wheat_prop, util, color = type)) + 
+  geom_line() + 
+  facet_wrap(~risk) + 
+ # geom_vline(xintercept = nash_eq$wheat_prop) +
+  theme_minimal()
+```
+
+*Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
+```{r}
+test_dat %>%
+  gather(type, utility, payoff_exp:payoff_max) %>%
+  ggplot(aes(wheat_prop, utility)) + 
+  geom_line(aes(color = type)) +
+  facet_wrap(~risk)
+```
+
 ```{r}
@@
 }
+
+tibble(risk = seq(0, 1, .01),
+       ratio = calc_ratio(risk)) %>%
+  ggplot(aes(risk, ratio)) +
+  geom_line() + 
+  theme_classic()
 ```
 
-Simulate a time series with varying drought frequency.
+## Crop Allocation Decisions
+
+These equations are iteratively applied to the entire span of the climate-model output.
+The result is that farmers calculate the optimal land-use strategy that, \emph{to the best of their knowledge and given their preferences}, uses information from the past.
+
+# Climate Risks and Risk Perception
+
+Using the bias-corrected climate model output, I divided each model year into dry years and wet years.
+A dry year was any year where less than 300mm of rain fell during the wet season (October-March), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
+
+Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years
+
+\begin{equation}
+    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
+\end{equation} 
+where $t$ is the current time step.
+The 50 year time span was selected to approximate the accumulated observational knowledge of an individual farmer and their immediate household.
+As a result, this approach does not allow for accumulated social learning, although it could easily be extended to do so in future studies.
+
+Start with a population of agents with the same subjective beliefs about drought risk.
+
 ```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
+agents <- tibble(id = 1:100, 
+                 alpha = 1.5,
+                 beta = 1.5,
+                 risk = alpha / (alpha + beta))
+```
+
+```{r}
+agents
 ```
@@
   agents %>%
-    # farmer's turn
+    # farmer's turn to decide what to plant
       mutate(yield_exp = expected_utility(risk),
@@
                               barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)),
+         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
          wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
       select(-prediction) %>%
-  # nature's turn
+  # nature's turn to decide realized yields
     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
@@
 
-Run an experiment.
+Similar to above, but satisficing
+```{r}
+learn_satisficing <- function(agents, drought, aspiration) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
+
+  agents %>%
+     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
+            new_prop = round(runif(n()), 2),
+           wheat_prop = if_else(payoff > aspiration, wheat_prop, new_prop)) %>%
+    select(-new_prop)
+}
+```
+
+## Bayesian Updating
+
+```{r}
+prior <- function(m,n){
+a = n * m
+b = n * (1 - m)
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+tibble(year = 1:20,
+       risk = .5) %>%
+  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line() +
+  facet_wrap( ~year) +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+
+#prior(.5, ) %>% plot
+```
+
+```{r}
+likelihood <- function(N,Y){
+a <- Y + 1
+b <- N - Y + 1
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+#likelihood(1, 1)%>% plot
+```
+
+```{r}
+posterior <- function(m,n,N,Y){
+a <- Y + (n*m) -1
+b <- N - Y + (n*(1-m)) - 1
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+#dbeta(seq(0,1,0.005), -1, 0)
+#prior(.5, 2) %>% plot
+#posterior(.5, 3, 1, 1) %>% plot
+```
+
+## Climate Modeling and Bias Correction
+
+Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
+TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
+It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
+orbit, greenhouse gasses, glacial meltwater flux).
+The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
+
+![Example model precipitation outputs from the TraCE-21k simulation at 9.5ka BP.](figures/trace_white.png)
+
+The computational complexity of the TraCE-21k simulation requires it to be run at a coarse (~200km) spatial resolution.
+At this resolution, key hydroclimate processes such as individual cloud formation cannot be simulated directly, leading to biases in the distribution of modeled precipitation.
+In order to reveal climate patterns at regional scales, outputs from this simulation must be statistically downscaled and bias corrected.
+The cumulative distribution function transform (CDF-t) method is a skillful yet conceptually simple technique to downscale and bias-corrected GCM data [@Michelangeli2009].
+The technique works directly with the CDFs of observed and simulated climate variables, rather than individual observations.
+A function is first derived that transforms the CDF of GCM data to match the CDF of weather-station observations from the same period.
+The same function is then used to transform the CDFs of GCM data from different time periods, accounting for the changes in the CDFs from one simulation period to another (Figure 3).
+CDF-t is nonparametric, and thus more data-driven than weather generators and related statistical downscaling methods that rely to a greater degree on external assumptions about the nature of the data.
+
+![CDFs for simulated and observed precipitation in the present day, and simulated and bias-corrected precipitation in the past.](figures/cdf.png)
+
+TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
+This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
+
+\section{Results and Discussion}
+
+\begin{figure}[htp]
+\centering
+\includegraphics[width=16cm]{precon.png}
+\caption{Bias-corrected wet season (October-March) precipitation accumulations from the past 10,000 years of TraCE-21k simulation.
+1 point = 1 year, color corresponds to whether that year falls bellow the thresholds for wheat or barley crop failure.}
+\end{figure}
+
+The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
+8ka BP (Figure 5).
+The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
+
+The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
+Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
+
+![Annual risk of wheat crop failure due to drought, averaged by century.](figures/wheatrisk.png)
+
+Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
+$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
+This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
+Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
+Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
+
+Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
+Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
+Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
+
+During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
+Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
+This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
+This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
+
+Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
+To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
+Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
+Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
+
+Run an experiment. Simulate a time series with varying drought frequency.
+```{r}
+droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
+```
+
 ```{r}
@@
 ```{r}
+test3 <- tibble(agent = 1:100) %>%
+  mutate(wheat_prop = round(runif(n()), 2))
+
+simulation2 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .8), .init = test3) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+
+simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
+simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
+```
+
+```{r}
 simulation %>%
@@
 
+# References
+
 ```{r}
 knitr::knit_exit()
+```
+
+Now incorporate realistic drought variability from simulations.
+
+```{r}
+library(raster)
+library(CDFt)
+```
+
+```{r}
+konya <- matrix(c(32.5, 37.9), nrow = 1)
+
+sim_present <- brick('~/Downloads/b40.20th.track1.1deg.005.cam2.h0.PRECT.185001-200512.nc') %>%
+  raster::extract(konya, df = TRUE) %>%
+  dplyr::select(-ID) %>%
+  gather(time, rainfall) %>%
+  mutate(rainfall = rainfall * 2.628e+9,
+         year = rep(1850:2005, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
+         month = rep(1:12, times = 156)) %>%
+  dplyr::select(-time) %>%
+  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
+  filter(between(water_year, 1932, 2005)) %>%
+  group_by(water_year) %>%
+  summarise(precip = sum(rainfall))
+
+sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
+  raster::extract(konya, df = TRUE) %>%
+  dplyr::select(-ID) %>%
+  gather(time, rainfall) %>%
+  mutate(rainfall = rainfall * 2.628e+9,
+         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
+         month = rep(1:12, times = 516)) %>%
+   dplyr::select(-time) %>%
+  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
+  group_by(water_year) %>%
+  summarise(precip = sum(rainfall))
+```
+
+```{r}
+obs <- read_table('~/Downloads/pa17244.dat', 
+           comment = "#", 
+           col_names = c('year', 1:12), 
+           na = c('-999.9', '-888.8')) %>%
+  gather(month, precip, `1`:`12`) %>%
+ # filter(!(month %in% c(4:10)) & year < 2006) %>%
+  group_by(month) %>%
+  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
+  ungroup %>%
+  mutate(month = as.numeric(month),
+         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
+    filter(between(water_year, 1932, 2005)) %>%
+  group_by(water_year) %>%
+  summarise(precip = sum(precip))
+```
+
+```{r}
+obs_past <- CDFt(obs$precip, sim_present$precip, sim_past$precip)$DS
+
+plot(obs_past)
+
+obs_past < 300
 ```
@@
 
-```{r}
-
-update_prop <- function(x, aspiration) {
-  x %>%
-    mutate(new_prop = round(runif(n()), 2),
-           wheat_prop = if_else(yield > aspiration, wheat_prop, new_prop)) %>%
-    select(-new_prop)
-}
-
-test3 <- tibble(agent = 1:1000) %>%
-  mutate(wheat_prop = round(runif(n()), 2)) %>%
-  calc_payoff(drought = TRUE)
-test3 %>% update_prop(aspiration = .6)
-
-accumulate((runif(25) < .3), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .5), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(as.numeric(time), yield, group = agent)) + geom_line(alpha = .2)
-
-```
+## Satisficing learners
 
@@
 ```
-
-So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
-
-```{r}
-test_dat <- expand_grid(wheat_prop = (0:10)/10,
-       risk = seq(.1, .9, 0.1)) %>%
-  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
-         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
-         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
-         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
-         payoff_exp = util_dry + util_wet,
-         payoff_max = pmax(util_wheat, util_barley))
-```
-
-```{r}
-nash_eq2 <- tibble(risk = seq(0.1, 0.9, 0.1),
-                   wheat_prop = pmax(round(((1 - risk) * barley_wet - risk * barley_dry) / (risk * wheat_dry + (1 - risk) * barley_wet - ((1 - risk) * wheat_wet + risk * barley_dry)), 2), 0),
-                  utility = round(risk * wheat_dry * wheat_prop + risk * barley_dry * (1 - wheat_prop), 2))
-
-test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  gather(type, utility, c(util_dry, util_wet)) %>%
-    ggplot(aes(wheat_prop, utility)) + 
-  geom_line(aes(color = type)) + 
-  facet_wrap(~risk) + 
-  geom_hline(data = nash_eq2, aes(yintercept = utility), alpha = .3, linetype = 2) + 
-  geom_vline(data = nash_eq2, aes(xintercept = wheat_prop), alpha = 0.3,linetype = 2) +
-  geom_point(data = nash_eq2) +
-  theme_classic()
-```
-
-```{r}
-test_dat %>%
-  select(-(util_dry:util_wet)) %>%
-  gather(type, util, util_wheat:util_barley) %>%
-  ggplot(aes(wheat_prop, util, color = type)) + 
-  geom_line() + 
-  facet_wrap(~risk) + 
-  geom_vline(xintercept = nash_eq$wheat_prop)
-```
-
-Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.
-```{r}
-test_dat %>%
-  gather(type, utility, payoff_exp:payoff_max) %>%
-  ggplot(aes(wheat_prop, utility)) + 
-  geom_line(aes(color = type)) +
-  facet_wrap(~risk)
-```
-
-So both the Wald and Savage regret criteria tell us to choose barley all the time.
-
-If we're using the empirical frequency distirubtion of nature's moves, then we are playing a fictitious game.
-The strategy of planting the crop with the highest expected yield, given the frequency distribution of nature is known, will only work when nature is stationary.
-
-We could think of other algorithms that would minimize the cumulative or average regret of a player over many years in a situation with nonstationary rainfall, but we have to remain within cognitive limits.
 
@@
 # so this works as long as you're between 0 and 1
-```
-
-```{r}
-
-prior <- function(m,n){
-a = n * m
-b = n * (1 - m)
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-prior(.5, 2) %>% plot
-```
-
-```{r}
-likelihood <- function(N,Y){
-a <- Y + 1
-b <- N - Y + 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-likelihood(1, 1)%>% plot
-```
-
-```{r}
-posterior <- function(m,n,N,Y){
-a <- Y + (n*m) -1
-b <- N - Y + (n*(1-m)) - 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
--1
-1 - 0 - 1
-
-dbeta(seq(0,1,0.005), -1, 0)
-prior(.5, 2) %>% plot
-posterior(.5, 3, 1, 1) %>% plot
-```
-
-so now we get the same yields, but the proportions are different
-```{r}
-(1 *  0.625 - 2 * .375) * .375 * .625
-```
-
-```{r}
-1- .375
-```
-
-```{r}
--1 * 0.0625 * (1 - 0.0625)
 ```
@@
   geom_hline(yintercept = yield_max_barley * kcal_barley) +
-  geom_hline(yintercept = 1016525* 5, color ='red') +
+  geom_hline(yintercept = 1016525 * 5, color ='red') +
   geom_vline(xintercept = 0.3)
@@
   filter(drought == TRUE)
-```
-
-```{r}
-# generate some forgetting curves
-forgetting_curves <- tibble(x = c(.01, .1, 1, 0)) %>%
-  mutate(memory = map(x, ~exp(-1 * . *1:50))) %>%
-  unnest(cols= c(memory)) %>%
-  group_by(x) %>%
-  mutate(time = 1:n())
-
-ggplot(forgetting_curves, aes(time, memory)) +
-  geom_line(aes(color = as.factor(x)), size = 1.2) +
-  labs(title = 'Forgetting curves', x = 'Years', y = 'Relative memory strength\n') +
-  scale_color_viridis_d(name = 'Decay rate') +
-  theme_classic()
 ```

### 2020-01-20 risk-percption.Rmd
add state of nature

@@
 knitr::opts_chunk$set(echo = FALSE)
+library(raster)
 library(tidyverse)
+library(CDFt)
 ```
@@
 
-||Drought Year|Normal Year|
-|-------|-----|-----|----|
-|Barley Yield|0.93|1.18|
-|Wheat Yield|0|1.60|
+||Dry Year|Normal Year|Wet Year|
+|-------|-----|-----|----|----|
+|Barley Yield|0.93|1.18||
+|Wheat Yield|0|1.60||
 
@@
 ```{r}
-wheat_dry <- 0
-wheat_wet <- 1.6
-barley_dry <- 0.93
-barley_wet <- 1.18
+#wheat_dry <- 0
+#wheat_wet <- 1.6
+#barley_dry <- 0.93
+#barley_wet <- 1.18
+
+wheat_dry <- 30
+wheat_normal <- 60
+wheat_wet <- 70
+
+barley_dry <- 50
+barley_normal <- 50
+barley_wet <- 40
 ```
@@
 ```{r}
-expected_utility <- function(risk) {
-  risk * (wheat_dry + barley_dry) + (1 - risk) * (wheat_wet + barley_wet) 
+expected_yield_wheat <- function(p_dry, p_wet) {
+  p_dry * wheat_dry + p_wet * wheat_wet + (1 - p_wet - p_dry) * wheat_normal 
+}
+
+expected_yield_barley <- function(p_dry, p_wet) {
+  p_dry * barley_dry + p_wet * barley_wet + (1 - p_wet - p_dry) * barley_normal 
 }
@@
 ```{r}
-expected_utility(0.2) # assuming a drought risk of 0.2
+expected_yield_wheat(p_dry = 0.3, p_wet = 0.2)
+expected_yield_barley(p_dry = 0.3, p_wet = 0.2)
+```
+
+The risk averse MAXIMIN pure strategy is to plant barley all the time, and the risk indifferent farmer plants wheat for its high yields
+
+```{r}
+yield_var_wheat <- function(p_dry, p_wet) {
+  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
+  p_dry * (wheat_dry - exp_wheat) ^ 2 +
+    p_wet * (wheat_wet - exp_wheat) ^ 2 +
+    (1 - p_dry - p_wet) * (wheat_normal - exp_wheat) ^ 2
+}
+
+yield_var_barley <- function(p_dry, p_wet) {
+  exp_barley <- expected_yield_barley(p_dry, p_wet)
+  p_dry * (barley_dry - exp_barley) ^ 2 +
+    p_wet * (barley_wet - exp_barley) ^ 2 +
+    (1 - p_dry - p_wet) * (barley_normal - exp_barley) ^ 2
+}
+
+yield_var_wheat(0.3, 0.2)
+yield_var_barley(0.3, 0.2)
+```
+
+```{r}
+yield_mad_wheat <- function(p_dry, p_wet) {
+  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
+  p_dry * abs(wheat_dry - exp_wheat) +
+    p_wet * abs(wheat_wet - exp_wheat) +
+    (1 - p_dry - p_wet) * abs(wheat_normal - exp_wheat)
+  }
+
+yield_mad_barley <- function(p_dry, p_wet) {
+  exp_barley <- expected_yield_barley(p_dry, p_wet)
+  p_dry * abs(barley_dry - exp_barley) +
+    p_wet * abs(barley_wet - exp_barley) +
+    (1 - p_dry - p_wet) * abs(barley_normal - exp_barley)
+}
+
+yield_mad_wheat(0.3, 0.2)
+yield_mad_barley(0.3, 0.2)
 ```
@@
        dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
+       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
       wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
@@
   ggplot(aes(wheat_prop, utility)) +
-  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
-  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
+ # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
+#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
   geom_line(aes(color = type), size = 1.2) +
-  geom_point(data = nash_eq, color = 'black', size = 3) +
-  geom_text(data = nash_eq, 
-            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
-            nudge_x = 0.1, nudge_y = 0.05) +
+#  geom_point(data = nash_eq, color = 'black', size = 3) +
+ # geom_text(data = nash_eq, 
+#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
+#            nudge_x = 0.1, nudge_y = 0.05) +
+  theme_classic()
+```
+
+```{r}
+tibble(wheat_prop = seq(0, 1, .1),
+       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
+       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
+      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet,
+      yield_exp = dry_exp + normal_exp + wet_exp, 
+      lower_limit = pmin(dry_exp, normal_exp,wet_exp)) %>%
+  ggplot(aes(yield_exp, lower_limit)) +
+ # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
+#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
+  geom_line(aes(color = type), size = 1.2) +
+#  geom_point(data = nash_eq, color = 'black', size = 3) +
+ # geom_text(data = nash_eq, 
+#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
+#            nudge_x = 0.1, nudge_y = 0.05) +
   theme_classic()
@@
 ```{r}
-library(raster)
-library(CDFt)
-```
-
-```{r}
 konya <- matrix(c(32.5, 37.9), nrow = 1)
@@
   raster::extract(konya, df = TRUE) %>%
-  dplyr::select(-ID) %>%
+  select(-ID) %>%
   gather(time, rainfall) %>%
@@
          month = rep(1:12, times = 156)) %>%
-  dplyr::select(-time) %>%
+  select(-time) %>%
   mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
   filter(between(water_year, 1932, 2005)) %>%
+   filter(!(month %in% c(5:9))) %>%
   group_by(water_year) %>%
@@
          month = rep(1:12, times = 516)) %>%
-   dplyr::select(-time) %>%
+  select(-time) %>%
   mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
+   filter(!(month %in% c(5:9))) %>%
   group_by(water_year) %>%
@@
   gather(month, precip, `1`:`12`) %>%
- # filter(!(month %in% c(4:10)) & year < 2006) %>%
   group_by(month) %>%
@@
     filter(between(water_year, 1932, 2005)) %>%
+ filter(!(month %in% c(5:9))) %>%
   group_by(water_year) %>%
@@
 
-obs_past < 300
+tibble(droughts = (obs_past < 250)[1:500]) %>%
+  mutate(time = rep(1:20, each = 25))%>%
+  group_by(time) %>%
+  summarise(test = sum(droughts)/25)
 ```
+
+```{r}
+simulation <- accumulate((obs_past < 250), ~learn(.x, .y), .init = agents) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+```
+
+```{r, echo = TRUE}
+ggplot(simulation, aes(time, risk)) +
+  geom_point(alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
+```
+
+so in this case drought risk is low enough all the time you never get any diversification
+
+~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
@@
 
-so we're imagining a yield of 1 for barley and 2 for wheat, with equal prportions of each. then in the next time step we'd revise the proportion to take away .25
+so we're imagining a yield of 1 for barley and 2 for wheat, with equal proportions of each. then in the next time step we'd revise the proportion to take away .25
 ```{r}
@@
   geom_hline(yintercept = yield_max_barley * kcal_barley) +
-  geom_hline(yintercept = 1016525 * 5, color ='red') +
-  geom_vline(xintercept = 0.3)
+  geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
+  geom_vline(xintercept = 0.3) +
+  geom_vline(xintercept = .2)
 ```

### 2020-01-22 risk-percption.Rmd
new payoff matrix

@@
 ```{r}
-#wheat_dry <- 0
-#wheat_wet <- 1.6
-#barley_dry <- 0.93
-#barley_wet <- 1.18
+wheat_dry <- 0
+wheat_normal <- 1.5
+wheat_wet <- 1.7
 
-wheat_dry <- 30
-wheat_normal <- 60
-wheat_wet <- 70
+barley_dry <- 0.8
+barley_normal <- 1
+barley_wet <- 0.5
+```
 
-barley_dry <- 50
-barley_normal <- 50
-barley_wet <- 40
+```{r}
+readxl::read_xls('41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
+  select(`Site Location`, Taxon, `ky cal BP`, `GY (t*ha-1)`) %>%
+  filter(!is.na(`GY (t*ha-1)`)) %>% 
+  group_by(`Site Location`, Taxon) %>%
+  summarise(yield_mean = mean(`GY (t*ha-1)`)) %>%
+  spread(Taxon, yield_mean)
 ```

### 2020-01-22 risk-percption.Rmd
add may to growing season

@@
   filter(between(water_year, 1932, 2005)) %>%
-   filter(!(month %in% c(5:9))) %>%
+   filter(!(month %in% c(6:9))) %>%
   group_by(water_year) %>%

### 2020-01-22 risk-percption.Rmd
remove extra plots

@@
 indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
-                 utility = risk * wheat_dry + (1 - risk) * wheat_wet)
+                 yield = risk * wheat_dry + (1 - risk) * wheat_wet)
+indiff_label <- paste0('(', round(indiff$risk, 2), ', ', round(indiff$yield, 2), ')')
 
@@
       wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
-  gather(type, utility, barley_exp:wheat_exp) %>%
-  ggplot(aes(risk, utility)) +
-  geom_hline(yintercept = indiff$utility, linetype = 2)+
+  gather(type, yield, barley_exp:wheat_exp) %>%
+  ggplot(aes(risk, yield)) +
+  geom_hline(yintercept = indiff$yield, linetype = 2)+
   geom_vline(xintercept = indiff$risk, linetype = 2) +
@@
   geom_point(data = indiff, color = 'black', size = 3) +
-  geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
+  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
   theme_classic() +
@@
 
+```{r}
+qplot(x = c(300, 400, 500), y = c(0, 1.5, 2), geom = 'line') +
+  geom_line(aes(y = c(0.8, 1, 0.5)))
+```
+
 ```{r, echo = FALSE}
@@
   ggplot(aes(wheat_prop, utility)) +
- # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
-#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
+  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
+  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
   geom_line(aes(color = type), size = 1.2) +
-#  geom_point(data = nash_eq, color = 'black', size = 3) +
- # geom_text(data = nash_eq, 
-#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
-#            nudge_x = 0.1, nudge_y = 0.05) +
+  geom_point(data = nash_eq, color = 'black', size = 3) +
+  geom_text(data = nash_eq, 
+            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
+            nudge_x = 0.1, nudge_y = 0.05) +
   theme_classic()
@@
 #  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
-  geom_line(aes(color = type), size = 1.2) +
+  geom_line( size = 1.2) +
 #  geom_point(data = nash_eq, color = 'black', size = 3) +
@@
 
-```{r}
-test_dat <- expand_grid(wheat_prop = (0:10)/10,
-       risk = seq(.1, .9, 0.1)) %>%
-  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
-         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
-         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
-         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
-         payoff_exp = util_dry + util_wet,
-         payoff_max = pmax(util_wheat, util_barley))
-```
-
-```{r}
-nash_eq2 <- tibble(risk = seq(0.1, 0.9, 0.1),
-                   wheat_prop = pmax(round(((1 - risk) * barley_wet - risk * barley_dry) / (risk * wheat_dry + (1 - risk) * barley_wet - ((1 - risk) * wheat_wet + risk * barley_dry)), 2), 0),
-                  utility = round(risk * wheat_dry * wheat_prop + risk * barley_dry * (1 - wheat_prop), 2))
-
-test_dat %>%
-  select(-(util_wheat:util_barley)) %>%
-  gather(type, utility, c(util_dry, util_wet)) %>%
-    ggplot(aes(wheat_prop, utility)) + 
-  geom_line(aes(color = type)) + 
-  facet_wrap(~risk) + 
-  geom_hline(data = nash_eq2, aes(yintercept = utility), alpha = .3, linetype = 2) + 
-  geom_vline(data = nash_eq2, aes(xintercept = wheat_prop), alpha = 0.3,linetype = 2) +
-  geom_point(data = nash_eq2) +
-  theme_minimal()
-# hmmm, on second thought this might not make any sense, what's actually informative here is tha maximum value, which is the same as saying whether it is above or below the indifference threshold.
-but beyond that the idea of the nash equilibrium at different risk levels doesn't make sense, becuase then the nash equilibrium is, under the hood, adding another risk assessment to nature's moves.
-```
-
-```{r}
-test_dat %>%
-  select(-(util_dry:util_wet)) %>%
-  gather(type, util, util_wheat:util_barley) %>%
-  ggplot(aes(wheat_prop, util, color = type)) + 
-  geom_line() + 
-  facet_wrap(~risk) + 
- # geom_vline(xintercept = nash_eq$wheat_prop) +
-  theme_minimal()
-```
-
 *Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
-```{r}
-test_dat %>%
-  gather(type, utility, payoff_exp:payoff_max) %>%
-  ggplot(aes(wheat_prop, utility)) + 
-  geom_line(aes(color = type)) +
-  facet_wrap(~risk)
-```
 
@@
   mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-   filter(!(month %in% c(5:9))) %>%
+   filter(!(month %in% c(6:9))) %>%
   group_by(water_year) %>%
@@
     filter(between(water_year, 1932, 2005)) %>%
- filter(!(month %in% c(5:9))) %>%
+ filter(!(month %in% c(6:9))) %>%
   group_by(water_year) %>%

### 2020-01-23 risk-percption.Rmd
remove unnecessary calculations

@@
 ---
-title: "Crop diversification in a game against nature"
+title: "Crop diversification in a game against nature in a mediterranean agro-ecosystem"
 author: "Nick Gauthier"
@@
 In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,Marston2011190].
-Wheat generally has higher yields but is sensitive to water deficits, while barley is less productive but drought tolerant.
+Wheat generally has higher yields but is sensitive to water deficits, while barley has lower yields and is drought tolerant.
 Planting a mix of wheat and barley, either in the same plot or in a combination of plots, is an effective means of diversifying grain supplies.
@@
 
-The risk averse MAXIMIN pure strategy is to plant barley all the time, and the risk indifferent farmer plants wheat for its high yields
-
 ```{r}
-yield_var_wheat <- function(p_dry, p_wet) {
-  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
-  p_dry * (wheat_dry - exp_wheat) ^ 2 +
-    p_wet * (wheat_wet - exp_wheat) ^ 2 +
-    (1 - p_dry - p_wet) * (wheat_normal - exp_wheat) ^ 2
-}
-
-yield_var_barley <- function(p_dry, p_wet) {
-  exp_barley <- expected_yield_barley(p_dry, p_wet)
-  p_dry * (barley_dry - exp_barley) ^ 2 +
-    p_wet * (barley_wet - exp_barley) ^ 2 +
-    (1 - p_dry - p_wet) * (barley_normal - exp_barley) ^ 2
-}
-
-yield_var_wheat(0.3, 0.2)
-yield_var_barley(0.3, 0.2)
+expand_grid(wheat_prop = seq(0, 1, .01),
+       risk = seq(.1, .9, .1)) %>%
+  mutate(wheat_exp = wheat_prop * expected_yield_wheat(risk,0),
+      barley_exp = (1 - wheat_prop) * expected_yield_barley(risk, 0),
+      exp_yield = wheat_exp + barley_exp) %>%
+  ggplot(aes(wheat_prop, exp_yield, group = risk, color = risk)) +
+  geom_line()
 ```
 
-```{r}
-yield_mad_wheat <- function(p_dry, p_wet) {
-  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
-  p_dry * abs(wheat_dry - exp_wheat) +
-    p_wet * abs(wheat_wet - exp_wheat) +
-    (1 - p_dry - p_wet) * abs(wheat_normal - exp_wheat)
-  }
-
-yield_mad_barley <- function(p_dry, p_wet) {
-  exp_barley <- expected_yield_barley(p_dry, p_wet)
-  p_dry * abs(barley_dry - exp_barley) +
-    p_wet * abs(barley_wet - exp_barley) +
-    (1 - p_dry - p_wet) * abs(barley_normal - exp_barley)
-}
-
-yield_mad_wheat(0.3, 0.2)
-yield_mad_barley(0.3, 0.2)
-```
+The risk averse MAXIMIN pure strategy is to plant barley all the time, and the risk indifferent farmer plants wheat for its high yields
 
@@
 So the intersection point of the next plot is the mixed strategy equilbrium (indifference point) in a risk-free environment
-
-```{r}
-qplot(x = c(300, 400, 500), y = c(0, 1.5, 2), geom = 'line') +
-  geom_line(aes(y = c(0.8, 1, 0.5)))
-```
 
@@
 ```{r}
-tibble(wheat_prop = seq(0, 1, .1),
-       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
-       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
-      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet,
-      yield_exp = dry_exp + normal_exp + wet_exp, 
-      lower_limit = pmin(dry_exp, normal_exp,wet_exp)) %>%
-  ggplot(aes(yield_exp, lower_limit)) +
- # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
-#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
-  geom_line( size = 1.2) +
-#  geom_point(data = nash_eq, color = 'black', size = 3) +
- # geom_text(data = nash_eq, 
-#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
-#            nudge_x = 0.1, nudge_y = 0.05) +
-  theme_classic()
-```
-
-```{r}
 # farmer's ne
@@
 *Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
-
-```{r}
-calc_ratio <- function(risk) {
-  pmax(0, (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1)))
-}
-
-tibble(risk = seq(0, 1, .01),
-       ratio = calc_ratio(risk)) %>%
-  ggplot(aes(risk, ratio)) +
-  geom_line() + 
-  theme_classic()
-```
 

### 2020-01-23 risk-percption.Rmd
pure strategy plots

@@
 
-||Dry Year|Normal Year|Wet Year|
-|-------|-----|-----|----|----|
-|Barley Yield|0.93|1.18||
-|Wheat Yield|0|1.60||
+||Dry Year|Normal Year|
+|-------|-----|-----|----|
+|Barley Yield|0.93|1.18|
+|Wheat Yield|0|1.60|
 
@@
 wheat_dry <- 0
-wheat_normal <- 1.5
-wheat_wet <- 1.7
+wheat_normal <- 1.6
 
-barley_dry <- 0.8
-barley_normal <- 1
-barley_wet <- 0.5
+barley_dry <- 0.93
+barley_normal <- 1.18
 ```
@@
 
-* Maximax
-2. Maximize *subjective* expected utility
-3. Satisficing learning -- win-stay lose-move
-4. Nash equilibrium -- minimax (risk free)
-5. Fictitious play learning rule -- use empirical frequency distribution.
-A. Good against environment when environment is stationary but not when its nonstationary.
-B. Bayesian belief based learning
-
 The pure maximin strategy is to plant barley:
-```{r}
-max(wheat_dry, barley_dry)
-```
 
@@
 
+The expected yield is thus calculated as.
+```{r}
+expected_yield_wheat <- function(risk) {
+  risk * wheat_dry + (1 - risk) * wheat_normal
+}
+
+expected_yield_barley <- function(risk) {
+  risk * barley_dry + (1 - risk) * barley_normal
+}
+
+expected_yield <- function(risk) {
+  expected_yield_wheat(risk) + expected_yield_barley(risk)
+}
+```
+
 Given these payoffs, what's the point of indifference between the two crop types?
-```{r, echo = FALSE}
-indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
-                 yield = risk * wheat_dry + (1 - risk) * wheat_wet)
+```{r, echo = FALSE, fig.cap=title = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
+indiff <- tibble(risk = (barley_normal - wheat_normal) / (wheat_dry - wheat_normal - barley_dry + barley_normal),
+                 yield = risk * wheat_dry + (1 - risk) * wheat_normal)
 indiff_label <- paste0('(', round(indiff$risk, 2), ', ', round(indiff$yield, 2), ')')
@@
 tibble(risk = seq(0, 1, .1),
-      barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
-      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
-  gather(type, yield, barley_exp:wheat_exp) %>%
+      Barley = expected_yield_barley(risk),
+      Wheat = expected_yield_wheat(risk)) %>%
+  gather(type, yield, Barley:Wheat) %>%
   ggplot(aes(risk, yield)) +
@@
   theme_classic() +
-  ggtitle('Point of indifference between wheat and barley', 'Given varying drought risk')
-```
-
-The expected yield is thus calculated as.
-```{r}
-expected_yield_wheat <- function(p_dry, p_wet) {
-  p_dry * wheat_dry + p_wet * wheat_wet + (1 - p_wet - p_dry) * wheat_normal 
-}
-
-expected_yield_barley <- function(p_dry, p_wet) {
-  p_dry * barley_dry + p_wet * barley_wet + (1 - p_wet - p_dry) * barley_normal 
-}
-```
-
-```{r}
-expected_yield_wheat(p_dry = 0.3, p_wet = 0.2)
-expected_yield_barley(p_dry = 0.3, p_wet = 0.2)
-```
-
-```{r}
-expand_grid(wheat_prop = seq(0, 1, .01),
-       risk = seq(.1, .9, .1)) %>%
-  mutate(wheat_exp = wheat_prop * expected_yield_wheat(risk,0),
-      barley_exp = (1 - wheat_prop) * expected_yield_barley(risk, 0),
-      exp_yield = wheat_exp + barley_exp) %>%
-  ggplot(aes(wheat_prop, exp_yield, group = risk, color = risk)) +
-  geom_line()
-```
-
-The risk averse MAXIMIN pure strategy is to plant barley all the time, and the risk indifferent farmer plants wheat for its high yields
-
-The maximin strategy here is to plant barley all the time, but this entails a huge opportunity cost over planting wheat or a mix of the two.
-So the intersection point of the next plot is the mixed strategy equilbrium (indifference point) in a risk-free environment
-
-```{r, echo = FALSE}
-nash_eq <- tibble(wheat_prop = round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2),
-                  utility = round(wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop), 2))
-
-tibble(wheat_prop = seq(0, 1, .1),
-       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
-       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
-      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
-  gather(type, utility, dry_exp:wet_exp) %>%
-  ggplot(aes(wheat_prop, utility)) +
-  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
-  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
-  geom_line(aes(color = type), size = 1.2) +
-  geom_point(data = nash_eq, color = 'black', size = 3) +
-  geom_text(data = nash_eq, 
-            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
-            nudge_x = 0.1, nudge_y = 0.05) +
-  theme_classic()
+  scale_color_discrete(name = 'Crop') +
+  labs(x = 'Drought probability', y = 'Expected yield')
 ```
@@
 # farmer's ne
-pmax(0, round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2))
+pmax(0, round((barley_normal - barley_dry) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2))
 #nature's ne
-round((barley_wet - wheat_wet) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2)
+round((barley_normal - wheat_normal) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2)
 ```
@@
   if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet
+  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
 
@@
     # farmer's turn to decide what to plant
-      mutate(yield_exp = expected_utility(risk),
-         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
-         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
+      mutate(yield_exp = expected_yield(risk),
+         barley_exp = expected_yield_barley(risk),
+         wheat_exp = expected_yield_wheat(risk),
+         expected_yield = barley_exp + wheat_exp,
          crop = case_when(barley_exp > wheat_exp ~ 'barley',
@@
 }
+```
+
+Run an experiment. Simulate a time series with varying drought frequency.
+```{r}
+droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
+```
+
+```{r}
+simulation <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+```
+
+```{r, echo = TRUE}
+ggplot(simulation, aes(time, risk)) +
+  geom_point(alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
+```
+
+```{r, echo = TRUE}
+simulation %>%
+  group_by(time, crop) %>%
+  tally() %>%
+ggplot(aes(time, n, fill = crop)) +
+  geom_bar(width = 1, stat = 'identity') +
+  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
+  theme_classic()
+```
+
+These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
+```{r}
+ggplot(simulation, aes(time, wheat_prop)) +
+  geom_point(alpha = .3) +
+  theme_classic()
+```
+
+```{r}
+simulation %>%
+  group_by(time) %>%
+  summarise(famine = sum(payoff < .1)) %>%
+ggplot(aes(time, famine)) +
+  geom_line() +
+  theme_classic()
+```
+
+```{r}
+simulation %>%
+  group_by(time) %>%
+  summarise(payoff = mean(payoff)) %>%
+ggplot(aes(time,payoff)) +
+  geom_point(aes(color = time)) +
+  theme_classic()
+```
+
+```{r}
+library(ggridges)
+simulation %>%
+  mutate(time_step = 5 * round(time/5),
+         famine = payoff == 0) %>%
+  ggplot(aes(famine, as.factor(time_step))) +
+  geom_density_ridges(rel_min_height = 0.005) +
+  theme_ridges()
+```
+
+make this a function to calculate payoffs for all wheat or all barley strategies as a comparison
+```{r}
+tibble(drought = droughts,
+       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
+       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
+       equal = (barley + wheat) / 2) %>%
+  mutate(time = 1:n()) %>%
+  gather(crop, yield, barley:equal) %>%
+  ggplot(aes(time, yield, group = crop, color = crop)) +
+  geom_point()
+```
+
+```{r}
+tibble(drought = droughts,
+       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
+       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
+       equal = (barley + wheat) / 2) %>%
+  mutate(time = 1:n()) %>%
+  gather(crop, yield, barley:equal) %>%
+  mutate(famine = yield == 0) %>%
+  ggplot(aes(time, famine, group = crop, color = crop)) +
+  geom_line()
+```
+
+```{r}
+test3 <- tibble(agent = 1:100) %>%
+  mutate(wheat_prop = round(runif(n()), 2))
+
+simulation2 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+
+simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
+simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
 ```
@@
 Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
-
-Run an experiment. Simulate a time series with varying drought frequency.
-```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
-```
-
-```{r}
-simulation <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
-
-```{r, echo = TRUE}
-ggplot(simulation, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r, echo = TRUE}
-simulation %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
-```
-
-These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
-```{r}
-ggplot(simulation, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
-```
-
-```{r}
-test3 <- tibble(agent = 1:100) %>%
-  mutate(wheat_prop = round(runif(n()), 2))
-
-simulation2 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .8), .init = test3) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-
-simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
-simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
-```
-
-```{r}
-simulation %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-```
 

### 2020-01-24 risk-percption.Rmd
trace data

@@
 konya <- matrix(c(32.5, 37.9), nrow = 1)
-
-sim_present <- brick('~/Downloads/b40.20th.track1.1deg.005.cam2.h0.PRECT.185001-200512.nc') %>%
-  raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, rainfall) %>%
-  mutate(rainfall = rainfall * 2.628e+9,
-         year = rep(1850:2005, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 156)) %>%
-  select(-time) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-  filter(between(water_year, 1932, 2005)) %>%
-   filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  summarise(precip = sum(rainfall))
-
-sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
-  raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, rainfall) %>%
-  mutate(rainfall = rainfall * 2.628e+9,
-         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 516)) %>%
-  select(-time) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-   filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  summarise(precip = sum(rainfall))
+konya <- matrix(c(33.5138, 36.2765), nrow = 1)
 ```
@@
 ```{r}
-obs <- read_table('~/Downloads/pa17244.dat', 
-           comment = "#", 
-           col_names = c('year', 1:12), 
-           na = c('-999.9', '-888.8')) %>%
-  gather(month, precip, `1`:`12`) %>%
-  group_by(month) %>%
-  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
-  ungroup %>%
-  mutate(month = as.numeric(month),
-         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
-    filter(between(water_year, 1932, 2005)) %>%
- filter(!(month %in% c(6:9))) %>%
+precc <- map(list.files('Data/PRECC', full.names = TRUE)[-12],
+             ~brick(.) %>%
+               raster::extract(konya, df = TRUE) %>%
+  select(-ID) %>%
+  gather(time, precc) ) %>%
+  bind_rows()
+precl <- map(list.files('Data/PRECL', full.names = TRUE)[-12],
+             ~brick(.) %>%
+               raster::extract(konya, df = TRUE) %>%
+  select(-ID) %>%
+  gather(time, precl) ) %>%
+  bind_rows()
+
+sim_past <- left_join(precc, precl) %>%
+  mutate(rainfall = (precc + precl) * 2.628e+9,
+         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
+         month = rep(1:12, times = 5200)) %>%
+  select(-time, -precl, -precc) %>%
+  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
+  # filter(!(month %in% c(6:9))) %>%
   group_by(water_year) %>%
-  summarise(precip = sum(precip))
+  add_tally() %>%
+filter(n == 12) %>%
+  summarise(precip = sum(rainfall))
+
+plot(sim_past)
 ```
@@
 ```{r}
-obs_past <- CDFt(obs$precip, sim_present$precip, sim_past$precip)$DS
+plot(sim_past)
+```
 
-plot(obs_past)
-
-tibble(droughts = (obs_past < 250)[1:500]) %>%
-  mutate(time = rep(1:20, each = 25))%>%
+```{r}
+tibble(droughts = (sim_past[1:5150, 2] < 250)) %>%
+  mutate(time = rep(1:206, each = 25))%>%
   group_by(time) %>%
@@
 ```{r}
-simulation <- accumulate((obs_past < 250), ~learn(.x, .y), .init = agents) %>% 
+simulation <- accumulate((sim_past$precip < 300), ~learn(.x, .y), .init = agents) %>% 
   bind_rows(.id = 'time') %>%

### 2020-01-24 risk-percption.Rmd
trace drought risk plot

@@
 konya <- matrix(c(32.5, 37.9), nrow = 1)
-konya <- matrix(c(33.5138, 36.2765), nrow = 1)
+#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
 ```
@@
   mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-  # filter(!(month %in% c(6:9))) %>%
+   filter(!(month %in% c(6:9))) %>%
   group_by(water_year) %>%
   add_tally() %>%
-filter(n == 12) %>%
-  summarise(precip = sum(rainfall))
+filter(n == 8) %>%
+  summarise(precip = sum(rainfall)) %>%
+  mutate(drought = precip < 300)
+```
 
-plot(sim_past)
+```{r, fig.height =3, fig.width = 9}
+sim_past %>%
+  filter(between(water_year, -9500, -5500)) %>%
+  mutate(time_step = ceiling(water_year / 50),
+         year = time_step * 50) %>%
+  group_by(year) %>%
+  summarise(risk = sum(drought) / 50) %>%
+  ggplot(aes(year, risk, fill = risk, color = risk)) +
+  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, width = 1.2) +
+  geom_col()  +
+  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
+  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
+  theme_minimal() +
+  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
 ```
@@
 ```{r}
-plot(sim_past)
-```
-
-```{r}
-tibble(droughts = (sim_past[1:5150, 2] < 250)) %>%
-  mutate(time = rep(1:206, each = 25))%>%
-  group_by(time) %>%
-  summarise(test = sum(droughts)/25)
-```
-
-```{r}
-simulation <- accumulate((sim_past$precip < 300), ~learn(.x, .y), .init = agents) %>% 
+simulation <- accumulate(sim_past$drought, ~learn(.x, .y), .init = agents) %>% 
   bind_rows(.id = 'time') %>%

### 2020-01-24 risk-percption.Rmd
remove maximax calculation uisng lpsolve

@@
 
-~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+moran_birth_death <- function(agents, r = 1) {
+```{r}
+load('../Silvanus/R/sysdata.rda')
+life_table
+fertility_shape <- 4.579627
+fertility_scale <- 0.1
 
-maximax
+die <- function(individuals, food_ratio = 0.7) {
+  individuals %>%
+        left_join(life_table, by = 'age') %>%
+
 ```{r}
-risk <- 0.9
-wheat_exp <- wheat_wet * (1 - risk) + wheat_dry * risk
-barley_exp <- barley_wet * (1 - risk) + barley_dry * risk
-obj.fun <- c(wheat_exp, barley_exp)
-
-constr=matrix(c(1,1,1,0,0,1, wheat_exp, barley_exp),ncol = 2, byrow = TRUE) #Loading the constraints
-constr.dir=c("=",">=",">=", ">=")
-constr.rhs=c(1,0,0, .5)
-
-mod=lp("min",obj.fun,constr,constr.dir,constr.rhs) #Using lp() to solve our problem
-mod$solution
+reduce(1:600, ~reproduce(.x) %>% die , .init = agents2 %>% mutate(age = sample.int(80, size = 150, replace = TRUE))) %>%
 ```

### 2020-01-24 risk-percption.Rmd
birth-death processes

@@
 ```{r}
-agents <- tibble(id = 1:100, 
+agents <- tibble(id = 1:150, 
                  alpha = 1.5,
@@
 
-so in this case drought risk is low enough all the time you never get any diversification
+```{r}
+agents2 <- agents %>%
+  select(id) %>%
+  mutate(age = 1)
 
 moran_birth_death <- function(agents, r = 1) {
+  n <- nrow(agents)
+  birth_index <- sample.int(n, size = r)
+  death_index <- sample.int(n, size = r)
+  newborn <- agents %>%
+    filter(row_number() %in% birth_index) %>%
+    mutate(age = 1)
+
+  agents %>%
+    filter(!row_number() %in% death_index) %>%
+    mutate(age  = age + 1) %>%
+    bind_rows(newborn)
+}
+
+reduce(1:1000, ~moran_birth_death(., r = 4), .init = agents2) %>%
+  pull(age) %>%hist
+```
+
 ```{r}
@@
 fertility_scale <- 0.1
+survivor_scale <- 0.1
+```
+
+```{r}
+reproduce <- function(individuals, food_ratio = 0.7) {
+  individuals %>%
+    left_join(life_table, by = 'age') %>% 
+    mutate(fertility_reduction = pgamma(food_ratio, shape = fertility_shape, scale = fertility_scale),
+           reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
+    filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
+    tally %>% # count births per household
+    uncount(n) %>% # repeat rows based on birth count per household
+    mutate(age = 0) %>%
+    bind_rows(individuals, .) %>%
+    ungroup # check for filled na's here?
+}
 
@@
         left_join(life_table, by = 'age') %>% 
+    mutate(survival_reduction = pgamma(pmin(1, food_ratio), shape = survival_shape, scale = survivor_scale),
+           survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
+    filter(survived == TRUE) %>%
+    select(-c(fertility_rate:survived)) %>%
+    mutate(age = age + 1)
+}
+
+```
 
@@
 reduce(1:600, ~reproduce(.x) %>% die , .init = agents2 %>% mutate(age = sample.int(80, size = 150, replace = TRUE))) %>%
+  pull(age) %>% hist
 ```
+
+so in this case drought risk is low enough all the time you never get any diversification
+
+~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 

### 2020-01-24 risk-percption.Rmd
rearrange sections

@@
 
+## Bayesian Updating
+
+```{r}
+prior <- function(m,n){
+a = n * m
+b = n * (1 - m)
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+tibble(year = 1:20,
+       risk = .5) %>%
+  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line() +
+  facet_wrap( ~year) +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+
+#prior(.5, ) %>% plot
+```
+
+```{r}
+likelihood <- function(N,Y){
+a <- Y + 1
+b <- N - Y + 1
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+#likelihood(1, 1)%>% plot
+```
+
+```{r}
+posterior <- function(m,n,N,Y){
+a <- Y + (n*m) -1
+b <- N - Y + (n*(1-m)) - 1
+dom <- seq(0,1,0.005)
+val <- dbeta(dom,a,b)
+return(data.frame('x'=dom, 'y'=val))
+}
+
+#dbeta(seq(0,1,0.005), -1, 0)
+#prior(.5, 2) %>% plot
+#posterior(.5, 3, 1, 1) %>% plot
+```
+
 Allow the agents to learn the drought risk over time. Also need to calculate the actual payoffs for each agent.
@@
 
-## Bayesian Updating
-
 ```{r}
-prior <- function(m,n){
-a = n * m
-b = n * (1 - m)
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-tibble(year = 1:20,
-       risk = .5) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line() +
-  facet_wrap( ~year) +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-
-#prior(.5, ) %>% plot
+load('../Silvanus/R/sysdata.rda')
+life_table
+fertility_shape <- 4.579627
+fertility_scale <- 0.1
+survivor_scale <- 0.1
 ```
@@
 ```{r}
-likelihood <- function(N,Y){
-a <- Y + 1
-b <- N - Y + 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
+reproduce <- function(individuals, food_ratio = 0.7) {
+  individuals %>%
+    left_join(life_table, by = 'age') %>% 
+    mutate(fertility_reduction = pgamma(food_ratio, shape = fertility_shape, scale = fertility_scale),
+           reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
+    filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
+    tally %>% # count births per household
+    uncount(n) %>% # repeat rows based on birth count per household
+    mutate(age = 0) %>%
+    bind_rows(individuals, .) %>%
+    ungroup # check for filled na's here?
 }
 
-#likelihood(1, 1)%>% plot
+die <- function(individuals, food_ratio = 0.7) {
+  individuals %>%
+        left_join(life_table, by = 'age') %>% 
+    mutate(survival_reduction = pgamma(pmin(1, food_ratio), shape = survival_shape, scale = survivor_scale),
+           survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
+    filter(survived == TRUE) %>%
+    select(-c(fertility_rate:survived)) %>%
+    mutate(age = age + 1)
+}
+
 ```
@@
 ```{r}
-posterior <- function(m,n,N,Y){
-a <- Y + (n*m) -1
-b <- N - Y + (n*(1-m)) - 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-#dbeta(seq(0,1,0.005), -1, 0)
-#prior(.5, 2) %>% plot
-#posterior(.5, 3, 1, 1) %>% plot
+reduce(1:600, ~reproduce(.x) %>% die , .init = agents2 %>% mutate(age = sample.int(80, size = 150, replace = TRUE))) %>%
+  pull(age) %>% hist
 ```
@@
 
-The computational complexity of the TraCE-21k simulation requires it to be run at a coarse (~200km) spatial resolution.
-At this resolution, key hydroclimate processes such as individual cloud formation cannot be simulated directly, leading to biases in the distribution of modeled precipitation.
-In order to reveal climate patterns at regional scales, outputs from this simulation must be statistically downscaled and bias corrected.
-The cumulative distribution function transform (CDF-t) method is a skillful yet conceptually simple technique to downscale and bias-corrected GCM data [@Michelangeli2009].
-The technique works directly with the CDFs of observed and simulated climate variables, rather than individual observations.
-A function is first derived that transforms the CDF of GCM data to match the CDF of weather-station observations from the same period.
-The same function is then used to transform the CDFs of GCM data from different time periods, accounting for the changes in the CDFs from one simulation period to another (Figure 3).
-CDF-t is nonparametric, and thus more data-driven than weather generators and related statistical downscaling methods that rely to a greater degree on external assumptions about the nature of the data.
-
-![CDFs for simulated and observed precipitation in the present day, and simulated and bias-corrected precipitation in the past.](figures/cdf.png)
-
 TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
 This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
-
-\section{Results and Discussion}
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=16cm]{precon.png}
-\caption{Bias-corrected wet season (October-March) precipitation accumulations from the past 10,000 years of TraCE-21k simulation.
-1 point = 1 year, color corresponds to whether that year falls bellow the thresholds for wheat or barley crop failure.}
-\end{figure}
-
-The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure 5).
-The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
-
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
-Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
-
-![Annual risk of wheat crop failure due to drought, averaged by century.](figures/wheatrisk.png)
-
-Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
-$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
-This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
-Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
-Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
-
-Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
-Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
-Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
-
-During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
-Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
-This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
-This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
-
-Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
-To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
-Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
-Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
-
-Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
-
-# References
-
-```{r}
-knitr::knit_exit()
-```
 
@@
 
+\section{Results and Discussion}
+
+\begin{figure}[htp]
+\centering
+\includegraphics[width=16cm]{precon.png}
+\caption{Bias-corrected wet season (October-March) precipitation accumulations from the past 10,000 years of TraCE-21k simulation.
+1 point = 1 year, color corresponds to whether that year falls bellow the thresholds for wheat or barley crop failure.}
+\end{figure}
+
+The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
+8ka BP (Figure 5).
+The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
+
+The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
+Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
+
+![Annual risk of wheat crop failure due to drought, averaged by century.](figures/wheatrisk.png)
+
+Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
+$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
+This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
+Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
+Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
+
+Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
+Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
+Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
+
+During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
+Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
+This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
+This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
+
+Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
+To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
+Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
+Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
+
+Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
+
+# References
+
+```{r}
+knitr::knit_exit()
+```
+
 ```{r}
@@
   pull(age) %>%hist
-```
-
-```{r}
-load('../Silvanus/R/sysdata.rda')
-life_table
-fertility_shape <- 4.579627
-fertility_scale <- 0.1
-survivor_scale <- 0.1
-```
-
-```{r}
-reproduce <- function(individuals, food_ratio = 0.7) {
-  individuals %>%
-    left_join(life_table, by = 'age') %>% 
-    mutate(fertility_reduction = pgamma(food_ratio, shape = fertility_shape, scale = fertility_scale),
-           reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
-    filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
-    tally %>% # count births per household
-    uncount(n) %>% # repeat rows based on birth count per household
-    mutate(age = 0) %>%
-    bind_rows(individuals, .) %>%
-    ungroup # check for filled na's here?
-}
-
-die <- function(individuals, food_ratio = 0.7) {
-  individuals %>%
-        left_join(life_table, by = 'age') %>% 
-    mutate(survival_reduction = pgamma(pmin(1, food_ratio), shape = survival_shape, scale = survivor_scale),
-           survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
-    filter(survived == TRUE) %>%
-    select(-c(fertility_rate:survived)) %>%
-    mutate(age = age + 1)
-}
-
-```
-
-```{r}
-reduce(1:600, ~reproduce(.x) %>% die , .init = agents2 %>% mutate(age = sample.int(80, size = 150, replace = TRUE))) %>%
-  pull(age) %>% hist
 ```

### 2020-01-24 risk-percption.Rmd
fresh priors for newborns

@@
     uncount(n) %>% # repeat rows based on birth count per household
-    mutate(age = 0) %>%
+    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta)) %>%
     bind_rows(individuals, .) %>%
@@
 
+Ok so next we need to have the birth and death stuff run with all the prior information in it.
+dont need to propogate the prior exactly, just need to add in a new weak prior.
+at least propogate the risk number, but not the prior?
 ```{r}

### 2020-01-24 risk-percption.Rmd
incorporate payoffs into population dynamics

@@
 
-```{r}
-readxl::read_xls('41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
+```{r, eval = FALSE}
+readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
   select(`Site Location`, Taxon, `ky cal BP`, `GY (t*ha-1)`) %>%
@@
 Given these payoffs, what's the point of indifference between the two crop types?
-```{r, echo = FALSE, fig.cap=title = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
+```{r, echo = FALSE, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
 indiff <- tibble(risk = (barley_normal - wheat_normal) / (wheat_dry - wheat_normal - barley_dry + barley_normal),
@@
                  beta = 1.5,
-                 risk = alpha / (alpha + beta))
+                 risk = alpha / (alpha + beta),
+                 age = sample.int(80, size = 150, replace = TRUE),
+                 crop = 'barley')
 ```
@@
 ```{r}
-simulation <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
+simulation1 <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
   bind_rows(.id = 'time') %>%
@@
 ```{r, echo = TRUE}
-ggplot(simulation, aes(time, risk)) +
+ggplot(simulation1, aes(time, risk)) +
   geom_point(alpha = .1) +
@@
 
-```{r, echo = TRUE}
-simulation %>%
+```{r}
+simulation2 <- accumulate(droughts, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+```
+
+```{r}
+ggplot(simulation2, aes(time, risk)) +
+  geom_point(alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
+```
+
+```{r}
+simulation1 %>%
   group_by(time, crop) %>%
@@
 ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity') +
+  geom_bar(width = 1, stat = 'identity', position = 'fill') +
+  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
+  theme_classic()
+```
+
+```{r}
+simulation2 %>%
+  group_by(time, crop) %>%
+  tally() %>%
+ggplot(aes(time, n, fill = crop)) +
+  geom_bar(width = 1, stat = 'identity', position = 'fill') +
   scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
@@
 These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
+
+Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals
 ```{r}
-ggplot(simulation, aes(time, wheat_prop)) +
+ggplot(simulation1, aes(time, wheat_prop)) +
+  geom_point(alpha = .3) +
+  theme_classic()
+
+ggplot(simulation2, aes(time, wheat_prop)) +
   geom_point(alpha = .3) +
@@
 ```{r}
-simulation %>%
+simulation1 %>%
   group_by(time) %>%
-  summarise(famine = sum(payoff < .1)) %>%
+  summarise(famine = sum(payoff == 0)) %>%
+ggplot(aes(time, famine)) +
+  geom_line() +
+  theme_classic()
+simulation2 %>%
+  group_by(time) %>%
+  summarise(famine = sum(payoff == 0, na.rm = TRUE)) %>%
 ggplot(aes(time, famine)) +
@@
 ```{r}
-simulation %>%
+simulation1 %>%
   group_by(time) %>%
   summarise(payoff = mean(payoff)) %>%
+ggplot(aes(time,payoff)) +
+  geom_point(aes(color = time)) +
+  theme_classic()
+
+simulation2 %>%
+  group_by(time) %>%
+  summarise(payoff = mean(payoff, na.rm = TRUE)) %>%
 ggplot(aes(time,payoff)) +
@@
 library(ggridges)
-simulation %>%
+simulation1 %>%
   mutate(time_step = 5 * round(time/5),
          famine = payoff == 0) %>%
-  ggplot(aes(famine, as.factor(time_step))) +
+  ggplot(aes(wheat_prop, as.factor(time_step))) +
+  geom_density_ridges(rel_min_height = 0.005) +
+  theme_ridges()
+
+simulation2 %>%
+  mutate(time_step = 5 * round(time/5),
+         famine = payoff == 0) %>%
+  ggplot(aes(wheat_prop, as.factor(time_step))) +
   geom_density_ridges(rel_min_height = 0.005) +
@@
 
-simulation2 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
+simulation3 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
   bind_rows(.id = 'time') %>%
@@
 ```{r}
-reproduce <- function(individuals, food_ratio = 0.7) {
+reproduce <- function(individuals) {
   individuals %>%
     left_join(life_table, by = 'age') %>% 
-    mutate(fertility_reduction = pgamma(food_ratio, shape = fertility_shape, scale = fertility_scale),
+    mutate(fertility_reduction = pgamma(payoff, shape = fertility_shape, scale = fertility_scale),
            reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
@@
     uncount(n) %>% # repeat rows based on birth count per household
-    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta)) %>%
+    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta), crop = 'barley') %>%
     bind_rows(individuals, .) %>%
@@
 
-die <- function(individuals, food_ratio = 0.7) {
+die <- function(individuals) {
   individuals %>%
         left_join(life_table, by = 'age') %>% 
-    mutate(survival_reduction = pgamma(pmin(1, food_ratio), shape = survival_shape, scale = survivor_scale),
+    mutate(survival_reduction = pgamma(pmin(1, payoff), shape = survival_shape, scale = survivor_scale),
            survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
@@
 ```{r}
-reduce(1:600, ~reproduce(.x) %>% die , .init = agents2 %>% mutate(age = sample.int(80, size = 150, replace = TRUE))) %>%
+reduce(1:600, ~reproduce(.x) %>% die , .init = agents) %>%
   pull(age) %>% hist
@@
 
-```{r, fig.height =3, fig.width = 9}
+```{r, fig.height =3, fig.width = 9, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by century.'}
 sim_past %>%
@@
 ```{r}
-simulation <- accumulate(sim_past$drought, ~learn(.x, .y), .init = agents) %>% 
+simulation4 <- accumulate(sim_past$drought, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
   bind_rows(.id = 'time') %>%
@@
 
+So the problem now is the population dynamics are getting in the way ... and that it'd be nice to have something akin to the moran process to ensure fixed population
 ```{r, echo = TRUE}
-ggplot(simulation, aes(time, risk)) +
+ggplot(simulation4, aes(time, risk)) +
   geom_point(alpha = .1) +
@@
   theme_classic()
+```
+
+```{r}
+simulation4 %>%
+  group_by(time) %>%
+  tally %>%
+  plot
+```
+
+so we have to connect the food ratio to the payoffs? maybe have the barley yield be ~ .8 food ratio
+
+```{r}
+.5/.8
 ```
@@
 Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
-
-![Annual risk of wheat crop failure due to drought, averaged by century.](figures/wheatrisk.png)
 
@@
 ```{r}
-agents2 <- agents %>%
-  select(id) %>%
-  mutate(age = 1)
-
 moran_birth_death <- function(agents, r = 1) {
@@
 ```
-
-```{r}
-farm <- function(households){
-  households %>%
-    mutate(yield = climatic_yield, #* n_inhabitants ^ labor_elasticity,
-           yield_memory = yield, #map2(yield_memory, yield, remember),
-           harvest = land * yield * .5 - land * sowing_rate) #%>%  # *.5 is for fallow
-  #select(-yield)
-}
-
-remember <- function(yield_memory, yield){
-  # rnorm(1, yield, yield * 0.0333) %>%  #memory is fuzzy
-  append(yield_memory[-length(yield_memory)], yield, after = 0) # remove the last entry in the vector and add new yield to the begining
-}
-
-#Agents use the peak-end rule when accessing memory.
-peak_end <- function(x){
-  map_dbl(x, ~mean(c(.x[1], min(.x))))
-}
-
-```
-
-```{r}
-obs.na <- read.csv('salihli_p.csv')[10:620, 5]  / 10  # raw data in mm*10. subset time series as water years (i.e. first observation is october, last is september)
-
-get.winter <- function(x){seq(0, length(x) - 12, 12) %>%
-    sapply(function(y) y + 1:6) %>%
-    as.numeric %>% extract(x, .)}# function to get Oct-Mar values from
-obs<- obs.na %>% 
-  rev %>% 
-  na.locf %>% 
-  rev %>% 
-  `+`(na.locf(obs.na)) %>% 
-  `/`(2) %>%  # fill NA with avg of adjacent values
-  get.winter
-
-# Now import TraCE-21k GCM data.
-Extract a time series for the grid cell covering the Salihli weather station.
-Add the model-output convective and large-scale precipitation variables to get total precipitation.
-Convert from m/s to mm/month.
-Subset time series to the calibration period (10 BP - AD 1940) and prediction period (10000 BP - 11 BP).
-
-salihli.pt <- c(28.14, 38.48) %>% matrix(nrow = 1) %>% SpatialPoints  # create a point in Salihli to extract data from GCM
-
-pc <- list.files('GCM Data/PRECC', full.names = T) %>%  # get a list of all GCM output files
-  lapply(brick) %>%                                 # import GCM files as raster bricks
-  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
-  unlist                                            # combine into a single vector
-
-pl <- list.files('GCM Data/PRECL', full.names = T) %>%  # get a list of all GCM output files
-  lapply(brick) %>%                                 # import GCM files as raster bricks
-  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
-  unlist                                            # combine into a single vector
-
-pt <- pc %>% add(pl) %>% multiply_by(60 * 60 * 24 * 30 * 1000)  # add precip components and convert from m/s to mm/month
-
-cal <- pt %>% extract(122278:122877) %>% get.winter
-
-pred <- pt %>% extract(2398:122397) %>% get.winter
-
-pred.byfifty <- pred %>% seq_along %>% divide_by(300) %>% ceiling %>% split(pred,.)
-
-calc.cdf <- function(x){
-  pred <- x
-  ct <- CDFt(obs, cal, pred)
-  ct$DS
-}
-monthly.recon <- lapply(pred.byfifty, calc.cdf) %>% unlist
-annual.recon <- monthly.recon %>% seq_along %>% divide_by(6) %>% ceiling %>% split(monthly.recon,.) %>%
-  lapply(sum) %>% unlist %>% data.frame(date = seq(-10000,-1,1), p = ., color = cut(., c(0,200,300, 800)))
-
-```

### 2020-01-25 risk-percption.Rmd
more plots of subjective probability

@@
 barley_normal <- 1.18
-```
-
-```{r, eval = FALSE}
-readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
-  select(`Site Location`, Taxon, `ky cal BP`, `GY (t*ha-1)`) %>%
-  filter(!is.na(`GY (t*ha-1)`)) %>% 
-  group_by(`Site Location`, Taxon) %>%
-  summarise(yield_mean = mean(`GY (t*ha-1)`)) %>%
-  spread(Taxon, yield_mean)
 ```
@@
 
+## Climate Modeling and Bias Correction
+
+Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
+TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
+It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
+orbit, greenhouse gasses, glacial meltwater flux).
+The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
+
+TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
+This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
+
+Now incorporate realistic drought variability from simulations.
+
+```{r}
+konya <- matrix(c(32.5, 37.9), nrow = 1)
+#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
+```
+
+```{r}
+precc <- map(list.files('Data/PRECC', full.names = TRUE)[-12],
+             ~brick(.) %>%
+               raster::extract(konya, df = TRUE) %>%
+  select(-ID) %>%
+  gather(time, precc) ) %>%
+  bind_rows()
+precl <- map(list.files('Data/PRECL', full.names = TRUE)[-12],
+             ~brick(.) %>%
+               raster::extract(konya, df = TRUE) %>%
+  select(-ID) %>%
+  gather(time, precl) ) %>%
+  bind_rows()
+
+sim_past <- left_join(precc, precl) %>%
+  mutate(rainfall = (precc + precl) * 2.628e+9,
+         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
+         month = rep(1:12, times = 5200)) %>%
+  select(-time, -precl, -precc) %>%
+  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
+   filter(!(month %in% c(6:9))) %>%
+  group_by(water_year) %>%
+  add_tally() %>%
+filter(n == 8) %>%
+  summarise(precip = sum(rainfall)) %>%
+  mutate(drought = precip < 300)
+```
+
+```{r, fig.height =3, fig.width = 9, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period'}
+sim_past %>%
+  filter(between(water_year, -9500, -5500)) %>%
+  mutate(time_step = ceiling(water_year / 50),
+         year = time_step * 50) %>%
+  group_by(year) %>%
+  summarise(risk = sum(drought) / 50) %>%
+  ggplot(aes(year, risk, fill = risk, color = risk)) +
+  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, width = 1.2) +
+  geom_col()  +
+  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
+  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
+  theme_minimal() +
+  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
+```
+
 Start with a population of agents with the same subjective beliefs about drought risk.
@@
 ```{r}
-prior <- function(m,n){
-a = n * m
-b = n * (1 - m)
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
+prior <- function(m, n) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, n * m, n * (1 - m))
+  tibble('x' = domain, 'y' = val)
 }
 
-tibble(year = 1:20,
+tibble(year = seq(5, 50, 5),
        risk = .5) %>%
@@
   ggplot(aes(x, y)) +
-  geom_line() +
-  facet_wrap( ~year) +
+  geom_line(aes(group = year, color = year), size = 1.2) +
+  scale_color_viridis_c() +
   theme_bw() +
   ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-
-#prior(.5, ) %>% plot
-```
-
-```{r}
-likelihood <- function(N,Y){
-a <- Y + 1
-b <- N - Y + 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-#likelihood(1, 1)%>% plot
-```
-
-```{r}
-posterior <- function(m,n,N,Y){
-a <- Y + (n*m) -1
-b <- N - Y + (n*(1-m)) - 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-#dbeta(seq(0,1,0.005), -1, 0)
-#prior(.5, 2) %>% plot
-#posterior(.5, 3, 1, 1) %>% plot
 ```
@@
 ```{r}
-learn <- function(agents, drought) {
+learn1 <- function(agents, drought) {
   if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
@@
            beta = beta + 1 - drought,
-           mu = alpha / (alpha + beta),
+          # mu = alpha / (alpha + beta),
+           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
+}
+
+learn2 <- function(agents, drought) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
+
+  agents %>%
+    # farmer's turn to decide what to plant
+      mutate(yield_exp = expected_yield(risk),
+         barley_exp = expected_yield_barley(risk),
+         wheat_exp = expected_yield_wheat(risk),
+         expected_yield = barley_exp + wheat_exp,
+         crop = case_when(barley_exp > wheat_exp ~ 'barley',
+                              barley_exp < wheat_exp ~ 'wheat',
+                              barley_exp == wheat_exp ~ 'either'),
+         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
+         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
+      select(-prediction) %>%
+  # nature's turn to decide realized yields
+    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
+           alpha = if_else(payoff < .5, 1, drought + alpha),
+           beta = if_else(payoff < .5, 1, beta + 1 - drought),
+          # mu = alpha / (alpha + beta),
            risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
@@
 
-Similar to above, but satisficing
-```{r}
-learn_satisficing <- function(agents, drought, aspiration) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
-
-  agents %>%
-     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-            new_prop = round(runif(n()), 2),
-           wheat_prop = if_else(payoff > aspiration, wheat_prop, new_prop)) %>%
-    select(-new_prop)
-}
-```
-
-Run an experiment. Simulate a time series with varying drought frequency.
+Run an experiment. Simulate a time series with varying drought frequency. This is a stepwise nonstationary process
 ```{r}
@@
 ```{r}
-simulation1 <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
+prior2 <- function(a, b) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, a, b)
+  tibble('x' = domain, 'y' = val)
+}
+
+plot(prior2( 0, 0));plot(prior2( .5, .5));plot(prior2( 1, 1))
+
+accumulate(droughts, 
+           ~mutate(.x, alpha = .y + alpha, beta = beta + 1 - .y), 
+           .init = tibble(alpha = 1.5, beta = 1.5)) %>%
+  bind_rows(.id = 'time') %>%
+    mutate(time = as.numeric(time), belief = map2(alpha, beta, ~prior2(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(color = time, group = time), size = .8) +
+    geom_vline(xintercept = c(0.1, 0.5, 0.2), linetype = 2, color = 'red', size = .8) +
+  scale_color_viridis_c() +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+```
+
+```{r}
+simulation1 <- accumulate(droughts, ~learn1(.x, .y), .init = agents) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+
+simulation2 <- accumulate(droughts, ~learn2(.x, .y), .init = agents) %>% 
   bind_rows(.id = 'time') %>%
@@
 
+Redo these plots so its a single person, one who was born in a drought period then experiences an ameliorarion and another who experiences a shift to drought conditions later in life?
+basically this section neds to work more with individuals and explore that stuff
+
 ```{r, echo = TRUE}
+test1 <- tibble(time = 1:100,
+       risk = c(rep(0.1, 25), rep(0.5, 25), rep(0.2, 50)))
+
 ggplot(simulation1, aes(time, risk)) +
+  geom_line(data = test1) +
   geom_point(alpha = .1) +
@@
   theme_classic()
+```
+
+```{r}
+ggplot(simulation2, aes(time, risk)) +
+  geom_line(data = test1) +
+  geom_point(alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
+
 ```
@@
 ```{r}
-test3 <- tibble(agent = 1:100) %>%
-  mutate(wheat_prop = round(runif(n()), 2))
-
-simulation3 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-
-simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
-simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
-```
-
-```{r}
 load('../Silvanus/R/sysdata.rda')
@@
     left_join(life_table, by = 'age') %>% 
-    mutate(fertility_reduction = pgamma(payoff, shape = fertility_shape, scale = fertility_scale),
+    mutate(fertility_reduction = pgamma(pmin(1, payoff / 0.5), shape = fertility_shape, scale = fertility_scale),
            reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
     filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
-    tally %>% # count births per household
-    uncount(n) %>% # repeat rows based on birth count per household
-    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta), crop = 'barley') %>%
-    bind_rows(individuals, .) %>%
-    ungroup # check for filled na's here?
+    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta)) %>%
+    select(-c(fertility_rate:reproduced)) %>%
+    bind_rows(individuals, .)
 }
@@
         left_join(life_table, by = 'age') %>% 
-    mutate(survival_reduction = pgamma(pmin(1, payoff), shape = survival_shape, scale = survivor_scale),
+    mutate(survival_reduction = pgamma(pmin(1, payoff / 0.5), shape = survival_shape, scale = survivor_scale),
            survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
@@
 }
-
 ```
@@
 
-## Climate Modeling and Bias Correction
-
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
-TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
-It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
-orbit, greenhouse gasses, glacial meltwater flux).
-The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
-
-![Example model precipitation outputs from the TraCE-21k simulation at 9.5ka BP.](figures/trace_white.png)
-
-TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
-This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
-
-Now incorporate realistic drought variability from simulations.
+Another question here about initialization ... namely that we are starting with 80 year olds with no beliefs about the system?
 
 ```{r}
-konya <- matrix(c(32.5, 37.9), nrow = 1)
-#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
-```
-
-```{r}
-precc <- map(list.files('Data/PRECC', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precc) ) %>%
-  bind_rows()
-precl <- map(list.files('Data/PRECL', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precl) ) %>%
-  bind_rows()
-
-sim_past <- left_join(precc, precl) %>%
-  mutate(rainfall = (precc + precl) * 2.628e+9,
-         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 5200)) %>%
-  select(-time, -precl, -precc) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-   filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  add_tally() %>%
-filter(n == 8) %>%
-  summarise(precip = sum(rainfall)) %>%
-  mutate(drought = precip < 300)
-```
-
-```{r, fig.height =3, fig.width = 9, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by century.'}
-sim_past %>%
-  filter(between(water_year, -9500, -5500)) %>%
-  mutate(time_step = ceiling(water_year / 50),
-         year = time_step * 50) %>%
-  group_by(year) %>%
-  summarise(risk = sum(drought) / 50) %>%
-  ggplot(aes(year, risk, fill = risk, color = risk)) +
-  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, width = 1.2) +
-  geom_col()  +
-  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
-  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
-  theme_minimal() +
-  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
-```
-
-```{r}
-simulation4 <- accumulate(sim_past$drought, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
+simulation3 <- accumulate(sim_past$drought, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
   bind_rows(.id = 'time') %>%
@@
 
-So the problem now is the population dynamics are getting in the way ... and that it'd be nice to have something akin to the moran process to ensure fixed population
-```{r, echo = TRUE}
-ggplot(simulation4, aes(time, risk)) +
+So the problem now is the population dynamics are getting in the way ...
+and that it'd be nice to have something akin to the moran process to ensure fixed population.
+OR recognize the importance of population dynamics and have the full continuous yields from the medland functions and just determine the chunked payoffs from there?
+The problem now is that if we set the food requirement to calculate the food ratio to something like 0.5, then there is no benefit to the increased yields and hence no real tradeoff between yields and risk management.
+perhaps just record the cumulative surplus and have that be something
+```{r}
+ggplot(simulation3, aes(time, risk)) +
   geom_point(alpha = .1) +
@@
 ```{r}
-simulation4 %>%
+ggplot(simulation3, aes(time, payoff)) +
+  geom_point(alpha = .1) +
+  #geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic()
+```
+
+```{r}
+simulation3 %>%
   group_by(time) %>%
@@
 
-```{r}
-library(ggridges)
-accumulate((runif(25) < .2), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .6), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(wheat_prop, y = as.numeric(time), group = as.numeric(time))) +
-  geom_density_ridges() +
-  theme_minimal()
-```
-
-```{r}
-# this is for nature
-calcNE <- function(risk){
-  (118 - 25 * risk) / (278 - 185 * risk)
-}
-
-plot(0:10, calcNE((0:10)/10))
-```
-
 so we're imagining a yield of 1 for barley and 2 for wheat, with equal proportions of each. then in the next time step we'd revise the proportion to take away .25
@@
 # so this works as long as you're between 0 and 1
+
+```
+
+```{r, eval = FALSE}
+readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
+  rename(site = `Site Location`, taxon = Taxon, age = `ky cal BP`, yield = `GY (t*ha-1)`) %>%
+  select(site, taxon, age, yield) %>%
+  #filter(!is.na(yield)) %>%
+  mutate(yield = if_else(is.na(yield), 0, yield)) %>%
+    ggplot(aes(yield)) +
+  geom_histogram(aes(fill = taxon)) +
+  facet_wrap(~taxon, nrow = 2)
+
+readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
+  rename(site = `Site Location`, taxon = Taxon, age = `ky cal BP`, yield = `GY (t*ha-1)`) %>%
+  select(site, taxon, age, yield) %>%
+  filter(!is.na(yield)) %>%
+  filter(site == 'Akarçay Tepe') %>%
+  mutate(yield = yield * 1000,
+         yield = if_else(taxon == 'Hordeum vulgare', yield * 3000, yield * 3540)) %>%
+group_by(taxon) %>%
+  summarise(yield_mean = mean(yield), yield_sd = sd(yield)) %>%
+  mutate(yield = map2(yield_mean, yield_sd, ~rnorm(100000, .x, .y))) %>%
+  select(taxon, yield) %>%
+  unnest(cols = c(yield)) %>% 
+  mutate(yield = if_else(yield > 0, yield, 0)) %>%
+  ggplot(aes(yield)) +
+  geom_histogram(aes(fill = taxon)) +
+  facet_wrap(~taxon, nrow = 2)
 ```
@@
 ```{r}
-yield_max_barley <- 2500
-yield_max_wheat <- 3500
-kcal_barley <- 3000
-kcal_wheat <- 3540
+rnorm(10000, 1582521,	780883.5) %>% hist
+rnorm(10000, 3060657,	1461350.0) %>% hist
+```
+
+```{r}
+yield_max_barley <- 2.500
+yield_max_wheat <- 3.500
+kcal_barley <- 1 #3000
+kcal_wheat <- 3540 / 3000
 
@@
   geom_hline(yintercept = yield_max_barley * kcal_barley) +
-  geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
-  geom_vline(xintercept = 0.3) +
-  geom_vline(xintercept = .2)
+  #geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
+  geom_vline(xintercept = 0.3)
 ```
@@
 ```
+
+Similar to above, but satisficing
+```{r}
+learn_satisficing <- function(agents, drought, aspiration) {
+  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
+  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
+
+  agents %>%
+     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
+            new_prop = round(runif(n()), 2),
+           wheat_prop = if_else(payoff > aspiration, wheat_prop, new_prop)) %>%
+    select(-new_prop)
+}
+```
+
+```{r}
+test3 <- tibble(agent = 1:100) %>%
+  mutate(wheat_prop = round(runif(n()), 2))
+
+simulation3 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
+  bind_rows(.id = 'time') %>%
+  mutate(time = as.numeric(time)) %>%
+  filter(time != 1)
+
+simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
+simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
+```

### 2020-02-07 manuscript/manuscript.Rmd
pull analysis into new manuscript and combine with manuscript text

--- 2020-01-25 risk-percption.Rmd
@@
 ---
-title: "Crop diversification in a game against nature in a mediterranean agro-ecosystem"
-author: "Nick Gauthier"
-date: ""
-bibliography: refs.bib 
-output: html_document
+title: The dynamics of risk perception in a Mediterranean agroecosystem
+subtitle: 
+titlerunning: The dynamics of risk perception in a Mediterranean agroecosystem
+authorrunning: Gauthier
+thanks: | 
+    Grants or other notes about the article that should go on the front 
+    page should be placed here. General acknowledgments should be placed at the
+    end of the article.
+
+authors: 
+- name: Nicolas Gauthier 1
+  address: Laboratory of Tree-Ring Research & School of Geography and Development, University of Arizona
+  email: ngauthier@email.arizona.edu
+
+keywords:
+- key
+- dictionary
+- word
+
+#PACS: 
+#- PAC1
+#- superPAC
+
+MSC:
+- MSC code 1
+- MSC code 2
+
+abstract: |
+  Small-scale agriculturalists in the Mediterranean Basin rely on multiple strategies including diversification, intensification, and storage to maintain a stable food supply in the face of environmental uncertainty.
+Each of these strategies requires farmers to make specific resource allocation decisions in response to environmental risks and is thus sensitive to variability in both the spatiotemporal pattern of risk and the ability of farmers to perceive that pattern.
+In this chapter, I present an agent-based model of a Mediterranean agroecosystem.
+By driving the model with realistic environmental dynamics derived from simulations of mid-Holocene Mediterranean climate, and by allowing the psychology of risk perception to vary among individual farmers, I explore the hidden vulnerabilities of traditional risk-management strategies to periods of rapid climate change.
+I show that even when farmers are able to manage risk “optimally” in light of past experience, unanticipated changes in the spatiotemporal pattern of rainfall can still lead to major food shortfalls.
+
+bibliography: bibliography.bib
+bibstyle: spbasic
+# bibstyle options spbasic(default), spphys, spmpsci
+output: rticles::springer_article
 ---
@@
 library(tidyverse)
-library(CDFt)
 ```
 
-# Introduction
+```{r}
+set.seed(1111)
+```
+
+# Introduction {#intro}
 
@@
 
-![Conceptual model of crop diversification as a regulatory feedback system.
-Uncertain rainfall leads to uncertain returns from land-use systems.
-To keep food supplies stable, farmers must monitor year-to-year variations and decide what combination of crops works best to minimize that variance.](figures/regulatory_feedback.png)
+The secret to the success of earliest farmers on the mediterrenan littoral was in their manipulation of the local ecosystem to manage environmental risks.
+Early farmers constructed a complex agroecosystem to manage risk.
+They needed to maintain stable crop yields in face of uncertain drought risks.
+Yield stability is very important for traditional agriculturlal commmunities not connected to larger bulk food networks (Abbo et al 2010)
+neolithic crop package -- weiss zohary 2011
+cite abbo et al 2010? for more stuff about ladraces
+Wheat/barley systems
+example of system that combines low yield, low wirks crop with high yield high risk crop, an effective context paut et al 2019
+These strategies work when risks are knonw and constant, but may be more susceptible to periods of changing climateic risks
+so understanding how people of the past adapted to change is important to us today
+
 Are these crop-diversification strategies vulnerable to the same dynamics as other social-ecological systems with similar regulatory feedbacks mechanisms?
@@
 
+Questions:
+How would the earliest farmers have made decisions (dealt with uncertainty?) about crop diversification?
+ie what were the different algorithms? -- better understand the decision context
+What are the consequences of theses different decisions (i.e. how well did they perform) at different points in the history of the mediterranean
+How were farmers able to detect changes in the environment, rather than unconsciously adapt to them?
+what are the different consequences we'd expect for each
+THIS is the most important question because with it I can frame the results around and individual's perception of droughts in the neolithic, and the populaiton-level distribution of drought perceptions at given times
+
 I couple a climate model with high temporal resolution to a simple game theoretic model of agricultural decision making under uncertainty, in order to examine the performance of optimal and suboptimal wheat-barley diversification practices.
@@
 
-# Decision-Making in a Game Against Nature
+# Decision-making in a game against nature {#sec:1}
 
-Assume two different crops, a low yield drought tolerant crop (barley) and a high yield drought susceptible crop (wheat).
+The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a game against nature.
+We can represent this simplified decision context as a "game" in a game theorietic context.
+cite luce raiffa 1989 for game against nature, and milnor 1952, Agrawal1968, cassidy1971, gould 1963
+The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertianty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years.
+payoff matrix goes here
+in reality continuous, but cite behavioral econ for people often intuitively solve an easier problem when faced with a complex real-world situation
+The exact values in the payoff matrix here are less important than the relative payoffs in each quadrant.
+We can think of this game as being the culturally-inherited object. our question is how do individuals make decisions here?
+so in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio -- both individual and social learning influence perceptions of eindividual vs social learning tucker 2007
 
@@
 
-||Dry Year|Normal Year|
-|-------|-----|-----|----|
-|Barley Yield|0.93|1.18|
-|Wheat Yield|0|1.60|
+\begin{table}
+\centering
+\caption{Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].}
+\begin{tabular}{|l|l|l|}
+\hline
 
-Table: Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].
+ & Dry Year & Normal Year \\ \hline
+Barley Yield & 0.93 & 1.18 \\ \hline
+Wheat Yield & 0 & 1.60 \\ \hline
+
+\end{tabular}
+\end{table}
 
@@
 
-Optimal wheat/barley proportions can be calculated using two different objective functions representing different decision making types of farmers.
+How should an early farmer make a decision in light ot this uncertainty?
+The basic logic of crop diversitication is Portfolio of crops ala modern portfolio theory: Blank2001
+Sometimes best to think of nature as a sentient opponent out to get you, and play strategically based on that assumption. Gould 1963,  Beckenkamp 2008
+quick overview of criteria without risk, then say that a better option is to estimate the probabilities of the different state of nature
+Agrawal1968 talks about the strategies in the context of agriculture being a game against nature
+here I can slip in the cool thing about acting like the weather is out to get you being a good idea in states of complete uncertainty
+transition -- the problem becomes easier if you have at least someI idea of nature's moves , because then you cna work to get the highest yeilds given the risk of drought
+Decision strategies as choosing different points -- plot of decisions
 
-## Strategies
+\begin{align}
+a^2+b^2=c^2
+\end{align}
 
-1. Wald Maximin -- pessimist, maximize the worst case scenario
-2. Savage Regret -- minimize regret
-3. Hurewicz - balance optimism and pessimism
-4. Laplace or "naive" - 50/50
-
-So both the Wald and Savage regret criteria tell us to choose barley all the time.
-
-If we're using the empirical frequency distirubtion of nature's moves, then we are playing a fictitious game.
-The strategy of planting the crop with the highest expected yield, given the frequency distribution of nature is known, will only work when nature is stationary.
-
-We could think of other algorithms that would minimize the cumulative or average regret of a player over many years in a situation with nonstationary rainfall, but we have to remain within cognitive limits.
-
-The pure maximin strategy is to plant barley:
-
-Alternatively, yield maximizing farmers simply seek to maximize the function for their expected utility given all possible states of nature
-\begin{equation}
-    E\left[ U_{all} \right] = E \left[U_{dry} \right] \times P_{dry}^{*} + E \left[U_{wet} \right] \times \left(1-P_{dry}^{*}\right).
-\end{equation}
-
-Given this information on the potential payoffs for planting each crop type, and information on the (perceived) probability of a drought occurring, the crop-diversification decision practice can be thought of as a simple two-player game against nature [@GOULD1963a].
-The expected yields for a farmer given a fixed proportion of barley and wheat can be calculated as
-
-\begin{equation}
-    E \left[Y_{dry} \right] = W_{dry} \times  P + B_{dry} \times (1 - P)
-\end{equation}
-if a drought occurs and likewise
-\begin{equation}
-    E \left[Y_{wet} \right] = W_{wet} \times P + B_{wet} \times (1 - P)
-\end{equation}
-
-if a drought does not occur. This specification allows us to calculate the optimal proportion of wheat to barley one should plant in a diversified strategy as a mixed strategy profile.
-
-Risk minimizing farmers play the equivalent of a mixed strategy Nash equilibrium, setting the expected yields in drought and normal years equal to each other, weighted by risk, and solving for the wheat/barley proportion
-\begin{equation}
-    E \left[Y_{dry} \right] \times P_{dry}^{*} = E \left[U_{wet} \right] \times \left(1- P_{dry}^{*}\right), 
-\end{equation}
-where $P_{dry}^{*}$ is the \emph{perceived}, not realized, risk of a drought in a particular year.
-The solution to this function represents the ratio of wheat to barley that would make a farmer indifferent to whether a drought would occur in a given year.
-
-The expected yield is thus calculated as.
 ```{r}
@@
 
-Given these payoffs, what's the point of indifference between the two crop types?
 ```{r, echo = FALSE, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
@@
 
+If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility
+If the probability distributin of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
+In the language of game theory, this strategy is knoan as a "game of ficticious play" against nature.
+This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
+This strategy works well when the environment is stationary -- nature plays from a fixed probability distirubtion, but is vulnerable to environmental nonstationarity.
+That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because you're too stuck in your ways while the world changes around you.
+Thus it is important to understand the dynamics of risk perception -- how subjective risks, and thus subjective expected utility -- rise and fall in uncertain environments.
+
+# The dynamics of risk perception {#sec:2}
+
 ```{r}
-# farmer's ne
-pmax(0, round((barley_normal - barley_dry) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2))
-#nature's ne
-round((barley_normal - wheat_normal) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2)
+prior <- function(m, n) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, n * m, n * (1 - m))
+  tibble('x' = domain, 'y' = val)
+}
+
+tibble(year = seq(5, 50, 5),
+       risk = .5) %>%
+  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = year, color = year), size = 1.2) +
+  scale_color_viridis_c() +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+
+tibble(year = 50,
+       risk = seq(.1,.9, .1)) %>%
+  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = risk, color = risk), size = 1.2) +
+  scale_color_viridis_c() +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
 ```
 
-0.3111 is also the indifference point for the farmer with respect to crops.
-That is, if we imagine nature playing strategically, a drought would occur with probability 0.31111 because that would make nature indifferent to whether the farmer planted wheat or barley.
-
-when you sample from the expected value, use that to choose a strategy. but if you sample lots, use that to create a mixed streateg
-
-So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
-
-*Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
-
-## Crop Allocation Decisions
-
-These equations are iteratively applied to the entire span of the climate-model output.
-The result is that farmers calculate the optimal land-use strategy that, \emph{to the best of their knowledge and given their preferences}, uses information from the past.
+# Drought risks in the middle Holocene
 
@@
 
-```{r}
-precc <- map(list.files('Data/PRECC', full.names = TRUE)[-12],
+```{r, cache = TRUE}
+precc <- map(list.files('../Data/PRECC', full.names = TRUE)[-12],
              ~brick(.) %>%
@@
   bind_rows()
-precl <- map(list.files('Data/PRECL', full.names = TRUE)[-12],
+precl <- map(list.files('../Data/PRECL', full.names = TRUE)[-12],
              ~brick(.) %>%
@@
 
-sim_past <- left_join(precc, precl) %>%
+sim_past <- left_join(precc, precl, by = 'time') %>%
   mutate(rainfall = (precc + precl) * 2.628e+9,
@@
 
-```{r, fig.height =3, fig.width = 9, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period'}
+```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period'}
 sim_past %>%
@@
   ggplot(aes(year, risk, fill = risk, color = risk)) +
-  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, width = 1.2) +
+  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
   geom_col()  +
@@
 
-Start with a population of agents with the same subjective beliefs about drought risk.
-
+# The Model
 ```{r}
@@
 ```
-
-```{r}
-agents
-```
-
-## Bayesian Updating
-
-```{r}
-prior <- function(m, n) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, n * m, n * (1 - m))
-  tibble('x' = domain, 'y' = val)
-}
-
-tibble(year = seq(5, 50, 5),
-       risk = .5) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = year, color = year), size = 1.2) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-```
-
-Allow the agents to learn the drought risk over time. Also need to calculate the actual payoffs for each agent.
 
@@
 }
+
+# or keep risk parameter constant and reduce the effective sample size, in other words keep the ratio of alpha to beta but decrease the absoulte value
 ```
@@
 }
-
-plot(prior2( 0, 0));plot(prior2( .5, .5));plot(prior2( 1, 1))
 
@@
 
-```{r, echo = TRUE}
+```{r}
 test1 <- tibble(time = 1:100,
@@
 
-```
-
-```{r}
-simulation2 <- accumulate(droughts, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
 ```
@@
 
-```{r}
-load('../Silvanus/R/sysdata.rda')
-life_table
-fertility_shape <- 4.579627
-fertility_scale <- 0.1
-survivor_scale <- 0.1
-```
-
-```{r}
-reproduce <- function(individuals) {
-  individuals %>%
-    left_join(life_table, by = 'age') %>% 
-    mutate(fertility_reduction = pgamma(pmin(1, payoff / 0.5), shape = fertility_shape, scale = fertility_scale),
-           reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
-    filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
-    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta)) %>%
-    select(-c(fertility_rate:reproduced)) %>%
-    bind_rows(individuals, .)
-}
-
-die <- function(individuals) {
-  individuals %>%
-        left_join(life_table, by = 'age') %>% 
-    mutate(survival_reduction = pgamma(pmin(1, payoff / 0.5), shape = survival_shape, scale = survivor_scale),
-           survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
-    filter(survived == TRUE) %>%
-    select(-c(fertility_rate:survived)) %>%
-    mutate(age = age + 1)
-}
-```
-
-Ok so next we need to have the birth and death stuff run with all the prior information in it.
-dont need to propogate the prior exactly, just need to add in a new weak prior.
-at least propogate the risk number, but not the prior?
-```{r}
-reduce(1:600, ~reproduce(.x) %>% die , .init = agents) %>%
-  pull(age) %>% hist
-```
-
-Another question here about initialization ... namely that we are starting with 80 year olds with no beliefs about the system?
-
-```{r}
-simulation3 <- accumulate(sim_past$drought, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
-
-So the problem now is the population dynamics are getting in the way ...
-and that it'd be nice to have something akin to the moran process to ensure fixed population.
-OR recognize the importance of population dynamics and have the full continuous yields from the medland functions and just determine the chunked payoffs from there?
-The problem now is that if we set the food requirement to calculate the food ratio to something like 0.5, then there is no benefit to the increased yields and hence no real tradeoff between yields and risk management.
-perhaps just record the cumulative surplus and have that be something
-```{r}
-ggplot(simulation3, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-ggplot(simulation3, aes(time, payoff)) +
-  geom_point(alpha = .1) +
-  #geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation3 %>%
-  group_by(time) %>%
-  tally %>%
-  plot
-```
-
-so we have to connect the food ratio to the payoffs? maybe have the barley yield be ~ .8 food ratio
-
-```{r}
-.5/.8
-```
-
 \section{Results and Discussion}
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=16cm]{precon.png}
-\caption{Bias-corrected wet season (October-March) precipitation accumulations from the past 10,000 years of TraCE-21k simulation.
-1 point = 1 year, color corresponds to whether that year falls bellow the thresholds for wheat or barley crop failure.}
-\end{figure}
 
@@
 # References
-
-```{r}
-knitr::knit_exit()
-```
-
-```{r}
-moran_birth_death <- function(agents, r = 1) {
-  n <- nrow(agents)
-  birth_index <- sample.int(n, size = r)
-  death_index <- sample.int(n, size = r)
-  newborn <- agents %>%
-    filter(row_number() %in% birth_index) %>%
-    mutate(age = 1)
-
-  agents %>%
-    filter(!row_number() %in% death_index) %>%
-    mutate(age  = age + 1) %>%
-    bind_rows(newborn)
-}
-
-reduce(1:1000, ~moran_birth_death(., r = 4), .init = agents2) %>%
-  pull(age) %>%hist
-```
-
-so in this case drought risk is low enough all the time you never get any diversification
-
-~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-## Satisficing learners
-
-so we're imagining a yield of 1 for barley and 2 for wheat, with equal proportions of each. then in the next time step we'd revise the proportion to take away .25
-```{r}
-reallocate <- function(wheat_prop, risk) {
-  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
-}
-
-accumulate(1:10, ~reallocate(.x), .init = .4) %>% plot
-
-# so this works as long as you're between 0 and 1
-
-```
-
-```{r, eval = FALSE}
-readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
-  rename(site = `Site Location`, taxon = Taxon, age = `ky cal BP`, yield = `GY (t*ha-1)`) %>%
-  select(site, taxon, age, yield) %>%
-  #filter(!is.na(yield)) %>%
-  mutate(yield = if_else(is.na(yield), 0, yield)) %>%
-    ggplot(aes(yield)) +
-  geom_histogram(aes(fill = taxon)) +
-  facet_wrap(~taxon, nrow = 2)
-
-readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
-  rename(site = `Site Location`, taxon = Taxon, age = `ky cal BP`, yield = `GY (t*ha-1)`) %>%
-  select(site, taxon, age, yield) %>%
-  filter(!is.na(yield)) %>%
-  filter(site == 'Akarçay Tepe') %>%
-  mutate(yield = yield * 1000,
-         yield = if_else(taxon == 'Hordeum vulgare', yield * 3000, yield * 3540)) %>%
-group_by(taxon) %>%
-  summarise(yield_mean = mean(yield), yield_sd = sd(yield)) %>%
-  mutate(yield = map2(yield_mean, yield_sd, ~rnorm(100000, .x, .y))) %>%
-  select(taxon, yield) %>%
-  unnest(cols = c(yield)) %>% 
-  mutate(yield = if_else(yield > 0, yield, 0)) %>%
-  ggplot(aes(yield)) +
-  geom_histogram(aes(fill = taxon)) +
-  facet_wrap(~taxon, nrow = 2)
-```
-
-```{r}
-rnorm(10000, 1582521,	780883.5) %>% hist
-rnorm(10000, 3060657,	1461350.0) %>% hist
-```
-
-```{r}
-yield_max_barley <- 2.500
-yield_max_wheat <- 3.500
-kcal_barley <- 1 #3000
-kcal_wheat <- 3540 / 3000
-
-calc_yield <- function(rainfall, crop = 'wheat') {
-  if (crop == 'wheat') {
-    yield <- yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
-  }
-
-  if(crop == 'barley') {
-    yield <- yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51) 
-  }
-
-  return(yield)
-}
-
-tibble(rainfall = (0:200) / 100,
-       barley = calc_yield(rainfall, 'barley'),
-       wheat = calc_yield(rainfall, 'wheat')) %>%
-  gather(crop, yield, barley:wheat) %>%
-  ggplot(aes(rainfall, yield, color = crop)) + 
-  geom_line() +
-  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
-  geom_hline(yintercept = yield_max_barley * kcal_barley) +
-  #geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
-  geom_vline(xintercept = 0.3)
-```
-
-How do we define the "drought year" cutoff? Below 300mm there isn't enough wheat growing in a hectare to feed a family of 5, which seems appropriate.
-```{r}
-test <- tibble(time = 1:100,
-       rain = rnorm(100, mean = 1, sd = 1) %>% pmax(0),
-       yield = calc_yield(rain))
-
-ggplot(test, aes(time, yield)) + geom_point() + geom_hline(yintercept = 1016525)
-
-test %>%
-  mutate(drought = if_else(yield < 5e+6, TRUE, FALSE))%>%
-  filter(drought == TRUE)
-```
-
-Similar to above, but satisficing
-```{r}
-learn_satisficing <- function(agents, drought, aspiration) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
-
-  agents %>%
-     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-            new_prop = round(runif(n()), 2),
-           wheat_prop = if_else(payoff > aspiration, wheat_prop, new_prop)) %>%
-    select(-new_prop)
-}
-```
-
-```{r}
-test3 <- tibble(agent = 1:100) %>%
-  mutate(wheat_prop = round(runif(n()), 2))
-
-simulation3 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-
-simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
-simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
-```

### 2020-02-07 manuscript/manuscript.tex
pull analysis into new manuscript and combine with manuscript text

--- 2020-02-07 manuscript/manuscript.Rmd
@@
----
-title: The dynamics of risk perception in a Mediterranean agroecosystem
-subtitle: 
-titlerunning: The dynamics of risk perception in a Mediterranean agroecosystem
-authorrunning: Gauthier
-thanks: | 
-    Grants or other notes about the article that should go on the front 
-    page should be placed here. General acknowledgments should be placed at the
-    end of the article.
+% !TeX program = pdfLaTeX
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage[utf8]{inputenc}
 
-authors: 
-- name: Nicolas Gauthier 1
-  address: Laboratory of Tree-Ring Research & School of Geography and Development, University of Arizona
-  email: ngauthier@email.arizona.edu
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-keywords:
-- key
-- dictionary
-- word
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+% \journalname{myjournal}
+%
 
-#PACS: 
-#- PAC1
-#- superPAC
+%% load any required packages here
 
-MSC:
-- MSC code 1
-- MSC code 2
+\begin{document}
 
-abstract: |
-  Small-scale agriculturalists in the Mediterranean Basin rely on multiple strategies including diversification, intensification, and storage to maintain a stable food supply in the face of environmental uncertainty.
-Each of these strategies requires farmers to make specific resource allocation decisions in response to environmental risks and is thus sensitive to variability in both the spatiotemporal pattern of risk and the ability of farmers to perceive that pattern.
-In this chapter, I present an agent-based model of a Mediterranean agroecosystem.
-By driving the model with realistic environmental dynamics derived from simulations of mid-Holocene Mediterranean climate, and by allowing the psychology of risk perception to vary among individual farmers, I explore the hidden vulnerabilities of traditional risk-management strategies to periods of rapid climate change.
-I show that even when farmers are able to manage risk “optimally” in light of past experience, unanticipated changes in the spatiotemporal pattern of rainfall can still lead to major food shortfalls.
+\title{The dynamics of risk perception in a Mediterranean agroecosystem \thanks{Grants or other notes about the article that should go on the front page
+should be placed here. General acknowledgments should be placed at the
+end of the article.} }
 
-bibliography: bibliography.bib
-bibstyle: spbasic
-# bibstyle options spbasic(default), spphys, spmpsci
-output: rticles::springer_article
----
+    \titlerunning{Risk perception in a Mediterranean agroecosystem}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-library(raster)
-library(tidyverse)
-```
+\author{  Nicolas Gauthier 1 \and  }
 
-```{r}
-set.seed(1111)
-```
+    \authorrunning{ Gauthier }
 
-# Introduction {#intro}
+\institute{
+        Nicolas Gauthier 1 \at
+     Laboratory of Tree-Ring Research \& School of Geography and Development,
+ University of Arizona \\
+     \email{\href{mailto:ngauthier@email.arizona.edu}{\nolinkurl{ngauthier@email.arizona.edu}}}  %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+    }
 
-Over the past 10,000 years, subsistence farmers in the Mediterranean basin have developed sophisticated strategies to maintain stable food supplies given uncertain rainfall.
-Their suite of strategies include practices like crop diversification, storage, and exchange [@Halstead1989]).
-These risk-management strategies all represent forms of \emph{regulatory feedback} used to manage complex socio-ecological systems.
-Regulatory feedbacks convert a system's high-variance inputs into low-variance outputs by dynamically monitoring the outputs and adjusting internal properties of the system [@Anderies2015].
-Risk-managing strategies require farmers to make specific resource allocation decisions in response to environmental risks, and as regulatory feedbacks they are thus sensitive to variability in both the spatiotemporal patterns of risk and the ability of farmers to perceive and act on those patterns.
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
-Crop diversification is an excellent example of a widespread and effective regulatory feedback strategy (Figure 1).
-Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@helmers2001separating,Anderies2006].
-In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,Marston2011190].
-Wheat generally has higher yields but is sensitive to water deficits, while barley has lower yields and is drought tolerant.
-Planting a mix of wheat and barley, either in the same plot or in a combination of plots, is an effective means of diversifying grain supplies.
-By adjusting the ratio of wheat to barley, farmers can adapt to changing drought risks.
+\maketitle
 
-The secret to the success of earliest farmers on the mediterrenan littoral was in their manipulation of the local ecosystem to manage environmental risks.
-Early farmers constructed a complex agroecosystem to manage risk.
-They needed to maintain stable crop yields in face of uncertain drought risks.
-Yield stability is very important for traditional agriculturlal commmunities not connected to larger bulk food networks (Abbo et al 2010)
-neolithic crop package -- weiss zohary 2011
-cite abbo et al 2010? for more stuff about ladraces
-Wheat/barley systems
-example of system that combines low yield, low wirks crop with high yield high risk crop, an effective context paut et al 2019
-These strategies work when risks are knonw and constant, but may be more susceptible to periods of changing climateic risks
-so understanding how people of the past adapted to change is important to us today
+\begin{abstract}
+Small-scale agriculturalists in the Mediterranean Basin rely on multiple
+strategies including diversification, intensification, and storage to
+maintain a stable food supply in the face of environmental uncertainty.
+Each of these strategies requires farmers to make specific resource
+allocation decisions in response to environmental risks and is thus
+sensitive to variability in both the spatiotemporal pattern of risk and
+the ability of farmers to perceive that pattern. In this chapter, I
+present an agent-based model of a Mediterranean agroecosystem. By
+driving the model with realistic environmental dynamics derived from
+simulations of mid-Holocene Mediterranean climate, and by allowing the
+psychology of risk perception to vary among individual farmers, I
+explore the hidden vulnerabilities of traditional risk-management
+strategies to periods of rapid climate change. I show that even when
+farmers are able to manage risk ``optimally'' in light of past
+experience, unanticipated changes in the spatiotemporal pattern of
+rainfall can still lead to major food shortfalls.
+\\
+\keywords{
+        key \and
+        dictionary \and
+        word \and
+    }
 
-Are these crop-diversification strategies vulnerable to the same dynamics as other social-ecological systems with similar regulatory feedbacks mechanisms?
-One way to address this is to model the influence of imperfect monitoring and biased decision making in uncertain environments.
-In this study, I accomplish this by answering two main questions:
+    \subclass{
+                    MSC code 1 \and
+                    MSC code 2 \and
+            }
 
-1. What was the temporal pattern of climate and climate change during the Holocene in the eastern Mediterranean?
-2. How well could crop diversification strategies cope with this pattern, given farmers’ imperfect risk perception and decision-making?
+\end{abstract}
 
-Questions:
-How would the earliest farmers have made decisions (dealt with uncertainty?) about crop diversification?
-ie what were the different algorithms? -- better understand the decision context
-What are the consequences of theses different decisions (i.e. how well did they perform) at different points in the history of the mediterranean
-How were farmers able to detect changes in the environment, rather than unconsciously adapt to them?
-what are the different consequences we'd expect for each
-THIS is the most important question because with it I can frame the results around and individual's perception of droughts in the neolithic, and the populaiton-level distribution of drought perceptions at given times
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
-I couple a climate model with high temporal resolution to a simple game theoretic model of agricultural decision making under uncertainty, in order to examine the performance of optimal and suboptimal wheat-barley diversification practices.
-I model the year-to-year crop diversification strategies made by farmers as an iterated game of fictitious play against nature [@GOULD1963a].
-Farmers are boundedly rational, in that they seek to maximize their objective functions given the perceived probabilities of different actions by nature, but work with flawed mental models of climate-related risks due to imperfect recall of past events.
-Furthermore, by using a climate model to represent the actions of nature, I can more precisely capture the characteristic autocorrelated patterns of rainfall variability, rather than simply drawing rainfall values from a static distribution.
-This framework allows me to address the sensitivity of diversification strategies to changes and climatic variability and imperfect human perceptions of those changes.
+\hypertarget{intro}{%
+\section{Introduction}\label{intro}}
 
-# Decision-making in a game against nature {#sec:1}
+Over the past 10,000 years, subsistence farmers in the Mediterranean
+basin have developed sophisticated strategies to maintain stable food
+supplies given uncertain rainfall. Their suite of strategies include
+practices like crop diversification, storage, and exchange (Halstead and
+O'Shea 1989)). These risk-management strategies all represent forms of
+\emph{regulatory feedback} used to manage complex socio-ecological
+systems. Regulatory feedbacks convert a system's high-variance inputs
+into low-variance outputs by dynamically monitoring the outputs and
+adjusting internal properties of the system ({\textbf{???}}).
+Risk-managing strategies require farmers to make specific resource
+allocation decisions in response to environmental risks, and as
+regulatory feedbacks they are thus sensitive to variability in both the
+spatiotemporal patterns of risk and the ability of farmers to perceive
+and act on those patterns.
 
-The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a game against nature.
-We can represent this simplified decision context as a "game" in a game theorietic context.
-cite luce raiffa 1989 for game against nature, and milnor 1952, Agrawal1968, cassidy1971, gould 1963
-The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertianty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years.
-payoff matrix goes here
-in reality continuous, but cite behavioral econ for people often intuitively solve an easier problem when faced with a complex real-world situation
-The exact values in the payoff matrix here are less important than the relative payoffs in each quadrant.
-We can think of this game as being the culturally-inherited object. our question is how do individuals make decisions here?
-so in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio -- both individual and social learning influence perceptions of eindividual vs social learning tucker 2007
+Crop diversification is an excellent example of a widespread and
+effective regulatory feedback strategy (Figure 1). Relying on a mix of
+food types with different climatic tolerances is an efficient way to
+maintain a robust food supply ({\textbf{???}},Anderies2006). In the
+Mediterranean, land-use strategies involving a diversified portfolio of
+wheat and barley have been employed by even the earliest sedentary
+farmers, and continue to be used to this day (Gould
+1963,Slafer1999,Abbo2009a,Marston2011190). Wheat generally has higher
+yields but is sensitive to water deficits, while barley has lower yields
+and is drought tolerant. Planting a mix of wheat and barley, either in
+the same plot or in a combination of plots, is an effective means of
+diversifying grain supplies. By adjusting the ratio of wheat to barley,
+farmers can adapt to changing drought risks.
 
-Estimates of yield volume (t/ha) for each crop type in wet and dry years were derived from isotopic studies of ancient wheat and barley samples from archaeological sites in the Mediterranean [@Slafer1999] (Table 1).
+The secret to the success of earliest farmers on the mediterrenan
+littoral was in their manipulation of the local ecosystem to manage
+environmental risks. Early farmers constructed a complex agroecosystem
+to manage risk. They needed to maintain stable crop yields in face of
+uncertain drought risks. Yield stability is very important for
+traditional agriculturlal commmunities not connected to larger bulk food
+networks (Abbo et al 2010) neolithic crop package -- weiss zohary 2011
+cite abbo et al 2010? for more stuff about ladraces Wheat/barley systems
+example of system that combines low yield, low wirks crop with high
+yield high risk crop, an effective context paut et al 2019 These
+strategies work when risks are knonw and constant, but may be more
+susceptible to periods of changing climateic risks so understanding how
+people of the past adapted to change is important to us today
+
+Are these crop-diversification strategies vulnerable to the same
+dynamics as other social-ecological systems with similar regulatory
+feedbacks mechanisms? One way to address this is to model the influence
+of imperfect monitoring and biased decision making in uncertain
+environments. In this study, I accomplish this by answering two main
+questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  What was the temporal pattern of climate and climate change during the
+  Holocene in the eastern Mediterranean?
+\item
+  How well could crop diversification strategies cope with this pattern,
+  given farmers' imperfect risk perception and decision-making?
+\end{enumerate}
+
+Questions: How would the earliest farmers have made decisions (dealt
+with uncertainty?) about crop diversification? ie what were the
+different algorithms? -- better understand the decision context What are
+the consequences of theses different decisions (i.e.~how well did they
+perform) at different points in the history of the mediterranean How
+were farmers able to detect changes in the environment, rather than
+unconsciously adapt to them? what are the different consequences we'd
+expect for each THIS is the most important question because with it I
+can frame the results around and individual's perception of droughts in
+the neolithic, and the populaiton-level distribution of drought
+perceptions at given times
+
+I couple a climate model with high temporal resolution to a simple game
+theoretic model of agricultural decision making under uncertainty, in
+order to examine the performance of optimal and suboptimal wheat-barley
+diversification practices. I model the year-to-year crop diversification
+strategies made by farmers as an iterated game of fictitious play
+against nature (Gould 1963). Farmers are boundedly rational, in that
+they seek to maximize their objective functions given the perceived
+probabilities of different actions by nature, but work with flawed
+mental models of climate-related risks due to imperfect recall of past
+events. Furthermore, by using a climate model to represent the actions
+of nature, I can more precisely capture the characteristic
+autocorrelated patterns of rainfall variability, rather than simply
+drawing rainfall values from a static distribution. This framework
+allows me to address the sensitivity of diversification strategies to
+changes and climatic variability and imperfect human perceptions of
+those changes.
+
+\hypertarget{sec:1}{%
+\section{Decision-making in a game against nature}\label{sec:1}}
+
+The basic decision-making problem facing a farmer seeking to diversify
+their crops can be thought of as a game against nature. We can represent
+this simplified decision context as a ``game'' in a game theorietic
+context. cite luce raiffa 1989 for game against nature, and milnor 1952,
+Agrawal1968, cassidy1971, gould 1963 The ``game'' in this context is the
+farmer's decision of which crops to plant and in what proportions, given
+uncertianty in the future ``state-of-nature'', and nature varies between
+several possible states such as dry and wet years. payoff matrix goes
+here in reality continuous, but cite behavioral econ for people often
+intuitively solve an easier problem when faced with a complex real-world
+situation The exact values in the payoff matrix here are less important
+than the relative payoffs in each quadrant. We can think of this game as
+being the culturally-inherited object. our question is how do
+individuals make decisions here? so in this case the cultural norm would
+be to mix wheat and barley, but the learning is in the exact ratio --
+both individual and social learning influence perceptions of eindividual
+vs social learning tucker 2007
+
+Estimates of yield volume (t/ha) for each crop type in wet and dry years
+were derived from isotopic studies of ancient wheat and barley samples
+from archaeological sites in the Mediterranean (Araus, Slafer, and
+Romagosa 1999) (Table 1).
 
@@
 
-```{r}
-wheat_dry <- 0
-wheat_normal <- 1.6
-
-barley_dry <- 0.93
-barley_normal <- 1.18
-```
-
 How should an early farmer make a decision in light ot this uncertainty?
-The basic logic of crop diversitication is Portfolio of crops ala modern portfolio theory: Blank2001
-Sometimes best to think of nature as a sentient opponent out to get you, and play strategically based on that assumption. Gould 1963,  Beckenkamp 2008
-quick overview of criteria without risk, then say that a better option is to estimate the probabilities of the different state of nature
-Agrawal1968 talks about the strategies in the context of agriculture being a game against nature
-here I can slip in the cool thing about acting like the weather is out to get you being a good idea in states of complete uncertainty
-transition -- the problem becomes easier if you have at least someI idea of nature's moves , because then you cna work to get the highest yeilds given the risk of drought
-Decision strategies as choosing different points -- plot of decisions
+The basic logic of crop diversitication is Portfolio of crops ala modern
+portfolio theory: Blank2001 Sometimes best to think of nature as a
+sentient opponent out to get you, and play strategically based on that
+assumption. Gould 1963, Beckenkamp 2008 quick overview of criteria
+without risk, then say that a better option is to estimate the
+probabilities of the different state of nature Agrawal1968 talks about
+the strategies in the context of agriculture being a game against nature
+here I can slip in the cool thing about acting like the weather is out
+to get you being a good idea in states of complete uncertainty
+transition -- the problem becomes easier if you have at least someI idea
+of nature's moves , because then you cna work to get the highest yeilds
+given the risk of drought Decision strategies as choosing different
+points -- plot of decisions
 
@@
 
-```{r}
-expected_yield_wheat <- function(risk) {
-  risk * wheat_dry + (1 - risk) * wheat_normal
-}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Expected wheat and barley yields under increasing drought risk
+with the point of indifference highlighted.}
+\end{figure}
 
-expected_yield_barley <- function(risk) {
-  risk * barley_dry + (1 - risk) * barley_normal
-}
+If an individual can learn about their environment, such as the risk of
+drought in any given year, then they can behave more rationally by
+trying to maximize their subjective expected utility If the probability
+distributin of nature's moves is known, the farmer can choose the crop
+mix that simply maximizes the expected crop yields given the empirical
+frequency distirubtion of nature's moves. In the language of game
+theory, this strategy is knoan as a ``game of ficticious play'' against
+nature. This is an effective risk-managing strategy, but like all
+similar strategies that adapt to a specific pattern of variability, it
+is vulnerable to changes in the pattern of variability (Janssen eta al
+2007) This strategy works well when the environment is stationary --
+nature plays from a fixed probability distirubtion, but is vulnerable to
+environmental nonstationarity. That is, when the mean or higher order
+moments of the rainfall distribution shift, playing a game of fictitious
+play can backfire because you're too stuck in your ways while the world
+changes around you. Thus it is important to understand the dynamics of
+risk perception -- how subjective risks, and thus subjective expected
+utility -- rise and fall in uncertain environments.
 
-expected_yield <- function(risk) {
-  expected_yield_wheat(risk) + expected_yield_barley(risk)
-}
-```
+\hypertarget{sec:2}{%
+\section{The dynamics of risk perception}\label{sec:2}}
 
-```{r, echo = FALSE, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
-indiff <- tibble(risk = (barley_normal - wheat_normal) / (wheat_dry - wheat_normal - barley_dry + barley_normal),
-                 yield = risk * wheat_dry + (1 - risk) * wheat_normal)
-indiff_label <- paste0('(', round(indiff$risk, 2), ', ', round(indiff$yield, 2), ')')
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
 
-tibble(risk = seq(0, 1, .1),
-      Barley = expected_yield_barley(risk),
-      Wheat = expected_yield_wheat(risk)) %>%
-  gather(type, yield, Barley:Wheat) %>%
-  ggplot(aes(risk, yield)) +
-  geom_hline(yintercept = indiff$yield, linetype = 2)+
-  geom_vline(xintercept = indiff$risk, linetype = 2) +
-  geom_line(aes(color = type), size = 1.2) +
-  geom_point(data = indiff, color = 'black', size = 3) +
-  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
-  theme_classic() +
-  scale_color_discrete(name = 'Crop') +
-  labs(x = 'Drought probability', y = 'Expected yield')
-```
+\hypertarget{drought-risks-in-the-middle-holocene}{%
+\section{Drought risks in the middle
+Holocene}\label{drought-risks-in-the-middle-holocene}}
 
-If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility
-If the probability distributin of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
-In the language of game theory, this strategy is knoan as a "game of ficticious play" against nature.
-This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
-This strategy works well when the environment is stationary -- nature plays from a fixed probability distirubtion, but is vulnerable to environmental nonstationarity.
-That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because you're too stuck in your ways while the world changes around you.
-Thus it is important to understand the dynamics of risk perception -- how subjective risks, and thus subjective expected utility -- rise and fall in uncertain environments.
+\hypertarget{climate-risks-and-risk-perception}{%
+\section{Climate Risks and Risk
+Perception}\label{climate-risks-and-risk-perception}}
 
-# The dynamics of risk perception {#sec:2}
+Using the bias-corrected climate model output, I divided each model year
+into dry years and wet years. A dry year was any year where less than
+300mm of rain fell during the wet season (October-March), the threshold
+below which wheat crops will generally fail ({\textbf{???}}), and a
+wet/normal year was defined as any year above this threshold.
 
-```{r}
-prior <- function(m, n) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, n * m, n * (1 - m))
-  tibble('x' = domain, 'y' = val)
-}
-
-tibble(year = seq(5, 50, 5),
-       risk = .5) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = year, color = year), size = 1.2) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-
-tibble(year = 50,
-       risk = seq(.1,.9, .1)) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = risk, color = risk), size = 1.2) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-```
-
-# Drought risks in the middle Holocene
-
-# Climate Risks and Risk Perception
-
-Using the bias-corrected climate model output, I divided each model year into dry years and wet years.
-A dry year was any year where less than 300mm of rain fell during the wet season (October-March), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
-
-Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years
+Given the modeled patterns of wet and dry years, drought risk for any
+particular year was defined as the proportion of the previous 50 years
+that were dry years
 
@@
     P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
-\end{equation} 
-where $t$ is the current time step.
-The 50 year time span was selected to approximate the accumulated observational knowledge of an individual farmer and their immediate household.
-As a result, this approach does not allow for accumulated social learning, although it could easily be extended to do so in future studies.
+\end{equation} where \(t\) is the current time step. The 50 year time
+span was selected to approximate the accumulated observational knowledge
+of an individual farmer and their immediate household. As a result, this
+approach does not allow for accumulated social learning, although it
+could easily be extended to do so in future studies.
 
-## Climate Modeling and Bias Correction
+\hypertarget{climate-modeling-and-bias-correction}{%
+\subsection{Climate Modeling and Bias
+Correction}\label{climate-modeling-and-bias-correction}}
 
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
-TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
-It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
-orbit, greenhouse gasses, glacial meltwater flux).
-The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
+Estimates of drought risk in the eastern Mediterranean during the the
+past 10,000 years were derived from outputs of the TraCE-21k
+paleoclimate simulation available on the National Center for Atmospheric
+Research's Earth System Grid repository
+(\url{https://www.earthsystemgrid.org}). TraCE-21K is a state-of-the-art
+simulation that uses a coupled atmosphere-ocean general circulation
+model (GCM) to recreate the transient response of the global climate
+system to orbital parameters and greenhouse gasses over the past 22,000
+years from the Last Glacial Maximum to the present (Figure 2) (He 2011).
+It generates physically consistent spatiotemporal climate dynamics,
+driven by current best estimates of climate forcings (e.g.~orbit,
+greenhouse gasses, glacial meltwater flux). The model simulates these
+dynamics on a six hourly timescale, and model outputs are archived at a
+monthly resolution.
 
-TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
-This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
+TraCE-21k simulation outputs for the past 10,000 years were
+bias-corrected using the CDF-t method and observed precipitation from a
+weather station in near the town of Salihli in western Turkey. This
+location was selected due to its long (\textasciitilde{}70yr)
+observational record, the representativeness of western Anatolian
+climate for the greater eastern Mediterranean, and its proximity to
+archaeological sites with comparative evidence of crop diversification
+strategies.
 
@@
 
-```{r}
-konya <- matrix(c(32.5, 37.9), nrow = 1)
-#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
-```
+\begin{verbatim}
+## Loading required namespace: ncdf4
+\end{verbatim}
 
-```{r, cache = TRUE}
-precc <- map(list.files('../Data/PRECC', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precc) ) %>%
-  bind_rows()
-precl <- map(list.files('../Data/PRECL', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precl) ) %>%
-  bind_rows()
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{Annual risk of wheat crop failure due to drought, averaged by
+fifty-year period}
+\end{figure}
 
-sim_past <- left_join(precc, precl, by = 'time') %>%
-  mutate(rainfall = (precc + precl) * 2.628e+9,
-         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 5200)) %>%
-  select(-time, -precl, -precc) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-   filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  add_tally() %>%
-  filter(n == 8) %>%
-  summarise(precip = sum(rainfall)) %>%
-  mutate(drought = precip < 300)
-```
+\hypertarget{the-model}{%
+\section{The Model}\label{the-model}}
 
-```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period'}
-sim_past %>%
-  filter(between(water_year, -9500, -5500)) %>%
-  mutate(time_step = ceiling(water_year / 50),
-         year = time_step * 50) %>%
-  group_by(year) %>%
-  summarise(risk = sum(drought) / 50) %>%
-  ggplot(aes(year, risk, fill = risk, color = risk)) +
-  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
-  geom_col()  +
-  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
-  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
-  theme_minimal() +
-  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
-```
+Run an experiment. Simulate a time series with varying drought
+frequency. This is a stepwise nonstationary process
 
-# The Model
-```{r}
-agents <- tibble(id = 1:150, 
-                 alpha = 1.5,
-                 beta = 1.5,
-                 risk = alpha / (alpha + beta),
-                 age = sample.int(80, size = 150, replace = TRUE),
-                 crop = 'barley')
-```
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
 
-```{r}
-learn1 <- function(agents, drought) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
+Redo these plots so its a single person, one who was born in a drought
+period then experiences an ameliorarion and another who experiences a
+shift to drought conditions later in life? basically this section neds
+to work more with individuals and explore that stuff
 
-  agents %>%
-    # farmer's turn to decide what to plant
-      mutate(yield_exp = expected_yield(risk),
-         barley_exp = expected_yield_barley(risk),
-         wheat_exp = expected_yield_wheat(risk),
-         expected_yield = barley_exp + wheat_exp,
-         crop = case_when(barley_exp > wheat_exp ~ 'barley',
-                              barley_exp < wheat_exp ~ 'wheat',
-                              barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
-         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
-      select(-prediction) %>%
-  # nature's turn to decide realized yields
-    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-           alpha = drought + alpha,
-           beta = beta + 1 - drought,
-          # mu = alpha / (alpha + beta),
-           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
-}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
 
-learn2 <- function(agents, drought) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
 
-  agents %>%
-    # farmer's turn to decide what to plant
-      mutate(yield_exp = expected_yield(risk),
-         barley_exp = expected_yield_barley(risk),
-         wheat_exp = expected_yield_wheat(risk),
-         expected_yield = barley_exp + wheat_exp,
-         crop = case_when(barley_exp > wheat_exp ~ 'barley',
-                              barley_exp < wheat_exp ~ 'wheat',
-                              barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
-         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
-      select(-prediction) %>%
-  # nature's turn to decide realized yields
-    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-           alpha = if_else(payoff < .5, 1, drought + alpha),
-           beta = if_else(payoff < .5, 1, beta + 1 - drought),
-          # mu = alpha / (alpha + beta),
-           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
-}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}
 
-# or keep risk parameter constant and reduce the effective sample size, in other words keep the ratio of alpha to beta but decrease the absoulte value
-```
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-17-1.pdf}
 
-Run an experiment. Simulate a time series with varying drought frequency. This is a stepwise nonstationary process
-```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
-```
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
 
-```{r}
-prior2 <- function(a, b) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, a, b)
-  tibble('x' = domain, 'y' = val)
-}
+These two plots begin to look more alike the more you sample, which
+makes sense because its an ergodic system.
 
-accumulate(droughts, 
-           ~mutate(.x, alpha = .y + alpha, beta = beta + 1 - .y), 
-           .init = tibble(alpha = 1.5, beta = 1.5)) %>%
-  bind_rows(.id = 'time') %>%
-    mutate(time = as.numeric(time), belief = map2(alpha, beta, ~prior2(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(color = time, group = time), size = .8) +
-    geom_vline(xintercept = c(0.1, 0.5, 0.2), linetype = 2, color = 'red', size = .8) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-```
+Note in these plots above and below we're summing all the wheat and
+barley in the population, compared to below which shows the wheat
+proportions of individuals
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-2.pdf}
 
-```{r}
-simulation1 <- accumulate(droughts, ~learn1(.x, .y), .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-2.pdf}
 
-simulation2 <- accumulate(droughts, ~learn2(.x, .y), .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-2.pdf}
 
-Redo these plots so its a single person, one who was born in a drought period then experiences an ameliorarion and another who experiences a shift to drought conditions later in life?
-basically this section neds to work more with individuals and explore that stuff
+\begin{verbatim}
+## Picking joint bandwidth of 0.0146
+\end{verbatim}
 
-```{r}
-test1 <- tibble(time = 1:100,
-       risk = c(rep(0.1, 25), rep(0.5, 25), rep(0.2, 50)))
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}
 
-ggplot(simulation1, aes(time, risk)) +
-  geom_line(data = test1) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
+\begin{verbatim}
+## Picking joint bandwidth of 0.0204
+\end{verbatim}
 
-```{r}
-ggplot(simulation2, aes(time, risk)) +
-  geom_line(data = test1) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-2.pdf}
 
-```
+make this a function to calculate payoffs for all wheat or all barley
+strategies as a comparison
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
 
-```{r}
-ggplot(simulation2, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
-```
-
-```{r}
-simulation2 %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
-```
-
-These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
-
-Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals
-```{r}
-ggplot(simulation1, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
-
-ggplot(simulation2, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-simulation2 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0, na.rm = TRUE)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
-  group_by(time) %>%
-  summarise(payoff = mean(payoff)) %>%
-ggplot(aes(time,payoff)) +
-  geom_point(aes(color = time)) +
-  theme_classic()
-
-simulation2 %>%
-  group_by(time) %>%
-  summarise(payoff = mean(payoff, na.rm = TRUE)) %>%
-ggplot(aes(time,payoff)) +
-  geom_point(aes(color = time)) +
-  theme_classic()
-```
-
-```{r}
-library(ggridges)
-simulation1 %>%
-  mutate(time_step = 5 * round(time/5),
-         famine = payoff == 0) %>%
-  ggplot(aes(wheat_prop, as.factor(time_step))) +
-  geom_density_ridges(rel_min_height = 0.005) +
-  theme_ridges()
-
-simulation2 %>%
-  mutate(time_step = 5 * round(time/5),
-         famine = payoff == 0) %>%
-  ggplot(aes(wheat_prop, as.factor(time_step))) +
-  geom_density_ridges(rel_min_height = 0.005) +
-  theme_ridges()
-```
-
-make this a function to calculate payoffs for all wheat or all barley strategies as a comparison
-```{r}
-tibble(drought = droughts,
-       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
-       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
-       equal = (barley + wheat) / 2) %>%
-  mutate(time = 1:n()) %>%
-  gather(crop, yield, barley:equal) %>%
-  ggplot(aes(time, yield, group = crop, color = crop)) +
-  geom_point()
-```
-
-```{r}
-tibble(drought = droughts,
-       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
-       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
-       equal = (barley + wheat) / 2) %>%
-  mutate(time = 1:n()) %>%
-  gather(crop, yield, barley:equal) %>%
-  mutate(famine = yield == 0) %>%
-  ggplot(aes(time, famine, group = crop, color = crop)) +
-  geom_line()
-```
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-24-1.pdf}
 
@@
 
-The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure 5).
-The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
+The bias-corrected wet season precipitation outputs from TraCE-21k show
+a relatively stationary distribution over the past 10,000 years, with
+slightly drier conditions during the early-mid Holocene transition at
+c.a. 8ka BP (Figure 5). The utility of a diversified wheat-barley crop
+is immediately apparent; rainfall often fails to meet the 300mm
+threshold for a productive wheat crop during the entire 10,000 year
+span, but drops below the threshold for barley only a handful of years.
 
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
-Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
+The risk of crop failure due to drought in any given year varies between
+10\% and 20\%, punctuated by one to four century-long events where risk
+drops as low as 6\% and or exceeds 25\% (Figure 6). Centuries of high
+drought risk cluster around known periods of regional and global climate
+deterioration, such as at 8.6ka BP and 4.2ka BP.
 
-Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
-$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
-This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
-Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
-Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
+Given the average drought risks estimated from TraCE-21k and assuming
+perfect monitoring and recall of recent droughts (i.e. \(\lambda = 0\)),
+farmers intent on minimizing the risk of crop failure would be expected
+to plant about 65\% barley to 35\% wheat on average. This proportion is
+within the range estimated from macrobotanical remains at a nearby
+archaeological site ({\textbf{???}}). Yield maximizing farmers instead
+play a pure strategy profile, planting wheat if the perceived drought
+risk is less than 0.68 and barley if it is more. Because the
+GCM-simulated drought risk never rises above this threshold, a
+yield-maximizing farmer will always plant a wheat monocrop.
 
-Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
-Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
-Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
+Examining the long-term dynamics of crop yields among all
+decision-making strategies and psychological profiles reveals
+variability in yields due to different decision making preferences is
+far greater variability due to differences in risk perception (Figure
+7). Risk-minimizing strategies sacrifice productivity for
+predictability, consistently lowering the mean and variance of crop
+yields over time. Risk-minimizing strategies are more sensitive to risk
+perception than yield-maximizing strategies; varying the memory decay
+rate parameter \(\lambda\) had no effect on the optimal crop allocation
+for yield maximizers.
 
-During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
-Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
-This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
-This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
+During periods of climatic stability, allowing past experiences to
+influence decision making helps farmers minimize the impacts of
+\emph{predictable} drought. But past experiences are less informative
+during periods of rapid climate change, and even farmers who manage risk
+``optimally'' experience major food shortfalls. Climatically-induced
+variability in food supplies consistently surpasses that from
+differences in risk-perception psychology, but not that from different
+risk aversion preferences. This finding suggests that the efficacy of
+risk managing strategies that rely on regulatory feedbacks is limited to
+periods of relative climatic stability. This is consistent with the
+observation from the Roman Period Mediterranean that while minor food
+crises were common in the ancient world, extreme famines were rare
+(Garnsey 1989).
 
-Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
-To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
-Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
-Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
+Statistically downscaling palaeoclimate simulations and coupling them to
+game theoretic models of decision making under uncertainty is a simple
+way to better understand the patterning of climate changes and the
+regulatory feedbacks farmers use to adapt to them. To further explore
+these patterns, future work should incorporate additional
+risk-minimizing strategies such as storage and exchange. Food storage
+and exchange can also be thought of as forms of diversification, the
+former in the time and the latter in space ({\textbf{???}}). Food
+exchange would be of particular interest from a game theoretic
+perspective, as incomplete information with respect to an exchange
+partner's crop diversification and storage practices can lead to
+situations of moral hazard.
 
-Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
+Think about the importance of the population's age structure, becuase
+that's going to impact the collective memory of drought.
 
-# References
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Slafer1999}{}%
+Araus, J. L., G. A. Slafer, and I. Romagosa. 1999. ``Durum wheat and
+barley yields in antiquity estimated from 13C discrimination of
+archaeological grains: a case study from the Western Mediterranean
+Basin.'' \emph{Australian Journal of Plant Physiology} 26 (4): 345.
+\url{https://doi.org/10.1071/PP98141}.
+
+\leavevmode\hypertarget{ref-Garnsey1989}{}%
+Garnsey, Peter. 1989. \emph{Famine and Food Supply in the Graeco-Roman
+World: Responses to Risk and Crisis}. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-GOULD1963a}{}%
+Gould, Peter R. 1963. ``Man Against His Environment: A Game Theoretic
+Framework.'' \emph{Annals of the Association of American Geographers} 53
+(3): 290--97. \url{https://doi.org/10.1111/j.1467-8306.1963.tb00450.x}.
+
+\leavevmode\hypertarget{ref-Halstead1989}{}%
+Halstead, Paul, and John O'Shea. 1989. ``Introduction: cultural
+responses to risk and uncertainty.'' In \emph{Bad Year Economics:
+Cultural Responses to Risk and Uncertainty}, edited by Paul Halstead and
+John O'Shea, 1--7. Cambridge: Cambridge University Press.
+\url{https://doi.org/10.2307/281441}.
+
+\leavevmode\hypertarget{ref-He2011}{}%
+He, Feng. 2011. ``Simulating Transient Climate Evolution of the Last
+Deglatiation with CCSM3.'' PhD thesis, Madison.
+
+\bibliographystyle{spbasic}
+\bibliography{bibliography.bib}
+
+\end{document}

### 2020-02-07 risk-percption.Rmd
changepoint

@@
 
+```{r}
+devtools::install_github('nfultz/binomialbcp')
+library(binomialbcp)
+data("kansas")
+kansas
+
+t12 <- sim_past %>%
+  filter(between(water_year, -9500, -5500)) %>%
+  mutate(time_step = ceiling(water_year / 50),
+         year = time_step * 50) %>%
+  group_by(year) %>%
+  summarise(risk = sum(drought)) %>% pull(risk)
+binomialbcp(t12, rep(50, length(t12))) %>% plot
+
+1 - c(71,18,18)/225
+```
+
 Start with a population of agents with the same subjective beliefs about drought risk.
@@
   geom_line(aes(group = year, color = year), size = 1.2) +
+  scale_color_viridis_c() +
+  theme_bw() +
+  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+
+tibble(year = 50,
+       risk = seq(.1,.9, .1)) %>%
+  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = risk, color = risk), size = 1.2) +
   scale_color_viridis_c() +

### 2020-02-11 manuscript/manuscript.Rmd
reorder trace21k section, add risk dynamics

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
 library(raster)
+library(ncdf4)
 library(tidyverse)
-```
-
-```{r}
-set.seed(1111)
+set.seed(1000)
 ```
@@
 
-1. What was the temporal pattern of climate and climate change during the Holocene in the eastern Mediterranean?
-2. How well could crop diversification strategies cope with this pattern, given farmers’ imperfect risk perception and decision-making?
-
-Questions:
-How would the earliest farmers have made decisions (dealt with uncertainty?) about crop diversification?
-ie what were the different algorithms? -- better understand the decision context
-What are the consequences of theses different decisions (i.e. how well did they perform) at different points in the history of the mediterranean
-How were farmers able to detect changes in the environment, rather than unconsciously adapt to them?
-what are the different consequences we'd expect for each
-THIS is the most important question because with it I can frame the results around and individual's perception of droughts in the neolithic, and the populaiton-level distribution of drought perceptions at given times
+1. How likely are droughts to occur each year in the eastern Mediterreanean, and how did these risks change over the Holocene?
+2. How might Neolithic farmers have percieved these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
 
@@
 The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a game against nature.
-We can represent this simplified decision context as a "game" in a game theorietic context.
-cite luce raiffa 1989 for game against nature, and milnor 1952, Agrawal1968, cassidy1971, gould 1963
-The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertianty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years.
-payoff matrix goes here
-in reality continuous, but cite behavioral econ for people often intuitively solve an easier problem when faced with a complex real-world situation
+We can represent this simplified decision context as a "game" in a game theorietic context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould 1963).
+The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertianty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years (Table 1).
+In reality continuous, but people often intuitively solve an easier problem when faced with a complex real-world situation (cite behavioral econ or simon).
 The exact values in the payoff matrix here are less important than the relative payoffs in each quadrant.
-We can think of this game as being the culturally-inherited object. our question is how do individuals make decisions here?
-so in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio -- both individual and social learning influence perceptions of eindividual vs social learning tucker 2007
-
-Estimates of yield volume (t/ha) for each crop type in wet and dry years were derived from isotopic studies of ancient wheat and barley samples from archaeological sites in the Mediterranean [@Slafer1999] (Table 1).
+We can think of this game as being the culturally-inherited object.
+Our question is how do individuals make decisions here?
+So in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio.
 
@@
 \centering
-\caption{Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].}
+\caption{Estimates of yield volume (t/ha) for ancient wheat and barley varieties derived from [@Slafer1999].}
 \begin{tabular}{|l|l|l|}
@@
 here I can slip in the cool thing about acting like the weather is out to get you being a good idea in states of complete uncertainty
-transition -- the problem becomes easier if you have at least someI idea of nature's moves , because then you cna work to get the highest yeilds given the risk of drought
+transition -- the problem becomes easier if you have at least some idea of nature's moves, because then you cna work to get the highest yeilds given the risk of drought.
 Decision strategies as choosing different points -- plot of decisions
@@
 
-```{r}
+```{r, fig.height = 3, fig.width = 5, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
 expected_yield_wheat <- function(risk) {
@@
 }
-```
 
-```{r, echo = FALSE, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
 indiff <- tibble(risk = (barley_normal - wheat_normal) / (wheat_dry - wheat_normal - barley_dry + barley_normal),
@@
   ggplot(aes(risk, yield)) +
-  geom_hline(yintercept = indiff$yield, linetype = 2)+
-  geom_vline(xintercept = indiff$risk, linetype = 2) +
+  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .5)+
+  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .5) +
   geom_line(aes(color = type), size = 1.2) +
@@
   theme_classic() +
-  scale_color_discrete(name = 'Crop') +
-  labs(x = 'Drought probability', y = 'Expected yield')
+  scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield')
 ```
 
-If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility
+If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility.
 If the probability distributin of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
-In the language of game theory, this strategy is knoan as a "game of ficticious play" against nature.
+In the language of game theory, this strategy is known as a "game of ficticious play" against nature.
 This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
@@
 
-# The dynamics of risk perception {#sec:2}
+# Drought risks in eastern Mediterranean {#sec:2}
 
-```{r}
-prior <- function(m, n) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, n * m, n * (1 - m))
-  tibble('x' = domain, 'y' = val)
-}
-
-tibble(year = seq(5, 50, 5),
-       risk = .5) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = year, color = year), size = 1.2) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-
-tibble(year = 50,
-       risk = seq(.1,.9, .1)) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = risk, color = risk), size = 1.2) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-```
-
-# Drought risks in the middle Holocene
-
-# Climate Risks and Risk Perception
-
-Using the bias-corrected climate model output, I divided each model year into dry years and wet years.
-A dry year was any year where less than 300mm of rain fell during the wet season (October-March), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
-
-Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years
-
-\begin{equation}
-    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
-\end{equation} 
-where $t$ is the current time step.
-The 50 year time span was selected to approximate the accumulated observational knowledge of an individual farmer and their immediate household.
-As a result, this approach does not allow for accumulated social learning, although it could easily be extended to do so in future studies.
-
-## Climate Modeling and Bias Correction
-
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
-TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
+Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research.
+TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present [@He2011].
 It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
@@
 
-TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
-This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
+TraCE-21k simulation outputs for the past 10,000 years were extracted from the grid cell overlaying the northern Levant and southeastern Turkey.
+This location was selected due to its proximity to archaeological sites with comparative evidence of crop diversification strategies.
+This region includes some of the earliest evidence of crop domestication, including sites like Jericho and Can Hasan.
 
-Now incorporate realistic drought variability from simulations.
+Using the climate model output, I divided each model year into dry years and wet years.
+A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
+Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years.
+Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (mean, variance of dry years) but also the higher order dynamics like serial autocorrelation in the perisistence of wet and dry years.
+This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a challenge to simple risk-managing strategies that assume climatic risks are fixed.
 
@@
   select(-ID) %>%
-  gather(time, precc) ) %>%
+  gather(time, precc)) %>%
   bind_rows()
@@
 
-```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period'}
+```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
+The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
 sim_past %>%
@@
 ```
+
+# The dynamics of risk perception {#sec:3}
+
+Risk Perception
+In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
+An individual's perceived risk entails more than simply the "objective" risk in the external environment [@Tucker2013].
+A person's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
+Likewise, the social responses to risk will refelct the distribution of perceived risks in a population, and by extension the collective memory of past droughts will have consequences for the population's survival and evolution.
+
+Belief-based agents
+We can capture the influences of risk perception computationally by simulating a population of "belief-based agents".
+These agents develop beliefs about the risk of drought in their environment through personal experience. 
+Each agent's perceived drought risks will change their subjective expected crop yields from planting different ratios of wheat to barley.
+
+Bayesian brains
+We can efficiently simulate these population-level risks perceptions simulating a population of agents with Bayesian brains (Kahvalti et al 2019 double check).
+Bayesian learning allows for a minimal representation of the environemnt, so that each agent's beliefs can be encoded in two parameters -- defining the Beta distribution over their prior beliefs -- rather than the record of all observed years.
+Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+Our brains do not record every bit of perceived informaiton in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
+Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
+
+Bayesian learning from bernoulli observations.
+We can treat this problem elegantly using bayesian statistics using online bayesian learning from bernoulli objersbations (Bissiri 2010).
+Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
+bernoulli distribution) with paramter theta reprenting the risk of drought.
+Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1 (Figure XX).
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief.
+What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
+As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
+
+A simple "fixed" bayesian learner thus gets more certain with age.
+For young agents with weak priors, the information of each new year can strongly influence their beliefs.
+For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
+As was the case with the simple game of ficitious play above, this strategy is fine when the environment is stable and risks do not change.
+But when there is volatility in the environment, and risks can change, it becomes important to change your mind.
+Being too inflexible in one's priors can lead being too optomisitc when things have really changed for the worse, and alternately being two conservative when things finally do get better.
+So from a Bayesian learning standpoint, how should agents choose to believe new information.
+
+```{r, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
+prior <- function(m, n) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, n * m, n * (1 - m))
+  tibble('x' = domain, 'y' = val)
+}
+
+tibble(Age = seq(5, 50, 5),
+       risk = .5) %>%
+  mutate(belief = map2(risk, Age, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = Age, color = Age), size = 1.2, alpha = .8) +
+  scale_color_viridis_c(guide = 'legend') +
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
+  theme_classic()
+```
+
+But what if nature isn't stationary?
+That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
+The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
+In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
+Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
+Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
+
+A simple way to capture this behavior computaitonally is with satisficing learning (Crandall and goodrich 2008).
+The basic win-stay lose-shift algorithm works well (Bonawitz et al 2011).
+As long as things are going good and agents don't experience any food shortfalls, they are static bayesina learners, making decisions from a prior belief that is updated by only one observation each year.
+But if the agent does experience to many shortfalls, they will realise their prior beliefs are no longer relevant and they "throw out" their prior and reset to a diffuse prior.
 
@@
 
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+
 # References

### 2020-02-12 manuscript/manuscript.Rmd
schematic update with age

@@
 Bayesian brains
-We can efficiently simulate these population-level risks perceptions simulating a population of agents with Bayesian brains (Kahvalti et al 2019 double check).
+We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 double check).
 Bayesian learning allows for a minimal representation of the environemnt, so that each agent's beliefs can be encoded in two parameters -- defining the Beta distribution over their prior beliefs -- rather than the record of all observed years.
@@
 But when there is volatility in the environment, and risks can change, it becomes important to change your mind.
-Being too inflexible in one's priors can lead being too optomisitc when things have really changed for the worse, and alternately being two conservative when things finally do get better.
+Being too inflexible in one's priors can lead being too optimistic when things have really changed for the worse, and alternately being too conservative when things finally do get better.
 So from a Bayesian learning standpoint, how should agents choose to believe new information.
@@
 tibble(Age = seq(5, 50, 5),
-       risk = .5) %>%
+       risk = 0.5) %>%
   mutate(belief = map2(risk, Age, ~prior(.x, .y))) %>%
@@
   scale_color_viridis_c(guide = 'legend') +
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
+  theme_classic()
+
+tibble(alpha = c(15 * 0.2, 50 * 0.2, 15 * 0.2 + 25 * 0.4, 50 * 0.2 + 25 * 0.4),
+       beta = c(15 * 0.8, 50 * 0.8, 15 * 0.8 + 25 * 0.6, 50 * 0.8 + 25 * 0.6),
+       time = c('before', 'before', 'after', 'after'),
+       age = c('young', 'old', 'young', 'old')) %>%
+  mutate(belief = map2(alpha, beta, ~prior2(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = time, color = time), size = 1.2, alpha = .8) +
+  scale_color_viridis_d(guide = 'legend') +
+  facet_wrap(~age) +
   labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
@@
 
-A simple way to capture this behavior computaitonally is with satisficing learning (Crandall and goodrich 2008).
+A simple way to capture this behavior computationally is with satisficing learning (Crandall and goodrich 2008).
 The basic win-stay lose-shift algorithm works well (Bonawitz et al 2011).
@@
 
-Run an experiment. Simulate a time series with varying drought frequency. This is a stepwise nonstationary process
-```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
-```
+Simulate a time series with varying drought frequency. This is a stepwise nonstationary process. Setup an experiment exploring 6 scenarios.
 
-```{r}
+This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the firstplace.
+The law of small numbers is very important here. The range of each person's experience is so small -- every individual's perceived risk will be a small and noisy sample of the true value.
+Older people are much more certain, even though they are wrong.
+```{r, fig.height=3, fig.width=5}
 prior2 <- function(a, b) {
@@
 
-accumulate(droughts, 
-           ~mutate(.x, alpha = .y + alpha, beta = beta + 1 - .y), 
-           .init = tibble(alpha = 1.5, beta = 1.5)) %>%
-  bind_rows(.id = 'time') %>%
-    mutate(time = as.numeric(time), belief = map2(alpha, beta, ~prior2(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(color = time, group = time), size = .8) +
-    geom_vline(xintercept = c(0.1, 0.5, 0.2), linetype = 2, color = 'red', size = .8) +
-  scale_color_viridis_c() +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
+droughts <-   c(runif(40) < 0.2, runif(20) < 0.4, runif(40) < 0.2)
 ```
@@
 simulation1 <- accumulate(droughts, ~learn1(.x, .y), .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-
-simulation2 <- accumulate(droughts, ~learn2(.x, .y), .init = agents) %>% 
   bind_rows(.id = 'time') %>%
@@
 
-```{r}
+```{r, fig.height = 6, fig.width = 6}
 test1 <- tibble(time = 1:100,
-       risk = c(rep(0.1, 25), rep(0.5, 25), rep(0.2, 50)))
+       exp1 = rep(0.2, 100),
+       exp2 = rep(0.4, 100),
+       exp3 = c(rep(0.2, )))
 
 ggplot(simulation1, aes(time, risk)) +
-  geom_line(data = test1) +
+  #geom_line(data = test1) +
   geom_point(alpha = .1) +
   geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  geom_smooth(color = 'darkorange') +
   theme_classic()

### 2020-02-12 manuscript/manuscript.Rmd
remove win-stay lose-shift
at least temporarily

@@
   theme_classic()
+```
+
+But what if nature isn't stationary?
+That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
+The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
+In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
+Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
+Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
+
+```{r, fig.width = 5, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
+prior2 <- function(a, b) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, a, b)
+  tibble('x' = domain, 'y' = val)
+}
 
@@
   geom_line(aes(group = time, color = time), size = 1.2, alpha = .8) +
-  scale_color_viridis_d(guide = 'legend') +
+  #scale_color_viridis_d(guide = 'legend') +
   facet_wrap(~age) +
@@
 ```
-
-But what if nature isn't stationary?
-That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
-The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
-In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
-Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
-Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
-
-A simple way to capture this behavior computationally is with satisficing learning (Crandall and goodrich 2008).
-The basic win-stay lose-shift algorithm works well (Bonawitz et al 2011).
-As long as things are going good and agents don't experience any food shortfalls, they are static bayesina learners, making decisions from a prior belief that is updated by only one observation each year.
-But if the agent does experience to many shortfalls, they will realise their prior beliefs are no longer relevant and they "throw out" their prior and reset to a diffuse prior.
 
@@
 
-learn2 <- function(agents, drought) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
-
-  agents %>%
-    # farmer's turn to decide what to plant
-      mutate(yield_exp = expected_yield(risk),
-         barley_exp = expected_yield_barley(risk),
-         wheat_exp = expected_yield_wheat(risk),
-         expected_yield = barley_exp + wheat_exp,
-         crop = case_when(barley_exp > wheat_exp ~ 'barley',
-                              barley_exp < wheat_exp ~ 'wheat',
-                              barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
-         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
-      select(-prediction) %>%
-  # nature's turn to decide realized yields
-    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-           alpha = if_else(payoff < .5, 1, drought + alpha),
-           beta = if_else(payoff < .5, 1, beta + 1 - drought),
-          # mu = alpha / (alpha + beta),
-           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
-}
-
 # or keep risk parameter constant and reduce the effective sample size, in other words keep the ratio of alpha to beta but decrease the absoulte value
@@
 Older people are much more certain, even though they are wrong.
-```{r, fig.height=3, fig.width=5}
-prior2 <- function(a, b) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, a, b)
-  tibble('x' = domain, 'y' = val)
-}
+```{r}
 

### 2020-02-12 manuscript/manuscript.Rmd
adjust drought length

@@
 library(tidyverse)
-set.seed(1000)
 ```
@@
 
-# Drought risks in eastern Mediterranean {#sec:2}
+# Holocene drought risks in the eastern Mediterranean {#sec:2}
 
@@
 
-Risk Perception
 In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
@@
 
-Belief-based agents
-We can capture the influences of risk perception computationally by simulating a population of "belief-based agents".
+We can capture the influences of risk perception computationally by simulating a population of "belief-based agents."
 These agents develop beliefs about the risk of drought in their environment through personal experience. 
@@
 
-Bayesian brains
+## Bayesian brains
 We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 double check).
@@
 
-Bayesian learning from bernoulli observations.
+### Bayesian learning from bernoulli observations
 We can treat this problem elegantly using bayesian statistics using online bayesian learning from bernoulli objersbations (Bissiri 2010).
@@
 
-But what if nature isn't stationary?
+### The problem of nonstationarity
 That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
@@
 
-# The Model
+# Risk perception in a Mediterranean agroecosystem
+
+So what are the consequences of these cognitive processes for early farmers in the Mediterranean.
+
 ```{r}
@@
                  risk = alpha / (alpha + beta),
-                 age = sample.int(80, size = 150, replace = TRUE),
+                 #age = sample.int(80, size = 150, replace = TRUE),
                  crop = 'barley')
@@
 ```{r}
-learn1 <- function(agents, drought) {
+learn <- function(agents, drought) {
   if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
@@
 }
-
-# or keep risk parameter constant and reduce the effective sample size, in other words keep the ratio of alpha to beta but decrease the absoulte value
 ```
@@
 Older people are much more certain, even though they are wrong.
+
 ```{r}
+make_droughts <- function() c(runif(25) < 0.2, runif(25) < 0.4, runif(25) < 0.2)
 
-droughts <-   c(runif(40) < 0.2, runif(20) < 0.4, runif(40) < 0.2)
+set.seed(11111)
+droughts <- make_droughts()
 ```
@@
 ```{r}
-simulation1 <- accumulate(droughts, ~learn1(.x, .y), .init = agents) %>% 
+run_sim <- function(droughts) {
+  accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
     bind_rows(.id = 'time') %>%
     mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
+    filter(time > 1)
+}
+
+sim1 <- run_sim(droughts)
 ```
 
-Redo these plots so its a single person, one who was born in a drought period then experiences an ameliorarion and another who experiences a shift to drought conditions later in life?
-basically this section neds to work more with individuals and explore that stuff
+```{r, fig.height = 4, fig.width = 6}
+test1 <- tibble(time = 1:75,
+       risk = c(rep(0.2, 25), rep(0.4, 25), rep(0.2, 25)))
 
-```{r, fig.height = 6, fig.width = 6}
-test1 <- tibble(time = 1:100,
-       exp1 = rep(0.2, 100),
-       exp2 = rep(0.4, 100),
-       exp3 = c(rep(0.2, )))
-
-ggplot(simulation1, aes(time, risk)) +
-  #geom_line(data = test1) +
+ggplot(sim1, aes(time, risk)) +
+  geom_line(data = test1) +
   geom_point(alpha = .1) +
   geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  geom_smooth(color = 'darkorange') +
+ # geom_smooth(color = 'darkorange') +
   theme_classic()
@@
 ```{r}
-ggplot(simulation2, aes(time, risk)) +
-  geom_line(data = test1) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-
-```
-
-```{r}
-ggplot(simulation2, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
+sim1 %>%
   group_by(time, crop) %>%
@@
 ```{r}
-simulation2 %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
+library(furrr)
+plan(multisession)
+
+sim2 <- tibble(rep = 1:500,
+       droughts = replicate(500, make_droughts(), simplify = FALSE)) %>%
+  mutate(results = future_map(droughts, run_sim)) %>%
+  select(-droughts) %>%
+  unnest(results)
 ```
@@
 
-Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals
+Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals.
+
 ```{r}
-ggplot(simulation1, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
+sim2 %>%
+  group_by(rep, time) %>%
+  summarise(test = mean(payoff))%>%
+  summarise(test = mean(test)) %>%
+  #summarise(famine = sum(payoff == 0) / 150) %>%
+  #summarise(famine = mean(famine)) %>%
+  ggplot(aes(test)) +
+  geom_histogram()
+#ggplot(aes(time, famine)) +
+#  geom_line() +
+#  theme_classic()
 
-ggplot(simulation2, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
+sim2 %>%
+  group_by(rep, time) %>%
+  summarise(test = mean(payoff))%>%
+  ggplot(aes(time, test, color = rep)) +
+  geom_point(alpha = .5)
 ```
@@
 ```{r}
-simulation1 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-simulation2 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0, na.rm = TRUE)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
+sim1 %>%
   group_by(time) %>%
   summarise(payoff = mean(payoff)) %>%
-ggplot(aes(time,payoff)) +
-  geom_point(aes(color = time)) +
-  theme_classic()
-
-simulation2 %>%
-  group_by(time) %>%
-  summarise(payoff = mean(payoff, na.rm = TRUE)) %>%
 ggplot(aes(time,payoff)) +
@@
 library(ggridges)
-simulation1 %>%
-  mutate(time_step = 5 * round(time/5),
-         famine = payoff == 0) %>%
-  ggplot(aes(wheat_prop, as.factor(time_step))) +
-  geom_density_ridges(rel_min_height = 0.005) +
-  theme_ridges()
-
-simulation2 %>%
+sim1 %>%
   mutate(time_step = 5 * round(time/5),
@@
 
-\section{Results and Discussion}
+# Discussion and Conclusions
 

### 2020-02-13 manuscript/manuscript.Rmd
refine prior update plot

@@
 library(tidyverse)
+library(patchwork)
 ```
@@
 
-```{r, fig.width = 5, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
+```{r, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
 prior2 <- function(a, b) {
@@
        beta = c(15 * 0.8, 50 * 0.8, 15 * 0.8 + 25 * 0.6, 50 * 0.8 + 25 * 0.6),
-       time = c('before', 'before', 'after', 'after'),
-       age = c('young', 'old', 'young', 'old')) %>%
+       time = factor(c('prior', 'prior', 'posterior', 'posterior'), levels = c('prior', 'posterior')),
+       age = factor(c('15-year-old agent', '50-year-old agent', '15-year-old agent', '50-year-old agent'), levels = c('15-year-old agent', '50-year-old agent'))) %>%
   mutate(belief = map2(alpha, beta, ~prior2(.x, .y))) %>%
@@
   geom_line(aes(group = time, color = time), size = 1.2, alpha = .8) +
-  #scale_color_viridis_d(guide = 'legend') +
-  facet_wrap(~age) +
-  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
-  theme_classic()
+scale_color_manual(values = c('purple', 'orange'), name = '') +
+    facet_wrap(~age, strip.position = 'bottom') +
+    theme_classic() +
+  theme(strip.background = element_blank(),
+        legend.position = c(.43, .5),
+        strip.placement = "outside") +
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density')
 ```
@@
 
-```{r, fig.height = 4, fig.width = 6}
-test1 <- tibble(time = 1:75,
-       risk = c(rep(0.2, 25), rep(0.4, 25), rep(0.2, 25)))
+```{r, fig.height = 8, fig.width = 7}
 
-ggplot(sim1, aes(time, risk)) +
-  geom_line(data = test1) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
- # geom_smooth(color = 'darkorange') +
+a <- ggplot(sim1) +
+  geom_rect(aes(xmin = 25, xmax = 50, ymin = -Inf, ymax = Inf), fill = NA, color = 'black', alpha = .002) +
+  geom_point(aes(time, risk), alpha = .1) +
+  geom_hline(yintercept = indiff$risk, color = 'red') +
   theme_classic()
-```
 
-```{r}
-sim1 %>%
+b <- sim1 %>%
   group_by(time, crop) %>%
@@
   theme_classic()
+
+a / b
 ```
 
-```{r}
+```{reval=FALSE}
 library(furrr)

### 2020-02-13 manuscript/manuscript.Rmd
change color theme to PuOr
also add data directory to gitignore

@@
   theme_classic() +
-  scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
+  #scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
+  scale_color_manual(values = c('#e66101', '#5e3c99'), name = '') +
   labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield')
@@
   summarise(risk = sum(drought) / 50) %>%
+  filter(year != min(year)) %>%
   ggplot(aes(year, risk, fill = risk, color = risk)) +
@@
   geom_col()  +
-  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
-  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
+  scale_fill_distiller(palette = 'PuOr') +
+  scale_color_distiller(palette = 'PuOr')+
   theme_minimal() +
@@
   geom_line(aes(group = Age, color = Age), size = 1.2, alpha = .8) +
-  scale_color_viridis_c(guide = 'legend') +
+  scale_color_distiller(palette = 'PuOr', guide = 'legend') +
   labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
@@
   geom_line(aes(group = time, color = time), size = 1.2, alpha = .8) +
-scale_color_manual(values = c('purple', 'orange'), name = '') +
+scale_color_manual(values = c('#5e3c99', '#e66101'), name = '') +
     facet_wrap(~age, strip.position = 'bottom') +
@@
 a <- ggplot(sim1) +
-  geom_rect(aes(xmin = 25, xmax = 50, ymin = -Inf, ymax = Inf), fill = NA, color = 'black', alpha = .002) +
+  geom_rect(aes(xmin = 25, xmax = 50, ymin = -Inf, ymax = Inf), fill = 'lightgray', alpha = .1) +
   geom_point(aes(time, risk), alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red') +
-  theme_classic()
+  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
+  theme_classic() +
+  labs(x = 'years', y = 'Perceived drought risk')
 
@@
   geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
+  #scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
+  #  scale_fill_manual(values = c('#e66101', '#5e3c99'), name = '') +
+    scale_fill_manual(values = c('#f1a340', '#998ec3'), name = '') + 
+  labs(x = 'Years', y = 'Wheat proportion') +
   theme_classic()

### 2020-02-13 manuscript/manuscript.Rmd
remove extra visualization code

@@
 
-```{r}
-sim2 %>%
-  group_by(rep, time) %>%
-  summarise(test = mean(payoff))%>%
-  summarise(test = mean(test)) %>%
-  #summarise(famine = sum(payoff == 0) / 150) %>%
-  #summarise(famine = mean(famine)) %>%
-  ggplot(aes(test)) +
-  geom_histogram()
-#ggplot(aes(time, famine)) +
-#  geom_line() +
-#  theme_classic()
-
-sim2 %>%
-  group_by(rep, time) %>%
-  summarise(test = mean(payoff))%>%
-  ggplot(aes(time, test, color = rep)) +
-  geom_point(alpha = .5)
-```
-
-```{r}
-sim1 %>%
-  group_by(time) %>%
-  summarise(payoff = mean(payoff)) %>%
-ggplot(aes(time,payoff)) +
-  geom_point(aes(color = time)) +
-  theme_classic()
-```
-
-```{r}
-library(ggridges)
-sim1 %>%
-  mutate(time_step = 5 * round(time/5),
-         famine = payoff == 0) %>%
-  ggplot(aes(wheat_prop, as.factor(time_step))) +
-  geom_density_ridges(rel_min_height = 0.005) +
-  theme_ridges()
-```
-
-make this a function to calculate payoffs for all wheat or all barley strategies as a comparison
-```{r}
-tibble(drought = droughts,
-       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
-       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
-       equal = (barley + wheat) / 2) %>%
-  mutate(time = 1:n()) %>%
-  gather(crop, yield, barley:equal) %>%
-  ggplot(aes(time, yield, group = crop, color = crop)) +
-  geom_point()
-```
-
-```{r}
-tibble(drought = droughts,
-       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
-       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
-       equal = (barley + wheat) / 2) %>%
-  mutate(time = 1:n()) %>%
-  gather(crop, yield, barley:equal) %>%
-  mutate(famine = yield == 0) %>%
-  ggplot(aes(time, famine, group = crop, color = crop)) +
-  geom_line()
-```
-
 # Discussion and Conclusions

### 2020-02-13 manuscript_overflow.rmd
remove extra visualization code

--- 2020-01-24 risk-percption.Rmd
@@
 ---
-title: "Crop diversification in a game against nature in a mediterranean agro-ecosystem"
+title: "manuscript overflow"
 author: "Nick Gauthier"
-date: ""
-bibliography: refs.bib 
+date: "2/12/2020"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-library(raster)
-library(tidyverse)
-library(CDFt)
-```
-
-# Introduction
-
-Over the past 10,000 years, subsistence farmers in the Mediterranean basin have developed sophisticated strategies to maintain stable food supplies given uncertain rainfall.
-Their suite of strategies include practices like crop diversification, storage, and exchange [@Halstead1989]).
-These risk-management strategies all represent forms of \emph{regulatory feedback} used to manage complex socio-ecological systems.
-Regulatory feedbacks convert a system's high-variance inputs into low-variance outputs by dynamically monitoring the outputs and adjusting internal properties of the system [@Anderies2015].
-Risk-managing strategies require farmers to make specific resource allocation decisions in response to environmental risks, and as regulatory feedbacks they are thus sensitive to variability in both the spatiotemporal patterns of risk and the ability of farmers to perceive and act on those patterns.
-
-Crop diversification is an excellent example of a widespread and effective regulatory feedback strategy (Figure 1).
-Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@helmers2001separating,Anderies2006].
-In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,Marston2011190].
-Wheat generally has higher yields but is sensitive to water deficits, while barley has lower yields and is drought tolerant.
-Planting a mix of wheat and barley, either in the same plot or in a combination of plots, is an effective means of diversifying grain supplies.
-By adjusting the ratio of wheat to barley, farmers can adapt to changing drought risks.
-
-![Conceptual model of crop diversification as a regulatory feedback system.
-Uncertain rainfall leads to uncertain returns from land-use systems.
-To keep food supplies stable, farmers must monitor year-to-year variations and decide what combination of crops works best to minimize that variance.](figures/regulatory_feedback.png)
-Are these crop-diversification strategies vulnerable to the same dynamics as other social-ecological systems with similar regulatory feedbacks mechanisms?
-One way to address this is to model the influence of imperfect monitoring and biased decision making in uncertain environments.
-In this study, I accomplish this by answering two main questions:
-
-1. What was the temporal pattern of climate and climate change during the Holocene in the eastern Mediterranean?
-2. How well could crop diversification strategies cope with this pattern, given farmers’ imperfect risk perception and decision-making?
-
-I couple a climate model with high temporal resolution to a simple game theoretic model of agricultural decision making under uncertainty, in order to examine the performance of optimal and suboptimal wheat-barley diversification practices.
-I model the year-to-year crop diversification strategies made by farmers as an iterated game of fictitious play against nature [@GOULD1963a].
-Farmers are boundedly rational, in that they seek to maximize their objective functions given the perceived probabilities of different actions by nature, but work with flawed mental models of climate-related risks due to imperfect recall of past events.
-Furthermore, by using a climate model to represent the actions of nature, I can more precisely capture the characteristic autocorrelated patterns of rainfall variability, rather than simply drawing rainfall values from a static distribution.
-This framework allows me to address the sensitivity of diversification strategies to changes and climatic variability and imperfect human perceptions of those changes.
-
-# Decision-Making in a Game Against Nature
-
-Assume two different crops, a low yield drought tolerant crop (barley) and a high yield drought susceptible crop (wheat).
-
-Estimates of yield volume (t/ha) for each crop type in wet and dry years were derived from isotopic studies of ancient wheat and barley samples from archaeological sites in the Mediterranean [@Slafer1999] (Table 1).
-
-||Dry Year|Normal Year|
-|-------|-----|-----|----|
-|Barley Yield|0.93|1.18|
-|Wheat Yield|0|1.60|
-
-Table: Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].
-
-```{r}
-wheat_dry <- 0
-wheat_normal <- 1.6
-
-barley_dry <- 0.93
-barley_normal <- 1.18
-```
-
-```{r, eval = FALSE}
-readxl::read_xls('Data/41467_2014_BFncomms4953_MOESM580_ESM.xls', range = 'A2:O370') %>%
-  select(`Site Location`, Taxon, `ky cal BP`, `GY (t*ha-1)`) %>%
-  filter(!is.na(`GY (t*ha-1)`)) %>% 
-  group_by(`Site Location`, Taxon) %>%
-  summarise(yield_mean = mean(`GY (t*ha-1)`)) %>%
-  spread(Taxon, yield_mean)
-```
-
-Optimal wheat/barley proportions can be calculated using two different objective functions representing different decision making types of farmers.
-
-## Strategies
-
-1. Wald Maximin -- pessimist, maximize the worst case scenario
-2. Savage Regret -- minimize regret
-3. Hurewicz - balance optimism and pessimism
-4. Laplace or "naive" - 50/50
-
-So both the Wald and Savage regret criteria tell us to choose barley all the time.
-
-If we're using the empirical frequency distirubtion of nature's moves, then we are playing a fictitious game.
-The strategy of planting the crop with the highest expected yield, given the frequency distribution of nature is known, will only work when nature is stationary.
-
-We could think of other algorithms that would minimize the cumulative or average regret of a player over many years in a situation with nonstationary rainfall, but we have to remain within cognitive limits.
-
-The pure maximin strategy is to plant barley:
-
-Alternatively, yield maximizing farmers simply seek to maximize the function for their expected utility given all possible states of nature
-\begin{equation}
-    E\left[ U_{all} \right] = E \left[U_{dry} \right] \times P_{dry}^{*} + E \left[U_{wet} \right] \times \left(1-P_{dry}^{*}\right).
-\end{equation}
-
-Given this information on the potential payoffs for planting each crop type, and information on the (perceived) probability of a drought occurring, the crop-diversification decision practice can be thought of as a simple two-player game against nature [@GOULD1963a].
-The expected yields for a farmer given a fixed proportion of barley and wheat can be calculated as
-
-\begin{equation}
-    E \left[Y_{dry} \right] = W_{dry} \times  P + B_{dry} \times (1 - P)
-\end{equation}
-if a drought occurs and likewise
-\begin{equation}
-    E \left[Y_{wet} \right] = W_{wet} \times P + B_{wet} \times (1 - P)
-\end{equation}
-
-if a drought does not occur. This specification allows us to calculate the optimal proportion of wheat to barley one should plant in a diversified strategy as a mixed strategy profile.
-
-Risk minimizing farmers play the equivalent of a mixed strategy Nash equilibrium, setting the expected yields in drought and normal years equal to each other, weighted by risk, and solving for the wheat/barley proportion
-\begin{equation}
-    E \left[Y_{dry} \right] \times P_{dry}^{*} = E \left[U_{wet} \right] \times \left(1- P_{dry}^{*}\right), 
-\end{equation}
-where $P_{dry}^{*}$ is the \emph{perceived}, not realized, risk of a drought in a particular year.
-The solution to this function represents the ratio of wheat to barley that would make a farmer indifferent to whether a drought would occur in a given year.
-
-The expected yield is thus calculated as.
-```{r}
-expected_yield_wheat <- function(risk) {
-  risk * wheat_dry + (1 - risk) * wheat_normal
-}
-
-expected_yield_barley <- function(risk) {
-  risk * barley_dry + (1 - risk) * barley_normal
-}
-
-expected_yield <- function(risk) {
-  expected_yield_wheat(risk) + expected_yield_barley(risk)
-}
-```
-
-Given these payoffs, what's the point of indifference between the two crop types?
-```{r, echo = FALSE, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
-indiff <- tibble(risk = (barley_normal - wheat_normal) / (wheat_dry - wheat_normal - barley_dry + barley_normal),
-                 yield = risk * wheat_dry + (1 - risk) * wheat_normal)
-indiff_label <- paste0('(', round(indiff$risk, 2), ', ', round(indiff$yield, 2), ')')
-
-tibble(risk = seq(0, 1, .1),
-      Barley = expected_yield_barley(risk),
-      Wheat = expected_yield_wheat(risk)) %>%
-  gather(type, yield, Barley:Wheat) %>%
-  ggplot(aes(risk, yield)) +
-  geom_hline(yintercept = indiff$yield, linetype = 2)+
-  geom_vline(xintercept = indiff$risk, linetype = 2) +
-  geom_line(aes(color = type), size = 1.2) +
-  geom_point(data = indiff, color = 'black', size = 3) +
-  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
-  theme_classic() +
-  scale_color_discrete(name = 'Crop') +
-  labs(x = 'Drought probability', y = 'Expected yield')
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 ```{r}
-# farmer's ne
-pmax(0, round((barley_normal - barley_dry) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2))
-#nature's ne
-round((barley_normal - wheat_normal) / (wheat_dry + barley_normal - (wheat_normal + barley_dry)), 2)
-```
+sim2 %>%
+  group_by(rep, time) %>%
+  summarise(test = mean(payoff))%>%
+  summarise(test = mean(test)) %>%
+  #summarise(famine = sum(payoff == 0) / 150) %>%
+  #summarise(famine = mean(famine)) %>%
+  ggplot(aes(test)) +
+  geom_histogram()
+#ggplot(aes(time, famine)) +
+#  geom_line() +
+#  theme_classic()
 
-0.3111 is also the indifference point for the farmer with respect to crops.
-That is, if we imagine nature playing strategically, a drought would occur with probability 0.31111 because that would make nature indifferent to whether the farmer planted wheat or barley.
-
-when you sample from the expected value, use that to choose a strategy. but if you sample lots, use that to create a mixed streateg
-
-So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
-
-*Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
-
-## Crop Allocation Decisions
-
-These equations are iteratively applied to the entire span of the climate-model output.
-The result is that farmers calculate the optimal land-use strategy that, \emph{to the best of their knowledge and given their preferences}, uses information from the past.
-
-# Climate Risks and Risk Perception
-
-Using the bias-corrected climate model output, I divided each model year into dry years and wet years.
-A dry year was any year where less than 300mm of rain fell during the wet season (October-March), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
-
-Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years
-
-\begin{equation}
-    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
-\end{equation} 
-where $t$ is the current time step.
-The 50 year time span was selected to approximate the accumulated observational knowledge of an individual farmer and their immediate household.
-As a result, this approach does not allow for accumulated social learning, although it could easily be extended to do so in future studies.
-
-Start with a population of agents with the same subjective beliefs about drought risk.
-
-```{r}
-agents <- tibble(id = 1:150, 
-                 alpha = 1.5,
-                 beta = 1.5,
-                 risk = alpha / (alpha + beta),
-                 age = sample.int(80, size = 150, replace = TRUE),
-                 crop = 'barley')
+sim2 %>%
+  group_by(rep, time) %>%
+  summarise(test = mean(payoff))%>%
+  ggplot(aes(time, test, color = rep)) +
+  geom_point(alpha = .5)
 ```
@@
 ```{r}
-agents
-```
-
-## Bayesian Updating
-
-```{r}
-prior <- function(m,n){
-a = n * m
-b = n * (1 - m)
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-tibble(year = 1:20,
-       risk = .5) %>%
-  mutate(belief = map2(risk, year, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line() +
-  facet_wrap( ~year) +
-  theme_bw() +
-  ggtitle('Subjective beliefs about drought risk', 'Beta prior, varying sample size')
-
-#prior(.5, ) %>% plot
-```
-
-```{r}
-likelihood <- function(N,Y){
-a <- Y + 1
-b <- N - Y + 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-#likelihood(1, 1)%>% plot
-```
-
-```{r}
-posterior <- function(m,n,N,Y){
-a <- Y + (n*m) -1
-b <- N - Y + (n*(1-m)) - 1
-dom <- seq(0,1,0.005)
-val <- dbeta(dom,a,b)
-return(data.frame('x'=dom, 'y'=val))
-}
-
-#dbeta(seq(0,1,0.005), -1, 0)
-#prior(.5, 2) %>% plot
-#posterior(.5, 3, 1, 1) %>% plot
-```
-
-Allow the agents to learn the drought risk over time. Also need to calculate the actual payoffs for each agent.
-
-```{r}
-learn <- function(agents, drought) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
-
-  agents %>%
-    # farmer's turn to decide what to plant
-      mutate(yield_exp = expected_yield(risk),
-         barley_exp = expected_yield_barley(risk),
-         wheat_exp = expected_yield_wheat(risk),
-         expected_yield = barley_exp + wheat_exp,
-         crop = case_when(barley_exp > wheat_exp ~ 'barley',
-                              barley_exp < wheat_exp ~ 'wheat',
-                              barley_exp == wheat_exp ~ 'either'),
-         prediction = map2(alpha, beta, ~rbeta(100, .x, .y)), # sample 100 as if agent is thinking in percentages
-         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
-      select(-prediction) %>%
-  # nature's turn to decide realized yields
-    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-           alpha = drought + alpha,
-           beta = beta + 1 - drought,
-           mu = alpha / (alpha + beta),
-           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
-}
-```
-
-Similar to above, but satisficing
-```{r}
-learn_satisficing <- function(agents, drought, aspiration) {
-  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
-  if (drought == 'FALSE') wheat_yield <- wheat_normal; barley_yield <- barley_normal
-
-  agents %>%
-     mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-            new_prop = round(runif(n()), 2),
-           wheat_prop = if_else(payoff > aspiration, wheat_prop, new_prop)) %>%
-    select(-new_prop)
-}
-```
-
-Run an experiment. Simulate a time series with varying drought frequency.
-```{r}
-droughts <- c((runif(25) < .1),(runif(25) < .5), runif(50) < .2)
-```
-
-```{r}
-simulation1 <- accumulate(droughts, ~learn(.x, .y), .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
-
-```{r, echo = TRUE}
-ggplot(simulation1, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation2 <- accumulate(droughts, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
-
-```{r}
-ggplot(simulation2, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
-```
-
-```{r}
-simulation2 %>%
-  group_by(time, crop) %>%
-  tally() %>%
-ggplot(aes(time, n, fill = crop)) +
-  geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  theme_classic()
-```
-
-These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
-
-Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals
-```{r}
-ggplot(simulation1, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
-
-ggplot(simulation2, aes(time, wheat_prop)) +
-  geom_point(alpha = .3) +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-simulation2 %>%
-  group_by(time) %>%
-  summarise(famine = sum(payoff == 0, na.rm = TRUE)) %>%
-ggplot(aes(time, famine)) +
-  geom_line() +
-  theme_classic()
-```
-
-```{r}
-simulation1 %>%
+sim1 %>%
   group_by(time) %>%
   summarise(payoff = mean(payoff)) %>%
-ggplot(aes(time,payoff)) +
-  geom_point(aes(color = time)) +
-  theme_classic()
-
-simulation2 %>%
-  group_by(time) %>%
-  summarise(payoff = mean(payoff, na.rm = TRUE)) %>%
 ggplot(aes(time,payoff)) +
@@
 ```
-
-```{r}
-test3 <- tibble(agent = 1:100) %>%
-  mutate(wheat_prop = round(runif(n()), 2))
-
-simulation3 <- accumulate(droughts, ~learn_satisficing(.x, .y, aspiration = .5), .init = test3) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-
-simulation2 %>% ggplot(aes(as.numeric(time), payoff, group = agent)) + geom_point(alpha = .2)
-simulation2 %>% ggplot(aes(as.numeric(time), wheat_prop, group = agent)) + geom_point(alpha = .2)
-```
-
-```{r}
-load('../Silvanus/R/sysdata.rda')
-life_table
-fertility_shape <- 4.579627
-fertility_scale <- 0.1
-survivor_scale <- 0.1
-```
-
-```{r}
-reproduce <- function(individuals) {
-  individuals %>%
-    left_join(life_table, by = 'age') %>% 
-    mutate(fertility_reduction = pgamma(payoff, shape = fertility_shape, scale = fertility_scale),
-           reproduced = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>% # divide by two to make everyone female
-    filter(reproduced == TRUE) %>% # if nrows == 0, will give a (for some reason unsuppressible) warning
-    tally %>% # count births per household
-    uncount(n) %>% # repeat rows based on birth count per household
-    mutate(age = 0, alpha = 1.5, beta = 1.5, risk = alpha / (alpha + beta), crop = 'barley') %>%
-    bind_rows(individuals, .) %>%
-    ungroup # check for filled na's here?
-}
-
-die <- function(individuals) {
-  individuals %>%
-        left_join(life_table, by = 'age') %>% 
-    mutate(survival_reduction = pgamma(pmin(1, payoff), shape = survival_shape, scale = survivor_scale),
-           survived = rbernoulli(n(), survival_rate * survival_reduction)) %>%
-    filter(survived == TRUE) %>%
-    select(-c(fertility_rate:survived)) %>%
-    mutate(age = age + 1)
-}
-
-```
-
-Ok so next we need to have the birth and death stuff run with all the prior information in it.
-dont need to propogate the prior exactly, just need to add in a new weak prior.
-at least propogate the risk number, but not the prior?
-```{r}
-reduce(1:600, ~reproduce(.x) %>% die , .init = agents) %>%
-  pull(age) %>% hist
-```
-
-## Climate Modeling and Bias Correction
-
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research's Earth System Grid repository (https://www.earthsystemgrid.org).
-TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to orbital parameters and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present (Figure 2) [@He2011].
-It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
-orbit, greenhouse gasses, glacial meltwater flux).
-The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
-
-![Example model precipitation outputs from the TraCE-21k simulation at 9.5ka BP.](figures/trace_white.png)
-
-TraCE-21k simulation outputs for the past 10,000 years were bias-corrected using the CDF-t method and observed precipitation from a weather station in near the town of Salihli in western Turkey.
-This location was selected due to its long (~70yr) observational record, the representativeness of western Anatolian climate for the greater eastern Mediterranean, and its proximity to archaeological sites with comparative evidence of crop diversification strategies.
-
-Now incorporate realistic drought variability from simulations.
-
-```{r}
-konya <- matrix(c(32.5, 37.9), nrow = 1)
-#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
-```
-
-```{r}
-precc <- map(list.files('Data/PRECC', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precc) ) %>%
-  bind_rows()
-precl <- map(list.files('Data/PRECL', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precl) ) %>%
-  bind_rows()
-
-sim_past <- left_join(precc, precl) %>%
-  mutate(rainfall = (precc + precl) * 2.628e+9,
-         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 5200)) %>%
-  select(-time, -precl, -precc) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-   filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  add_tally() %>%
-filter(n == 8) %>%
-  summarise(precip = sum(rainfall)) %>%
-  mutate(drought = precip < 300)
-```
-
-```{r, fig.height =3, fig.width = 9, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by century.'}
-sim_past %>%
-  filter(between(water_year, -9500, -5500)) %>%
-  mutate(time_step = ceiling(water_year / 50),
-         year = time_step * 50) %>%
-  group_by(year) %>%
-  summarise(risk = sum(drought) / 50) %>%
-  ggplot(aes(year, risk, fill = risk, color = risk)) +
-  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, width = 1.2) +
-  geom_col()  +
-  scale_fill_distiller(palette = 'YlOrRd', direction = 1) +
-  scale_color_distiller(palette = 'YlOrRd', direction = 1)+
-  theme_minimal() +
-  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
-```
-
-```{r}
-simulation4 <- accumulate(sim_past$drought, ~learn(.x, .y) %>% reproduce %>% die, .init = agents) %>% 
-  bind_rows(.id = 'time') %>%
-  mutate(time = as.numeric(time)) %>%
-  filter(time != 1)
-```
-
-So the problem now is the population dynamics are getting in the way ... and that it'd be nice to have something akin to the moran process to ensure fixed population
-```{r, echo = TRUE}
-ggplot(simulation4, aes(time, risk)) +
-  geom_point(alpha = .1) +
-  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
-  theme_classic()
-```
-
-```{r}
-simulation4 %>%
-  group_by(time) %>%
-  tally %>%
-  plot
-```
-
-so we have to connect the food ratio to the payoffs? maybe have the barley yield be ~ .8 food ratio
-
-```{r}
-.5/.8
-```
-
-\section{Results and Discussion}
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=16cm]{precon.png}
-\caption{Bias-corrected wet season (October-March) precipitation accumulations from the past 10,000 years of TraCE-21k simulation.
-1 point = 1 year, color corresponds to whether that year falls bellow the thresholds for wheat or barley crop failure.}
-\end{figure}
-
-The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure 5).
-The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
-
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
-Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
-
-Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
-$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
-This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
-Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
-Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
-
-Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
-Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
-Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
-
-During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
-Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
-This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
-This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
-
-Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
-To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
-Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
-Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
-
-Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
-
-# References
-
-```{r}
-knitr::knit_exit()
-```
-
-```{r}
-moran_birth_death <- function(agents, r = 1) {
-  n <- nrow(agents)
-  birth_index <- sample.int(n, size = r)
-  death_index <- sample.int(n, size = r)
-  newborn <- agents %>%
-    filter(row_number() %in% birth_index) %>%
-    mutate(age = 1)
-
-  agents %>%
-    filter(!row_number() %in% death_index) %>%
-    mutate(age  = age + 1) %>%
-    bind_rows(newborn)
-}
-
-reduce(1:1000, ~moran_birth_death(., r = 4), .init = agents2) %>%
-  pull(age) %>%hist
-```
-
-so in this case drought risk is low enough all the time you never get any diversification
-
-~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-## Satisficing learners
-
-```{r}
-library(ggridges)
-accumulate((runif(25) < .2), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .6), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(wheat_prop, y = as.numeric(time), group = as.numeric(time))) +
-  geom_density_ridges() +
-  theme_minimal()
-```
-
-```{r}
-# this is for nature
-calcNE <- function(risk){
-  (118 - 25 * risk) / (278 - 185 * risk)
-}
-
-plot(0:10, calcNE((0:10)/10))
-```
-
-so we're imagining a yield of 1 for barley and 2 for wheat, with equal proportions of each. then in the next time step we'd revise the proportion to take away .25
-```{r}
-reallocate <- function(wheat_prop, risk) {
-  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
-}
-
-accumulate(1:10, ~reallocate(.x), .init = .4) %>% plot
-
-# so this works as long as you're between 0 and 1
-```
-
-```{r}
-yield_max_barley <- 2500
-yield_max_wheat <- 3500
-kcal_barley <- 3000
-kcal_wheat <- 3540
-
-calc_yield <- function(rainfall, crop = 'wheat') {
-  if (crop == 'wheat') {
-    yield <- yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
-  }
-
-  if(crop == 'barley') {
-    yield <- yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51) 
-  }
-
-  return(yield)
-}
-
-tibble(rainfall = (0:200) / 100,
-       barley = calc_yield(rainfall, 'barley'),
-       wheat = calc_yield(rainfall, 'wheat')) %>%
-  gather(crop, yield, barley:wheat) %>%
-  ggplot(aes(rainfall, yield, color = crop)) + 
-  geom_line() +
-  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
-  geom_hline(yintercept = yield_max_barley * kcal_barley) +
-  geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
-  geom_vline(xintercept = 0.3) +
-  geom_vline(xintercept = .2)
-```
-
-How do we define the "drought year" cutoff? Below 300mm there isn't enough wheat growing in a hectare to feed a family of 5, which seems appropriate.
-```{r}
-test <- tibble(time = 1:100,
-       rain = rnorm(100, mean = 1, sd = 1) %>% pmax(0),
-       yield = calc_yield(rain))
-
-ggplot(test, aes(time, yield)) + geom_point() + geom_hline(yintercept = 1016525)
-
-test %>%
-  mutate(drought = if_else(yield < 5e+6, TRUE, FALSE))%>%
-  filter(drought == TRUE)
-```

### 2020-02-18 manuscript/manuscript.Rmd
preliminary .tex file

@@
 output: rticles::springer_article
+header-includes:
+- \usepackage{float}
+- \floatplacement{figure}{tb}
 ---
@@
   scale_color_manual(values = c('#e66101', '#5e3c99'), name = '') +
-  labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield')
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield (t/ha)')
 ```
@@
 
-```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
+```{r, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
 The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
@@
 
-### Bayesian learning from bernoulli observations
-We can treat this problem elegantly using bayesian statistics using online bayesian learning from bernoulli objersbations (Bissiri 2010).
-Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
-bernoulli distribution) with paramter theta reprenting the risk of drought.
-Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1 (Figure XX).
-We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief.
-What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
-As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
-
-A simple "fixed" bayesian learner thus gets more certain with age.
-For young agents with weak priors, the information of each new year can strongly influence their beliefs.
-For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
-As was the case with the simple game of ficitious play above, this strategy is fine when the environment is stable and risks do not change.
-But when there is volatility in the environment, and risks can change, it becomes important to change your mind.
-Being too inflexible in one's priors can lead being too optimistic when things have really changed for the worse, and alternately being too conservative when things finally do get better.
-So from a Bayesian learning standpoint, how should agents choose to believe new information.
-
 ```{r, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
@@
 ```
+
+### Bayesian learning from bernoulli observations
+We can treat this problem elegantly using bayesian statistics using online bayesian learning from bernoulli objersbations (Bissiri 2010).
+Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
+bernoulli distribution) with paramter theta reprenting the risk of drought.
+Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1 (Figure XX).
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief.
+What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
+As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
+
+A simple "fixed" bayesian learner thus gets more certain with age.
+For young agents with weak priors, the information of each new year can strongly influence their beliefs.
+For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
+As was the case with the simple game of ficitious play above, this strategy is fine when the environment is stable and risks do not change.
+But when there is volatility in the environment, and risks can change, it becomes important to change your mind.
+Being too inflexible in one's priors can lead being too optimistic when things have really changed for the worse, and alternately being too conservative when things finally do get better.
+So from a Bayesian learning standpoint, how should agents choose to believe new information.
 
@@
 ```
-
-# Risk perception in a Mediterranean agroecosystem
 
@@
   tally() %>%
-ggplot(aes(time, n, fill = crop)) +
+ggplot(aes(time, n, fill = crop, color = crop)) +
   geom_bar(width = 1, stat = 'identity', position = 'fill') +
@@
     scale_fill_manual(values = c('#f1a340', '#998ec3'), name = '') + 
+    scale_color_manual(values = c('#f1a340', '#998ec3'), name = '') + 
     labs(x = 'Years', y = 'Wheat proportion') +
@@
 a / b
-```
-
-```{reval=FALSE}
-library(furrr)
-plan(multisession)
-
-sim2 <- tibble(rep = 1:500,
-       droughts = replicate(500, make_droughts(), simplify = FALSE)) %>%
-  mutate(results = future_map(droughts, run_sim)) %>%
-  select(-droughts) %>%
-  unnest(results)
 ```

### 2020-02-18 manuscript/manuscript.tex
preliminary .tex file

@@
 
+\usepackage{float}
+\floatplacement{figure}{tb}
+
 \begin{document}
@@
 
-    \titlerunning{Risk perception in a Mediterranean agroecosystem}
+    \titlerunning{The dynamics of risk perception in a Mediterranean agroecosystem}
 
@@
 \item
-  What was the temporal pattern of climate and climate change during the
-  Holocene in the eastern Mediterranean?
+  How likely are droughts to occur each year in the eastern
+  Mediterreanean, and how did these risks change over the Holocene?
 \item
-  How well could crop diversification strategies cope with this pattern,
-  given farmers' imperfect risk perception and decision-making?
+  How might Neolithic farmers have percieved these changing drought
+  risks, and what were the consequences for farming communities with
+  respect to their collective ability to manage risk?
 \end{enumerate}
-
-Questions: How would the earliest farmers have made decisions (dealt
-with uncertainty?) about crop diversification? ie what were the
-different algorithms? -- better understand the decision context What are
-the consequences of theses different decisions (i.e.~how well did they
-perform) at different points in the history of the mediterranean How
-were farmers able to detect changes in the environment, rather than
-unconsciously adapt to them? what are the different consequences we'd
-expect for each THIS is the most important question because with it I
-can frame the results around and individual's perception of droughts in
-the neolithic, and the populaiton-level distribution of drought
-perceptions at given times
 
@@
 this simplified decision context as a ``game'' in a game theorietic
-context. cite luce raiffa 1989 for game against nature, and milnor 1952,
-Agrawal1968, cassidy1971, gould 1963 The ``game'' in this context is the
-farmer's decision of which crops to plant and in what proportions, given
-uncertianty in the future ``state-of-nature'', and nature varies between
-several possible states such as dry and wet years. payoff matrix goes
-here in reality continuous, but cite behavioral econ for people often
-intuitively solve an easier problem when faced with a complex real-world
-situation The exact values in the payoff matrix here are less important
-than the relative payoffs in each quadrant. We can think of this game as
-being the culturally-inherited object. our question is how do
-individuals make decisions here? so in this case the cultural norm would
-be to mix wheat and barley, but the learning is in the exact ratio --
-both individual and social learning influence perceptions of eindividual
-vs social learning tucker 2007
-
-Estimates of yield volume (t/ha) for each crop type in wet and dry years
-were derived from isotopic studies of ancient wheat and barley samples
-from archaeological sites in the Mediterranean (Araus, Slafer, and
-Romagosa 1999) (Table 1).
+context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould
+1963). The ``game'' in this context is the farmer's decision of which
+crops to plant and in what proportions, given uncertianty in the future
+``state-of-nature'', and nature varies between several possible states
+such as dry and wet years (Table 1). In reality continuous, but people
+often intuitively solve an easier problem when faced with a complex
+real-world situation (cite behavioral econ or simon). The exact values
+in the payoff matrix here are less important than the relative payoffs
+in each quadrant. We can think of this game as being the
+culturally-inherited object. Our question is how do individuals make
+decisions here? So in this case the cultural norm would be to mix wheat
+and barley, but the learning is in the exact ratio.
 
@@
 to get you being a good idea in states of complete uncertainty
-transition -- the problem becomes easier if you have at least someI idea
-of nature's moves , because then you cna work to get the highest yeilds
-given the risk of drought Decision strategies as choosing different
+transition -- the problem becomes easier if you have at least some idea
+of nature's moves, because then you cna work to get the highest yeilds
+given the risk of drought. Decision strategies as choosing different
 points -- plot of decisions
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
 \caption{Expected wheat and barley yields under increasing drought risk
@@
 drought in any given year, then they can behave more rationally by
-trying to maximize their subjective expected utility If the probability
+trying to maximize their subjective expected utility. If the probability
 distributin of nature's moves is known, the farmer can choose the crop
@@
 frequency distirubtion of nature's moves. In the language of game
-theory, this strategy is knoan as a ``game of ficticious play'' against
+theory, this strategy is known as a ``game of ficticious play'' against
 nature. This is an effective risk-managing strategy, but like all
@@
 \hypertarget{sec:2}{%
-\section{The dynamics of risk perception}\label{sec:2}}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
-
-\hypertarget{drought-risks-in-the-middle-holocene}{%
-\section{Drought risks in the middle
-Holocene}\label{drought-risks-in-the-middle-holocene}}
-
-\hypertarget{climate-risks-and-risk-perception}{%
-\section{Climate Risks and Risk
-Perception}\label{climate-risks-and-risk-perception}}
-
-Using the bias-corrected climate model output, I divided each model year
-into dry years and wet years. A dry year was any year where less than
-300mm of rain fell during the wet season (October-March), the threshold
-below which wheat crops will generally fail ({\textbf{???}}), and a
-wet/normal year was defined as any year above this threshold.
-
-Given the modeled patterns of wet and dry years, drought risk for any
-particular year was defined as the proportion of the previous 50 years
-that were dry years
-
-\begin{equation}
-    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
-\end{equation} where \(t\) is the current time step. The 50 year time
-span was selected to approximate the accumulated observational knowledge
-of an individual farmer and their immediate household. As a result, this
-approach does not allow for accumulated social learning, although it
-could easily be extended to do so in future studies.
-
-\hypertarget{climate-modeling-and-bias-correction}{%
-\subsection{Climate Modeling and Bias
-Correction}\label{climate-modeling-and-bias-correction}}
+\section{Holocene drought risks in the eastern
+Mediterranean}\label{sec:2}}
 
@@
 paleoclimate simulation available on the National Center for Atmospheric
-Research's Earth System Grid repository
-(\url{https://www.earthsystemgrid.org}). TraCE-21K is a state-of-the-art
-simulation that uses a coupled atmosphere-ocean general circulation
-model (GCM) to recreate the transient response of the global climate
-system to orbital parameters and greenhouse gasses over the past 22,000
-years from the Last Glacial Maximum to the present (Figure 2) (He 2011).
-It generates physically consistent spatiotemporal climate dynamics,
-driven by current best estimates of climate forcings (e.g.~orbit,
-greenhouse gasses, glacial meltwater flux). The model simulates these
-dynamics on a six hourly timescale, and model outputs are archived at a
-monthly resolution.
+Research. TraCE-21K is a state-of-the-art simulation that uses a coupled
+atmosphere-ocean general circulation model (GCM) to recreate the
+transient response of the global climate system to changes in the
+Earth's orbit and greenhouse gasses over the past 22,000 years from the
+Last Glacial Maximum to the present (He 2011). It generates physically
+consistent spatiotemporal climate dynamics, driven by current best
+estimates of climate forcings (e.g.~orbit, greenhouse gasses, glacial
+meltwater flux). The model simulates these dynamics on a six hourly
+timescale, and model outputs are archived at a monthly resolution.
 
-TraCE-21k simulation outputs for the past 10,000 years were
-bias-corrected using the CDF-t method and observed precipitation from a
-weather station in near the town of Salihli in western Turkey. This
-location was selected due to its long (\textasciitilde{}70yr)
-observational record, the representativeness of western Anatolian
-climate for the greater eastern Mediterranean, and its proximity to
+TraCE-21k simulation outputs for the past 10,000 years were extracted
+from the grid cell overlaying the northern Levant and southeastern
+Turkey. This location was selected due to its proximity to
 archaeological sites with comparative evidence of crop diversification
-strategies.
+strategies. This region includes some of the earliest evidence of crop
+domestication, including sites like Jericho and Can Hasan.
 
-Now incorporate realistic drought variability from simulations.
-
-\begin{verbatim}
-## Loading required namespace: ncdf4
-\end{verbatim}
+Using the climate model output, I divided each model year into dry years
+and wet years. A dry year was any year where less than 300mm of rain
+fell during the wet season (October-May), the threshold below which
+wheat crops will generally fail ({\textbf{???}}), and a wet/normal year
+was defined as any year above this threshold. Given the modeled patterns
+of wet and dry years, drought risk for any particular year was defined
+as the proportion of the previous 50 years that were dry years. Using
+the simulated paleoclimate, rather than modern-day weather station
+observations or paleoclimate proxies, allows us to estimate not just the
+first order statistics of the climate system (mean, variance of dry
+years) but also the higher order dynamics like serial autocorrelation in
+the perisistence of wet and dry years. This presents a much more
+realistic picture of the inherent year-to-year uncertainty in the
+climate system, and presents a challenge to simple risk-managing
+strategies that assume climatic risks are fixed.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
 \caption{Annual risk of wheat crop failure due to drought, averaged by
-fifty-year period}
+fifty-year period. The dashed line indicates the level of risk beyond
+which one would plant barley over wheat to maximize subjective expected
+yields, given the payoffs defined in Table 1.}
 \end{figure}
 
-\hypertarget{the-model}{%
-\section{The Model}\label{the-model}}
+\hypertarget{sec:3}{%
+\section{The dynamics of risk perception}\label{sec:3}}
 
-Run an experiment. Simulate a time series with varying drought
-frequency. This is a stepwise nonstationary process
+In order to manage risk, decision-makers must be perceive what exactly
+that risk is in the local environment. An individual's perceived risk
+entails more than simply the ``objective'' risk in the external
+environment (Tucker et al. 2013). A person's perceived risk depends in
+large part on their past experience, and how those experiences shape
+their expectations for the future. Likewise, the social responses to
+risk will refelct the distribution of perceived risks in a population,
+and by extension the collective memory of past droughts will have
+consequences for the population's survival and evolution.
+
+We can capture the influences of risk perception computationally by
+simulating a population of ``belief-based agents.'' These agents develop
+beliefs about the risk of drought in their environment through personal
+experience. Each agent's perceived drought risks will change their
+subjective expected crop yields from planting different ratios of wheat
+to barley.
+
+\hypertarget{bayesian-brains}{%
+\subsection{Bayesian brains}\label{bayesian-brains}}
+
+We can efficiently simulate these population-level risk perceptions in a
+population of agents with Bayesian brains (Kahvalti et al 2019 double
+check). Bayesian learning allows for a minimal representation of the
+environemnt, so that each agent's beliefs can be encoded in two
+parameters -- defining the Beta distribution over their prior beliefs --
+rather than the record of all observed years. Not only does this
+approach have a clear computational efficiency, but for the same reason
+this reflects the processes actually going on in the heads of decision
+makers. Our brains do not record every bit of perceived informaiton in
+memory, rather it stores a ``compact encoding'' of that information
+which it uses for future decisions. Even if individuals aren't --
+consciously or unconsciously --- making these calculations -- the basic
+algorithmic problem faced by the brain, and solutions it has evolved,
+reflect the same basic sets of constraints (Sanborn and Chater 2016).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{Development of an individual's perceived drought risk with
+time, assuming a fixed drought risk of 0.5. Beliefs are represented as
+Beta probabiity distributions, and the increased certainty with age
+reflects the varying effective sample size of the Beta prior.}
+\end{figure}
+
+\hypertarget{bayesian-learning-from-bernoulli-observations}{%
+\subsubsection{Bayesian learning from bernoulli
+observations}\label{bayesian-learning-from-bernoulli-observations}}
+
+We can treat this problem elegantly using bayesian statistics using
+online bayesian learning from bernoulli objersbations (Bissiri 2010).
+Put simply, the probability of a drought occuring in any given year is
+treated as a coin flip (i.e.~bernoulli distribution) with paramter theta
+reprenting the risk of drought. Uncertainty in the value of theta can be
+represented with a beta distribution, constrained between 0 and 1
+(Figure XX). We can thus represent the exact information content of each
+individual agent's subjective experience of droughts using the diffusion
+of the prior belief. What we see here is how an indiviudal's perception
+of the risk of drought may start off diffuse at first (here as an
+uniformative flat prior in which all values of theta are equally likely.
+As each year of experience is observed, the agent updates its beliefs
+accordingly, balancing the information in this year's observation with
+the cumulative weight of experience.
+
+A simple ``fixed'' bayesian learner thus gets more certain with age. For
+young agents with weak priors, the information of each new year can
+strongly influence their beliefs. For older agents who have experienced
+many more years their priors will be stronger, and they will be less
+likely to update their beliefs when balancing the information from a
+single year with the many decades of accumulated experience. As was the
+case with the simple game of ficitious play above, this strategy is fine
+when the environment is stable and risks do not change. But when there
+is volatility in the environment, and risks can change, it becomes
+important to change your mind. Being too inflexible in one's priors can
+lead being too optimistic when things have really changed for the worse,
+and alternately being too conservative when things finally do get
+better. So from a Bayesian learning standpoint, how should agents choose
+to believe new information.
+
+\hypertarget{the-problem-of-nonstationarity}{%
+\subsubsection{The problem of
+nonstationarity}\label{the-problem-of-nonstationarity}}
+
+That is, what if the risk of a drought, theta, is not a constant but
+rather varies unpredictably in time? The statistical problem shifts then
+to one of change detection (o'reilly 2013, gallistel et al 2014). In
+other words do people simply adapt to changes -- as in the fixed
+bayesian learner and moving average approaches -- or do they detect
+them? Strict year-to-year updating can result in overly conservative
+risk perceptions, suggesting people are slow to change their beliefs
+when in reality they often quickly update them (Bissiri 2010). Indeed,
+the ``forgetful bayesian brain'' model seems to perform better, with a
+dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{Change in perceived drought risks in an older (50) and younger
+(15) agent before and after a 25-year dry period.}
+\end{figure}
+
+So what are the consequences of these cognitive processes for early
+farmers in the Mediterranean.
+
+Simulate a time series with varying drought frequency. This is a
+stepwise nonstationary process. Setup an experiment exploring 6
+scenarios.
+
+This suggests that people are more in danger of not realizing conditions
+are improving when they actually have -- the opportunity cost -- rather
+than failing to realize the change in the firstplace. The law of small
+numbers is very important here. The range of each person's experience is
+so small -- every individual's perceived risk will be a small and noisy
+sample of the true value. Older people are much more certain, even
+though they are wrong.
 
 \includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
-
-Redo these plots so its a single person, one who was born in a drought
-period then experiences an ameliorarion and another who experiences a
-shift to drought conditions later in life? basically this section neds
-to work more with individuals and explore that stuff
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-17-1.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
 
@@
 barley in the population, compared to below which shows the wheat
-proportions of individuals
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-2.pdf}
+proportions of individuals.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-2.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-2.pdf}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.0146
-\end{verbatim}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.0204
-\end{verbatim}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-2.pdf}
-
-make this a function to calculate payoffs for all wheat or all barley
-strategies as a comparison
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-24-1.pdf}
-
-\section{Results and Discussion}
+\hypertarget{discussion-and-conclusions}{%
+\section{Discussion and Conclusions}\label{discussion-and-conclusions}}
 
@@
 
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+
 \hypertarget{references}{%
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Slafer1999}{}%
-Araus, J. L., G. A. Slafer, and I. Romagosa. 1999. ``Durum wheat and
-barley yields in antiquity estimated from 13C discrimination of
-archaeological grains: a case study from the Western Mediterranean
-Basin.'' \emph{Australian Journal of Plant Physiology} 26 (4): 345.
-\url{https://doi.org/10.1071/PP98141}.
-
 \leavevmode\hypertarget{ref-Garnsey1989}{}%
@@
 
+\leavevmode\hypertarget{ref-Tucker2013}{}%
+Tucker, Bram, Jaovola Tombo, Tsiazonera, Patricia Hajasoa, Charlotte
+Nagnisaha, Vorisoa Rene Lahitoka, and Christian Zahatsy. 2013. ``Beyond
+Mean and Variance: Economic Risk Versus Perceived Risk of Farming,
+Foraging, and Fishing Activities in Southwestern Madagascar.''
+\emph{Human Ecology} 41 (3). Springer US: 393--407.
+\url{https://doi.org/10.1007/s10745-013-9563-2}.
+
 \bibliographystyle{spbasic}

### 2020-02-19 manuscript/manuscript.Rmd
reorder sections

@@
 authorrunning: Gauthier
-thanks: | 
-    Grants or other notes about the article that should go on the front 
-    page should be placed here. General acknowledgments should be placed at the
-    end of the article.
 
@@
   address: Laboratory of Tree-Ring Research & School of Geography and Development, University of Arizona
-  email: ngauthier@email.arizona.edu
+  email: ngauthier@arizona.edu
 
 keywords:
-- key
-- dictionary
-- word
-
-#PACS: 
-#- PAC1
-#- superPAC
-
-MSC:
-- MSC code 1
-- MSC code 2
+- drought risk
+- belief-based agents
+- crop diversification
+- rapid climate change
 
@@
 example of system that combines low yield, low wirks crop with high yield high risk crop, an effective context paut et al 2019
-These strategies work when risks are knonw and constant, but may be more susceptible to periods of changing climateic risks
+These strategies work when risks are knonw and constant, but may be more susceptible to periods of changing climatic risks
 so understanding how people of the past adapted to change is important to us today
@@
 We can represent this simplified decision context as a "game" in a game theorietic context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould 1963).
-The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertianty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years (Table 1).
+The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years (Table 1).
 In reality continuous, but people often intuitively solve an easier problem when faced with a complex real-world situation (cite behavioral econ or simon).
@@
 
-How should an early farmer make a decision in light ot this uncertainty?
+How should an early farmer make a decision in light of this uncertainty?
 The basic logic of crop diversitication is Portfolio of crops ala modern portfolio theory: Blank2001
@@
 If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility.
-If the probability distributin of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
+If the probability distribution of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
 In the language of game theory, this strategy is known as a "game of ficticious play" against nature.
 This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
+
 This strategy works well when the environment is stationary -- nature plays from a fixed probability distirubtion, but is vulnerable to environmental nonstationarity.
-That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because you're too stuck in your ways while the world changes around you.
+That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because as agents get too stuck in their ways while the world changes around them.
 Thus it is important to understand the dynamics of risk perception -- how subjective risks, and thus subjective expected utility -- rise and fall in uncertain environments.
 
-# Holocene drought risks in the eastern Mediterranean {#sec:2}
+# Modeling perceived risks with Bayesian agents {#sec:3}
+
+In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
+An individual's perceived risk entails more than simply the "objective" risk in the external environment [@Tucker2013].
+A person's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
+Likewise, the social responses to risk will refelct the distribution of perceived risks in a population, and by extension the collective memory of past droughts will have consequences for the population's survival and evolution.
+
+We can capture the influences of risk perception computationally by simulating a population of "belief-based agents."
+These agents develop beliefs about the risk of drought in their environment through personal experience. 
+Each agent's perceived drought risks will change their subjective expected crop yields from planting different ratios of wheat to barley.
+
+We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 ???double check).
+Bayesian learning allows for a minimal representation of the environment, so that each agent's beliefs can be encoded in two parameters -- defining the Beta distribution over their prior beliefs -- rather than the record of all observed years.
+Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+
+Our brains do not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
+Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
+
+We can treat this problem elegantly using Bayesian statistics using online Bayesian learning from bernoulli observations (Bissiri 2010).
+Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
+bernoulli distribution) with parameter theta reprenting the risk of drought.
+Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1.
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief \@ref(beta-prior).
+What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
+As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
+
+```{r beta-prior, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
+prior <- function(m, n) {
+  domain <- seq(0,1,0.005)
+  val <- dbeta(domain, n * m, n * (1 - m))
+  tibble('x' = domain, 'y' = val)
+}
+
+tibble(Age = seq(5, 50, 5),
+       risk = 0.5) %>%
+  mutate(belief = map2(risk, Age, ~prior(.x, .y))) %>%
+  unnest(cols = c(belief)) %>%
+  ggplot(aes(x, y)) +
+  geom_line(aes(group = Age, color = Age), size = 1.2, alpha = .8) +
+  scale_color_distiller(palette = 'PuOr', guide = 'legend') +
+  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
+  theme_classic()
+```
+
+# Drought risks in the eastern Mediterranean {#sec:2}
 
@@
 TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present [@He2011].
-It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of climate forcings (e.g.
+It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of external climate drivers (e.g.
 orbit, greenhouse gasses, glacial meltwater flux).
@@
 
-TraCE-21k simulation outputs for the past 10,000 years were extracted from the grid cell overlaying the northern Levant and southeastern Turkey.
+TraCE-21k simulation outputs for the past 10,000 years were extracted from the grid cell covering the northern Levant and southeastern Turkey.
 This location was selected due to its proximity to archaeological sites with comparative evidence of crop diversification strategies.
@@
 This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a challenge to simple risk-managing strategies that assume climatic risks are fixed.
+
+Drought risks ranged from 1 in every 10 to 5 in 10 over the Holocene.
 
@@
   geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
-  geom_col()  +
+  geom_col(width = 38)  +
   scale_fill_distiller(palette = 'PuOr') +
@@
 
-# The dynamics of risk perception {#sec:3}
+The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
+8ka BP (Figure 5).
+The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
 
-In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
-An individual's perceived risk entails more than simply the "objective" risk in the external environment [@Tucker2013].
-A person's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
-Likewise, the social responses to risk will refelct the distribution of perceived risks in a population, and by extension the collective memory of past droughts will have consequences for the population's survival and evolution.
-
-We can capture the influences of risk perception computationally by simulating a population of "belief-based agents."
-These agents develop beliefs about the risk of drought in their environment through personal experience. 
-Each agent's perceived drought risks will change their subjective expected crop yields from planting different ratios of wheat to barley.
-
-## Bayesian brains
-We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 double check).
-Bayesian learning allows for a minimal representation of the environemnt, so that each agent's beliefs can be encoded in two parameters -- defining the Beta distribution over their prior beliefs -- rather than the record of all observed years.
-Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
-Our brains do not record every bit of perceived informaiton in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
-Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
-
-```{r, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
-Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
-prior <- function(m, n) {
-  domain <- seq(0,1,0.005)
-  val <- dbeta(domain, n * m, n * (1 - m))
-  tibble('x' = domain, 'y' = val)
-}
-
-tibble(Age = seq(5, 50, 5),
-       risk = 0.5) %>%
-  mutate(belief = map2(risk, Age, ~prior(.x, .y))) %>%
-  unnest(cols = c(belief)) %>%
-  ggplot(aes(x, y)) +
-  geom_line(aes(group = Age, color = Age), size = 1.2, alpha = .8) +
-  scale_color_distiller(palette = 'PuOr', guide = 'legend') +
-  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
-  theme_classic()
-```
-
-### Bayesian learning from bernoulli observations
-We can treat this problem elegantly using bayesian statistics using online bayesian learning from bernoulli objersbations (Bissiri 2010).
-Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
-bernoulli distribution) with paramter theta reprenting the risk of drought.
-Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1 (Figure XX).
-We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief.
-What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
-As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
+The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
+Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
 
@@
 
-### The problem of nonstationarity
-That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
-The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
-In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
-Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
-Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
+# Bayesian learning and the dynamics of risk perception in the Neolithic {#sec:4}
 
@@
 
-So what are the consequences of these cognitive processes for early farmers in the Mediterranean.
+So what are the consequences of these cognitive processes for early farmers in the Mediterranean?
+
+Simulate a time series with varying drought frequency, based off the empirical drought risks derived from the paleoclimate simulation.
+
+This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the firstplace.
+The law of small numbers is very important here. The range of each person's experience is so small -- every individual's perceived risk will be a small and noisy sample of the true value.
+Older people are much more certain, even though they are wrong.
+
+That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
+The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
+In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
+Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
+Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
+
+That is, the above figure assumes a noise free environment with a single learner. Next we can look at a population of learners in a noisy environment
+
+Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater than variability due to differences in risk perception (Figure 7).
+Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
+Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
+^^ change above to talk about individual uncertainty in drought risk is often greater than the actual variability in drought risk, given the limited life experience an individual draws from
+
+During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
+Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
+This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
+This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
 
@@
                  crop = 'barley')
-```
 
-```{r}
 learn <- function(agents, drought) {
@@
 }
-```
 
-Simulate a time series with varying drought frequency. This is a stepwise nonstationary process. Setup an experiment exploring 6 scenarios.
-
-This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the firstplace.
-The law of small numbers is very important here. The range of each person's experience is so small -- every individual's perceived risk will be a small and noisy sample of the true value.
-Older people are much more certain, even though they are wrong.
-
-```{r}
 make_droughts <- function() c(runif(25) < 0.2, runif(25) < 0.4, runif(25) < 0.2)
@@
 droughts <- make_droughts()
-```
 
-```{r}
 run_sim <- function(droughts) {
@@
 ```{r, fig.height = 8, fig.width = 7}
-
 a <- ggplot(sim1) +
@@
 
-# Discussion and Conclusions
-
-The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure 5).
-The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
-
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
-Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
+# Conclusion
 
@@
 Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
-
-Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater variability due to differences in risk perception (Figure 7).
-Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
-Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
-
-During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
-Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
-This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
-This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
 

### 2020-02-20 manuscript/manuscript.Rmd
working figure crossref

@@
 # bibstyle options spbasic(default), spphys, spmpsci
-output: rticles::springer_article
+output:
+  bookdown::pdf_book:
+    base_format: rticles::springer_article
+    number_sections: yes
 header-includes:
@@
 
+Our brains do not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
+Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
+
 We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 ???double check).
@@
 
-Our brains do not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
-Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
-
-We can treat this problem elegantly using Bayesian statistics using online Bayesian learning from bernoulli observations (Bissiri 2010).
-Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
-bernoulli distribution) with parameter theta reprenting the risk of drought.
-Uncertainty in the value of theta can be represented with a beta distribution, constrained between 0 and 1.
-We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief \@ref(beta-prior).
-What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely.
-As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of experience.
-
-```{r beta-prior, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+```{r prior, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
 Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
@@
 
-# Drought risks in the eastern Mediterranean {#sec:2}
+We can treat this problem elegantly using Bayesian statistics using online Bayesian learning from Bernoulli observations (Bissiri 2010).
+Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
+Bernoulli distribution) with parameter $\theta$ representing the risk of drought.
+Uncertainty in the value of $\theta$ can be represented with a Beta distribution falling between 0 and 1.
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief (Figure \@ref(fig:prior)).
+What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely).
+As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of past experience.
 
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from outputs of the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research.
+# Holocene drought risk in the eastern Mediterranean {#sec:2}
+
+Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research.
 TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present [@He2011].
@@
 
-```{r, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
+```{r risk, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
 The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
@@
 The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure 5).
+8ka BP (Figure \@ref(fig:risk)).
 The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
 
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure 6).
+The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure \@ref(fig:risk)).
 Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
@@
 
-# Bayesian learning and the dynamics of risk perception in the Neolithic {#sec:4}
+# The dynamics of risk perception {#sec:4}
 
@@
 ```
+
+What's important to remember here is that both agents are acting rationally based on their prior subjective beliefs, they only differ in those beliefs.
+So the conservatism of the older agent and the flexibility of the yonger agent are not wrong per se, they simply reflect differet rational approaches to the world.
 

### 2020-02-21 manuscript/manuscript.Rmd
lock trace section, reorder prior1 and prior2 (again)

@@
 
-# Modeling perceived risks with Bayesian agents {#sec:3}
+# Early to mid-Holocene drought risks in the eastern Mediterranean {#sec:2}
+
+In order to estimate drought risks in the past, we cannot rely on present-day observations alone.
+Not only does precipitation vary from year-to-year, but it is also subject to centennial and millennial-scale trends and cycles throughout the Holocene that are unresolved in the observational period.
+Estimating not only the risks of drought in the eastern Mediterranean is therefore not sufficient, and this information must be supplemented with estimates of the *change* in drought risks over time.
+Climate dynamics are nonlinear, nonstationary, and nonergodic, which means sudden, unpredictable variability is the norm in the region rather than the exception.
+Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (e.g.
+the mean and variance of dry years) but also the higher order patterns such as the serial perisistence of wet and dry years.
+This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
+
+## Paleoclimate simulation
+
+Estimates of changing Holocene drought risks were derived from the TraCE-21ka paleoclimate simulation [@He2011].
+TraCE-21ka is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gas concentrations from the Last Glacial Maximum to the present.
+The simulation generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of external climate drivers (e.g.
+orbit, greenhouse gasses, glacial meltwater flux).
+The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
+
+Monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
+This location was selected to sample climate patterns typical for key archaeological sites in the region such as Çatalhöyük and Aşıklı Höyük, and for the eastern Mediterranean more broadly.
+```{r}
+konya <- matrix(c(32.5, 37.9), nrow = 1) # point at which to sample TraCE
+```
+
+```{r}
+# convective precipitation
+precc <- map(
+  list.files('../Data/PRECC', full.names = TRUE)[-12],
+  ~ brick(.) %>%
+    raster::extract(konya, df = TRUE) %>%
+    select(-ID) %>%
+    gather(time, precc)
+) %>%
+  bind_rows()
+
+# large-scale precipitation
+precl <- map(
+  list.files('../Data/PRECL', full.names = TRUE)[-12],
+  ~ brick(.) %>%
+    raster::extract(konya, df = TRUE) %>%
+    select(-ID) %>%
+    gather(time, precl)
+) %>%
+  bind_rows()
+```
+
+## Estimating drought risks
+
+Using the climate model output, I divided each model year into dry years and normal years.
+A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a normal year was defined as any year above this threshold.
+Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\theta_{empirical}$for any particular year was defined as the proportion of the previous 50 years that were dry year as
+
+\begin{equation}
+    \theta_{empirical} = \frac{\sum_{n=t-1}^{t-50} P_n < 300}{50},
+\end{equation}
+
+where $P_n$ is the growing season precipitation accumulation in millimeters for year $n$.
+
+```{r}
+sim_past <- left_join(precc, precl, by = 'time') %>%
+  mutate(
+    rainfall = (precc + precl) * 2.628e+9, # unit conversion
+    year = rep(-10200:-5001, each = 12),
+    # add hardcode dates instead of using ncdf time because of date shift with averaging period
+    month = rep(1:12, times = 5200)
+  ) %>%
+  select(-time,-precl,-precc) %>%
+  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>%
+  filter(!(month %in% c(6:9))) %>%
+  group_by(water_year) %>%
+  add_tally() %>%
+  filter(n == 8) %>%
+  summarise(precip = sum(rainfall)) %>%
+  mutate(drought = precip < 300) %>%
+  filter(between(water_year, -9500, -5500)) %>%
+  mutate(time_step = ceiling(water_year / 50),
+         year = time_step * 50) %>%
+  group_by(year) %>%
+  summarise(risk = sum(drought) / 50) %>%
+  filter(year != min(year))
+
+sim_past %>% summary
+```
+
+The simulated risks of crop failure due to drought ranged between 10\% and 46\% during the period from 9.5ka to 5.5ka, with a median risk of 24\% (Figure \@ref(fig:risk)).
+This means that, on average, a Neolithic farmer in Central Anatolia could expect their wheat crops to fail two or three times a decade, punctuated by even drier periods in which wheat crops could be expected to fail roughly every other year.
+The simulation also reveals a long-term trend of decreasing drought risks, in particular with higher drought risk in the early Holocene giving way to lower drought risk in the midle Holocene.
+
+```{r risk, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought aggregated by fifty-year period.
+The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields (after Figure 1), given the payoffs defined in Table 1.'}
+sim_past %>%
+  ggplot(aes(year, risk, fill = risk, color = risk)) +
+  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
+  geom_col(width = 38)  +
+  scale_fill_distiller(palette = 'PuOr') +
+  scale_color_distiller(palette = 'PuOr')+
+  theme_minimal() +
+  labs(x = 'Years BP', y = 'Drought risk') + theme(legend.position = "none")
+```
+
+The TraCE-21ka simulation confirms that drought risk in the eastern Mediterranean was non-stationary and, in fact, quite volatile during periods of climatic disruption in the early Holocene.
+This volatility would have had severe consequences for early farming communities whose risk-managing practices depended so heavily on accurately perceived local climatic risks.
+
+# Modeling risk perceptions with Bayesian agents {#sec:3}
 
 In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
-An individual's perceived risk entails more than simply the "objective" risk in the external environment [@Tucker2013].
-A person's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
+An individual's perceived risk entails more than simply the "objective" empirical risk in the local environment [@Tucker2013].
+A individual's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
 Likewise, the social responses to risk will refelct the distribution of perceived risks in a population, and by extension the collective memory of past droughts will have consequences for the population's survival and evolution.
 
-We can capture the influences of risk perception computationally by simulating a population of "belief-based agents."
-These agents develop beliefs about the risk of drought in their environment through personal experience. 
-Each agent's perceived drought risks will change their subjective expected crop yields from planting different ratios of wheat to barley.
+We can capture the influences of risk perception computationally by simulating a population of "belief-based agents." These agents develop beliefs about the risk of drought in their environment through personal experience.
+Each agent's perceived drought risk will influences their decision-making by altering their subjective expected crop yields from planting different ratios of wheat to barley.
 
@@
 
+## Bayesian learning from Bernoulli observations
+
+Bayesian statistics provides an elegant solution to the compact encoding problem.
+Instead of recording each year in memory, a Bayesian agent can compactly store their beliefs in the form of a prior distirbution
 We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 ???double check).
@@
 Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+
+We can treat this problem elegantly using Bayesian statistics.
+(Bissiri 2010).
+The probability of a drought occuring in any given year is treated as a coin flip (i.e.
+Bernoulli distribution) with parameter $\theta$ representing the risk of drought.
+Uncertainty in the value of $\theta$ can be represented with a Beta distribution falling between 0 and 1.
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief (Figure \@ref(fig:prior)).
+What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely).
+As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of past experience.
 
@@
 
-We can treat this problem elegantly using Bayesian statistics using online Bayesian learning from Bernoulli observations (Bissiri 2010).
-Put simply, the probability of a drought occuring in any given year is treated as a coin flip (i.e.
-Bernoulli distribution) with parameter $\theta$ representing the risk of drought.
-Uncertainty in the value of $\theta$ can be represented with a Beta distribution falling between 0 and 1.
-We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief (Figure \@ref(fig:prior)).
-What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely).
-As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of past experience.
-
-# Holocene drought risk in the eastern Mediterranean {#sec:2}
-
-Estimates of drought risk in the eastern Mediterranean during the the past 10,000 years were derived from the TraCE-21k paleoclimate simulation available on the National Center for Atmospheric Research.
-TraCE-21K is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model (GCM) to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gasses over the past 22,000 years from the Last Glacial Maximum to the present [@He2011].
-It generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of external climate drivers (e.g.
-orbit, greenhouse gasses, glacial meltwater flux).
-The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
-
-TraCE-21k simulation outputs for the past 10,000 years were extracted from the grid cell covering the northern Levant and southeastern Turkey.
-This location was selected due to its proximity to archaeological sites with comparative evidence of crop diversification strategies.
-This region includes some of the earliest evidence of crop domestication, including sites like Jericho and Can Hasan.
-
-Using the climate model output, I divided each model year into dry years and wet years.
-A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a wet/normal year was defined as any year above this threshold.
-Given the modeled patterns of wet and dry years, drought risk for any particular year was defined as the proportion of the previous 50 years that were dry years.
-Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (mean, variance of dry years) but also the higher order dynamics like serial autocorrelation in the perisistence of wet and dry years.
-This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a challenge to simple risk-managing strategies that assume climatic risks are fixed.
-
-Drought risks ranged from 1 in every 10 to 5 in 10 over the Holocene.
-
-```{r}
-konya <- matrix(c(32.5, 37.9), nrow = 1)
-#konya <- matrix(c(33.5138, 36.2765), nrow = 1)
-```
-
-```{r, cache = TRUE}
-precc <- map(list.files('../Data/PRECC', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precc)) %>%
-  bind_rows()
-precl <- map(list.files('../Data/PRECL', full.names = TRUE)[-12],
-             ~brick(.) %>%
-               raster::extract(konya, df = TRUE) %>%
-  select(-ID) %>%
-  gather(time, precl) ) %>%
-  bind_rows()
-
-sim_past <- left_join(precc, precl, by = 'time') %>%
-  mutate(rainfall = (precc + precl) * 2.628e+9,
-         year = rep(-10200:-5001, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
-         month = rep(1:12, times = 5200)) %>%
-  select(-time, -precl, -precc) %>%
-  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
-  filter(!(month %in% c(6:9))) %>%
-  group_by(water_year) %>%
-  add_tally() %>%
-  filter(n == 8) %>%
-  summarise(precip = sum(rainfall)) %>%
-  mutate(drought = precip < 300)
-```
-
-```{r risk, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period.
-The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
-sim_past %>%
-  filter(between(water_year, -9500, -5500)) %>%
-  mutate(time_step = ceiling(water_year / 50),
-         year = time_step * 50) %>%
-  group_by(year) %>%
-  summarise(risk = sum(drought) / 50) %>%
-  filter(year != min(year)) %>%
-  ggplot(aes(year, risk, fill = risk, color = risk)) +
-  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
-  geom_col(width = 38)  +
-  scale_fill_distiller(palette = 'PuOr') +
-  scale_color_distiller(palette = 'PuOr')+
-  theme_minimal() +
-  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
-```
-
-The bias-corrected wet season precipitation outputs from TraCE-21k show a relatively stationary distribution over the past 10,000 years, with slightly drier conditions during the early-mid Holocene transition at c.a.
-8ka BP (Figure \@ref(fig:risk)).
-The utility of a diversified wheat-barley crop is immediately apparent; rainfall often fails to meet the 300mm threshold for a productive wheat crop during the entire 10,000 year span, but drops below the threshold for barley only a handful of years.
-
-The risk of crop failure due to drought in any given year varies between 10\% and 20\%, punctuated by one to four century-long events where risk drops as low as 6\% and or exceeds 25\% (Figure \@ref(fig:risk)).
-Centuries of high drought risk cluster around known periods of regional and global climate deterioration, such as at 8.6ka BP and 4.2ka BP.
-
 A simple "fixed" bayesian learner thus gets more certain with age.
@@
 
-# The dynamics of risk perception {#sec:4}
-
-```{r, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
+```{r update, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
 prior2 <- function(a, b) {
@@
 ```
+
+# Risk management and the dynamics of risk perception {#sec:4}
+
+Its important to note that, although these perceptions can be biased away from the true climatic drought risk, they still reflect perfectly rational beliefs.
 

### 2020-02-25 manuscript/manuscript.Rmd
update cls file

@@
 keywords:
+- belief-based agents
 - drought risk
-- belief-based agents
 - crop diversification
@@
 bibliography: bibliography.bib
-bibstyle: spbasic
+bibstyle: Formato_bibliografia_Springer
 # bibstyle options spbasic(default), spphys, spmpsci
@@
 They needed to maintain stable crop yields in face of uncertain drought risks.
-Yield stability is very important for traditional agriculturlal commmunities not connected to larger bulk food networks (Abbo et al 2010)
+Yield stability is very important for traditional agriculturlal communities not connected to larger bulk food networks (Abbo et al 2010)
 neolithic crop package -- weiss zohary 2011
@@
 
-1. How likely are droughts to occur each year in the eastern Mediterreanean, and how did these risks change over the Holocene?
-2. How might Neolithic farmers have percieved these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
+1. How likely are droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
+2. How might Neolithic farmers have perceived these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
 
@@
 
-# Early to mid-Holocene drought risks in the eastern Mediterranean {#sec:2}
+# Early to mid-Holocene drought risks {#sec:2}
 
@@
 Estimating not only the risks of drought in the eastern Mediterranean is therefore not sufficient, and this information must be supplemented with estimates of the *change* in drought risks over time.
-Climate dynamics are nonlinear, nonstationary, and nonergodic, which means sudden, unpredictable variability is the norm in the region rather than the exception.
+Climate dynamics are non-linear, non-stationary, and non-ergodic, which means sudden, unpredictable variability is the norm in the region rather than the exception.
 Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (e.g.
-the mean and variance of dry years) but also the higher order patterns such as the serial perisistence of wet and dry years.
+the mean and variance of dry years) but also the higher order patterns such as the serial persistence of wet and dry years.
 This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
@@
 A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a normal year was defined as any year above this threshold.
-Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\theta_{empirical}$for any particular year was defined as the proportion of the previous 50 years that were dry year as
+Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\hat\theta$for any particular year was defined as the proportion of the previous 50 years that were dry year as
 
 \begin{equation}
-    \theta_{empirical} = \frac{\sum_{n=t-1}^{t-50} P_n < 300}{50},
+    \hat\theta = \frac{\sum_{n=t-1}^{t-50} P_n < 300}{50},
 \end{equation}
@@
   filter(year != min(year)) 
-
-sim_past %>% summary
 ```

### 2020-02-25 manuscript/manuscript.Rmd
prior description

@@
     year = rep(-10200:-5001, each = 12),
-    # add hardcode dates instead of using ncdf time because of date shift with averaging period
+    # hardcode dates instead of using ncdf time because of date shift with averaging period
     month = rep(1:12, times = 5200)
@@
 This means that, on average, a Neolithic farmer in Central Anatolia could expect their wheat crops to fail two or three times a decade, punctuated by even drier periods in which wheat crops could be expected to fail roughly every other year.
-The simulation also reveals a long-term trend of decreasing drought risks, in particular with higher drought risk in the early Holocene giving way to lower drought risk in the midle Holocene.
+The simulation also reveals a long-term trend of decreasing drought risks, in particular with higher drought risk in the early Holocene giving way to lower drought risk in the middle Holocene.
 
@@
 
-# Modeling risk perceptions with Bayesian agents {#sec:3}
+# Modeling risk perception {#sec:3}
 
-In order to manage risk, decision-makers must be perceive what exactly that risk is in the local environment.
-An individual's perceived risk entails more than simply the "objective" empirical risk in the local environment [@Tucker2013].
-A individual's perceived risk depends in large part on their past experience, and how those experiences shape their expectations for the future.
-Likewise, the social responses to risk will refelct the distribution of perceived risks in a population, and by extension the collective memory of past droughts will have consequences for the population's survival and evolution.
+In order to properly manage drought risk, a farmer must first be able to perceive that risk.
+Yet, a farmer's perception of risk reflects more than just the objective, empirical risk observable in the world around them [@Tucker2013].
+Individual risk perception is inherently subjective -- influenced by a person's past experience of dry and wet years as filtered through memory -- and can reflect varying levels of uncertainty.
+Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks.
 
-We can capture the influences of risk perception computationally by simulating a population of "belief-based agents." These agents develop beliefs about the risk of drought in their environment through personal experience.
-Each agent's perceived drought risk will influences their decision-making by altering their subjective expected crop yields from planting different ratios of wheat to barley.
+But how best to model risk perception at the individual level?
+The human brain does not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decision-making.
+The tools of Bayesian statistics provides an elegant solution to the compact encoding problem.
+Instead of recording detailed memories of each year, a "Bayesian agent" can compactly store their beliefs about the world around them in the form of a prior probability distribution.
+These agents thus develop beliefs about the risk of drought in their environment through their personal experience of the weather, and their perceptions go on to impact their successive decision-making by altering their subjective expected crop yields from planting different crop mixes.
 
-Our brains do not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decisions.
-Even if individuals aren't -- consciously or unconsciously --- making these calculations -- the basic algorithmic problem faced by the brain, and solutions it has evolved, reflect the same basic sets of constraints (Sanborn and Chater 2016).
+## Prior beliefs and Bayesian Agents
 
-## Bayesian learning from Bernoulli observations
+It is possible to capture the influence of risk perception on decision making computationally by simulating a population of "belief-based" Bayesian agents.
+A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature.
 
-Bayesian statistics provides an elegant solution to the compact encoding problem.
-Instead of recording each year in memory, a Bayesian agent can compactly store their beliefs in the form of a prior distirbution
-We can efficiently simulate these population-level risk perceptions in a population of agents with Bayesian brains (Kahvalti et al 2019 ???double check).
-Bayesian learning allows for a minimal representation of the environment, so that each agent's beliefs can be encoded in two parameters -- defining the Beta distribution over their prior beliefs -- rather than the record of all observed years.
-Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+Here, the probability of a drought occurring in any given year can be treated as a coin flip (i.e.
+a draw from Bernoulli distribution) with parameter $\theta$ representing the drought probability.
+The beta distribution is a natural choice for representing knowledge about probabilities because it is constrained to fall between 0 and 1.
+Hence, an individual agent's prior belief about the value of $\theta$ can be represented as
+\begin{equation}
+  \theta \sim \textrm{Beta}(\alpha, \,\beta),
+\end{equation}
+a beta distribution with the parameters $\alpha$ and $\beta$ corresponding to the number of dry and wet years previously experienced by that agent.
+Varying these two parameters thus allows us to represent a variety of different personal experiences of drought risk.
+For example, if an agent recalls having lived through 5 dry years and 25 normal years, their prior belief about the chance of a drought occurring in the following year would be represented as a Beta(5, 25) distribution with mean value 0.2 equivalent to the empirical drought risk for that period, $\hat\theta$.
 
-We can treat this problem elegantly using Bayesian statistics.
-(Bissiri 2010).
-The probability of a drought occuring in any given year is treated as a coin flip (i.e.
-Bernoulli distribution) with parameter $\theta$ representing the risk of drought.
-Uncertainty in the value of $\theta$ can be represented with a Beta distribution falling between 0 and 1.
-We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of the prior belief (Figure \@ref(fig:prior)).
-What we see here is how an indiviudal's perception of the risk of drought may start off diffuse at first (here as an uniformative flat prior in which all values of theta are equally likely).
-As each year of experience is observed, the agent updates its beliefs accordingly, balancing the information in this year's observation with the cumulative weight of past experience.
+Using probabilities to represent agents' beliefs also allows for estimation of the *uncertainty* in those beliefs (Figure \@ref(fig:prior-prob)).
+For example, an agent who experienced 5 wet and 10 normal years and one who experienced 25 wet and 50 normal years would both agree that, on average, droughts occur 50\% of the time.
+Yet the latter agent would be much more certain in this belief because it is drawn from a larger range of experience (i.e.
+a larger sample size).
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of this prior belief.
+An agents perception of drought risk by start off diffuse at first -- represented as an uninformative flat prior in which all values of $\theta$ are equally plausible.
+As each year's weather is observed, the agent updates its beliefs accordingly by balancing the information in this year's observation with the cumulative weight of their past experience.
 
-```{r prior, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+```{r prior-prob, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
 Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
@@
 
-A simple "fixed" bayesian learner thus gets more certain with age.
+## Individual learning and Bayesian updating
+
+A simple "fixed" Bayesian learner thus becomes more certain in their beliefs with age.
+
 For young agents with weak priors, the information of each new year can strongly influence their beliefs.
 For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
-As was the case with the simple game of ficitious play above, this strategy is fine when the environment is stable and risks do not change.
-But when there is volatility in the environment, and risks can change, it becomes important to change your mind.
+As was the case with the simple game of fictitious play discussed in Section \@ref(sec:1), this strategy is fine when the environment is stable and risks do not change.
+But when there is volatility in the environment, and risks can change, it becomes important to change one's mind.
 Being too inflexible in one's priors can lead being too optimistic when things have really changed for the worse, and alternately being too conservative when things finally do get better.
-So from a Bayesian learning standpoint, how should agents choose to believe new information.
 
-```{r update, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
+For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever lived through wet years when droughts happen on average 2 out of every ten years.
+
+online learning, so only ever +1 or 0 in each case
+
+Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+
+Even if individuals aren't -- consciously or unconsciously -- making these calculations, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same basic constraints (Sanborn and Chater 2016).
+ We can treat this problem elegantly using Bayesian statistics (Bissiri 2010). 
+(Kahvalti et al 2019 ???)
+
+```{r posterior, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
 prior2 <- function(a, b) {
@@
 What's important to remember here is that both agents are acting rationally based on their prior subjective beliefs, they only differ in those beliefs.
-So the conservatism of the older agent and the flexibility of the yonger agent are not wrong per se, they simply reflect differet rational approaches to the world.
+So the conservatism of the older agent and the flexibility of the younger agent are not wrong per se, they simply reflect different rational approaches to the world.
 
@@
 
-This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the firstplace.
+This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the first place.
 The law of small numbers is very important here. The range of each person's experience is so small -- every individual's perceived risk will be a small and noisy sample of the true value.
@@
 
-```{r, fig.height = 8, fig.width = 7}
+```{r dynamics, fig.height = 8, fig.width = 7, fig.cap='The dynamics of risk perception.'}
 a <- ggplot(sim1) +

### 2020-02-25 manuscript/manuscript.Rmd
posterior

@@
 It is possible to capture the influence of risk perception on decision making computationally by simulating a population of "belief-based" Bayesian agents.
-A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature.
+A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature (Kahvalti et al 2019 and others????).
 
@@
 Using probabilities to represent agents' beliefs also allows for estimation of the *uncertainty* in those beliefs (Figure \@ref(fig:prior-prob)).
+A simple Bayesian agent becomes more certain in their beliefs with age.
 For example, an agent who experienced 5 wet and 10 normal years and one who experienced 25 wet and 50 normal years would both agree that, on average, droughts occur 50\% of the time.
@@
 We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of this prior belief.
-An agents perception of drought risk by start off diffuse at first -- represented as an uninformative flat prior in which all values of $\theta$ are equally plausible.
-As each year's weather is observed, the agent updates its beliefs accordingly by balancing the information in this year's observation with the cumulative weight of their past experience.
+How, then, should an agent update its beliefs in light of new experience?
 
@@
 
-A simple "fixed" Bayesian learner thus becomes more certain in their beliefs with age.
+As a Bayesian agent moves through time and observes each successive year's weather, it must be able to update its beliefs accordingly.
+It does does by comparing the information in this year's observation with the cumulative weight of their past experience.
+The agent combines its prior beliefs about drought risk with the likelihood of observing a drought this year in order to generate a posterior distribution representing its updated beliefs about the world.
 
-For young agents with weak priors, the information of each new year can strongly influence their beliefs.
+Crucially, the strength of one's prior beliefs determines how much weight is given to new information.
+For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever experienced a moist climate where droughts happen on average two out of every ten years.
+The mean value of $\theta$ for both agents would thus be 20\%, but the degree of certainty varies because the older agent is basing this inference on many more years of experience.
+Now assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years (Figure \@ref(fig:posterior)).
+Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though each agent experienced the exact same objective drought risk*.
+
+For young agents with weak priors, the information of each new year can thus strongly influence their beliefs.
 For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
-As was the case with the simple game of fictitious play discussed in Section \@ref(sec:1), this strategy is fine when the environment is stable and risks do not change.
-But when there is volatility in the environment, and risks can change, it becomes important to change one's mind.
-Being too inflexible in one's priors can lead being too optimistic when things have really changed for the worse, and alternately being too conservative when things finally do get better.
-
-For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever lived through wet years when droughts happen on average 2 out of every ten years.
-
-online learning, so only ever +1 or 0 in each case
-
-Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
-
-Even if individuals aren't -- consciously or unconsciously -- making these calculations, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same basic constraints (Sanborn and Chater 2016).
- We can treat this problem elegantly using Bayesian statistics (Bissiri 2010). 
-(Kahvalti et al 2019 ???)
+As was the case with the simple game of fictitious play discussed in Section \@ref(sec:1), having conservative beliefs is fine when the environment is stable and risks do not vary.
+But when there is volatility in the environment, and risks can indeed change, the ability to change one's mind is crucial.
+Being too inflexible in one's priors can lead being too optimistic when things really have changed for the worse, or alternately being too conservative when conditions ultimately improve.
 
@@
 # Risk management and the dynamics of risk perception {#sec:4}
+
+online learning, so only ever +1 or 0 in each case
+
+This updated posterior distribution becomes the agent's prior for the following year, and the process repeats itself.
+Formally, this iterative process is known as "online" learning from Bernoulli observations, where "online" refers to the sequential, year-to-year updating [@Bissiri2010].
 
@@
 
+Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
+
+Even if individuals aren't -- consciously or unconsciously -- making these calculations, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same basic constraints (Sanborn and Chater 2016).
+
 Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.

### 2020-03-02 manuscript/manuscript.Rmd
online learning

@@
 If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility.
-If the probability distribution of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distirubtion of nature's moves.
-In the language of game theory, this strategy is known as a "game of ficticious play" against nature.
+If the probability distribution of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distribution of nature's moves.
+In the language of game theory, this strategy is known as a game of "fictitious play" against nature.
 This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
 
-This strategy works well when the environment is stationary -- nature plays from a fixed probability distirubtion, but is vulnerable to environmental nonstationarity.
+This strategy works well when the environment is stationary -- nature plays from a fixed probability distribution, but is vulnerable to environmental non-stationarity.
 That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because as agents get too stuck in their ways while the world changes around them.
@@
 \end{equation} 
-
 where $P_n$ is the growing season precipitation accumulation in millimeters for year $n$.
@@
   scale_color_distiller(palette = 'PuOr', guide = 'legend') +
-  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density') +
+  labs(x = expression('Perceived drought risk (' * theta * ')'), y = 'Density') +
   theme_classic()
@@
 
-## Individual learning and Bayesian updating
+## Bayesian updating and the weight of past experience
 
@@
 
-Crucially, the strength of one's prior beliefs determines how much weight is given to new information.
+Crucially, the strength of one's prior beliefs determines how much weight is given to new information (Figure \@ref(fig:posterior)).
 For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever experienced a moist climate where droughts happen on average two out of every ten years.
 The mean value of $\theta$ for both agents would thus be 20\%, but the degree of certainty varies because the older agent is basing this inference on many more years of experience.
-Now assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years (Figure \@ref(fig:posterior)).
-Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though each agent experienced the exact same objective drought risk*.
+Now assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years -- not an uncommon occurrence in the simulation of early to mid-Holocene climate.
+Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though both experienced the exact same climate*.
 
 For young agents with weak priors, the information of each new year can thus strongly influence their beliefs.
-For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year with the many decades of accumulated experience.
-As was the case with the simple game of fictitious play discussed in Section \@ref(sec:1), having conservative beliefs is fine when the environment is stable and risks do not vary.
+For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year's weather with previous decades' worth of accumulated experience.
+Although neither agent may perceive the "true" climatic drought risk exactly, they nevertheless reflect perfectly rational beliefs about the world.
+Both agents have perfectly rational beliefs and differ only in their prior subjective beliefs - their relative conservatism or flexibility are not biases but rather varying perspectives on an inherently uncertain world.
+
+As was the case in the game of "fictitious play" discussed in Section \@ref(sec:1), having conservative beliefs is fine when the environment is stable and risks do not vary.
 But when there is volatility in the environment, and risks can indeed change, the ability to change one's mind is crucial.
@@
 
-```{r posterior, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought risks in an older (50) and younger (15) agent before and after a 25-year dry period."}
+```{r posterior, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought probability in an older (50) and younger (15) agent before and after a 25-year dry period."}
 prior2 <- function(a, b) {
@@
         strip.placement = "outside") +
-  labs(x = expression('Drought probability (' * theta * ')'), y = 'Density')
+  labs(x = expression('Perceived drought risk (' * theta * ')'), y = 'Density')
 ```
@@
 
-online learning, so only ever +1 or 0 in each case
+While the above example helps establish some basic intuition for how agents with varying subjective beliefs can perceive risk differently, it represents an idealized situation where agents update their beliefs retrospectively after a many-year dry period.
+Farmers, on the other hand, must update their planting decisions each year and continuously monitor the weather around them.
+The transient, year-to-year changes in perceived risks can thus have major consequences for how a population responds *during* a dry period.
+How do Bayesian learners perform in such an uncertain, unpredictable time?
 
-This updated posterior distribution becomes the agent's prior for the following year, and the process repeats itself.
+To explore these dynamics we can simulate a population of Bayesian agents responding to a random sequence of wet and dry years.
+Here, observations are made about the world sequentially, and the agent must continuously update its beliefs accordingly.
+As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception risk.
+Now, this update occurs every year and thus only a year's worth of new information is incorporated into the agent's beliefs each time step.
+The learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
 Formally, this iterative process is known as "online" learning from Bernoulli observations, where "online" refers to the sequential, year-to-year updating [@Bissiri2010].
-
-Its important to note that, although these perceptions can be biased away from the true climatic drought risk, they still reflect perfectly rational beliefs.
-
-What's important to remember here is that both agents are acting rationally based on their prior subjective beliefs, they only differ in those beliefs.
-So the conservatism of the older agent and the flexibility of the younger agent are not wrong per se, they simply reflect different rational approaches to the world.
-
-So what are the consequences of these cognitive processes for early farmers in the Mediterranean?
-
-Simulate a time series with varying drought frequency, based off the empirical drought risks derived from the paleoclimate simulation.
-
-This suggests that people are more in danger of not realizing conditions are improving when they actually have -- the opportunity cost -- rather than failing to realize the change in the first place.
-The law of small numbers is very important here. The range of each person's experience is so small -- every individual's perceived risk will be a small and noisy sample of the true value.
-Older people are much more certain, even though they are wrong.
-
-That is, what if the risk of a drought, theta, is not a constant but rather varies unpredictably in time?
-The statistical problem shifts then to one of change detection (o'reilly 2013, gallistel et al 2014).
-In other words do people simply adapt to changes -- as in the fixed bayesian learner and moving average approaches -- or do they detect them?
-Strict year-to-year updating can result in overly conservative risk perceptions, suggesting people are slow to change their beliefs when in reality they often quickly update them (Bissiri 2010).
-Indeed, the "forgetful bayesian brain" model seems to perform better, with a dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).
-
-That is, the above figure assumes a noise free environment with a single learner. Next we can look at a population of learners in a noisy environment
-
-Examining the long-term dynamics of crop yields among all decision-making strategies and psychological profiles reveals variability in yields due to different decision making preferences is far greater than variability due to differences in risk perception (Figure 7).
-Risk-minimizing strategies sacrifice productivity for predictability, consistently lowering the mean and variance of crop yields over time.
-Risk-minimizing strategies are more sensitive to risk perception than yield-maximizing strategies; varying the memory decay rate parameter $\lambda$ had no effect on the optimal crop allocation for yield maximizers.
-^^ change above to talk about individual uncertainty in drought risk is often greater than the actual variability in drought risk, given the limited life experience an individual draws from
-
-During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
-Climatically-induced variability in food supplies consistently surpasses that from differences in risk-perception psychology, but not that from different risk aversion preferences.
-This finding suggests that the efficacy of risk managing strategies that rely on regulatory feedbacks is limited to periods of relative climatic stability.
-This is consistent with the observation from the Roman Period Mediterranean that while minor food crises were common in the ancient world, extreme famines were rare [@Garnsey1989].
+In this learning environment, isolating signal from noise becomes critical for accurately perceiving evolving drought risks.
 
@@
                  risk = alpha / (alpha + beta),
-                 #age = sample.int(80, size = 150, replace = TRUE),
                  crop = 'barley')
+```
 
+```{r}
 learn <- function(agents, drought) {
@@
          wheat_exp = expected_yield_wheat(risk),
-         expected_yield = barley_exp + wheat_exp,
          crop = case_when(barley_exp > wheat_exp ~ 'barley',
@@
   # nature's turn to decide realized yields
-    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
-           alpha = drought + alpha,
+    mutate(alpha = drought + alpha,
            beta = beta + 1 - drought,
-          # mu = alpha / (alpha + beta),
            risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
 }
+```
 
+```{r}
 make_droughts <- function() c(runif(25) < 0.2, runif(25) < 0.4, runif(25) < 0.2)
@@
   theme_classic() +
-  labs(x = 'years', y = 'Perceived drought risk')
+  labs(x = 'Year', y = 'Perceived drought risk')
 
@@
   geom_bar(width = 1, stat = 'identity', position = 'fill') +
-  #scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
-  #  scale_fill_manual(values = c('#e66101', '#5e3c99'), name = '') +
     scale_fill_manual(values = c('#f1a340', '#998ec3'), name = '') + 
     scale_color_manual(values = c('#f1a340', '#998ec3'), name = '') + 
-    labs(x = 'Years', y = 'Wheat proportion') +
+    labs(x = 'Year', y = 'Wheat proportion') +
   theme_classic()
@@
 
-These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
+# Conclusion
 
-Note in these plots above and below we're summing all the wheat and barley in the population, compared to below which shows the wheat proportions of individuals.
-
-# Conclusion
+During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
 
@@
 Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
+
+1. drought risks varied between 10\% and 45\% throughout the Holocene
+2. risk perceptions are influenced  by individual life experience, which is necessarily limited and noisy
+3. people behave rationally given experience?
+
+1. How likely are droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
+2. How might Neolithic farmers have perceived these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
 
@@
 
+Examining the long-term dynamics of crop yields among all decision-making strategies and risk perceptions reveals variability in yields due to differences in risk perception outweigh variability due to changing objective drought risk alone.
+This occurs because of the limited life experience each individual is able to draw from.
+
 \setlength{\parindent}{-0.5in}

### 2020-03-03 manuscript/manuscript.Rmd
discussion

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, fig.width = 5, fig.asp = 0.618)
 library(raster)
@@
 The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a game against nature.
-We can represent this simplified decision context as a "game" in a game theorietic context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould 1963).
+We can represent this simplified decision context as a "game" in a game theoretic context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould 1963).
 The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years (Table 1).
@@
 We can think of this game as being the culturally-inherited object.
+So in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio.
 Our question is how do individuals make decisions here?
-So in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio.
 
@@
 
-```{r, fig.height = 3, fig.width = 5, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
+```{r, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
 expected_yield_wheat <- function(risk) {
@@
   ggplot(aes(risk, yield)) +
-  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .5)+
-  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .5) +
+  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .8)+
+  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .8) +
   geom_line(aes(color = type), size = 1.2) +
@@
 
-```{r risk, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought aggregated by fifty-year period.
+```{r risk, fig.asp = 0.3333, fig.cap = 'Annual risk of wheat crop failure due to drought aggregated by fifty-year period.
 The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields (after Figure 1), given the payoffs defined in Table 1.'}
@@
 Individual risk perception is inherently subjective -- influenced by a person's past experience of dry and wet years as filtered through memory -- and can reflect varying levels of uncertainty.
-Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks.
+Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks [@Moussaid2013].
 
@@
 A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature (Kahvalti et al 2019 and others????).
+This approach has a clear computational efficiency -- for both modelers and decision-makers -- and even if real-world decision makers are not Bayesians in a literal sense the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same constraints on information processing in the heads of decision-makers (Sanborn and Chater 2016).
 
@@
 
-```{r prior-prob, fig.width = 5, fig.height = 3, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+```{r prior-prob, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
 Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
@@
   ggplot(aes(x, y)) +
-  geom_line(aes(group = Age, color = Age), size = 1.2, alpha = .8) +
+  geom_line(aes(group = Age, color = Age), size = 1.2) +
   scale_color_distiller(palette = 'PuOr', guide = 'legend') +
@@
 Now assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years -- not an uncommon occurrence in the simulation of early to mid-Holocene climate.
-Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though both experienced the exact same climate*.
+Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though both experienced the same climate*.
 
@@
 
-```{r posterior, fig.width = 6, fig.height = 3, fig.cap = "Change in perceived drought probability in an older (50) and younger (15) agent before and after a 25-year dry period."}
+```{r posterior, fig.asp = 0.5, fig.cap = "Change in perceived drought probability in an older (50) and younger (15) agent before and after a 25-year dry period."}
 prior2 <- function(a, b) {
@@
   ggplot(aes(x, y)) +
-  geom_line(aes(group = time, color = time), size = 1.2, alpha = .8) +
+  geom_line(aes(group = time, color = time), size = 1.2, alpha = .9) +
 scale_color_manual(values = c('#5e3c99', '#e66101'), name = '') +
@@
       select(-prediction) %>%
-  # nature's turn to decide realized yields
+  # "nature's" turn to decide realized yields
     mutate(alpha = drought + alpha,
@@
 ```{r}
-make_droughts <- function() c(runif(25) < 0.2, runif(25) < 0.4, runif(25) < 0.2)
-
 set.seed(11111)
-droughts <- make_droughts()
+droughts <- c(runif(25) < 0.2, runif(25) < 0.4, runif(25) < 0.2)
 
@@
 
-```{r dynamics, fig.height = 8, fig.width = 7, fig.cap='The dynamics of risk perception.'}
+Now we can simulate a time series.
+As above, we'll use a scenario not uncommon in the paleoclimate simulation -- a baseline drought risk of 20\% punctuated by a 25 year dry period when risks rise to 40\%.
+The results of such a simulation reveal the importance of the subjective experiences of individual decision makers on the population-level perception of risk (Figure \@ref(fig:dynamics)a).
+
+Translating the uncertainty in drought risk to a particular crop diversification strategy (Figure \@ref(fig:dynamcs)b).
+
+```{r dynamics, fig.asp = 1.1, fig.cap='A: The dynamics of risk perception. B:'}
 a <- ggplot(sim1) +
-  geom_rect(aes(xmin = 25, xmax = 50, ymin = -Inf, ymax = Inf), fill = 'lightgray', alpha = .1) +
+  geom_rect(aes(xmin = 25, xmax = 50, ymin = -Inf, ymax = Inf), fill = 'lightgray', alpha = .2) +
   geom_point(aes(time, risk), alpha = .1) +
@@
 
-a / b
+a / b + plot_annotation(tag_levels = 'A')
 ```
@@
 
+In this chapter, I explored the consequences of individual heterogeneity in risk perception on the risk-management practices of a simulated Neolithic farming community in the eastern Mediterranean.
+I used a long-term paleoclimate simulation to estimate the changing risks of agricultural drought over a 4,000 year time period spanning the early to middle Holocene.
+Over this time span, the risk of a wheat crop failing from drought varied from once every ten years to nearly once every two years.
+Changes of such magnitude would have severely impacted Neolithic agroecosystems in the long run, but would have been difficult for any individual farmer to perceive in the short run.
+
+To explore these dynamics, I simulated a population of Bayesian "belief-based" agents who use their subjective perception of annual drought risks to decide what mix of crops will best manage those risks.
 During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
-But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” experience major food shortfalls.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” in light of their prior beliefs can experience food shortfalls.
+Cognitive diversity and life experience can be as or more important than the exact mix of crops planted for a population's long-run survival under extreme uncertainty.
 
-Given the average drought risks estimated from TraCE-21k and assuming perfect monitoring and recall of recent droughts (i.e.
-$\lambda = 0$), farmers intent on minimizing the risk of crop failure would be expected to plant about 65\% barley to 35\% wheat on average.
-This proportion is within the range estimated from macrobotanical remains at a nearby archaeological site [@Marston2011190].
-Yield maximizing farmers instead play a pure strategy profile, planting wheat if the perceived drought risk is less than 0.68 and barley if it is more.
-Because the GCM-simulated drought risk never rises above this threshold, a yield-maximizing farmer will always plant a wheat monocrop.
+These dynamics have implications for understanding risk management and food production in the Neolithic.
+Two points arise that this model helps clarify our understanding of the earliest farming communities in the Mediterranean and can inform future simulation work:
 
-1. drought risks varied between 10\% and 45\% throughout the Holocene
-2. risk perceptions are influenced  by individual life experience, which is necessarily limited and noisy
-3. people behave rationally given experience?
+1.
+**Risk perception is hard.** The climate system inherently chaotic.
+Annual forecasts are fundamentally uncertain, even in the era of modern supercomputers and numerical weather prediction.
+For prehistoric farmers, this uncertainty would have been a existential challenge.
+With the end of the Last Glacial Maximum and the advent of the Holocene, precipitation became increasingly volatile on multiple time scales.
+This uncertainty in the objective drought risk is only compounded by uncertainty in individual farmers' subjective perception of drought risk.
+Any individual would have experienced only a brief snapshot of this complex period.
+The rhythm of a single human lifespan is out of sync with the centennial to millennial scale oscillations in the climate system, so even the predictive value of one's own experience is itself unpredictable.
+This fundamental uncertainty would have influenced far more than the choice of which type of crops to plant -- and would have pervaded all kinds of decision-making under risk.
 
-1. How likely are droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
-2. How might Neolithic farmers have perceived these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
+2.
+**Individual risk perception has consequences for collective risk management.** Individuals of different ages may perceive the same dry period depending on their prior life experience.
+Younger individuals are more likely to perceive a run of dry years as a trend, rather than a temporary deviation from the norm, and older individuals  are likely to do the opposite.
+By extension, the age structure of a population will influence how quickly it is able to perceive and adapt to a changing climate.
+For example, a young, fast growing population will have a different collective memory of a past drought event then an older population.
+Likewise, baby booms and busts associated with famines, or any mortality crisis that afflicts specific age classes like warfare or epidemics, will alter the time horizon of that population's collective memory.
+Individual heterogeneity in risk perceptions can thus play a key role in broader social responses to climatic risks.
 
-Not only does this approach have a clear computational efficiency, but for the same reason this reflects the processes actually going on in the heads of decision makers.
-
-Even if individuals aren't -- consciously or unconsciously -- making these calculations, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same basic constraints (Sanborn and Chater 2016).
-
-Statistically downscaling palaeoclimate simulations and coupling them to game theoretic models of decision making under uncertainty is a simple way to better understand the patterning of climate changes and the regulatory feedbacks farmers use to adapt to them.
-To further explore these patterns, future work should incorporate additional risk-minimizing strategies such as storage and exchange.
-Food storage and exchange can also be thought of as forms of diversification, the former in the time and the latter in space [@Marston2011190].
-Food exchange would be of particular interest from a game theoretic perspective, as incomplete information with respect to an exchange partner's crop diversification and storage practices can lead to situations of moral hazard.
-
-Think about the importance of the population's age structure, becuase that's going to impact the collective memory of drought.
-
-Examining the long-term dynamics of crop yields among all decision-making strategies and risk perceptions reveals variability in yields due to differences in risk perception outweigh variability due to changing objective drought risk alone.
-This occurs because of the limited life experience each individual is able to draw from.
+Although social learning and cumulative cultural evolution has not been a focus of this paper (see [@Boyd and Richerson 1998, 2001]), these findings provide insight into the individual learning dynamics that underlay those broader social processes.
+Relating collective knowledge to individual cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007].
+This chapter has focused primarily on the physical and cognitive dimensions of risk and risk perception.
+The social context of risk and risk perception can be equally consequential [@Rogers1997].
+The balance of individual learning with social transmission determines the collective perception of risks and its impact on collective memory [@Moussaid2013, candia2019].
+Our finite lifespans ultimately keep the skill of individual learning low over the long term -- cumulative cultural evolution is required.
 

### 2020-03-05 manuscript/manuscript.Rmd
typos

@@
 authors: 
-- name: Nicolas Gauthier 1
+- name: Nicolas Gauthier
   address: Laboratory of Tree-Ring Research & School of Geography and Development, University of Arizona
@@
 
-Over the past 10,000 years, subsistence farmers in the Mediterranean basin have developed sophisticated strategies to maintain stable food supplies given uncertain rainfall.
-Their suite of strategies include practices like crop diversification, storage, and exchange [@Halstead1989]).
-These risk-management strategies all represent forms of \emph{regulatory feedback} used to manage complex socio-ecological systems.
-Regulatory feedbacks convert a system's high-variance inputs into low-variance outputs by dynamically monitoring the outputs and adjusting internal properties of the system [@Anderies2015].
-Risk-managing strategies require farmers to make specific resource allocation decisions in response to environmental risks, and as regulatory feedbacks they are thus sensitive to variability in both the spatiotemporal patterns of risk and the ability of farmers to perceive and act on those patterns.
+The distinctive climate and ecology of the Mediterranean basin afforded both challenges and opportunities to the earliest farming communities.
+Here, water is the primary limiting resource for traditional agroecosystems.
+Agricultural droughts -- where growing season precipitation is low enough to cause crop failures, are a constant threat.
+Precipitation is highly variable in space and time and droughts are difficult to predict with any certainty.
+How were Neolithic farmers able to adapt to, and even thrive in, such an uncertain environment?
 
-Crop diversification is an excellent example of a widespread and effective regulatory feedback strategy (Figure 1).
+Over the past 10,000 years, small-scale subsistence farmers have relied on a suite of strategies to maintain stable food supplies given uncertain rainfall.
+These strategies includes practices like crop diversification, storage, mobility, and exchange [@Halstead1989].
+Crop diversification in particular is an excellent example of a widespread and effective risk management strategy that is well suited to Mediterranean agroecosystems.
+In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,abbo2010,weiss zohary 2011,Marston2011190].
 Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@helmers2001separating,Anderies2006].
-In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,Marston2011190].
-Wheat generally has higher yields but is sensitive to water deficits, while barley has lower yields and is drought tolerant.
-Planting a mix of wheat and barley, either in the same plot or in a combination of plots, is an effective means of diversifying grain supplies.
-By adjusting the ratio of wheat to barley, farmers can adapt to changing drought risks.
+Wheat is high yielding but drought sensitive, while barley is lower yielding but drought tolerant.
+Planting a mix of high yield, high risk and low yield, low risk crops, either in the same plot or in a combination of plots, is an effective means of diversifying the annual supply of staple food crops [@paul et al 2019].
+By dynamically adjusting the ratio of wheat to barley in their fields, farmers can adapt to a variety local climate conditions with different drought risks.
 
-The secret to the success of earliest farmers on the mediterrenan littoral was in their manipulation of the local ecosystem to manage environmental risks.
-Early farmers constructed a complex agroecosystem to manage risk.
-They needed to maintain stable crop yields in face of uncertain drought risks.
-Yield stability is very important for traditional agriculturlal communities not connected to larger bulk food networks (Abbo et al 2010)
-neolithic crop package -- weiss zohary 2011
-cite abbo et al 2010? for more stuff about ladraces
-Wheat/barley systems
-example of system that combines low yield, low wirks crop with high yield high risk crop, an effective context paut et al 2019
-These strategies work when risks are knonw and constant, but may be more susceptible to periods of changing climatic risks
-so understanding how people of the past adapted to change is important to us today
+Risk-managing strategies like crop diversification require farmers to make specific resource allocation decisions in response to specific environmental risks, so are sensitive to variability in both the spatiotemporal patterns of risk and the ability of individual decision-makers to perceive and act on those patterns.
+Here, I focus on two main questions:
 
-Are these crop-diversification strategies vulnerable to the same dynamics as other social-ecological systems with similar regulatory feedbacks mechanisms?
-One way to address this is to model the influence of imperfect monitoring and biased decision making in uncertain environments.
-In this study, I accomplish this by answering two main questions:
+1. How likely were droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
 
-1. How likely are droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
-2. How might Neolithic farmers have perceived these changing drought risks, and what were the consequences for farming communities with respect to their collective ability to manage risk?
+2. How might Neolithic farmers have perceived these changing risks, and what were the consequences for Neolithic farmers' collective ability to manage them?
 
-I couple a climate model with high temporal resolution to a simple game theoretic model of agricultural decision making under uncertainty, in order to examine the performance of optimal and suboptimal wheat-barley diversification practices.
-I model the year-to-year crop diversification strategies made by farmers as an iterated game of fictitious play against nature [@GOULD1963a].
-Farmers are boundedly rational, in that they seek to maximize their objective functions given the perceived probabilities of different actions by nature, but work with flawed mental models of climate-related risks due to imperfect recall of past events.
-Furthermore, by using a climate model to represent the actions of nature, I can more precisely capture the characteristic autocorrelated patterns of rainfall variability, rather than simply drawing rainfall values from a static distribution.
-This framework allows me to address the sensitivity of diversification strategies to changes and climatic variability and imperfect human perceptions of those changes.
+To address these questions, I first use results from a long-term paleoclimate simulation to estimate the changing risk of agricultural droughts in the eastern Mediterranean over a 4,000 year period in the early to middle Holocene.
+Then, I use a simulated population of Bayesian "belief-based" agents to explore how well individual farmers would have been able to perceive these long-term changes in drought risks given their finite life experiences and limited capacity to process information.
+This computational approach allows for a more nuanced understanding of the vulnerability of risk-management strategies to unpredictable climatic variability.
 
@@
 
-The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a game against nature.
-We can represent this simplified decision context as a "game" in a game theoretic context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould 1963).
-The "game" in this context is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the future "state-of-nature", and nature varies between several possible states such as dry and wet years (Table 1).
-In reality continuous, but people often intuitively solve an easier problem when faced with a complex real-world situation (cite behavioral econ or simon).
+The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a "game against nature" [@LuceRaiffa1989, Milnor1952, Agrawal1968, cassidy1971, gould1963].
+The "game" in this context represents the farmer's decision of which crops to plant and in what proportions, given uncertainty in whether "nature" will be dry or wet in a given year (Table 1).
+Although  reality continuous, but people often intuitively solve an easier problem when faced with a complex real-world situation [@Simon1952].
 The exact values in the payoff matrix here are less important than the relative payoffs in each quadrant.
 We can think of this game as being the culturally-inherited object.
-So in this case the cultural norm would be to mix wheat and barley, but the learning is in the exact ratio.
+So in this case the cultural norm would be to mix wheat and barley, but individuals need to choose the precise ratio given their perception of risk.
 Our question is how do individuals make decisions here?
@@
   ggplot(aes(risk, yield)) +
-  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .8)+
-  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .8) +
+  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .6)+
+  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .6) +
   geom_line(aes(color = type), size = 1.2) +
@@
   theme_classic() +
-  #scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
   scale_color_manual(values = c('#e66101', '#5e3c99'), name = '') +
@@
 In the language of game theory, this strategy is known as a game of "fictitious play" against nature.
-This is an effective risk-managing strategy, but like all similar strategies that adapt to a specific pattern of variability, it is vulnerable to changes in the pattern of variability (Janssen eta al 2007)
+While this is can be an effective risk-management strategy, like all strategies adapted to a specific pattern of variability it is vulnerable to changes in the pattern of variability [@Janssen2007].
 
@@
 
-In order to estimate drought risks in the past, we cannot rely on present-day observations alone.
-Not only does precipitation vary from year-to-year, but it is also subject to centennial and millennial-scale trends and cycles throughout the Holocene that are unresolved in the observational period.
-Estimating not only the risks of drought in the eastern Mediterranean is therefore not sufficient, and this information must be supplemented with estimates of the *change* in drought risks over time.
-Climate dynamics are non-linear, non-stationary, and non-ergodic, which means sudden, unpredictable variability is the norm in the region rather than the exception.
+In order to estimate drought risks in the past, present-day weather observations alone are insufficient.
+Precipitation varies not only from year-to-year, but also on centennial to millennial-scale time scales that are unresolved in the contemporary observational record.
+Climate dynamics are non-linear, non-stationary, and non-ergodic, which means sudden, unpredictable variability is the norm rather than the exception.
+
 Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (e.g.
@@
 
-Estimates of changing Holocene drought risks were derived from the TraCE-21ka paleoclimate simulation [@He2011].
+Estimates of changing Holocene precipitaiton variability were derived from the TraCE-21ka paleoclimate simulation [@He2011].
 TraCE-21ka is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gas concentrations from the Last Glacial Maximum to the present.
 The simulation generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of external climate drivers (e.g.
-orbit, greenhouse gasses, glacial meltwater flux).
+Earth's orbit, greenhouse gasses, glacial meltwater flux).
 The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
@@
 A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a normal year was defined as any year above this threshold.
-Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\hat\theta$for any particular year was defined as the proportion of the previous 50 years that were dry year as
+Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\hat\theta$ for any particular year was defined as the proportion of the previous 50 years that were dry years:
 
@@
 The simulated risks of crop failure due to drought ranged between 10\% and 46\% during the period from 9.5ka to 5.5ka, with a median risk of 24\% (Figure \@ref(fig:risk)).
-This means that, on average, a Neolithic farmer in Central Anatolia could expect their wheat crops to fail two or three times a decade, punctuated by even drier periods in which wheat crops could be expected to fail roughly every other year.
+On average, a Neolithic farmer in Central Anatolia could expect their wheat crops to fail two or three times a decade, punctuated by even drier periods in which wheat crops could be expected to fail roughly every other year.
 The simulation also reveals a long-term trend of decreasing drought risks, in particular with higher drought risk in the early Holocene giving way to lower drought risk in the middle Holocene.
@@
 Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks [@Moussaid2013].
+But how best to model risk perception at the individual level?
 
-But how best to model risk perception at the individual level?
+## Prior beliefs and Bayesian agents
+
 The human brain does not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decision-making.
-The tools of Bayesian statistics provides an elegant solution to the compact encoding problem.
-Instead of recording detailed memories of each year, a "Bayesian agent" can compactly store their beliefs about the world around them in the form of a prior probability distribution.
-These agents thus develop beliefs about the risk of drought in their environment through their personal experience of the weather, and their perceptions go on to impact their successive decision-making by altering their subjective expected crop yields from planting different crop mixes.
+Integrating the tools of agent-based modeling and Bayesian probability provide an elegant means of representing this process.
+A Bayesian "belief-based" agent is an one whose subjective beliefs can be represented as a probability distribution over possible states of nature [@Kahvalti et al 2019 and others????].
+This approach has a clear computational efficiency -- for both modelers and decision-makers -- and even if real-world decision makers are not Bayesians in a literal sense, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same constraints on information processing in the minds of decision-makers [@Sanborn and Chater 2016].
 
-## Prior beliefs and Bayesian Agents
-
-It is possible to capture the influence of risk perception on decision making computationally by simulating a population of "belief-based" Bayesian agents.
-A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature (Kahvalti et al 2019 and others????).
-This approach has a clear computational efficiency -- for both modelers and decision-makers -- and even if real-world decision makers are not Bayesians in a literal sense the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same constraints on information processing in the heads of decision-makers (Sanborn and Chater 2016).
-
-Here, the probability of a drought occurring in any given year can be treated as a coin flip (i.e.
+We can examine the dynamics of risk perception using a population of Bayesian agents that develop beliefs about the risk of drought in their environment through their personal experience of the weather, and these perceptions impact their decision-making by altering their subjective expected crop yields from planting different crop mixes.
+Here, the probability of a drought occurring in any given year is treated as a coin flip (i.e.
 a draw from Bernoulli distribution) with parameter $\theta$ representing the drought probability.
 The beta distribution is a natural choice for representing knowledge about probabilities because it is constrained to fall between 0 and 1.
-Hence, an individual agent's prior belief about the value of $\theta$ can be represented as
+Hence, an individual agent's prior belief about the plausivle value of $\theta$ can be represented as
 \begin{equation}
@@
 a beta distribution with the parameters $\alpha$ and $\beta$ corresponding to the number of dry and wet years previously experienced by that agent.
-Varying these two parameters thus allows us to represent a variety of different personal experiences of drought risk.
-For example, if an agent recalls having lived through 5 dry years and 25 normal years, their prior belief about the chance of a drought occurring in the following year would be represented as a Beta(5, 25) distribution with mean value 0.2 equivalent to the empirical drought risk for that period, $\hat\theta$.
+Varying these two parameters thus allows for the representation of a variety of different personal experiences of drought risk.
+For example, if an agent recalls having lived through 5 dry years and 25 normal years, their prior belief about the chance of a drought occurring in the following year would be represented as a Beta(5, 25) distribution with mean value 0.2 equivalent to the empirical drought risk ($\hat\theta$) for that period.
 
@@
 ```{r prior-prob, fig.cap = "Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
-Beliefs are represented as Beta probabiity distributions, and the increased certainty with age reflects the varying effective sample size of the Beta prior."}
+Beliefs are represented as beta distributions, and the increased certainty with age reflects the varying effective sample size of the beta prior."}
 prior <- function(m, n) {
@@
 
-As a Bayesian agent moves through time and observes each successive year's weather, it must be able to update its beliefs accordingly.
+As a Bayesian agent must be able to update its beliefs as it moves through time and observes each successive year's weather.
 It does does by comparing the information in this year's observation with the cumulative weight of their past experience.
-The agent combines its prior beliefs about drought risk with the likelihood of observing a drought this year in order to generate a posterior distribution representing its updated beliefs about the world.
+The agent combines its prior beliefs about drought risk with the likelihood of having observing a drought in the current year in order to generate a posterior distribution representing its updated beliefs about the world.
 
 Crucially, the strength of one's prior beliefs determines how much weight is given to new information (Figure \@ref(fig:posterior)).
-For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever experienced a moist climate where droughts happen on average two out of every ten years.
-The mean value of $\theta$ for both agents would thus be 20\%, but the degree of certainty varies because the older agent is basing this inference on many more years of experience.
-Now assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years -- not an uncommon occurrence in the simulation of early to mid-Holocene climate.
+For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever experienced a normal climate where droughts happen on average two out of every ten years.
+The mean value of $\theta$ for both agents would thus be 20\%, but the degree of certainty varies because the older agent bases its inference on many more years of experience than the younger agent.
+Now, assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years -- not an uncommon occurrence in the simulation of early to mid-Holocene climate.
 Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though both experienced the same climate*.
@@
 For young agents with weak priors, the information of each new year can thus strongly influence their beliefs.
-For older agents who have experienced many more years their priors will be stronger, and they will be less likely to update their beliefs when balancing the information from a single year's weather with previous decades' worth of accumulated experience.
+But older agents -- having more experience and stronger priors -- will be less likely to update their beliefs when comparing the information from a single year's weather with previous decades' worth of accumulated experience.
 Although neither agent may perceive the "true" climatic drought risk exactly, they nevertheless reflect perfectly rational beliefs about the world.
@@
 
-As was the case in the game of "fictitious play" discussed in Section \@ref(sec:1), having conservative beliefs is fine when the environment is stable and risks do not vary.
-But when there is volatility in the environment, and risks can indeed change, the ability to change one's mind is crucial.
-Being too inflexible in one's priors can lead being too optimistic when things really have changed for the worse, or alternately being too conservative when conditions ultimately improve.
+As was the case in the game of "fictitious play" discussed in Section \@ref(sec:1), a degree of conservatism is warranted when the environment is stable and risks do not vary.
+But when the environment is volatile the ability to change one's mind is crucial.
+Being too inflexible in one's priors can lead being too optimistic when things really have changed for the worse, or too conservative when conditions ultimately improve.
 
@@
 
-While the above example helps establish some basic intuition for how agents with varying subjective beliefs can perceive risk differently, it represents an idealized situation where agents update their beliefs retrospectively after a many-year dry period.
+The example in the previous section established some basic intuitions for how agents with varying subjective beliefs can perceive risk differently.
+But it represents an idealized situation where agents update their beliefs retrospectively after a many-year dry period.
 Farmers, on the other hand, must update their planting decisions each year and continuously monitor the weather around them.
@@
 To explore these dynamics we can simulate a population of Bayesian agents responding to a random sequence of wet and dry years.
-Here, observations are made about the world sequentially, and the agent must continuously update its beliefs accordingly.
-As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception risk.
+Here, observations are made about the world sequentially, and the agent must continuously update its beliefs.
+As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception of risk.
 Now, this update occurs every year and thus only a year's worth of new information is incorporated into the agent's beliefs each time step.
-The learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
+
+This learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
 Formally, this iterative process is known as "online" learning from Bernoulli observations, where "online" refers to the sequential, year-to-year updating [@Bissiri2010].
@@
 
-Translating the uncertainty in drought risk to a particular crop diversification strategy (Figure \@ref(fig:dynamcs)b).
+Translating the uncertainty in drought risk to a particular crop diversification strategy (Figure \@ref(fig:dynamics)b).
 
@@
 
-In this chapter, I explored the consequences of individual heterogeneity in risk perception on the risk-management practices of a simulated Neolithic farming community in the eastern Mediterranean.
-I used a long-term paleoclimate simulation to estimate the changing risks of agricultural drought over a 4,000 year time period spanning the early to middle Holocene.
-Over this time span, the risk of a wheat crop failing from drought varied from once every ten years to nearly once every two years.
+In this chapter, I explored the consequences of individual heterogeneity in risk perception on the risk-management practices of a simulated Neolithic farming community.
+I used a long-term paleoclimate simulation over the eastern Mediterranean to estimate the changing risks of agricultural drought over a 4,000 year time period spanning the early to mid-Holocene.
+Over this time, wheat crops would have been expected to fail from drought between once every ten years to nearly once every two years.
 Changes of such magnitude would have severely impacted Neolithic agroecosystems in the long run, but would have been difficult for any individual farmer to perceive in the short run.
@@
 1.
-**Risk perception is hard.** The climate system inherently chaotic.
+**Risk perception is hard.** The climate system is inherently chaotic.
 Annual forecasts are fundamentally uncertain, even in the era of modern supercomputers and numerical weather prediction.
@@
 2.
-**Individual risk perception has consequences for collective risk management.** Individuals of different ages may perceive the same dry period depending on their prior life experience.
+**Individual risk perception has consequences for collective risk management.** Individuals of different ages may perceive the same dry period differently depending on their prior life experience.
 Younger individuals are more likely to perceive a run of dry years as a trend, rather than a temporary deviation from the norm, and older individuals  are likely to do the opposite.
@@
 
-Although social learning and cumulative cultural evolution has not been a focus of this paper (see [@Boyd and Richerson 1998, 2001]), these findings provide insight into the individual learning dynamics that underlay those broader social processes.
+Although social learning and cumulative cultural evolution have not been a focus of this paper (see [@Boyd and Richerson 1998, 2001]), these findings provide insight into the individual learning dynamics that underlay those broader social processes.
 Relating collective knowledge to individual cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007].
@@
 # References
+\noindent

### 2020-03-05 manuscript/manuscript.Rmd
decision-making

@@
 subtitle: 
-titlerunning: The dynamics of risk perception in a Mediterranean agroecosystem
+titlerunning: The dynamics of risk and risk perception in a Mediterranean agroecosystem
 authorrunning: Gauthier
@@
 The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a "game against nature" [@LuceRaiffa1989, Milnor1952, Agrawal1968, cassidy1971, gould1963].
-The "game" in this context represents the farmer's decision of which crops to plant and in what proportions, given uncertainty in whether "nature" will be dry or wet in a given year (Table 1).
-Although  reality continuous, but people often intuitively solve an easier problem when faced with a complex real-world situation [@Simon1952].
-The exact values in the payoff matrix here are less important than the relative payoffs in each quadrant.
-We can think of this game as being the culturally-inherited object.
-So in this case the cultural norm would be to mix wheat and barley, but individuals need to choose the precise ratio given their perception of risk.
-Our question is how do individuals make decisions here?
+In the context of decision theory, the "game" is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the weather (or the "state of nature") in a given year.
+Assume that a farmers are working with a simplified representation of reality and intuitively solve an easier problem when faced with a complex real-world situation [@Simon1952].
+That is, rather than a continuum of possible states of nature, only care about two categories -- years when its too dry to plant wheat and years when it is not.
+
+This simplified decision problem can take the form of a payoff matrix in which the rows represent the moves of the farmer (wheat or barley) and the columns are the moves of nature (dry or normal) (Table 1).
+The matrix is populated with realistic estimates of potential crop yields in each scenario -- using data from isotopic analysis of grain residues at Neolithic sites [@Slafer1999] -- but the absolute yields are less important for the decision-making problem then the relative sums of each row and column.
+What strategies might a farmer use to "win" this game and maximize their yields?
 
@@
 \centering
-\caption{Estimates of yield volume (t/ha) for ancient wheat and barley varieties derived from [@Slafer1999].}
+\caption{Estimates of yield volume (t/ha) for prehistoric wheat and barley varieties derived from (Slafer et al.
+1999).
+The absolute values here are less important for decision-making than the relative values across each row and column.}
 \begin{tabular}{|l|l|l|}
 \hline
-
  & Dry Year & Normal Year \\ \hline
@@
 Wheat Yield & 0 & 1.60 \\ \hline
-
 \end{tabular}
@@
 
-How should an early farmer make a decision in light of this uncertainty?
-The basic logic of crop diversitication is Portfolio of crops ala modern portfolio theory: Blank2001
-Sometimes best to think of nature as a sentient opponent out to get you, and play strategically based on that assumption. Gould 1963,  Beckenkamp 2008
-quick overview of criteria without risk, then say that a better option is to estimate the probabilities of the different state of nature
-Agrawal1968 talks about the strategies in the context of agriculture being a game against nature
-here I can slip in the cool thing about acting like the weather is out to get you being a good idea in states of complete uncertainty
-transition -- the problem becomes easier if you have at least some idea of nature's moves, because then you cna work to get the highest yeilds given the risk of drought.
-Decision strategies as choosing different points -- plot of decisions
+There are several decision criteria a farmer might use in this situation [@Agrawal1968].
+A risk neutral farmer seek only to maximize yields in a normal year by planting wheat and hoping for the best, as wheat is the highest yielding crop overall.
+But this strategy risks starvation in drought years.
 
-\begin{align}
-a^2+b^2=c^2
-\end{align}
+Instead it is often rational to assume "nature's" moves are decided by sentient being bent on one's ruin, and to play strategically based on that assumption [@Gould 1963, Beckenkamp 2008].
+In the particular form of the wheat-barley game presented in Table 1, a risk sensitive farmer would assume that droughts are inevitable and plant barley to guarantee a minimum acceptable harvest even in the worst case scenario.
+This cautious strategy has some benefits -- particularly in situations of complete uncertainty -- but farmers miss out on high yields they would have received from planting wheat if a drought ultimately does not occur.
 
-```{r, fig.cap = 'Expected wheat and barley yields under increasing drought risk with the point of indifference highlighted.'}
+Playing this "game" over many years allows farmers to learn how probable each state of nature is to occur and to adjust their choices accordingly.
+This strategy is known as playing a game of "fictitious play" against nature.
+If the empirical frequency distribution of wet and dry years is known, the farmer can multiply the crop yields in Table 1 by the probability of each state of nature occurring, and plant the crop with the highest *expected* yield.
+If the probability of a dry year is denoted $\theta$, then the expected barley yield is $\theta \times 0.93 + (1 - \theta) \times 1.18$ and the expected wheat yield is $\theta \times 0 + (1 - \theta) \times 1.60$.
+Barley is favored if $\theta$ is low and droughts are more likely and wheat is favored if $\theta$ high, and value of $\theta$ where expected barley yields equal expected wheat yields is known as the point of indifference (Figure \@ref{fig:yields}).
+If a farmer is unsure whether $\theta$ is above or below the point of indifference, they can plant a mix of crops that maximizes expected yields given their uncertainty in $\theta$.
+
+```{r yields, fig.cap = 'Expected wheat and barley yields under increasing drought risk.
+The point of indifference is highlighted, at which point planting either wheat or barley results in the same expected yield.'}
 expected_yield_wheat <- function(risk) {
@@
 
-If an individual can learn about their environment, such as the risk of drought in any given year, then they can behave more rationally by trying to maximize their subjective expected utility.
-If the probability distribution of nature's moves is known, the farmer can choose the crop mix that simply maximizes the expected crop yields given the empirical frequency distribution of nature's moves.
-In the language of game theory, this strategy is known as a game of "fictitious play" against nature.
-While this is can be an effective risk-management strategy, like all strategies adapted to a specific pattern of variability it is vulnerable to changes in the pattern of variability [@Janssen2007].
-
-This strategy works well when the environment is stationary -- nature plays from a fixed probability distribution, but is vulnerable to environmental non-stationarity.
-That is, when the mean or higher order moments of the rainfall distribution shift, playing a game of fictitious play can backfire because as agents get too stuck in their ways while the world changes around them.
-Thus it is important to understand the dynamics of risk perception -- how subjective risks, and thus subjective expected utility -- rise and fall in uncertain environments.
-
 # Early to mid-Holocene drought risks {#sec:2}
@@
 
-Estimates of changing Holocene precipitaiton variability were derived from the TraCE-21ka paleoclimate simulation [@He2011].
+Estimates of changing Holocene precipitation variability were derived from the TraCE-21ka paleoclimate simulation [@He2011].
 TraCE-21ka is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gas concentrations from the Last Glacial Maximum to the present.
@@
 Monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
-This location was selected to sample climate patterns typical for key archaeological sites in the region such as Çatalhöyük and Aşıklı Höyük, and for the eastern Mediterranean more broadly.
+This location was selected to capture climate variability typical for major Neolithic settlements in the region, such as Çatalhöyük and Aşıklı Höyük, and for the eastern Mediterranean more broadly.
+
 ```{r}
@@
 Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\hat\theta$ for any particular year was defined as the proportion of the previous 50 years that were dry years:
-
 \begin{equation}
@@
 
+Relating collective knowledge to individual cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007].
 Although social learning and cumulative cultural evolution have not been a focus of this paper (see [@Boyd and Richerson 1998, 2001]), these findings provide insight into the individual learning dynamics that underlay those broader social processes.
-Relating collective knowledge to individual cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007].
 This chapter has focused primarily on the physical and cognitive dimensions of risk and risk perception.

### 2020-03-05 manuscript/manuscript.Rmd
fig 5 explanation

@@
 
+## The payoff matrix
+
 This simplified decision problem can take the form of a payoff matrix in which the rows represent the moves of the farmer (wheat or barley) and the columns are the moves of nature (dry or normal) (Table 1).
@@
 
+## Maximizing subjective expected yields
+
 Playing this "game" over many years allows farmers to learn how probable each state of nature is to occur and to adjust their choices accordingly.
@@
 If the probability of a dry year is denoted $\theta$, then the expected barley yield is $\theta \times 0.93 + (1 - \theta) \times 1.18$ and the expected wheat yield is $\theta \times 0 + (1 - \theta) \times 1.60$.
-Barley is favored if $\theta$ is low and droughts are more likely and wheat is favored if $\theta$ high, and value of $\theta$ where expected barley yields equal expected wheat yields is known as the point of indifference (Figure \@ref{fig:yields}).
+Barley is favored if $\theta$ is low and droughts are more likely, wheat is favored if $\theta$ high, and the value of $\theta$ at which expected barley and wheat yields are the same is known as the point of indifference (Figure \@ref(fig:yields)).
 If a farmer is unsure whether $\theta$ is above or below the point of indifference, they can plant a mix of crops that maximizes expected yields given their uncertainty in $\theta$.
+Thus, the exact risk of a drought occurring in any given year, and farmers' perceptions of that risk, have major implications for the entire decision-making process
 
@@
 
-Using the simulated paleoclimate, rather than modern-day weather station observations or paleoclimate proxies, allows us to estimate not just the first order statistics of the climate system (e.g.
-the mean and variance of dry years) but also the higher order patterns such as the serial persistence of wet and dry years.
+Estimates of past climate variability derived from paleoclimate simulations provides a richer representation of not only the first order statistics of the climate system (e.g.
+the mean and variance of precipitation) but also the higher order patterns such as the serial persistence of wet and dry years.
 This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
@@
 The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
-
-Monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
+For this analysis, monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
 This location was selected to capture climate variability typical for major Neolithic settlements in the region, such as Çatalhöyük and Aşıklı Höyük, and for the eastern Mediterranean more broadly.
@@
 
-The example in the previous section established some basic intuitions for how agents with varying subjective beliefs can perceive risk differently.
+The previous example established some basic intuitions for how agents with varying subjective beliefs can perceive risk differently.
 But it represents an idealized situation where agents update their beliefs retrospectively after a many-year dry period.
@@
 
-To explore these dynamics we can simulate a population of Bayesian agents responding to a random sequence of wet and dry years.
-Here, observations are made about the world sequentially, and the agent must continuously update its beliefs.
+To explore these dynamics further we can simulate random sequences of wet and dry years and see how a population of Bayesian agents responds to this sequence in time.
+A 75-year period of low (20\%) drought risks is punctuated by an abrupt 25 years dry period during which drought risk doubles.
+Here, observations are made about the world sequentially, and the agents must continuously update their beliefs.
 As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception of risk.
-Now, this update occurs every year and thus only a year's worth of new information is incorporated into the agent's beliefs each time step.
-
+Now, however, this update occurs every year and thus only a year's worth of new information is incorporated into the agent's beliefs each time step.
 This learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
 Formally, this iterative process is known as "online" learning from Bernoulli observations, where "online" refers to the sequential, year-to-year updating [@Bissiri2010].
-In this learning environment, isolating signal from noise becomes critical for accurately perceiving evolving drought risks.
 
@@
 
-Now we can simulate a time series.
-As above, we'll use a scenario not uncommon in the paleoclimate simulation -- a baseline drought risk of 20\% punctuated by a 25 year dry period when risks rise to 40\%.
-The results of such a simulation reveal the importance of the subjective experiences of individual decision makers on the population-level perception of risk (Figure \@ref(fig:dynamics)a).
+The results of this simulation reveal the importance of the subjective experiences of individual decision makers on the population-level perception of risk (Figure \@ref(fig:dynamics)a).
+In this learning environment, isolating signal from noise becomes critical for accurately perceiving evolving drought risks.
+All agents begin the simulation with diffuse subjective beliefs about drought risk, but as they begin to learn the collective subjective beliefs approaches the objective risk level.
+The agents are slow to update their beliefs during the 25-year dry period because drought years, while increasingly frequent, are still few enough not to outweigh the agents prior beliefs.
+Some agents may perceive gradual changes, but the population as a whole does not perceive the change in climate until more than a decade after it has begun.
 
-Translating the uncertainty in drought risk to a particular crop diversification strategy (Figure \@ref(fig:dynamics)b).
+The agents are even slower to realize when the dry period is over.
+Indeed, the dry period made a strong enough imprint on the population's collective memory that the they perceive droughts to be much more probable than they actually are for decades after conditions have ameliorated once and for all.
+These biased perceptions have consequences for collective ability of the population to manage risk, because uncertainty in whether drought risks are above or below the indifference point informing the mix of crops to plant (Figure \@ref(fig:dynamics)b).
+Once again, the skewed perceptions of increased drought risks last long after the dry period ends, reflected in a much higher ratio of barley to wheat -- a more cautious crop mix -- than is rational given the objective risk of drought in the environment.
 
-```{r dynamics, fig.asp = 1.1, fig.cap='A: The dynamics of risk perception. B:'}
+```{r dynamics, fig.asp = 1, fig.cap='A) Perceived drought risks in a population of Bayesian agents before and after a drought (grey band).
+The dashed red line represents the level of risk at which an agent is indifferent between planting wheat and barley.
+B) Crop mix over time, calculated based on the degree of dispersion above and below the indifference level in A.'}
 a <- ggplot(sim1) +
@@
 
-To explore these dynamics, I simulated a population of Bayesian "belief-based" agents who use their subjective perception of annual drought risks to decide what mix of crops will best manage those risks.
+To explore these dynamics, I simulated a population of "belief-based" Bayesian agents who use their subjective perception of annual drought risks to decide what mix of crops will best manage those risks.
 During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.

### 2020-03-06 manuscript/manuscript.Rmd
typos

@@
 subtitle: 
-titlerunning: The dynamics of risk and risk perception in a Mediterranean agroecosystem
+titlerunning: The dynamics of risk perception in a Mediterranean agroecosystem
 authorrunning: Gauthier
@@
 keywords:
-- belief-based agents
+- Bayesian agents
+- game against nature
+- crop diversification
 - drought risk
-- crop diversification
 - rapid climate change
@@
 Each of these strategies requires farmers to make specific resource allocation decisions in response to environmental risks and is thus sensitive to variability in both the spatiotemporal pattern of risk and the ability of farmers to perceive that pattern.
-In this chapter, I present an agent-based model of a Mediterranean agroecosystem.
+In this chapter, I present a simple agent-based model of a Mediterranean agroecosystem.
 By driving the model with realistic environmental dynamics derived from simulations of mid-Holocene Mediterranean climate, and by allowing the psychology of risk perception to vary among individual farmers, I explore the hidden vulnerabilities of traditional risk-management strategies to periods of rapid climate change.
@@
 bibstyle: Formato_bibliografia_Springer
-# bibstyle options spbasic(default), spphys, spmpsci
 output:
@@
 The distinctive climate and ecology of the Mediterranean basin afforded both challenges and opportunities to the earliest farming communities.
-Here, water is the primary limiting resource for traditional agroecosystems.
-Agricultural droughts -- where growing season precipitation is low enough to cause crop failures, are a constant threat.
-Precipitation is highly variable in space and time and droughts are difficult to predict with any certainty.
-How were Neolithic farmers able to adapt to, and even thrive in, such an uncertain environment?
+Water was the primary limiting resource for these traditional agroecosystems.
+Agricultural droughts where growing season precipitation is low enough to cause crop failures are a constant threat.
+Precipitation is highly variable in space and time and droughts are difficult to predict with certainty.
+How were Neolithic farmers able to adapt to and even thrive in such an uncertain environment?
 
 Over the past 10,000 years, small-scale subsistence farmers have relied on a suite of strategies to maintain stable food supplies given uncertain rainfall.
-These strategies includes practices like crop diversification, storage, mobility, and exchange [@Halstead1989].
+These strategies include practices like crop diversification, storage, mobility, and exchange [@Halstead1989].
 Crop diversification in particular is an excellent example of a widespread and effective risk management strategy that is well suited to Mediterranean agroecosystems.
-In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by even the earliest sedentary farmers, and continue to be used to this day [@GOULD1963a,Slafer1999,Abbo2009a,abbo2010,weiss zohary 2011,Marston2011190].
-Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@helmers2001separating,Anderies2006].
+In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by the earliest sedentary farmers and continue to be used to this day [@GOULD1963a; @Slafer1999; @Abbo2010; @Weiss2011; @Marston2011].
+Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply [@Helmers2001;@Anderies2006].
 Wheat is high yielding but drought sensitive, while barley is lower yielding but drought tolerant.
-Planting a mix of high yield, high risk and low yield, low risk crops, either in the same plot or in a combination of plots, is an effective means of diversifying the annual supply of staple food crops [@paul et al 2019].
+Planting a mix of high-yield, high-risk and low-yield, low-risk crops, either in the same plot or in a combination of plots, is an effective means of diversifying the annual supply of staple food crops [@Paut2019].
 By dynamically adjusting the ratio of wheat to barley in their fields, farmers can adapt to a variety local climate conditions with different drought risks.
 
-Risk-managing strategies like crop diversification require farmers to make specific resource allocation decisions in response to specific environmental risks, so are sensitive to variability in both the spatiotemporal patterns of risk and the ability of individual decision-makers to perceive and act on those patterns.
+Risk-management strategies like crop diversification require farmers to make specific resource allocation decisions in response to specific environmental risks.
+As such, they are sensitive to variability in both the spatiotemporal patterns of risk and the ability of individual decision-makers to perceive and act on those patterns.
 Here, I focus on two main questions:
@@
 
-2. How might Neolithic farmers have perceived these changing risks, and what were the consequences for Neolithic farmers' collective ability to manage them?
+2. How would Neolithic farmers have perceived these changing risks, and what were the consequences for Neolithic farmers' collective ability to manage them?
 
-To address these questions, I first use results from a long-term paleoclimate simulation to estimate the changing risk of agricultural droughts in the eastern Mediterranean over a 4,000 year period in the early to middle Holocene.
-Then, I use a simulated population of Bayesian "belief-based" agents to explore how well individual farmers would have been able to perceive these long-term changes in drought risks given their finite life experiences and limited capacity to process information.
+To address these questions, I first use results from a long-term paleoclimate simulation to estimate the changing risk of agricultural droughts in the eastern Mediterranean over a 4,000-year period in the early to middle Holocene.
+Then, I use a simulated population of "belief-based" Bayesian agents to explore how well individual farmers would have been able to perceive these long-term changes in drought risks given their finite life experiences and limited capacity to process information.
 This computational approach allows for a more nuanced understanding of the vulnerability of risk-management strategies to unpredictable climatic variability.
@@
 
-The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a "game against nature" [@LuceRaiffa1989, Milnor1952, Agrawal1968, cassidy1971, gould1963].
+The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a "game against nature" [@Milnor1952; @GOULD1963a; @Agrawal1968; @Cassidy1971; @Luce1989].
 In the context of decision theory, the "game" is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the weather (or the "state of nature") in a given year.
-Assume that a farmers are working with a simplified representation of reality and intuitively solve an easier problem when faced with a complex real-world situation [@Simon1952].
-That is, rather than a continuum of possible states of nature, only care about two categories -- years when its too dry to plant wheat and years when it is not.
+Assume that farmers are working with a simplified representation of reality and intuitively solve an easier problem when faced with a complex real-world situation [@Simon1990].
+That is, rather than a continuum of possible states of nature, farmers only care about two categories -- years when it is too dry to plant wheat and years when it is not.
 
@@
 
-This simplified decision problem can take the form of a payoff matrix in which the rows represent the moves of the farmer (wheat or barley) and the columns are the moves of nature (dry or normal) (Table 1).
-The matrix is populated with realistic estimates of potential crop yields in each scenario -- using data from isotopic analysis of grain residues at Neolithic sites [@Slafer1999] -- but the absolute yields are less important for the decision-making problem then the relative sums of each row and column.
-What strategies might a farmer use to "win" this game and maximize their yields?
+This simplified decision problem can take the form of a payoff matrix in which the rows represent the moves of the farmer (wheat or barley) and the columns are the moves of nature (dry or normal, Table 1).
+The matrix is populated with realistic estimates of crop yields derived from isotopic analyses of grain residues at Neolithic sites [@Slafer1999].
+However, the absolute yields are less important for the decision-making problem than the relative sums of each row and column.
+Given this payoff matrix, what strategies might a farmer use to "win" this game and maximize their yields?
 
@@
 
-There are several decision criteria a farmer might use in this situation [@Agrawal1968].
-A risk neutral farmer seek only to maximize yields in a normal year by planting wheat and hoping for the best, as wheat is the highest yielding crop overall.
-But this strategy risks starvation in drought years.
+There are several decision criteria a farmer might use in this situation [@Milnor1952; @Agrawal1968].
+A risk neutral farmer seeks only to maximize yields in a normal year by planting wheat and hoping for the best, as wheat is the highest yielding crop overall.
+But, this strategy risks starvation in drought years.
 
-Instead it is often rational to assume "nature's" moves are decided by sentient being bent on one's ruin, and to play strategically based on that assumption [@Gould 1963, Beckenkamp 2008].
-In the particular form of the wheat-barley game presented in Table 1, a risk sensitive farmer would assume that droughts are inevitable and plant barley to guarantee a minimum acceptable harvest even in the worst case scenario.
-This cautious strategy has some benefits -- particularly in situations of complete uncertainty -- but farmers miss out on high yields they would have received from planting wheat if a drought ultimately does not occur.
+Instead it is often rational to assume "nature's" moves are decided by sentient being bent on one's ruin, and to play strategically as if the worst-case scenario will occur [@GOULD1963a; @Beckenkamp2008].
+In the particular form of the wheat-barley game presented in Table 1, a risk sensitive farmer would assume that droughts are inevitable and plant barley to guarantee a minimum acceptable harvest even in the worst-case scenario.
+This cautious strategy is beneficial in situations of complete uncertainty, but farmers miss out on the high yields they would have received from planting wheat if a drought ultimately does not occur.
 
-## Maximizing subjective expected yields
+## Subjective expected yields
 
@@
 This strategy is known as playing a game of "fictitious play" against nature.
-If the empirical frequency distribution of wet and dry years is known, the farmer can multiply the crop yields in Table 1 by the probability of each state of nature occurring, and plant the crop with the highest *expected* yield.
+If the empirical frequency distribution of wet and dry years is known, the farmer can multiply the crop yields in Table 1 by the probability of each state of nature occurring and plant the crop with the highest *expected* yield [@Upton1987].
 If the probability of a dry year is denoted $\theta$, then the expected barley yield is $\theta \times 0.93 + (1 - \theta) \times 1.18$ and the expected wheat yield is $\theta \times 0 + (1 - \theta) \times 1.60$.
 Barley is favored if $\theta$ is low and droughts are more likely, wheat is favored if $\theta$ high, and the value of $\theta$ at which expected barley and wheat yields are the same is known as the point of indifference (Figure \@ref(fig:yields)).
-If a farmer is unsure whether $\theta$ is above or below the point of indifference, they can plant a mix of crops that maximizes expected yields given their uncertainty in $\theta$.
-Thus, the exact risk of a drought occurring in any given year, and farmers' perceptions of that risk, have major implications for the entire decision-making process
+If a farmer is unsure whether $\theta$ is above or below the point of indifference, they can plant a mix of crops that maximizes expected yields given their uncertainty in $\theta$ [@Luce1989].
+Thus, the exact risk of a drought occurring in any given year and farmers' perceptions of that risk both have major implications for the decision-making process.
 
@@
 the mean and variance of precipitation) but also the higher order patterns such as the serial persistence of wet and dry years.
-This presents a much more realistic picture of the inherent year-to-year uncertainty in the climate system, and presents a realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
+These estimates present a more realistic picture of the inherent year-to-year uncertainty in the climate system and presents a more realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
 
@@
 Earth's orbit, greenhouse gasses, glacial meltwater flux).
-The model simulates these dynamics on a six hourly timescale, and model outputs are archived at a monthly resolution.
+The model simulates these dynamics on a six-hourly timescale, and model outputs are archived at a monthly resolution.
 For this analysis, monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
@@
 Using the climate model output, I divided each model year into dry years and normal years.
-A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997], and a normal year was defined as any year above this threshold.
+A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail [@Wilkinson1997] and a normal year was defined as any year above this threshold.
 Given the modeled patterns of normal and dry years, the "objective" climatic drought risk $\hat\theta$ for any particular year was defined as the proportion of the previous 50 years that were dry years:
@@
 ```{r risk, fig.asp = 0.3333, fig.cap = 'Annual risk of wheat crop failure due to drought aggregated by fifty-year period.
-The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields (after Figure 1), given the payoffs defined in Table 1.'}
+The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields (after Figure 1, given the payoffs in Table 1).'}
 sim_past %>%
@@
 In order to properly manage drought risk, a farmer must first be able to perceive that risk.
-Yet, a farmer's perception of risk reflects more than just the objective, empirical risk observable in the world around them [@Tucker2013].
-Individual risk perception is inherently subjective -- influenced by a person's past experience of dry and wet years as filtered through memory -- and can reflect varying levels of uncertainty.
-Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks [@Moussaid2013].
+Yet, a farmer's perception of risk reflects more than just the objective, empirical risk observable in the world around them [@Tucker2007;@Tucker2013;@Findlater2019].
+Individual risk perception is inherently subjective, influenced by a person's past experience of dry and wet years as filtered through memory, and can reflect varying levels of uncertainty.
+Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks [@Moussaid2013;@Amaral2020].
 But how best to model risk perception at the individual level?
@@
 
-The human brain does not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decision-making.
+The human brain does not record every bit of perceived information in memory, rather it stores a "compact encoding" of that information which it uses for future decision-making [@Gallistel2014].
 Integrating the tools of agent-based modeling and Bayesian probability provide an elegant means of representing this process.
-A Bayesian "belief-based" agent is an one whose subjective beliefs can be represented as a probability distribution over possible states of nature [@Kahvalti et al 2019 and others????].
-This approach has a clear computational efficiency -- for both modelers and decision-makers -- and even if real-world decision makers are not Bayesians in a literal sense, the basic algorithmic problem faced by the brain, and the solutions it has evolved, reflect the same constraints on information processing in the minds of decision-makers [@Sanborn and Chater 2016].
+A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature [@Cushman2009;@Pope2015].
+This approach has a clear computational efficiency for both modelers and decision-makers.
+Even if real-world decision makers are not Bayesian in a literal, mathematical, sense, the basic algorithmic problems faced by the brain and the solutions them which it has evolved reflect the same constraints on information processing in the minds of decision-makers [@Bonawitz2011;@Sanborn2016;@Khalvati2019].
 
-We can examine the dynamics of risk perception using a population of Bayesian agents that develop beliefs about the risk of drought in their environment through their personal experience of the weather, and these perceptions impact their decision-making by altering their subjective expected crop yields from planting different crop mixes.
-Here, the probability of a drought occurring in any given year is treated as a coin flip (i.e.
-a draw from Bernoulli distribution) with parameter $\theta$ representing the drought probability.
+We can examine the dynamics of risk perception using a population of Bayesian agents that develop beliefs about the risk of drought through their personal experience of the weather.
+These perceptions impact their decision-making by altering their subjective expected crop yields from planting different crop mixes.
+Here, occurrence of a drought in any given year is treated as a draw from Bernoulli distribution (i.e.
+a coin flip) with parameter $\theta$ representing the drought probability.
 The beta distribution is a natural choice for representing knowledge about probabilities because it is constrained to fall between 0 and 1.
-Hence, an individual agent's prior belief about the plausivle value of $\theta$ can be represented as
+Hence, an individual agent's prior belief about the plausible value of $\theta$ can be represented as
 \begin{equation}
@@
 
-Using probabilities to represent agents' beliefs also allows for estimation of the *uncertainty* in those beliefs (Figure \@ref(fig:prior-prob)).
+Using probabilities to represent agents' beliefs also allows for estimation of the uncertainty in those beliefs (Figure \@ref(fig:prior-prob)).
 A simple Bayesian agent becomes more certain in their beliefs with age.
@@
 As a Bayesian agent must be able to update its beliefs as it moves through time and observes each successive year's weather.
-It does does by comparing the information in this year's observation with the cumulative weight of their past experience.
-The agent combines its prior beliefs about drought risk with the likelihood of having observing a drought in the current year in order to generate a posterior distribution representing its updated beliefs about the world.
+It does so by comparing the information in this year's observation with the cumulative weight of their past experience.
+The agent combines its prior beliefs about drought risk with the likelihood of having observed a drought in the current year in order to generate a posterior distribution representing its updated beliefs about the world.
 
@@
 The mean value of $\theta$ for both agents would thus be 20\%, but the degree of certainty varies because the older agent bases its inference on many more years of experience than the younger agent.
-Now, assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years -- not an uncommon occurrence in the simulation of early to mid-Holocene climate.
-Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different *even though both experienced the same climate*.
+Now, assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years, not an uncommon occurrence in the simulation of early to mid-Holocene climate.
+Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different even though both experienced the same climate.
 
@@
 Although neither agent may perceive the "true" climatic drought risk exactly, they nevertheless reflect perfectly rational beliefs about the world.
-Both agents have perfectly rational beliefs and differ only in their prior subjective beliefs - their relative conservatism or flexibility are not biases but rather varying perspectives on an inherently uncertain world.
+Both agents have rational beliefs and differ only in their prior subjective beliefs.
+Their relative conservatism or flexibility are not biases, but rather varying perspectives on an inherently uncertain world.
 
-As was the case in the game of "fictitious play" discussed in Section \@ref(sec:1), a degree of conservatism is warranted when the environment is stable and risks do not vary.
-But when the environment is volatile the ability to change one's mind is crucial.
-Being too inflexible in one's priors can lead being too optimistic when things really have changed for the worse, or too conservative when conditions ultimately improve.
+As was the case in the game of fictitious play discussed in Section \@ref(sec:1), a degree of conservatism is warranted when the environment is stable and risks do not vary.
+But, when the environment is volatile, the ability to change one's mind is crucial [@Gallistel2014].
+Being too inflexible in one's priors can lead to decisions that are too optimistic when conditions really have changed for the worse.
+Conversely, inflexible priors can fail to take advantage of good conditions.
 
@@
 To explore these dynamics further we can simulate random sequences of wet and dry years and see how a population of Bayesian agents responds to this sequence in time.
-A 75-year period of low (20\%) drought risks is punctuated by an abrupt 25 years dry period during which drought risk doubles.
-Here, observations are made about the world sequentially, and the agents must continuously update their beliefs.
+A 75-year period of low (20\%) drought risks is punctuated by an abrupt 25-year dry period during which drought risk doubles.
+Here, observations are made about the world sequentially, and agents must continuously update their beliefs.
 As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception of risk.
-Now, however, this update occurs every year and thus only a year's worth of new information is incorporated into the agent's beliefs each time step.
+Now, however, agents update their beliefs every year and thus only a year's worth of new information is incorporated into the agents' priors at each time step.
 This learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
@@
 All agents begin the simulation with diffuse subjective beliefs about drought risk, but as they begin to learn the collective subjective beliefs approaches the objective risk level.
-The agents are slow to update their beliefs during the 25-year dry period because drought years, while increasingly frequent, are still few enough not to outweigh the agents prior beliefs.
+The agents are slow to update their beliefs during the 25-year dry period because drought years, while increasingly frequent, are still few enough not to outweigh their prior beliefs.
 Some agents may perceive gradual changes, but the population as a whole does not perceive the change in climate until more than a decade after it has begun.
@@
 These biased perceptions have consequences for collective ability of the population to manage risk, because uncertainty in whether drought risks are above or below the indifference point informing the mix of crops to plant (Figure \@ref(fig:dynamics)b).
-Once again, the skewed perceptions of increased drought risks last long after the dry period ends, reflected in a much higher ratio of barley to wheat -- a more cautious crop mix -- than is rational given the objective risk of drought in the environment.
+Once again, the skewed perceptions of increased drought risks last long after the dry period ends, reflected in a much higher ratio of barley to wheat -- a more cautious crop mix than is rational given the objective risk of drought in the environment.
 
@@
     scale_color_manual(values = c('#f1a340', '#998ec3'), name = '') + 
-    labs(x = 'Year', y = 'Wheat proportion') +
+    labs(x = 'Year', y = 'Proportion of wheat') +
   theme_classic()
@@
 In this chapter, I explored the consequences of individual heterogeneity in risk perception on the risk-management practices of a simulated Neolithic farming community.
-I used a long-term paleoclimate simulation over the eastern Mediterranean to estimate the changing risks of agricultural drought over a 4,000 year time period spanning the early to mid-Holocene.
+I used a long-term paleoclimate simulation over the eastern Mediterranean to estimate the changing risks of agricultural drought over a 4,000-year time period spanning the early to mid-Holocene.
 Over this time, wheat crops would have been expected to fail from drought between once every ten years to nearly once every two years.
-Changes of such magnitude would have severely impacted Neolithic agroecosystems in the long run, but would have been difficult for any individual farmer to perceive in the short run.
+Changes in risk of such magnitude would have severely impacted Neolithic agroecosystems in the long run but would have been difficult for any individual farmer to perceive in the short run.
 
@@
 But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk “optimally” in light of their prior beliefs can experience food shortfalls.
-Cognitive diversity and life experience can be as or more important than the exact mix of crops planted for a population's long-run survival under extreme uncertainty.
+Cognitive diversity and life experience can be as or more important than the exact mix of crops planted for a population's long-term survival under extreme uncertainty.
 
 These dynamics have implications for understanding risk management and food production in the Neolithic.
-Two points arise that this model helps clarify our understanding of the earliest farming communities in the Mediterranean and can inform future simulation work:
+Two points arise from this model that clarify our understanding of the earliest farming communities in the Mediterranean and can inform future simulation work:
 
 1.
-**Risk perception is hard.** The climate system is inherently chaotic.
+**Risk perception is difficult.** The climate system is inherently chaotic.
 Annual forecasts are fundamentally uncertain, even in the era of modern supercomputers and numerical weather prediction.
-For prehistoric farmers, this uncertainty would have been a existential challenge.
+For prehistoric farmers, this uncertainty would have been an existential challenge.
 With the end of the Last Glacial Maximum and the advent of the Holocene, precipitation became increasingly volatile on multiple time scales.
-This uncertainty in the objective drought risk is only compounded by uncertainty in individual farmers' subjective perception of drought risk.
+This uncertainty in the objective drought risk is compounded by uncertainty in individual farmers' subjective perception of drought risk.
 Any individual would have experienced only a brief snapshot of this complex period.
-The rhythm of a single human lifespan is out of sync with the centennial to millennial scale oscillations in the climate system, so even the predictive value of one's own experience is itself unpredictable.
-This fundamental uncertainty would have influenced far more than the choice of which type of crops to plant -- and would have pervaded all kinds of decision-making under risk.
+The rhythms of human lifespans are out of sync with the centennial to millennial scale oscillations in the climate system, so even the predictive value of one's own experience is itself unpredictable.
+This fundamental uncertainty would have influenced far more than the choice of which type of crops to plant and would have pervaded all kinds of decision-making under risk.
 
@@
 By extension, the age structure of a population will influence how quickly it is able to perceive and adapt to a changing climate.
-For example, a young, fast growing population will have a different collective memory of a past drought event then an older population.
-Likewise, baby booms and busts associated with famines, or any mortality crisis that afflicts specific age classes like warfare or epidemics, will alter the time horizon of that population's collective memory.
+For example, a young, fast growing population will have a different collective memory of a past drought event than an older population.
+Likewise, famines, warfare, epidemics, and other crises that afflict specific age classes will alter the time horizon of that population's collective memory.
 Individual heterogeneity in risk perceptions can thus play a key role in broader social responses to climatic risks.
 
-Relating collective knowledge to individual cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007].
-Although social learning and cumulative cultural evolution have not been a focus of this paper (see [@Boyd and Richerson 1998, 2001]), these findings provide insight into the individual learning dynamics that underlay those broader social processes.
+Relating collective knowledge to individual perception and cognition is essential for understanding human behavior in complex social-ecological systems [@Beratan2007; @Pope2015].
+Although social learning and cumulative cultural evolution have not been the focus of this chapter, these findings provide insight into the individual learning dynamics that underlay those broader social processes.
 This chapter has focused primarily on the physical and cognitive dimensions of risk and risk perception.
 The social context of risk and risk perception can be equally consequential [@Rogers1997].
-The balance of individual learning with social transmission determines the collective perception of risks and its impact on collective memory [@Moussaid2013, candia2019].
-Our finite lifespans ultimately keep the skill of individual learning low over the long term -- cumulative cultural evolution is required.
+The balance of individual learning with social transmission determines the collective perception of risks and its impact on collective memory [@Moussaid2013; @Candia2019].
+Our finite lifespans ultimately limit the skill of individual learning over the long term, and cumulative cultural evolution is necessary for continued survival.
 

### 2020-03-06 manuscript/manuscript.tex
typos

--- 2020-02-07 manuscript/manuscript.tex
@@
 
+\usepackage{float}
+\floatplacement{figure}{tb}
+
 \begin{document}
 
-\title{The dynamics of risk perception in a Mediterranean agroecosystem \thanks{Grants or other notes about the article that should go on the front page
-should be placed here. General acknowledgments should be placed at the
-end of the article.} }
+\title{The dynamics of risk perception in a Mediterranean agroecosystem }
 
-    \titlerunning{Risk perception in a Mediterranean agroecosystem}
+    \titlerunning{The dynamics of risk perception in a Mediterranean agroecosystem}
 
-\author{  Nicolas Gauthier 1 \and  }
+\author{  Nicolas Gauthier \and  }
 
@@
 \institute{
-        Nicolas Gauthier 1 \at
-     Laboratory of Tree-Ring Research \& School of Geography and Development,
- University of Arizona \\
-     \email{\href{mailto:ngauthier@email.arizona.edu}{\nolinkurl{ngauthier@email.arizona.edu}}}  %  \\
+        Nicolas Gauthier \at
+     Laboratory of Tree-Ring Research \& School of Geography and Development, University of Arizona \\
+     \email{\href{mailto:ngauthier@arizona.edu}{\nolinkurl{ngauthier@arizona.edu}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 \begin{abstract}
-Small-scale agriculturalists in the Mediterranean Basin rely on multiple
-strategies including diversification, intensification, and storage to
-maintain a stable food supply in the face of environmental uncertainty.
-Each of these strategies requires farmers to make specific resource
-allocation decisions in response to environmental risks and is thus
-sensitive to variability in both the spatiotemporal pattern of risk and
-the ability of farmers to perceive that pattern. In this chapter, I
-present an agent-based model of a Mediterranean agroecosystem. By
-driving the model with realistic environmental dynamics derived from
-simulations of mid-Holocene Mediterranean climate, and by allowing the
-psychology of risk perception to vary among individual farmers, I
-explore the hidden vulnerabilities of traditional risk-management
-strategies to periods of rapid climate change. I show that even when
-farmers are able to manage risk ``optimally'' in light of past
-experience, unanticipated changes in the spatiotemporal pattern of
-rainfall can still lead to major food shortfalls.
+Small-scale agriculturalists in the Mediterranean Basin rely on multiple strategies including diversification, intensification, and storage to maintain a stable food supply in the face of environmental uncertainty.
+Each of these strategies requires farmers to make specific resource allocation decisions in response to environmental risks and is thus sensitive to variability in both the spatiotemporal pattern of risk and the ability of farmers to perceive that pattern.
+In this chapter, I present a simple agent-based model of a Mediterranean agroecosystem.
+By driving the model with realistic environmental dynamics derived from simulations of mid-Holocene Mediterranean climate, and by allowing the psychology of risk perception to vary among individual farmers, I explore the hidden vulnerabilities of traditional risk-management strategies to periods of rapid climate change.
+I show that even when farmers are able to manage risk ``optimally'' in light of past experience, unanticipated changes in the spatiotemporal pattern of rainfall can still lead to major food shortfalls.
 \\
 \keywords{
-        key \and
-        dictionary \and
-        word \and
-    }
-
-    \subclass{
-                    MSC code 1 \and
-                    MSC code 2 \and
+        Bayesian agents \and
+        game against nature \and
+        crop diversification \and
+        drought risk \and
+        rapid climate change \and
     }
@@
 
-Over the past 10,000 years, subsistence farmers in the Mediterranean
-basin have developed sophisticated strategies to maintain stable food
-supplies given uncertain rainfall. Their suite of strategies include
-practices like crop diversification, storage, and exchange (Halstead and
-O'Shea 1989)). These risk-management strategies all represent forms of
-\emph{regulatory feedback} used to manage complex socio-ecological
-systems. Regulatory feedbacks convert a system's high-variance inputs
-into low-variance outputs by dynamically monitoring the outputs and
-adjusting internal properties of the system ({\textbf{???}}).
-Risk-managing strategies require farmers to make specific resource
-allocation decisions in response to environmental risks, and as
-regulatory feedbacks they are thus sensitive to variability in both the
-spatiotemporal patterns of risk and the ability of farmers to perceive
-and act on those patterns.
+The distinctive climate and ecology of the Mediterranean basin afforded both challenges and opportunities to the earliest farming communities.
+Water was the primary limiting resource for these traditional agroecosystems.
+Agricultural droughts where growing season precipitation is low enough to cause crop failures are a constant threat.
+Precipitation is highly variable in space and time and droughts are difficult to predict with certainty.
+How were Neolithic farmers able to adapt to and even thrive in such an uncertain environment?
 
-Crop diversification is an excellent example of a widespread and
-effective regulatory feedback strategy (Figure 1). Relying on a mix of
-food types with different climatic tolerances is an efficient way to
-maintain a robust food supply ({\textbf{???}},Anderies2006). In the
-Mediterranean, land-use strategies involving a diversified portfolio of
-wheat and barley have been employed by even the earliest sedentary
-farmers, and continue to be used to this day (Gould
-1963,Slafer1999,Abbo2009a,Marston2011190). Wheat generally has higher
-yields but is sensitive to water deficits, while barley has lower yields
-and is drought tolerant. Planting a mix of wheat and barley, either in
-the same plot or in a combination of plots, is an effective means of
-diversifying grain supplies. By adjusting the ratio of wheat to barley,
-farmers can adapt to changing drought risks.
+Over the past 10,000 years, small-scale subsistence farmers have relied on a suite of strategies to maintain stable food supplies given uncertain rainfall.
+These strategies include practices like crop diversification, storage, mobility, and exchange (Halstead and O'Shea 1989).
+Crop diversification in particular is an excellent example of a widespread and effective risk management strategy that is well suited to Mediterranean agroecosystems.
+In the Mediterranean, land-use strategies involving a diversified portfolio of wheat and barley have been employed by the earliest sedentary farmers and continue to be used to this day (Gould 1963; Araus, Slafer, and Romagosa 1999; Abbo, Lev-Yadun, and Gopher 2010; Weiss and Zohary 2011; Marston 2011).
+Relying on a mix of food types with different climatic tolerances is an efficient way to maintain a robust food supply (Helmers, Yamoah, and Varvel 2001; Anderies 2006).
+Wheat is high yielding but drought sensitive, while barley is lower yielding but drought tolerant.
+Planting a mix of high-yield, high-risk and low-yield, low-risk crops, either in the same plot or in a combination of plots, is an effective means of diversifying the annual supply of staple food crops (Paut, Sabatier, and Tchamitchian 2019).
+By dynamically adjusting the ratio of wheat to barley in their fields, farmers can adapt to a variety local climate conditions with different drought risks.
 
-The secret to the success of earliest farmers on the mediterrenan
-littoral was in their manipulation of the local ecosystem to manage
-environmental risks. Early farmers constructed a complex agroecosystem
-to manage risk. They needed to maintain stable crop yields in face of
-uncertain drought risks. Yield stability is very important for
-traditional agriculturlal commmunities not connected to larger bulk food
-networks (Abbo et al 2010) neolithic crop package -- weiss zohary 2011
-cite abbo et al 2010? for more stuff about ladraces Wheat/barley systems
-example of system that combines low yield, low wirks crop with high
-yield high risk crop, an effective context paut et al 2019 These
-strategies work when risks are knonw and constant, but may be more
-susceptible to periods of changing climateic risks so understanding how
-people of the past adapted to change is important to us today
-
-Are these crop-diversification strategies vulnerable to the same
-dynamics as other social-ecological systems with similar regulatory
-feedbacks mechanisms? One way to address this is to model the influence
-of imperfect monitoring and biased decision making in uncertain
-environments. In this study, I accomplish this by answering two main
-questions:
+Risk-management strategies like crop diversification require farmers to make specific resource allocation decisions in response to specific environmental risks.
+As such, they are sensitive to variability in both the spatiotemporal patterns of risk and the ability of individual decision-makers to perceive and act on those patterns.
+Here, I focus on two main questions:
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
-  What was the temporal pattern of climate and climate change during the
-  Holocene in the eastern Mediterranean?
+  How likely were droughts to occur each year in the eastern Mediterranean, and how did these risks change over the Holocene?
 \item
-  How well could crop diversification strategies cope with this pattern,
-  given farmers' imperfect risk perception and decision-making?
+  How would Neolithic farmers have perceived these changing risks, and what were the consequences for Neolithic farmers' collective ability to manage them?
 \end{enumerate}
 
-Questions: How would the earliest farmers have made decisions (dealt
-with uncertainty?) about crop diversification? ie what were the
-different algorithms? -- better understand the decision context What are
-the consequences of theses different decisions (i.e.~how well did they
-perform) at different points in the history of the mediterranean How
-were farmers able to detect changes in the environment, rather than
-unconsciously adapt to them? what are the different consequences we'd
-expect for each THIS is the most important question because with it I
-can frame the results around and individual's perception of droughts in
-the neolithic, and the populaiton-level distribution of drought
-perceptions at given times
-
-I couple a climate model with high temporal resolution to a simple game
-theoretic model of agricultural decision making under uncertainty, in
-order to examine the performance of optimal and suboptimal wheat-barley
-diversification practices. I model the year-to-year crop diversification
-strategies made by farmers as an iterated game of fictitious play
-against nature (Gould 1963). Farmers are boundedly rational, in that
-they seek to maximize their objective functions given the perceived
-probabilities of different actions by nature, but work with flawed
-mental models of climate-related risks due to imperfect recall of past
-events. Furthermore, by using a climate model to represent the actions
-of nature, I can more precisely capture the characteristic
-autocorrelated patterns of rainfall variability, rather than simply
-drawing rainfall values from a static distribution. This framework
-allows me to address the sensitivity of diversification strategies to
-changes and climatic variability and imperfect human perceptions of
-those changes.
+To address these questions, I first use results from a long-term paleoclimate simulation to estimate the changing risk of agricultural droughts in the eastern Mediterranean over a 4,000-year period in the early to middle Holocene.
+Then, I use a simulated population of ``belief-based'' Bayesian agents to explore how well individual farmers would have been able to perceive these long-term changes in drought risks given their finite life experiences and limited capacity to process information.
+This computational approach allows for a more nuanced understanding of the vulnerability of risk-management strategies to unpredictable climatic variability.
 
@@
 
-The basic decision-making problem facing a farmer seeking to diversify
-their crops can be thought of as a game against nature. We can represent
-this simplified decision context as a ``game'' in a game theorietic
-context. cite luce raiffa 1989 for game against nature, and milnor 1952,
-Agrawal1968, cassidy1971, gould 1963 The ``game'' in this context is the
-farmer's decision of which crops to plant and in what proportions, given
-uncertianty in the future ``state-of-nature'', and nature varies between
-several possible states such as dry and wet years. payoff matrix goes
-here in reality continuous, but cite behavioral econ for people often
-intuitively solve an easier problem when faced with a complex real-world
-situation The exact values in the payoff matrix here are less important
-than the relative payoffs in each quadrant. We can think of this game as
-being the culturally-inherited object. our question is how do
-individuals make decisions here? so in this case the cultural norm would
-be to mix wheat and barley, but the learning is in the exact ratio --
-both individual and social learning influence perceptions of eindividual
-vs social learning tucker 2007
+The basic decision-making problem facing a farmer seeking to diversify their crops can be thought of as a ``game against nature'' (Milnor 1952; Gould 1963; Agrawal and Heady 1968; Cassidy, Field, and Kirby 1971; Luce and Raiffa 1989).
+In the context of decision theory, the ``game'' is the farmer's decision of which crops to plant and in what proportions, given uncertainty in the weather (or the ``state of nature'') in a given year.
+Assume that farmers are working with a simplified representation of reality and intuitively solve an easier problem when faced with a complex real-world situation (Simon 1990).
+That is, rather than a continuum of possible states of nature, farmers only care about two categories -- years when it is too dry to plant wheat and years when it is not.
 
-Estimates of yield volume (t/ha) for each crop type in wet and dry years
-were derived from isotopic studies of ancient wheat and barley samples
-from archaeological sites in the Mediterranean (Araus, Slafer, and
-Romagosa 1999) (Table 1).
+\hypertarget{the-payoff-matrix}{%
+\subsection{The payoff matrix}\label{the-payoff-matrix}}
+
+This simplified decision problem can take the form of a payoff matrix in which the rows represent the moves of the farmer (wheat or barley) and the columns are the moves of nature (dry or normal, Table 1).
+The matrix is populated with realistic estimates of crop yields derived from isotopic analyses of grain residues at Neolithic sites (Araus, Slafer, and Romagosa 1999).
+However, the absolute yields are less important for the decision-making problem than the relative sums of each row and column.
+Given this payoff matrix, what strategies might a farmer use to ``win'' this game and maximize their yields?
 
@@
 \centering
-\caption{Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].}
+\caption{Estimates of yield volume (t/ha) for prehistoric wheat and barley varieties derived from (Slafer et al.
+1999).
+The absolute values here are less important for decision-making than the relative values across each row and column.}
 \begin{tabular}{|l|l|l|}
 \hline
-
  & Dry Year & Normal Year \\ \hline
@@
 
-How should an early farmer make a decision in light ot this uncertainty?
-The basic logic of crop diversitication is Portfolio of crops ala modern
-portfolio theory: Blank2001 Sometimes best to think of nature as a
-sentient opponent out to get you, and play strategically based on that
-assumption. Gould 1963, Beckenkamp 2008 quick overview of criteria
-without risk, then say that a better option is to estimate the
-probabilities of the different state of nature Agrawal1968 talks about
-the strategies in the context of agriculture being a game against nature
-here I can slip in the cool thing about acting like the weather is out
-to get you being a good idea in states of complete uncertainty
-transition -- the problem becomes easier if you have at least someI idea
-of nature's moves , because then you cna work to get the highest yeilds
-given the risk of drought Decision strategies as choosing different
-points -- plot of decisions
+There are several decision criteria a farmer might use in this situation (Milnor 1952; Agrawal and Heady 1968).
+A risk neutral farmer seeks only to maximize yields in a normal year by planting wheat and hoping for the best, as wheat is the highest yielding crop overall.
+But, this strategy risks starvation in drought years.
 
-\begin{align}
-a^2+b^2=c^2
-\end{align}
+Instead it is often rational to assume ``nature's'' moves are decided by sentient being bent on one's ruin, and to play strategically as if the worst-case scenario will occur (Gould 1963; Beckenkamp 2008).
+In the particular form of the wheat-barley game presented in Table 1, a risk sensitive farmer would assume that droughts are inevitable and plant barley to guarantee a minimum acceptable harvest even in the worst-case scenario.
+This cautious strategy is beneficial in situations of complete uncertainty, but farmers miss out on the high yields they would have received from planting wheat if a drought ultimately does not occur.
+
+\hypertarget{subjective-expected-yields}{%
+\subsection{Subjective expected yields}\label{subjective-expected-yields}}
+
+Playing this ``game'' over many years allows farmers to learn how probable each state of nature is to occur and to adjust their choices accordingly.
+This strategy is known as playing a game of ``fictitious play'' against nature.
+If the empirical frequency distribution of wet and dry years is known, the farmer can multiply the crop yields in Table 1 by the probability of each state of nature occurring and plant the crop with the highest \emph{expected} yield (Upton 1987).
+If the probability of a dry year is denoted \(\theta\), then the expected barley yield is \(\theta \times 0.93 + (1 - \theta) \times 1.18\) and the expected wheat yield is \(\theta \times 0 + (1 - \theta) \times 1.60\).
+Barley is favored if \(\theta\) is low and droughts are more likely, wheat is favored if \(\theta\) high, and the value of \(\theta\) at which expected barley and wheat yields are the same is known as the point of indifference (Figure \ref{fig:yields}).
+If a farmer is unsure whether \(\theta\) is above or below the point of indifference, they can plant a mix of crops that maximizes expected yields given their uncertainty in \(\theta\) (Luce and Raiffa 1989).
+Thus, the exact risk of a drought occurring in any given year and farmers' perceptions of that risk both have major implications for the decision-making process.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Expected wheat and barley yields under increasing drought risk
-with the point of indifference highlighted.}
+\includegraphics{manuscript_files/figure-latex/yields-1.pdf}
+\caption{\label{fig:yields}Expected wheat and barley yields under increasing drought risk.
+The point of indifference is highlighted, at which point planting either wheat or barley results in the same expected yield.}
 \end{figure}
 
-If an individual can learn about their environment, such as the risk of
-drought in any given year, then they can behave more rationally by
-trying to maximize their subjective expected utility If the probability
-distributin of nature's moves is known, the farmer can choose the crop
-mix that simply maximizes the expected crop yields given the empirical
-frequency distirubtion of nature's moves. In the language of game
-theory, this strategy is knoan as a ``game of ficticious play'' against
-nature. This is an effective risk-managing strategy, but like all
-similar strategies that adapt to a specific pattern of variability, it
-is vulnerable to changes in the pattern of variability (Janssen eta al
-2007) This strategy works well when the environment is stationary --
-nature plays from a fixed probability distirubtion, but is vulnerable to
-environmental nonstationarity. That is, when the mean or higher order
-moments of the rainfall distribution shift, playing a game of fictitious
-play can backfire because you're too stuck in your ways while the world
-changes around you. Thus it is important to understand the dynamics of
-risk perception -- how subjective risks, and thus subjective expected
-utility -- rise and fall in uncertain environments.
+\hypertarget{sec:2}{%
+\section{Early to mid-Holocene drought risks}\label{sec:2}}
 
-\hypertarget{sec:2}{%
-\section{The dynamics of risk perception}\label{sec:2}}
+In order to estimate drought risks in the past, present-day weather observations alone are insufficient.
+Precipitation varies not only from year-to-year, but also on centennial to millennial-scale time scales that are unresolved in the contemporary observational record.
+Climate dynamics are non-linear, non-stationary, and non-ergodic, which means sudden, unpredictable variability is the norm rather than the exception.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
+Estimates of past climate variability derived from paleoclimate simulations provides a richer representation of not only the first order statistics of the climate system (e.g.~the mean and variance of precipitation) but also the higher order patterns such as the serial persistence of wet and dry years.
+These estimates present a more realistic picture of the inherent year-to-year uncertainty in the climate system and presents a more realistic challenge to simple risk-managing strategies that assume climatic risks are fixed.
 
-\hypertarget{drought-risks-in-the-middle-holocene}{%
-\section{Drought risks in the middle
-Holocene}\label{drought-risks-in-the-middle-holocene}}
+\hypertarget{paleoclimate-simulation}{%
+\subsection{Paleoclimate simulation}\label{paleoclimate-simulation}}
 
-\hypertarget{climate-risks-and-risk-perception}{%
-\section{Climate Risks and Risk
-Perception}\label{climate-risks-and-risk-perception}}
+Estimates of changing Holocene precipitation variability were derived from the TraCE-21ka paleoclimate simulation (He 2011).
+TraCE-21ka is a state-of-the-art simulation that uses a coupled atmosphere-ocean general circulation model to recreate the transient response of the global climate system to changes in the Earth's orbit and greenhouse gas concentrations from the Last Glacial Maximum to the present.
+The simulation generates physically consistent spatiotemporal climate dynamics, driven by current best estimates of external climate drivers (e.g.~Earth's orbit, greenhouse gasses, glacial meltwater flux).
+The model simulates these dynamics on a six-hourly timescale, and model outputs are archived at a monthly resolution.
+For this analysis, monthly TraCE-21ka precipitation outputs were extracted from the 3.75° grid cell covering Central Anatolia.
+This location was selected to capture climate variability typical for major Neolithic settlements in the region, such as Çatalhöyük and Aşıklı Höyük, and for the eastern Mediterranean more broadly.
 
-Using the bias-corrected climate model output, I divided each model year
-into dry years and wet years. A dry year was any year where less than
-300mm of rain fell during the wet season (October-March), the threshold
-below which wheat crops will generally fail ({\textbf{???}}), and a
-wet/normal year was defined as any year above this threshold.
+\hypertarget{estimating-drought-risks}{%
+\subsection{Estimating drought risks}\label{estimating-drought-risks}}
 
-Given the modeled patterns of wet and dry years, drought risk for any
-particular year was defined as the proportion of the previous 50 years
-that were dry years
+Using the climate model output, I divided each model year into dry years and normal years.
+A dry year was any year where less than 300mm of rain fell during the wet season (October-May), the threshold below which wheat crops will generally fail (Wilkinson 1997) and a normal year was defined as any year above this threshold.
+Given the modeled patterns of normal and dry years, the ``objective'' climatic drought risk \(\hat\theta\) for any particular year was defined as the proportion of the previous 50 years that were dry years:
+\begin{equation}
+    \hat\theta = \frac{\sum_{n=t-1}^{t-50} P_n < 300}{50},
+\end{equation}
+where \(P_n\) is the growing season precipitation accumulation in millimeters for year \(n\).
 
-\begin{equation}
-    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
-\end{equation} where \(t\) is the current time step. The 50 year time
-span was selected to approximate the accumulated observational knowledge
-of an individual farmer and their immediate household. As a result, this
-approach does not allow for accumulated social learning, although it
-could easily be extended to do so in future studies.
-
-\hypertarget{climate-modeling-and-bias-correction}{%
-\subsection{Climate Modeling and Bias
-Correction}\label{climate-modeling-and-bias-correction}}
-
-Estimates of drought risk in the eastern Mediterranean during the the
-past 10,000 years were derived from outputs of the TraCE-21k
-paleoclimate simulation available on the National Center for Atmospheric
-Research's Earth System Grid repository
-(\url{https://www.earthsystemgrid.org}). TraCE-21K is a state-of-the-art
-simulation that uses a coupled atmosphere-ocean general circulation
-model (GCM) to recreate the transient response of the global climate
-system to orbital parameters and greenhouse gasses over the past 22,000
-years from the Last Glacial Maximum to the present (Figure 2) (He 2011).
-It generates physically consistent spatiotemporal climate dynamics,
-driven by current best estimates of climate forcings (e.g.~orbit,
-greenhouse gasses, glacial meltwater flux). The model simulates these
-dynamics on a six hourly timescale, and model outputs are archived at a
-monthly resolution.
-
-TraCE-21k simulation outputs for the past 10,000 years were
-bias-corrected using the CDF-t method and observed precipitation from a
-weather station in near the town of Salihli in western Turkey. This
-location was selected due to its long (\textasciitilde{}70yr)
-observational record, the representativeness of western Anatolian
-climate for the greater eastern Mediterranean, and its proximity to
-archaeological sites with comparative evidence of crop diversification
-strategies.
-
-Now incorporate realistic drought variability from simulations.
-
-\begin{verbatim}
-## Loading required namespace: ncdf4
-\end{verbatim}
+The simulated risks of crop failure due to drought ranged between 10\% and 46\% during the period from 9.5ka to 5.5ka, with a median risk of 24\% (Figure \ref{fig:risk}).
+On average, a Neolithic farmer in Central Anatolia could expect their wheat crops to fail two or three times a decade, punctuated by even drier periods in which wheat crops could be expected to fail roughly every other year.
+The simulation also reveals a long-term trend of decreasing drought risks, in particular with higher drought risk in the early Holocene giving way to lower drought risk in the middle Holocene.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{Annual risk of wheat crop failure due to drought, averaged by
-fifty-year period}
+\includegraphics{manuscript_files/figure-latex/risk-1.pdf}
+\caption{\label{fig:risk}Annual risk of wheat crop failure due to drought aggregated by fifty-year period.
+The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields (after Figure 1, given the payoffs in Table 1).}
 \end{figure}
 
-\hypertarget{the-model}{%
-\section{The Model}\label{the-model}}
+The TraCE-21ka simulation confirms that drought risk in the eastern Mediterranean was non-stationary and, in fact, quite volatile during periods of climatic disruption in the early Holocene.
+This volatility would have had severe consequences for early farming communities whose risk-managing practices depended so heavily on accurately perceived local climatic risks.
 
-Run an experiment. Simulate a time series with varying drought
-frequency. This is a stepwise nonstationary process
+\hypertarget{sec:3}{%
+\section{Modeling risk perception}\label{sec:3}}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
+In order to properly manage drought risk, a farmer must first be able to perceive that risk.
+Yet, a farmer's perception of risk reflects more than just the objective, empirical risk observable in the world around them (Tucker 2007; Tucker et al.
+2013; Findlater, Satterfield, and Kandlikar 2019).
+Individual risk perception is inherently subjective, influenced by a person's past experience of dry and wet years as filtered through memory, and can reflect varying levels of uncertainty.
+Likewise, the distribution of individuals' perceived risks within a population influences the collective perception of drought risks and the potential aggregate societal-level response to those risks (Moussaïd 2013; Amaral and Javarone 2020).
+But how best to model risk perception at the individual level?
 
-Redo these plots so its a single person, one who was born in a drought
-period then experiences an ameliorarion and another who experiences a
-shift to drought conditions later in life? basically this section neds
-to work more with individuals and explore that stuff
+\hypertarget{prior-beliefs-and-bayesian-agents}{%
+\subsection{Prior beliefs and Bayesian agents}\label{prior-beliefs-and-bayesian-agents}}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
+The human brain does not record every bit of perceived information in memory, rather it stores a ``compact encoding'' of that information which it uses for future decision-making (Gallistel et al.
+2014).
+Integrating the tools of agent-based modeling and Bayesian probability provide an elegant means of representing this process.
+A Bayesian agent is one whose subjective beliefs can be represented as a probability distribution over possible states of nature (Cushman and Macindoe 2009; Pope and Gimblett 2015).
+This approach has a clear computational efficiency for both modelers and decision-makers.
+Even if real-world decision makers are not Bayesian in a literal, mathematical, sense, the basic algorithmic problems faced by the brain and the solutions them which it has evolved reflect the same constraints on information processing in the minds of decision-makers (Bonawitz et al.
+2011; Sanborn and Chater 2016; Khalvati et al.
+2019).
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
+We can examine the dynamics of risk perception using a population of Bayesian agents that develop beliefs about the risk of drought through their personal experience of the weather.
+These perceptions impact their decision-making by altering their subjective expected crop yields from planting different crop mixes.
+Here, occurrence of a drought in any given year is treated as a draw from Bernoulli distribution (i.e.~a coin flip) with parameter \(\theta\) representing the drought probability.
+The beta distribution is a natural choice for representing knowledge about probabilities because it is constrained to fall between 0 and 1.
+Hence, an individual agent's prior belief about the plausible value of \(\theta\) can be represented as
+\begin{equation}
+  \theta \sim \textrm{Beta}(\alpha, \,\beta),
+\end{equation}
+a beta distribution with the parameters \(\alpha\) and \(\beta\) corresponding to the number of dry and wet years previously experienced by that agent.
+Varying these two parameters thus allows for the representation of a variety of different personal experiences of drought risk.
+For example, if an agent recalls having lived through 5 dry years and 25 normal years, their prior belief about the chance of a drought occurring in the following year would be represented as a Beta(5, 25) distribution with mean value 0.2 equivalent to the empirical drought risk (\(\hat\theta\)) for that period.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}
+Using probabilities to represent agents' beliefs also allows for estimation of the uncertainty in those beliefs (Figure \ref{fig:prior-prob}).
+A simple Bayesian agent becomes more certain in their beliefs with age.
+For example, an agent who experienced 5 wet and 10 normal years and one who experienced 25 wet and 50 normal years would both agree that, on average, droughts occur 50\% of the time.
+Yet the latter agent would be much more certain in this belief because it is drawn from a larger range of experience (i.e.~a larger sample size).
+We can thus represent the exact information content of each individual agent's subjective experience of droughts using the diffusion of this prior belief.
+How, then, should an agent update its beliefs in light of new experience?
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-17-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/prior-prob-1.pdf}
+\caption{\label{fig:prior-prob}Development of an individual's perceived drought risk with time, assuming a fixed drought risk of 0.5.
+Beliefs are represented as beta distributions, and the increased certainty with age reflects the varying effective sample size of the beta prior.}
+\end{figure}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
+\hypertarget{bayesian-updating-and-the-weight-of-past-experience}{%
+\subsection{Bayesian updating and the weight of past experience}\label{bayesian-updating-and-the-weight-of-past-experience}}
 
-These two plots begin to look more alike the more you sample, which
-makes sense because its an ergodic system.
+As a Bayesian agent must be able to update its beliefs as it moves through time and observes each successive year's weather.
+It does so by comparing the information in this year's observation with the cumulative weight of their past experience.
+The agent combines its prior beliefs about drought risk with the likelihood of having observed a drought in the current year in order to generate a posterior distribution representing its updated beliefs about the world.
 
-Note in these plots above and below we're summing all the wheat and
-barley in the population, compared to below which shows the wheat
-proportions of individuals
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-2.pdf}
+Crucially, the strength of one's prior beliefs determines how much weight is given to new information (Figure \ref{fig:posterior}).
+For example, assume two agents -- one aged 15 and the other aged 50 -- who have only ever experienced a normal climate where droughts happen on average two out of every ten years.
+The mean value of \(\theta\) for both agents would thus be 20\%, but the degree of certainty varies because the older agent bases its inference on many more years of experience than the younger agent.
+Now, assume the climate suddenly changes such that the drought risk is doubled to 40\% for the next 25 years, not an uncommon occurrence in the simulation of early to mid-Holocene climate.
+Because the prior beliefs of the agents were so different, their subjective beliefs after the drought are also different even though both experienced the same climate.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-2.pdf}
+For young agents with weak priors, the information of each new year can thus strongly influence their beliefs.
+But older agents -- having more experience and stronger priors -- will be less likely to update their beliefs when comparing the information from a single year's weather with previous decades' worth of accumulated experience.
+Although neither agent may perceive the ``true'' climatic drought risk exactly, they nevertheless reflect perfectly rational beliefs about the world.
+Both agents have rational beliefs and differ only in their prior subjective beliefs.
+Their relative conservatism or flexibility are not biases, but rather varying perspectives on an inherently uncertain world.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-1.pdf}
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-2.pdf}
+As was the case in the game of fictitious play discussed in Section \ref{sec:1}, a degree of conservatism is warranted when the environment is stable and risks do not vary.
+But, when the environment is volatile, the ability to change one's mind is crucial (Gallistel et al.
+2014).
+Being too inflexible in one's priors can lead to decisions that are too optimistic when conditions really have changed for the worse.
+Conversely, inflexible priors can fail to take advantage of good conditions.
 
-\begin{verbatim}
-## Picking joint bandwidth of 0.0146
-\end{verbatim}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/posterior-1.pdf}
+\caption{\label{fig:posterior}Change in perceived drought probability in an older (50) and younger (15) agent before and after a 25-year dry period.}
+\end{figure}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}
+\hypertarget{sec:4}{%
+\section{Risk management and the dynamics of risk perception}\label{sec:4}}
 
-\begin{verbatim}
-## Picking joint bandwidth of 0.0204
-\end{verbatim}
+The previous example established some basic intuitions for how agents with varying subjective beliefs can perceive risk differently.
+But it represents an idealized situation where agents update their beliefs retrospectively after a many-year dry period.
+Farmers, on the other hand, must update their planting decisions each year and continuously monitor the weather around them.
+The transient, year-to-year changes in perceived risks can thus have major consequences for how a population responds \emph{during} a dry period.
+How do Bayesian learners perform in such an uncertain, unpredictable time?
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-2.pdf}
+To explore these dynamics further we can simulate random sequences of wet and dry years and see how a population of Bayesian agents responds to this sequence in time.
+A 75-year period of low (20\%) drought risks is punctuated by an abrupt 25-year dry period during which drought risk doubles.
+Here, observations are made about the world sequentially, and agents must continuously update their beliefs.
+As in the previous section, an agent's prior belief about drought risk is updated in light of new experience to generate a posterior perception of risk.
+Now, however, agents update their beliefs every year and thus only a year's worth of new information is incorporated into the agents' priors at each time step.
+This learning process is iterative, as an agent's posterior distribution in one year becomes their prior following year and the updating process repeats itself.
+Formally, this iterative process is known as ``online'' learning from Bernoulli observations, where ``online'' refers to the sequential, year-to-year updating (Bissiri and Walker 2010).
 
-make this a function to calculate payoffs for all wheat or all barley
-strategies as a comparison
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
+The results of this simulation reveal the importance of the subjective experiences of individual decision makers on the population-level perception of risk (Figure \ref{fig:dynamics}a).
+In this learning environment, isolating signal from noise becomes critical for accurately perceiving evolving drought risks.
+All agents begin the simulation with diffuse subjective beliefs about drought risk, but as they begin to learn the collective subjective beliefs approaches the objective risk level.
+The agents are slow to update their beliefs during the 25-year dry period because drought years, while increasingly frequent, are still few enough not to outweigh their prior beliefs.
+Some agents may perceive gradual changes, but the population as a whole does not perceive the change in climate until more than a decade after it has begun.
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-24-1.pdf}
+The agents are even slower to realize when the dry period is over.
+Indeed, the dry period made a strong enough imprint on the population's collective memory that the they perceive droughts to be much more probable than they actually are for decades after conditions have ameliorated once and for all.
+These biased perceptions have consequences for collective ability of the population to manage risk, because uncertainty in whether drought risks are above or below the indifference point informing the mix of crops to plant (Figure \ref{fig:dynamics}b).
+Once again, the skewed perceptions of increased drought risks last long after the dry period ends, reflected in a much higher ratio of barley to wheat -- a more cautious crop mix than is rational given the objective risk of drought in the environment.
 
-\section{Results and Discussion}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/dynamics-1.pdf}
+\caption{\label{fig:dynamics}A) Perceived drought risks in a population of Bayesian agents before and after a drought (grey band).
+The dashed red line represents the level of risk at which an agent is indifferent between planting wheat and barley.
+B) Crop mix over time, calculated based on the degree of dispersion above and below the indifference level in A.}
+\end{figure}
 
-The bias-corrected wet season precipitation outputs from TraCE-21k show
-a relatively stationary distribution over the past 10,000 years, with
-slightly drier conditions during the early-mid Holocene transition at
-c.a. 8ka BP (Figure 5). The utility of a diversified wheat-barley crop
-is immediately apparent; rainfall often fails to meet the 300mm
-threshold for a productive wheat crop during the entire 10,000 year
-span, but drops below the threshold for barley only a handful of years.
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-The risk of crop failure due to drought in any given year varies between
-10\% and 20\%, punctuated by one to four century-long events where risk
-drops as low as 6\% and or exceeds 25\% (Figure 6). Centuries of high
-drought risk cluster around known periods of regional and global climate
-deterioration, such as at 8.6ka BP and 4.2ka BP.
+In this chapter, I explored the consequences of individual heterogeneity in risk perception on the risk-management practices of a simulated Neolithic farming community.
+I used a long-term paleoclimate simulation over the eastern Mediterranean to estimate the changing risks of agricultural drought over a 4,000-year time period spanning the early to mid-Holocene.
+Over this time, wheat crops would have been expected to fail from drought between once every ten years to nearly once every two years.
+Changes in risk of such magnitude would have severely impacted Neolithic agroecosystems in the long run but would have been difficult for any individual farmer to perceive in the short run.
 
-Given the average drought risks estimated from TraCE-21k and assuming
-perfect monitoring and recall of recent droughts (i.e. \(\lambda = 0\)),
-farmers intent on minimizing the risk of crop failure would be expected
-to plant about 65\% barley to 35\% wheat on average. This proportion is
-within the range estimated from macrobotanical remains at a nearby
-archaeological site ({\textbf{???}}). Yield maximizing farmers instead
-play a pure strategy profile, planting wheat if the perceived drought
-risk is less than 0.68 and barley if it is more. Because the
-GCM-simulated drought risk never rises above this threshold, a
-yield-maximizing farmer will always plant a wheat monocrop.
+To explore these dynamics, I simulated a population of ``belief-based'' Bayesian agents who use their subjective perception of annual drought risks to decide what mix of crops will best manage those risks.
+During periods of climatic stability, allowing past experiences to influence decision making helps farmers minimize the impacts of \emph{predictable} drought.
+But past experiences are less informative during periods of rapid climate change, and even farmers who manage risk ``optimally'' in light of their prior beliefs can experience food shortfalls.
+Cognitive diversity and life experience can be as or more important than the exact mix of crops planted for a population's long-term survival under extreme uncertainty.
 
-Examining the long-term dynamics of crop yields among all
-decision-making strategies and psychological profiles reveals
-variability in yields due to different decision making preferences is
-far greater variability due to differences in risk perception (Figure
-7). Risk-minimizing strategies sacrifice productivity for
-predictability, consistently lowering the mean and variance of crop
-yields over time. Risk-minimizing strategies are more sensitive to risk
-perception than yield-maximizing strategies; varying the memory decay
-rate parameter \(\lambda\) had no effect on the optimal crop allocation
-for yield maximizers.
+These dynamics have implications for understanding risk management and food production in the Neolithic.
+Two points arise from this model that clarify our understanding of the earliest farming communities in the Mediterranean and can inform future simulation work:
 
-During periods of climatic stability, allowing past experiences to
-influence decision making helps farmers minimize the impacts of
-\emph{predictable} drought. But past experiences are less informative
-during periods of rapid climate change, and even farmers who manage risk
-``optimally'' experience major food shortfalls. Climatically-induced
-variability in food supplies consistently surpasses that from
-differences in risk-perception psychology, but not that from different
-risk aversion preferences. This finding suggests that the efficacy of
-risk managing strategies that rely on regulatory feedbacks is limited to
-periods of relative climatic stability. This is consistent with the
-observation from the Roman Period Mediterranean that while minor food
-crises were common in the ancient world, extreme famines were rare
-(Garnsey 1989).
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Risk perception is difficult.} The climate system is inherently chaotic.
+Annual forecasts are fundamentally uncertain, even in the era of modern supercomputers and numerical weather prediction.
+For prehistoric farmers, this uncertainty would have been an existential challenge.
+With the end of the Last Glacial Maximum and the advent of the Holocene, precipitation became increasingly volatile on multiple time scales.
+This uncertainty in the objective drought risk is compounded by uncertainty in individual farmers' subjective perception of drought risk.
+Any individual would have experienced only a brief snapshot of this complex period.
+The rhythms of human lifespans are out of sync with the centennial to millennial scale oscillations in the climate system, so even the predictive value of one's own experience is itself unpredictable.
+This fundamental uncertainty would have influenced far more than the choice of which type of crops to plant and would have pervaded all kinds of decision-making under risk.
+\item
+  \textbf{Individual risk perception has consequences for collective risk management.} Individuals of different ages may perceive the same dry period differently depending on their prior life experience.
+Younger individuals are more likely to perceive a run of dry years as a trend, rather than a temporary deviation from the norm, and older individuals are likely to do the opposite.
+By extension, the age structure of a population will influence how quickly it is able to perceive and adapt to a changing climate.
+For example, a young, fast growing population will have a different collective memory of a past drought event than an older population.
+Likewise, famines, warfare, epidemics, and other crises that afflict specific age classes will alter the time horizon of that population's collective memory.
+Individual heterogeneity in risk perceptions can thus play a key role in broader social responses to climatic risks.
+\end{enumerate}
 
-Statistically downscaling palaeoclimate simulations and coupling them to
-game theoretic models of decision making under uncertainty is a simple
-way to better understand the patterning of climate changes and the
-regulatory feedbacks farmers use to adapt to them. To further explore
-these patterns, future work should incorporate additional
-risk-minimizing strategies such as storage and exchange. Food storage
-and exchange can also be thought of as forms of diversification, the
-former in the time and the latter in space ({\textbf{???}}). Food
-exchange would be of particular interest from a game theoretic
-perspective, as incomplete information with respect to an exchange
-partner's crop diversification and storage practices can lead to
-situations of moral hazard.
+Relating collective knowledge to individual perception and cognition is essential for understanding human behavior in complex social-ecological systems (Beratan 2007; Pope and Gimblett 2015).
+Although social learning and cumulative cultural evolution have not been the focus of this chapter, these findings provide insight into the individual learning dynamics that underlay those broader social processes.
+This chapter has focused primarily on the physical and cognitive dimensions of risk and risk perception.
+The social context of risk and risk perception can be equally consequential (Rogers 1997).
+The balance of individual learning with social transmission determines the collective perception of risks and its impact on collective memory (Moussaïd 2013; Candia et al.
+2019).
+Our finite lifespans ultimately limit the skill of individual learning over the long term, and cumulative cultural evolution is necessary for continued survival.
 
-Think about the importance of the population's age structure, becuase
-that's going to impact the collective memory of drought.
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
 
 \hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}}
+
+\noindent
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Abbo2010}{}%
+Abbo, Shahal, Simcha Lev-Yadun, and Avi Gopher.
+2010.
+``Yield stability: an agronomic perspective on the origin of Near Eastern agriculture.'' \emph{Vegetation History and Archaeobotany} 19 (2): 143--50.
+\url{https://doi.org/10.1007/s00334-009-0233-7}.
+
+\leavevmode\hypertarget{ref-Agrawal1968}{}%
+Agrawal, R.
+C., and E.
+O.
+Heady.
+1968.
+``Applications of Game Theory Models in Agriculture.'' \emph{Journal of Agricultural Economics} 19 (2): 207--18.
+\url{https://doi.org/10.1111/j.1477-9552.1968.tb01326.x}.
+
+\leavevmode\hypertarget{ref-Amaral2020}{}%
+Amaral, Marco A., and Marco A. Javarone. 2020. ``Heterogeneity in evolutionary games: an analysis of the risk perception.'' \emph{arXiv Preprint}, 1--7. \url{http://arxiv.org/abs/2002.06645}.
+
+\leavevmode\hypertarget{ref-Anderies2006}{}%
+Anderies, John Martin.
+2006.
+``Robustness, institutions, and large-scale change in social-ecological systems: the Hohokam of the Phoenix Basin.'' \emph{Journal of Institutional Economics} 2 (02): 133.
+\url{https://doi.org/10.1017/S1744137406000312}.
+
 \leavevmode\hypertarget{ref-Slafer1999}{}%
-Araus, J. L., G. A. Slafer, and I. Romagosa. 1999. ``Durum wheat and
-barley yields in antiquity estimated from 13C discrimination of
-archaeological grains: a case study from the Western Mediterranean
-Basin.'' \emph{Australian Journal of Plant Physiology} 26 (4): 345.
+Araus, J.
+L., G.
+A.
+Slafer, and I.
+Romagosa.
+1999.
+``Durum wheat and barley yields in antiquity estimated from 13C discrimination of archaeological grains: a case study from the Western Mediterranean Basin.'' \emph{Australian Journal of Plant Physiology} 26 (4): 345.
 \url{https://doi.org/10.1071/PP98141}.
 
-\leavevmode\hypertarget{ref-Garnsey1989}{}%
-Garnsey, Peter. 1989. \emph{Famine and Food Supply in the Graeco-Roman
-World: Responses to Risk and Crisis}. Cambridge University Press.
+\leavevmode\hypertarget{ref-Beckenkamp2008}{}%
+Beckenkamp, Martin. 2008. ``Playing Strategically against Nature? Decisions Viewed from a Game-Theoretic Frame.'' Bonn. \url{https://doi.org/10.2139/ssrn.1275287}.
+
+\leavevmode\hypertarget{ref-Beratan2007}{}%
+Beratan, Kathi K. 2007. ``A cognition-based view of decision processes in complex social-ecological systems.'' \emph{Ecology and Society} 12 (1). \url{https://doi.org/10.5751/es-02103-120127}.
+
+\leavevmode\hypertarget{ref-Bissiri2010}{}%
+Bissiri, Pier Giovanni, and Stephen G.
+Walker.
+2010.
+``On Bayesian learning from Bernoulli observations.'' \emph{Journal of Statistical Planning and Inference} 140 (11): 3520--30.
+\url{https://doi.org/10.1016/j.jspi.2010.05.023}.
+
+\leavevmode\hypertarget{ref-Bonawitz2011}{}%
+Bonawitz, Elizabeth, Stephanie Denison, Annie Chen, Alison Gopnik, and Thomas L Griffiths.
+2011.
+``A simple sequential algorithm for approximating bayesian inference.'' In \emph{Proceedings of the 33rd Annual Conference of the Cognitive Science Society}, 2463--8.
+\url{papers://a6591aa3-238d-4eda-a3d9-0181a4361186/Paper/p1438\%20http://palm.mindmodeling.org/cogsci2011/papers/0582/paper0582.pdf}.
+
+\leavevmode\hypertarget{ref-Candia2019}{}%
+Candia, Cristian, C.
+Jara-Figueroa, Carlos Rodriguez-Sickert, Albert László Barabási, and César A.
+Hidalgo.
+2019.
+``The universal decay of collective memory and attention.'' \emph{Nature Human Behaviour} 3 (1): 82--91.
+\url{https://doi.org/10.1038/s41562-018-0474-5}.
+
+\leavevmode\hypertarget{ref-Cassidy1971}{}%
+Cassidy, R.
+G., C.
+A.
+Field, and M.
+J.
+L.
+Kirby.
+1971.
+``Random payoff games with partial information: One person games against nature.'' \emph{Revue Française d'informatique et de Recherche Opérationnelle} 5 (3): 3--17.
+
+\leavevmode\hypertarget{ref-Cushman2009}{}%
+Cushman, Fiery, and O Macindoe.
+2009.
+``The coevolution of punishment and prosociality among learning agents.'' \emph{Proceedings of the Annual Conference of the Cognitive Science Society} 31 (31): 1774--9.
+\url{http://141.14.165.6/CogSci09/papers/374/paper374.pdf}.
+
+\leavevmode\hypertarget{ref-Findlater2019}{}%
+Findlater, Kieran M., Terre Satterfield, and Milind Kandlikar.
+2019.
+``Farmers' Risk-Based Decision Making Under Pervasive Uncertainty: Cognitive Thresholds and Hazy Hedging.'' \emph{Risk Analysis}.
+\url{https://doi.org/10.1111/risa.13290}.
+
+\leavevmode\hypertarget{ref-Gallistel2014}{}%
+Gallistel, C.
+R., Monika Krishan, Ye Liu, Reilly Miller, and Peter E.
+Latham.
+2014.
+``The perception of probability.'' \emph{Psychological Review} 121 (1): 96--123.
+\url{https://doi.org/10.1037/a0035232}.
 
 \leavevmode\hypertarget{ref-GOULD1963a}{}%
-Gould, Peter R. 1963. ``Man Against His Environment: A Game Theoretic
-Framework.'' \emph{Annals of the Association of American Geographers} 53
-(3): 290--97. \url{https://doi.org/10.1111/j.1467-8306.1963.tb00450.x}.
+Gould, Peter R.
+1963.
+``Man Against His Environment: A Game Theoretic Framework.'' \emph{Annals of the Association of American Geographers} 53 (3): 290--97.
+\url{https://doi.org/10.1111/j.1467-8306.1963.tb00450.x}.
 
 \leavevmode\hypertarget{ref-Halstead1989}{}%
-Halstead, Paul, and John O'Shea. 1989. ``Introduction: cultural
-responses to risk and uncertainty.'' In \emph{Bad Year Economics:
-Cultural Responses to Risk and Uncertainty}, edited by Paul Halstead and
-John O'Shea, 1--7. Cambridge: Cambridge University Press.
+Halstead, Paul, and John O'Shea.
+1989.
+``Introduction: cultural responses to risk and uncertainty.'' In \emph{Bad Year Economics: Cultural Responses to Risk and Uncertainty}, edited by Paul Halstead and John O'Shea, 1--7.
+Cambridge: Cambridge University Press.
 \url{https://doi.org/10.2307/281441}.
@@
 \leavevmode\hypertarget{ref-He2011}{}%
-He, Feng. 2011. ``Simulating Transient Climate Evolution of the Last
-Deglatiation with CCSM3.'' PhD thesis, Madison.
+He, Feng. 2011. ``Simulating Transient Climate Evolution of the Last Deglatiation with CCSM3.'' PhD thesis, Madison.
 
-\bibliographystyle{spbasic}
+\leavevmode\hypertarget{ref-Helmers2001}{}%
+Helmers, G.
+A., C.
+F.
+Yamoah, and G.
+E.
+Varvel.
+2001.
+``Separating the impacts of crop diversification and rotations on risk.'' \emph{Agronomy Journal} 93 (6): 1337--40.
+\url{https://doi.org/10.2134/agronj2001.1337}.
+
+\leavevmode\hypertarget{ref-Khalvati2019}{}%
+Khalvati, Koosha, Seongmin A.
+Park, Saghar Mirbagheri, Remi Philippe, Mariateresa Sestito, Jean-Claude Dreher, and Rajesh P.
+N.
+Rao.
+2019.
+``Modeling Other Minds: Bayesian Inference Explains Human Choices in Group Decision Making.'' \emph{Science Advances} in press (November): aax8783.
+\url{https://doi.org/10.1101/419515}.
+
+\leavevmode\hypertarget{ref-Luce1989}{}%
+Luce, R Duncan, and Howard Raiffa. 1989. \emph{Games and decisions: Introduction and critical survey}. Courier Corporation.
+
+\leavevmode\hypertarget{ref-Marston2011}{}%
+Marston, John M. 2011. ``Archaeological markers of agricultural risk management.'' \emph{Journal of Anthropological Archaeology} 30 (2): 190--205. \url{https://doi.org/10.1016/j.jaa.2011.01.002}.
+
+\leavevmode\hypertarget{ref-Milnor1952}{}%
+Milnor, John. 1952. ``Games against nature.'' The RAND Corporation.
+
+\leavevmode\hypertarget{ref-Moussaid2013}{}%
+Moussaïd, Mehdi. 2013. ``Opinion formation and the collective dynamics of risk perception.'' \emph{PLoS ONE} 8 (12): 1--8. \url{https://doi.org/10.1371/journal.pone.0084592}.
+
+\leavevmode\hypertarget{ref-Paut2019}{}%
+Paut, Raphaël, Rodolphe Sabatier, and Marc Tchamitchian.
+2019.
+``Reducing risk through crop diversification: An application of portfolio theory to diversified horticultural systems.'' \emph{Agricultural Systems} 168: 123--30.
+\url{https://doi.org/10.1016/j.agsy.2018.11.002}.
+
+\leavevmode\hypertarget{ref-Pope2015}{}%
+Pope, Aloah J., and Randy Gimblett.
+2015.
+``Linking Bayesian and agent-based models to simulate complex social-ecological systems in semi-arid regions.'' \emph{Frontiers in Environmental Science} 3 (AUG): 1--9.
+\url{https://doi.org/10.3389/fenvs.2015.00055}.
+
+\leavevmode\hypertarget{ref-Rogers1997}{}%
+Rogers, George O.
+1997.
+``The dynamics of risk perception: How does perceived risk respond to risk events?'' \emph{Risk Analysis} 17 (6): 745--57.
+\url{https://doi.org/10.1111/j.1539-6924.1997.tb01280.x}.
+
+\leavevmode\hypertarget{ref-Sanborn2016}{}%
+Sanborn, Adam N., and Nick Chater. 2016. ``Bayesian Brains without Probabilities.'' \emph{Trends in Cognitive Sciences} 20 (12): 883--93. \url{https://doi.org/10.1016/j.tics.2016.10.003}.
+
+\leavevmode\hypertarget{ref-Simon1990}{}%
+Simon, Herbert A. 1990. ``Bounded Rationality.'' In \emph{Utility and Probability}, 15--18. London: Palgrave Macmillan UK. \url{https://doi.org/10.1007/978-1-349-20568-4_5}.
+
+\leavevmode\hypertarget{ref-Tucker2007}{}%
+Tucker, Bram.
+2007.
+``Perception of Interannual Covariation and Strategies for Risk Reduction among Mikea of Madagascar: Individual and Social Learning.'' \emph{Human Nature} 18 (2): 162--80.
+\url{https://doi.org/10.1007/s12110-007-9007-z}.
+
+\leavevmode\hypertarget{ref-Tucker2013}{}%
+Tucker, Bram, Jaovola Tombo, Tsiazonera, Patricia Hajasoa, Charlotte Nagnisaha, Vorisoa Rene Lahitoka, and Christian Zahatsy.
+2013.
+``Beyond Mean and Variance: Economic Risk Versus Perceived Risk of Farming, Foraging, and Fishing Activities in Southwestern Madagascar.'' \emph{Human Ecology} 41 (3): 393--407.
+\url{https://doi.org/10.1007/s10745-013-9563-2}.
+
+\leavevmode\hypertarget{ref-Upton1987}{}%
+Upton, Martin. 1987. \emph{African farm management}. CUP Archive.
+
+\leavevmode\hypertarget{ref-Weiss2011}{}%
+Weiss, Ehud, and Daniel Zohary. 2011. ``The Neolithic Southwest Asian founder crops their biology and archaeobotany.'' \emph{Current Anthropology} 52 (SUPPL. 4). \url{https://doi.org/10.1086/658367}.
+
+\leavevmode\hypertarget{ref-Wilkinson1997}{}%
+Wilkinson, Tony J.
+1997.
+``Environmental Fluctuations, Agricultural Production and Collapse: A View from Bronze Age Upper Mesopotamia.'' In \emph{Third Millenium Bc Climate Change and Old World Collapse}, edited by Hasan Nüzhet Dalfes, G.
+Kukla, and Harvey Weiss, 66--106.
+Springer.
+\url{http://books.google.com/books?id=LHIRAQAAIAAJ}.
+
+\bibliographystyle{Formato\_bibliografia\_Springer}
 \bibliography{bibliography.bib}
